### 2018-07-29 doc/Paper/authors.tex
migration to desc

% Author list file generated with: mkauthlist 1.2.3 
% mkauthlist -j mnras -f -c LSST Dark Energy Science Collaboration --cntrb contributions.tex authors.csv authors.tex

\author[LSST Dark Energy Science Collaboration]{
\parbox{\textwidth}{
\Large
David Alonso$^{1,2}$,
Javier Sanchez$^{3}$,
An\v{z}e Slosar$^{4}$
\begin{center} (LSST Dark Energy Science Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
%\scriptsize
$^{1}$ School of Physics and Astronomy, Cardiff University, The Parade, Cardiff, CF24 3AA, United Kingdom\\
$^{2}$ Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom\\
$^{3}$ Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA\\
$^{4}$ Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA\\
}
}

### 2018-07-29 doc/doc_C_API.tex
migration to desc

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx} % Include figure files
\usepackage{fullpage}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amssymb,amsmath} 
\usepackage[colorlinks=true,linkcolor=Blue,linktoc=page,citecolor=Red,urlcolor=BrickRed]{hyperref}
%\usepackage{color}
%\usepackage[utf8x]{inputenc}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{listings}
%\usepackage{makeidx}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{epigraph}
\usepackage{titling}
\lstset{
language=C,
basicstyle=\footnotesize,
numbers=left,
numberstyle=\footnotesize,
stepnumber=1,
numbersep=5pt,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=false,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
escapeinside={\%*}{*)}
}
\pretitle{\begin{flushleft}\LARGE\bfseries\sffamily}
\posttitle{\par\end{flushleft}\vskip 0.5em\vspace{-7ex}}
\predate{\begin{flushleft}\sffamily\large}
\postdate{\par\end{flushleft}\vskip 0.5em\vspace{-2ex}}
\titleformat{\section}[hang]{\bfseries\sffamily\Large}
{\thesection}
{12pt}{\filright}

\titleformat{\subsection}[hang]{\bfseries\sffamily}
{\thesubsection}
{12pt}{\filright}

\titleformat{\subsubsection}[hang]{\bfseries}
{\thesubsubsection}
{12pt}{\filright}

\newcommand{\vari}[2]{\textcolor{OliveGreen}{#1}{#2}}
\newcommand{\macr}[1]{{\tt \textcolor{BrickRed}{#1}}}
\newcommand{\functi}[3]{{\tt\textcolor{OliveGreen}{#1}
    \textcolor{blue}{#2}(#3)}}

\newenvironment{funcdef}[2][void main(void)]
{\vspace{12pt}\par\noindent{\large{\bf #1}}\vspace{5pt}
 \par\noindent\leftskip=20pt{{#2}}\vspace{5pt}
 \par\noindent\leftskip=5pt\rightskip=5pt}
{\par\leftskip=0pt\rightskip=0pt\vspace{12pt}}

\title{NaMaster: C API and code documentation}

\begin{document}

\maketitle
\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}
\tableofcontents
\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}

\section{Introduction}

NaMaster is a C library, Python module and standalone program to compute full-sky angular cross-power spectra of masked, spin-0 and spin-2 fields with an arbitrary number of known contaminants using a pseudo-$C_\ell$ (aka MASTER) approach.

\subsection{Dependencies}
The following packages need to be installed before NaMaster.
\begin{itemize}
  \item {\bf GSL}: the GNU Scientific Library. This should be available in your usual software repositories (e.g. synaptic for linux), but you can also download and install it from \url{https://www.gnu.org/software/gsl/} (the installation follows the usual {\tt ./configure}, {\tt make}, {\tt make install} procedure.
  \item {\bf HEALPix}: the C HEALPix subroutines are also needed. HEALPix can be downloaded from \url{http://healpix.jpl.nasa.gov/}, and the C library can be compiled following the instructions after typing {\tt ./configure} and then {\tt make c-all}. The header and library files will then be placed in two local folders called {\tt include} and {\tt lib}. The user should then feel free to move these files to a different location.
  \item {\bf CFITSIO}: a FITS file subroutine library. This is used to read/write HEALPix maps in FITS format. {\tt cfitsio} can be downloaded from \url{http://heasarc.gsfc.nasa.gov/fitsio/fitsio.html}.
  \item {\bf Libsharp}: a fast C library for spherical harmonic transforms. Libsharp can be downloaded from its github repository: \url{https://github.com/dagss/libsharp}. After cloning it, you should run {\tt autoreconf -i} to generate the configuration file. Then run {\tt ./configure} and {\tt make}, which will compile the library and place all compilation products in the folder {\tt ./auto}. The corresponding header and library files should then be manually moved to the desired installation directories.
\end{itemize}

\subsection{Compilation and usage}\label{ssec:compile}
\subsubsection*{Installing the C library and binaries}
NaMaster uses {\tt autotools} for installation, which means that you should be able to install it by simply typing
\begin{verbatim}
   ~$ ./configure
   ~$ make
   ~$ make install
\end{verbatim}
If you don't have default admin privileges you may have to precede the last command by {\tt sudo}. If you don't have admin privileges at all (i.e. you can't modify the contents of standard directories such as {\tt /usr/lib}), you can still install NaMaster by substituting the first command by {\tt ./configure --prefix=/path/to/dir}, where {\tt /path/to/dir} should be the full path of the directory where you want to install this package. This will create three sub-directories: {\tt /path/to/dir/include}, {\tt /path/to/dir/lib}, {\tt /path/to/dir/bin}, where the header, library and binary files will be placed respectively.

Note that, if you don't have admin privileges, probably some of the dependencies listed in the previous sections will also be installed in non-standard paths. If that is the case, you should make sure the environment variables {\tt CPPFLAGS} and {\tt LDFLAGS} contain the corresponding {\tt -I/path/to/dir2/include} and {\tt -L/path/to/dir2/lib} tags that point to the directories where these dependencies are installed (see \href{https://www.gnu.org/software/make/manual/html_node/Implicit-Variables.html}{this link} for more details on {\tt make} implicit variables).

\subsection*{Installing the python wrapper}
NaMaster comes equipped with a python wrapper. This is installed by running
\begin{verbatim}
   ~$ python setup.py install
\end{verbatim}
Without admin privileges you can still make this work by running
\begin{verbatim}
   ~$ python setup.py install --user
\end{verbatim}
The python wrapper needs to link with NaMaster's library. If the latter was installed in a non-standard path (e.g. {\tt /path/to/install}), you'll need to pass the corresponding directory to {\tt setup.py}. You can do so by running
\begin{verbatim}
   ~$ python setup.py build_ext --library-dirs=/path/to/install/lib/
       --rpath=/path/to/install/lib/
\end{verbatim}
before the install commands above.

\subsection*{Linking with the C library}
If you want to use NaMaster on your own C code you'll need to be able to link with {\tt libnmt} (the NaMaster C library). There are two main things to do:
\begin{enumerate}
 \item Make sure to include the NaMaster header in any C file that makes use of any of the NaMaster subroutines:
       \begin{verbatim}
         #include <namaster.h>
       \end{verbatim}
 \item When compiling your code, make sure you link to {\tt libnmt} and all dependencies. In the simplest case, assuming you have written a C script called {\tt min\_code.c}, the following should work:
       \begin{verbatim}
         gcc -fopenm -I/path/to/nmt/include min_code.c -o min_code
             -L/path/to/nmt/lib -lnmt -lsharp -lfftpack -lc_utils
             -lchealpix -lcfitsio -lgsl -lgslcblas -lm
       \end{verbatim}
       where {\tt /path/to/nmt/include} and {\tt /path/to/nmt/lib} are the directories where {\tt namaster.h} and {\tt libnmt.so} are installed.
\end{enumerate}
Section \ref{sec:sample} below contains a fully working C script that calls the NaMaster library.

\section{NaMaster - the program}
NaMaster comes with its own executable that computes the pseudo-$C_\ell$ power spectrum of two input masked fields with possible contaminants. After installation (and assuming it has been installed in an accessible directory), the program can be invoked by typing {\tt namaster}.

The user interacts with {\tt namaster} by providing parameters as command-line arguments in the form of a pair: {\tt -<param\_name> <param>}. After typing
\begin{verbatim}
  ~$ namaster -h
\end{verbatim}
the user is presented with a list of all available input parameters:
\begin{verbatim}
Usage: namaster -<opt-name> <option>
Options:
  -map      -> path to file containing map(s)
  -map_2    -> path to file containing 2nd map(s) (optional)
  -mask     -> path to file containing mask
  -mask_2   -> path to file containing mask for 2nd map(s) (optional)
  -temp     -> path to file containing contaminant templates (optional)
  -temp_2   -> path to file containing contaminant templates
               for 2nd map(s) (optional)
  -pol      -> spin-0 (0) or spin-2 (1) input map(s)
  -pol_2    -> spin-0 (0) or spin-2 (1) 2nd input map(s)
  -pure_e   -> use pure E-modes for 1st maps? (0-> no or 1-> yes (default->no))
  -pure_b   -> use pure B-modes for 1st maps? (0-> no or 1-> yes (default->no))
  -pure_e_2 -> use pure E-modes for 2nd maps? (0-> no or 1-> yes (default->no))
  -pure_b_2 -> use pure B-modes for 2nd maps? (0-> no or 1-> yes (default->no))
  -cl_noise -> path to file containing noise Cl(s)
  -cl_guess -> path to file containing initial guess for the Cl(s)
  -coupling -> path to file containing coupling matrix (optional)
  -out      -> output filename
  -binning  -> path to file containing binning scheme
  -nlb      -> number of ells per bin (used only if -binning isn't used)
  -h        -> this help
\end{verbatim}

Some clarification is in order regarding some of these parameters:
\begin{itemize}
 \item {\tt map} and {\tt map\_2}: these arguments should point to FITS files containing the maps of the two fields to correlate. These files should contain one map if {\tt pol} (or {\tt pol\_2}) is 0 and two maps ($Q$ and $U$) if {\tt pol} (or {\tt pol\_2}) is 1. If {\tt map\_2} is not passed, the auto-correlation of {\tt map} will be computed (and all other {\tt \_2} options will be ignored).
 \item {\tt mask} and {\tt mask\_2}: these arguments should point to FITS files containing the masks of the two fields to correlate (one map per file).
 \item {\tt temp} and {\tt temp\_2}: these arguments should point to FITS files containing the contaminant templates you suspect are poluting your fields. The number of contaminants will be derived by the code based on the number of maps in the files. Note that an exception will occur if this number is not a multiple of the number of maps corresponding to the field's spin (see above).
 \item {\tt cl\_noise}: this file should contain the expected noise bias (see discussion below Eq. \ref{eq:master_full}) to be subtracted from the estimated power spectrum. The format of this file should be $N+1$ columns, with the first column being the multipole number $\ell$ (starting at $\ell=0$), and the next $N$ being the different component of the power spectrum. $N$ should correspond to the number of power spectra expected given the spins of the two fields being correlated. I.e. $N=1$ if {\tt pol=pol\_2=0} ($(TT)$), $N=2$ if {\tt pol=0}, and {\tt pol\_2=1} (or vice-versa, $(TE,TB)$), and $N=4$ if {\tt pol=pol\_2=1} ($(EE,EB,BE,BB)$). If this option is not passed, it is assumed to be $0$. This file should contain $3{\tt nside}$ rows, where {\tt nside} is the HEALPix resolution parameter of the input maps.
 \item {\tt cl\_guess}: proposal power spectrum (i.e. best guess of true power spectrum) used to compute the deprojection bias (same format as {\tt cl\_noise} above). This is only relevant for contaminated fields. If not passed, the proposal power spectrum is set to $0$.
 \item {\tt out}: path to output ASCII file. First column will correspond to the effective $\ell$ in the bandpower. All subsequent columns correspond to the different components of the power spectrum (as described above depending on the field spins).
 \item {\tt binning}: path to ASCII file describing the binning scheme defining the output bandpowers. This file should contain three columns: the first column should contain the bandpower index of the different multipoles, the second column should contain the multipoles and the third column should contain the weight corresponding to each multipole (the sum of weights in each bandpower is automatically normalized to $1$).
 \item {\tt nlb}: number of multipoles per bandpower. In this case the output power spectrum will be computed by binning multipoles in sets of {\tt nlb} from $\ell=2$ to $\ell=3{\tt nside}-1$.
\end{itemize}

\section{C documentation}
Important note: all HEALPix maps passed to NaMaster routines should be in RING order.

\subsection{Fields}
The definition of the fields to be correlated (including their masks and possible contaminants) is handled through a C structure called {\tt nmt\_field}. The following routines exist to manage this structure:

\begin{funcdef}[nmt\_field\_alloc]
 {\functi{nmt\_field *}{nmt\_field\_alloc}
 {\vari{long}{ nside},\vari{double *}{mask},\vari{int}{ pol},\vari{double **}{maps},\\\vari{int}{ ntemp},\vari{double ***}{temp},\vari{double *}{beam},\vari{int }{pure\_e},\vari{int }{pure\_b}}}
 This is the constructor for {\tt nmt\_field}. The input variables are:
 \begin{itemize}[noitemsep]
   \item {\tt nside}: the HEALPix resolution of all maps involved
   \item {\tt mask}: sky mask (as a single scalar HEALPix map).
   \item {\tt pol}: set to 0 if this is a spin-0 field. Set to 1 if it's a spin-2 field.
   \item {\tt maps}: set of maps corresponding to observed field. This would correspond to one map for a scalar field or two maps for spin-2 quantities (e.g. Q and U for polarization maps or $\gamma_1$, $\gamma_2$ for cosmic shear). The first dimension of this double array would correspond to the number of maps, while the second dimension runs through the different pixels of each map.
   \item {\tt ntemp}: number of contaminant templates for this field.
   \item {\tt temp}: contaminant templates as HEALPix maps. The first dimension should run through the different templates, the second dimension corresponds to the number of maps per template (e.g. 1 for spin-0 and 2 for spin-2) and the third dimension corresponds to the number of pixels.
   \item {\tt beam}: spherical harmonic transform of the instrumental beam (assumed to be rotationally symmetric - i.e. no $m$ dependence). If {\tt NULL}, no beam will be corrected for. Otherwise, this array should have $3\times{\tt nside}$ elements, corresponding to multipoles $\ell\in[0,3\,{\tt nside}-1]$.
   \item {\tt pure\_e}, {\tt pure\_b}: set to a value different from 0 if you want to use pure-$E/B$ modes for this field. Note that, in this case you should make sure the mask and its first derivatives are continuous on its boundary (see {\tt nmt\_apodize\_mask} below).
 \end{itemize}
 \end{funcdef}

\begin{funcdef}[nmt\_field\_read]
 {\functi{nmt\_field *}{nmt\_field\_read}
 {\vari{char *}{fname\_mask},\vari{char *}{fname\_maps},\vari{char *}{fname\_temp},\\\vari{char *}{fname\_beam},\vari{int }{pol},\vari{int }{pure\_e},\vari{int }{pure\_b}}}
 As {\tt nmt\_field\_alloc}, this returns a pointer to a {\tt nmt\_field} structure based on:
 \begin{itemize}[noitemsep]
   \item {\tt pol}: set to 0 if this is a spin-0 field. Set to 1 if it's a spin-2 field.
   \item {\tt fname\_mask}: file name pointing to a FITS file containing the sky mask (as a single scalar HEALPix map).
   \item {\tt fname\_maps}: file name pointing to a FITS file containing the maps of the observed field. This file should contain a single map for {\tt pol}=0 and two maps for {\tt pol}=1.
   \item {\tt fname\_temp}: file name pointing to a FITS file containing the contaminant templates as HEALPix maps. Each template should contain $N$ maps with $N=1$ for {\tt pol}=0 and $N=2$ for {\tt pol}=1. Pass ``none'' if no contaminants are needed (in which case deprojection and debiasing will not take place).
   \item {\tt fname\_beam}: file name pointing to an ASCII file containing the spherical harmonic transform of the instrumental beam (assumed to be rotationally symmetric - i.e. no $m$ dependence). If {\tt ``none''}, no beam will be corrected for. Otherwise, this file should contain two columns ($\ell$ and $b_\ell$) and $3\times{\tt nside}$ columns, corresponding to multipoles $\ell\in[0,3\,{\tt nside}-1]$.
   \item {\tt pure\_e}, {\tt pure\_b}: set to a value different from 0 if you want to use pure-$E/B$ modes for this field. Note that, in this case you should make sure the mask and its first derivatives are continuous on its boundary (see {\tt nmt\_apodize\_mask} below).
 \end{itemize}
\end{funcdef}

\begin{funcdef}[nmt\_field\_free]
 {\functi{void}{nmt\_field\_free}{\vari{nmt\_field *}{fl}}}
 This frees up all memory associated to a previously-allocated {\tt nmt\_field}.
\end{funcdef}

\subsection{Binning scheme}
The definition of bandpowers is managed through C structures called {\tt nmt\_binning\_scheme}. The following routines allow you to interact with this structure:

\begin{funcdef}[nmt\_bins\_constant]
 {\functi{nmt\_binning\_scheme *}{nmt\_bins\_constant}{\vari{int }{nlb},\vari{int }{lmax}}}
 Creates an {\tt nmt\_binning\_scheme} structure (and returns a pointer to it) where the bandpowers are constant intervals of {\tt nlb} multipoles with equal weights between $\ell=2$ and $\ell={\tt lmax}$.
\end{funcdef}

\begin{funcdef}[nmt\_bins\_create]
{\functi{nmt\_binning\_scheme *}{nmt\_bins\_create}{\vari{int }{nell},\vari{int *}{bpws},\vari{int *}{ells},\\\vari{double *}{weights},\vari{int }{lmax}}}
 Creates an {\tt nmt\_binning\_scheme} structure (and returns a pointer to it) with bandpowers defined by the following parameters:
 \begin{itemize}[noitemsep]
  \item {\tt ells}: array of multipole indices
  \item {\tt bpws}: array containing the band power each $\ell$ in {\tt ells} corresponds to.
  \item {\tt weights}: array containing the weight for each $\ell$ in {\tt ells}. These need not be normalized, but they will be normalized such that the sum of weights within each bandpower equals 1.
  \item {\tt nell}: number of elements in the three previous arrays.
  \item {\tt lmax}: all multipoles $\ell>{\tt lmax}$ will be ignored.
 \end{itemize}
\end{funcdef}

\begin{funcdef}[nmt\_bins\_read]
{\functi{nmt\_binning\_scheme *}{nmt\_bins\_read}{\vari{char *}{fname},\vari{int }{lmax}}}
 Creates an {\tt nmt\_binning\_scheme} structure (and returns a pointer to it) with bandpowers defined by the contents of an ASCII file with name {\tt fname}. This file should contain three columns corresponding to the arrays {\tt bpws}, {\tt ells} and {\tt weights} passed to {\tt nmt\_bins\_create}. All multipoles $\ell>{\tt lmax}$ will be ignored.
\end{funcdef}

\begin{funcdef}[nmt\_bins\_free]
{\functi{void}{nmt\_bins\_free}{\vari{nmt\_binning\_scheme *}{bin}}}
 Frees all memory associated with an allocated {\tt nmt\_binning\_scheme} structure.
\end{funcdef}

\begin{funcdef}[nmt\_bin\_cls]
{\functi{void}{nmt\_bin\_cls}{\vari{nmt\_binning\_scheme *}{bin},\vari{double **}{cls\_in},\vari{double **}{cls\_out},\vari{int }{ncls}}}
 Performs a binning operation:
 \begin{equation}
   B_k=\sum_{\ell\in\vec{\ell}_k}w_\ell C_\ell.
 \end{equation}
 Here, $C_\ell\rightarrow{\tt cls\_in}$ is a set of {\tt ncls} angular power spectra, $B_k\rightarrow{\tt cls\_out}$ is a set of bandpowers and $w_\ell$ and $\vec{\ell}_k$ are the weights and multipole ranges defining the binning scheme {\tt bin}. Both {\tt cls\_in} and {\tt cls\_out} should have been previously allocated. The first dimension of both {\tt cls\_in} and {\tt cls\_out} should run from 0 to ${\tt ncls}-1$. Their second dimension should correspond to the number of multipoles and bandpowers used to create {\tt bin} respectively (for instance, the latter can be accessed as {\tt bin->n\_bands}).
\end{funcdef}

\begin{funcdef}[nmt\_unbin\_cls]
{\functi{void}{nmt\_unbin\_cls}{\vari{nmt\_binning\_scheme *}{bin},\vari{double **}{cls\_in},\vari{double **}{cls\_out},\vari{int }{ncls}}}
 Performs a un-binning operation:
 \begin{equation}
   C_\ell=\sum_k\,B_k\Theta(\ell\in\vec{\ell}_k)
 \end{equation}
 Here, $B_k\rightarrow{\tt cls\_in}$ is a set of {\tt ncls} bandpowers, $C_\ell\rightarrow{\tt cls\_out}$ is a set of angular power spectra and $w_\ell$ and $\vec{\ell}_k$ are the weights and multipole ranges defining the binning scheme {\tt bin}. The function $\Theta(\ell\in\ell_k)$ is $1$ for all multipoles contained in the $k$-th bandpower and zero otherwise. Both {\tt cls\_in} and {\tt cls\_out} should have been previously allocated. The first dimension of both {\tt cls\_in} and {\tt cls\_out} should run from 0 to ${\tt ncls}-1$. Their second dimension should correspond to the number of bandpowers and multipoles used to create {\tt bin} respectively (for instance, the former can be accessed as {\tt bin->n\_bands}).
\end{funcdef}

\begin{funcdef}[nmt\_ell\_eff]
 {\functi{void}{nmt\_ell\_eff}{\vari{nmt\_binning\_scheme *}{bin},\vari{double *}{larr}}}
 This function returns, in the output array {\tt larr}, the effective multipole corresponding to each bandpower defined by {\tt bin}. This is computed as:
 \begin{equation}
   \ell^{\rm eff}_k=\sum_{\ell\in\vec{\ell}_k}\ell\,w_\ell,
 \end{equation}
 where $w_\ell$ are the bandpower weights. {\tt larr} should have been previously allocated to the number of bandpowers defined by {\tt bin}.
\end{funcdef}

\subsection{Pseudo-$C_\ell$}
The implementation of the pseudo-$C_\ell$ estimator can be split into the following steps:
\begin{enumerate}
  \item Clean up your best guess of the known contaminants in your data maps. This step can be skipped if you think your maps are clean of contaminants. This step is automatically carried out when initializing an {\tt nmt\_field} structure with template contaminants.
  \item Compute the cross-pseudo-$C_\ell$ of the cleaned maps $\tilde{C}^{\rm clean}_\ell$.
  \item Compute the bias on the pseudo-$C_\ell$ caused by the statistical residual contaminants $\tilde{C}^{\rm cont}_\ell$.
  \item Compute the mode-coupling matrix associated with the field masks ${\sf M}_{\ell\ell'}$.
  \item Compute the de-coupled and de-biased bandpowers:
        \begin{equation}\label{eq:master_full}
          B_k=\sum_{k'}(\mathcal{M})^{-1}_{kk'}\sum_{\ell\in\vec{\ell}_{k'}}w_{\ell'}\left[\tilde{C}^{\rm clean}_{\ell'}-\tilde{C}^{\rm cont}_{\ell'}-\tilde{C}^{\rm noise}_{\ell'}\right],
        \end{equation}
        where $\mathcal{M}$ is the binned coupling matrix:
        \begin{equation}
          \mathcal{M}_{kk'}\equiv\sum_{\ell\in\vec{\ell}_k}\sum_{\ell'\in\vec{\ell}_{k'}}w_\ell{\sf M}_{\ell\ell'}.
        \end{equation}
        When auto-correlating a field with noise, it is in general also desirable to remove the noise bias on the power spectrum. This has been included in Eq. \ref{eq:master_full} above as $\tilde{C}^{\rm noise}_\ell$. This should be the pseudo-$C_\ell$ of the noise component (i.e. the angular power spectrum of masked noise realizations), which can be computed from Monte-Carlo simulations (or analytically for sufficiently simple noise models).
\end{enumerate}
In NaMaster, these computations are carried out through a C structure called {\tt nmt\_workspace}. The relevant functions are described below. Once the pseudo-$C_\ell$ coupling matrix corresponding to two fields has been computed and stored in an {\tt nmt\_workspace}, the same workspace can be re-used to compute the pseudo-$C_\ell$ of any pair of fields as long as they have the same polarizations, masks and choices for $E$ and $B$ purification as the original fields.

Note that the input and output power spectra are given as 2D arrays. The first dimension runs through $N_{\rm spec}$, the number of different cross-spectra:
\begin{enumerate}
 \item For two spin-0 fields $f_1$ and $f_2$, $N_{\rm spec}=1$: $C_\ell=\left(C^{f_1f_2}_\ell\right)$.
 \item For a spin-0 field $f_1$ and a spin-2 field $f_2$, $N_{\rm spec}=2$: $C_\ell=\left(C^{f_1f_2^E}_\ell,C^{f_1f_2^B}_\ell\right)$, where $f_2^{E,B}$ are the $E$ and $B$-modes of $f_2$.
 \item For two spin-2 fields $f_1$ and $f_2$, $N_{\rm spec}=4$: $C_\ell=\left(C^{f_1^Ef_2^E}_\ell,C^{f_1^Ef_2^B}_\ell,C^{f_1^Bf_2^E}_\ell,C^{f_1^Bf_2^B}_\ell\right)$, where $f_x^{E,B}$ are the $E$ and $B$-modes of $f_x$.
\end{enumerate}
The second dimension runs through the different multipole indices or bandpowers. For power spectra, before binning into bandpowers, this index runs from 0 to $\ell_{\rm max}=3\,{\tt nside}-1$, where {\tt nside} is the HEALPix resolution of the fields. For bandpowers, this index runs through the number of different bandpowers defined by the associated {\tt nmt\_binning\_scale} structure.

\begin{funcdef}[nmt\_compute\_coupling\_matrix]
{\functi{nmt\_workspace *}{nmt\_compute\_coupling\_matrix}{\vari{nmt\_field *}{fl1},\vari{nmt\_field *}{fl2},\\\vari{nmt\_binning\_scheme *}{bin}}}
 Computes the coupling matrix and the binned coupling matrix for the two fields {\tt fl1} and {\tt fl2} and the binning scheme {\tt bin}. Note that the only information needed from the two fields is their masks and spins. These matrices are stored internally in the {\tt nmt\_workspace} structure returned by this function.
\end{funcdef}

\begin{funcdef}[nmt\_workspace\_write]
{\functi{void}{nmt\_workspace\_write}{\vari{nmt\_workspace *}{w},\vari{char *}{fname}}}
 Writes an {\tt nmt\_workspace} structure into a file {\tt fname} (using an internal binary format).
\end{funcdef}

\begin{funcdef}[nmt\_workspace\_read]
{\functi{nmt\_workspace *}{nmt\_workspace\_read}{\vari{char *}{fname}}}
 Returns a pointer to a {\tt nmt\_workspace} structured read from file {\tt fname}. This file should have been generated by {\tt nmt\_workspace\_write}. These two functions are useful when computing the power spectrum of several pairs of fields with the same pairs of masks, for which the coupling matrices only need to be computed once.
\end{funcdef}

\begin{funcdef}[nmt\_workspace\_free]
{\functi{void}{nmt\_workspace\_free}{\vari{nmt\_workspace *}{w}}}
 Frees up all memory associated with an {\tt nmt\_workspace} structure.
\end{funcdef}

\begin{funcdef}[nmt\_compute\_deprojection\_bias]
{\functi{void}{nmt\_compute\_deprojection\_bias}{\vari{nmt\_field *}{fl1},\vari{nmt\_field *}{fl2},\\\vari{double **}{cl\_proposal},\vari{double **}{cl\_bias}}}
 Estimates the bias to the cross-power spectrum of two fields {\tt fl1} and {\tt fl2} induced by the contaminant cleaning (i.e. $C^{\rm cont}_\ell$ in Eq. \ref{eq:master_full}). This is returned into the variable {\tt cl\_bias}, which should have been previously allocated (see description in the introduction to this section). The estimate of this bias depends on a guess for the true power spectrum of both fields, given by {\tt cl\_proposal}\footnote{Thus, the pseudo-$C_\ell$ can be thought of as a recursive algorithm, where the estimate of the true power spectrum in a previous iteration is used as a proposal for the computation of the contaminant bias in the next one.}. Note that this operation does not require knowledge of the mode-coupling matrix, and therefore no {\tt nmt\_workspace} structure is needed.
\end{funcdef}

\begin{funcdef}[nmt\_compute\_coupled\_cell]
{\functi{void}{nmt\_compute\_coupled\_cell}{\vari{nmt\_field *}{fl1},\vari{nmt\_field *}{fl2},\vari{double **}{cl\_out},\vari{int }{iter}}}
 This computes the full-sky angular cross-power spectrum of two masked fields {\tt fl1} and {\tt fl2} without aiming to deconvolve the mode-coupling matrix. Effectively, this is equivalent as calling the usual HEALPix {\tt anafast} routine on the masked and contaminant-cleaned maps. The coupled power spectrum is returned in {\tt cl\_out}, which should have been previously allocated (see description in the introduction to this section). The variable {\tt iter} corresponds to the number of iterations used to compute the spherical harmonic transform. A value of 0 will correspond to the fastest but most inaccurate computation. Higher values will yield more accurate results at high-$\ell$ (${\tt niter}=3$ is usually enough in most cases).

 Since no attempt is made to deconvolve the mode-coupling matrix, this function does not require a {\tt nmt\_workspace} structure.
\end{funcdef}

\begin{funcdef}[nmt\_decouple\_cl\_l]
{\functi{void}{nmt\_decouple\_cl\_l}{\vari{nmt\_workspace *w},\vari{double **}{cl\_in},\vari{double **}{cl\_noise\_in},\\\vari{double **}{cl\_bias},\vari{double **}{cl\_out}}}
 This function performs the operation in Eq. \ref{eq:master_full}: debiasing and decoupling of a power spectrum computed from {\tt nmt\_compute\_coupled\_cell}. The coupled power spectrum $\tilde{C}_\ell^{\rm clean}$ must be provided in {\tt cl\_in}. The contaminant bias $\tilde{C}_\ell^{\rm cont}$ (e.g. computed through {\tt nmt\_compute\_deprojection\_bias}) and noise bias $\tilde{C}_\ell^{\rm noise}$ must be provided through {\tt cl\_bias} and {\tt cl\_noise\_in} respectively. The mode-coupling matrix ${\sf M}$ (and its binned version $\mathcal{M}$) are stored within {\tt w}, and the de-coupled bandpowers are returned in {\tt cl\_out}.
\end{funcdef}

\begin{funcdef}[nmt\_compute\_power\_spectra]
{\functi{nmt\_workspace *}{nmt\_compute\_power\_spectra}{\vari{nmt\_field *}{fl1},\vari{nmt\_field *}{fl2},\\\vari{nmt\_binning\_scheme *}{bin},\vari{nmt\_workspace *}{w0},\vari{double **}{cl\_noise},\vari{double **}{cl\_proposal},\\\vari{double **}{cl\_out}}}
 Carries out steps 2-5 of the pseudo-$C_\ell$ estimator described in the introduction of this section.
 {\tt fl1} and {\tt fl2} are the two fields to correlate, {\tt bin} defines the output bandpowers, {\tt cl\_noise} is the noise bias, {\tt cl\_proposal} is the best guess for the true power spectrum needed to estimate the contaminant bias $\tilde{C}^{\rm cont}_\ell$ (see {\tt nmt\_compute\_deprojection\_bias}). The output bandpowers are stored in {\tt cl\_out}, which should have been pre-allocated.

 This function also accepts an input pointer to a {\tt nmt\_workspace} structure, {\tt w0}. If a {\tt NULL} pointer is passed, the function will compute the mode-coupling matrix and return a newly-allocated {\tt nmt\_workspace} structure containing this information. Otherwise, the function will skip this computation and use the mode-coupling matrix stored in {\tt w0}. In this latter case, the function would return a pointer to {\tt w0}. Note that a call to this function is equivalent to a successive call to {\tt nmt\_compute\_coupling\_matrix}, {\tt nmt\_compute\_deprojection\_bias}, {\tt nmt\_compute\_coupled\_cell} and {\tt nmt\_decouple\_cl\_l}.
\end{funcdef}

\begin{funcdef}[nmt\_couple\_cl\_l]
{\functi{void}{nmt\_couple\_cl\_l}{\vari{nmt\_workspace *}{w},\vari{double **}{cl\_in},\vari{double **}{cl\_out}}}
 Convolves an input power spectrum {\tt cl\_in} with the mode-coupling matrix stored in {\tt w}, and provides the output in {\tt cl\_out}. I.e.:
 \begin{equation}
   C^{\rm out}_\ell=\sum_{\ell'}{\sf M}_{\ell\ell'}C^{\rm in}_{\ell'}
 \end{equation}
 where $C^{\rm out}_\ell\rightarrow{\tt cl\_out}$ and $C^{\rm in}_\ell\rightarrow{\tt cl\_in}$.
\end{funcdef}

\subsection{Utility functions}
\begin{funcdef}[nmt\_apodize\_mask]
{\functi{void}{nmt\_apodize\_mask}{\vari{long }{nside},\vari{double *}{mask\_in},\vari{double *}{mask\_out},\vari{double }{aposize},\\\vari{char *}{apotype}}}
 This function apodizes an input mask, provided in {\tt mask\_in} as a HEALPix map, and stores the result in {\tt mask\_out}. The apodization is defined by an apodization scale {\tt aposize} (in degrees) and an apodization type {\tt apotype}. Three different apodization types are supported (in what follows $\theta_*$ will be the apodization scale {\tt aposize}):
 \begin{itemize}
  \item {\bf {\tt apotype}$=$``C1''}. All pixels are multiplied by a factor $f$ given by:
  \begin{equation}
    f=\left\{\begin{array}{cc}
              x-\sin(2\pi x)/(2\pi) & x<1\\
              1                     & {\rm otherwise}
             \end{array}\right.
  \end{equation}
  where $x\equiv\sqrt{(1-\cos\theta)/(1-\cos\theta_*)}$, and $\theta$ is the angular separation between the pixel and its closest masked pixel (i.e. the closest pixel where the mask is zero.
  \item {\bf {\tt apotype}$=$``C2''}. All pixels are multiplied by a factor $f$ given by:
  \begin{equation}
    f=\left\{\begin{array}{cc}
              \frac{1}{2}\left[1-\cos(\pi\,x)\right] & x<1\\
              1                     & {\rm otherwise}
             \end{array}\right.
  \end{equation}
  where $x\equiv\sqrt{(1-\cos\theta)/(1-\cos\theta_*)}$, and $\theta$ is the angular separation between the pixel and its closest masked pixel (i.e. the closest pixel where the mask is zero.
  \item {\bf {\tt apotype}$=$``Smooth''}. This apodization is carried out in three steps:
    \begin{enumerate}
      \item All pixels within a disc of radius $2.5\theta_*$ of a masked pixel (i.e. where the mask is zero) are masked.
      \item The resulting map is smoothed with a Gaussian window function with standard deviation $\sigma=\theta_*$.
      \item One final pass is made through all pixels to ensure that all pixels that were originally masked remain masked after the smoothing operation.
    \end{enumerate}
 \end{itemize}
\end{funcdef}

\newpage
\section{Sample program}\label{sec:sample}

Here's a simple code using the NaMaster library. This code takes, as command-line options, filenames for two FITS files containing HEALPix maps for a spin-0 field and a mask, and writes the MASTER estimate of its power spectrum into an ASCII file. This file can be found in {\tt test/sample.c}, and can be compiled as described in Section \ref{ssec:compile}.

\lstinputlisting[language=C]{../test/sample.c}

\end{document}

### 2018-07-29 doc/doc_scientific.tex
migration to desc

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx} % Include figure files
\usepackage{fullpage}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amssymb,amsmath} 
\usepackage[colorlinks=true,linkcolor=Blue,linktoc=page,citecolor=Red,urlcolor=BrickRed]{hyperref}
%\usepackage{color}
%\usepackage[utf8x]{inputenc}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{listings}
%\usepackage{makeidx}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{epigraph}
\usepackage{titling}
%\renewcommand{\familydefault}{\sfdefault}
\newcommand{\apj}{ApJ}
\newcommand{\apjs}{ApJS}
\newcommand{\mnras}{MNRAS}

\pretitle{\begin{flushleft}\LARGE\bfseries\sffamily}
\posttitle{\par\end{flushleft}\vskip 0.5em\vspace{-7ex}}
\predate{\begin{flushleft}\sffamily\large}
\postdate{\par\end{flushleft}\vskip 0.5em\vspace{-2ex}}
\titleformat{\section}[hang]{\bfseries\sffamily\Large}
{\thesection}
{12pt}{\filright}

\titleformat{\subsection}[hang]{\bfseries\sffamily}
{\thesubsection}
{12pt}{\filright}

\titleformat{\subsubsection}[hang]{\bfseries}
{\thesubsubsection}
{12pt}{\filright}

\newcommand{\Tr}{{\rm Tr}}
\newcommand{\nv}{\hat{\bf n}}
\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}

\title{NaMaster: Scientific Documentation}

\begin{document}
\maketitle
\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}
\tableofcontents
\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}

\section{Introduction}
NaMaster is a C library, python module and standalone program to compute the pseudo-$C_\ell$ estimator of the angular power spectrum between two masked and contaminated fields (this is also the so-called ``MASTER'' algorithm). The contents of this scientific documentation describe the algorithm, drawing heavily from the methods presented by \cite{2002ApJ...567....2H} and \cite{2017MNRAS.465.1847E}, extending their results to arbitrary cross-correlations between spin-0 and spin-2 fields (see also \cite{2003ApJS..148..161K}).

\section{Generalities and SHTs}
Let ${\bf a}(\nv)$ be a spin-$s_a$ quantity defined on the sphere. Then we define its spherical harmonic coefficients as:
\begin{equation}
 {\bf a}_{\ell m}\equiv{\rm SHT}({\bf a}(\nv))^{s_a}_{\ell m}\equiv\int d\nv\,\hat{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv),\hspace{12pt}
 {\bf a}(\nv)={\rm SHT}^{-1}({\bf a}_{\ell m})^{s_a}_{\nv}\equiv\sum_{\ell m}\,\hat{\sf Y}^{s_a}_{\ell m}(\nv)\,{\bf a}_{\ell m}.
\end{equation}
Note that here we will use a vector notation, such that for a complex spin-$s_a$ field $a$ we form the vector ${\bf a}\equiv({\rm Re}(a),{\rm Im}(a))$. The harmonic coefficients above are decomposed in a similar manner into $E$ and $B$ modes: ${\bf a}_{\ell m}\equiv(a^E_{\ell m},a^B_{\ell m})$. The spherical harmonic operators $\hat{\sf Y}^s$ are therefore matrix that we define in the following subsection.

\subsection{Spin-weighed spherical harmonics}
  Let $\eth$ and $\bar{\eth}$ be the following complex differential operators defined on the sphere when acting on a spin-$s$ quantity $f_s$:
  \begin{equation}
    \eth f_s\equiv-(\sin\theta)^s\left(\partial_\theta+i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^{-s}\,f_s(\theta,\varphi),\hspace{12pt}
    \bar{\eth} f_s\equiv-(\sin\theta)^{-s}\left(\partial_\theta-i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^{s}\,f_s(\theta,\varphi).
  \end{equation}
  The following properties can be easily derived for the action of these operators:
  \begin{itemize}
    \item If $f_s$ is a spin-$s$ quantity, $(f_s)^*$ is a spin-$(-s)$ quantity.
    \item $\eth f_s$ is a spin-$(s+1)$ quantity, and $\bar{\eth} f_s$ is a spin-$(s-1)$ quantity.
    \item $(\eth^nf_s)^*=\bar{\eth}^n(f_s)^*$
    \item $\eth(f\,g)=f\eth g+g\eth f$
    \item $\eth^2(f\,g)=f\eth^2g+g\eth^2f+\eth f\eth g$
  \end{itemize}

  We start by defining the spin-weighed spherical harmonics with spin $s\geq0$:
  \begin{equation}
    _sY_{\ell m}\equiv \alpha_{\ell,s} \eth^s Y_{\ell m},\hspace{6pt}
    _{-s}Y_{\ell m}\equiv \alpha_{\ell,s} (-1)^s\bar{\eth}^s Y_{\ell m},\hspace{6pt}
    \alpha_{l,s}\equiv\sqrt{\frac{(\ell-s)!}{(\ell+s)!}},
  \end{equation}
  which have the property: $(_sY_{\ell m})^*=(-1)^{s+m}\,_{-s}Y_{\ell-m}$. We then define the $E$-mode and $B$-mode spherical harmonic vectors as:
  \begin{align}
    &_s{\bf Y}^E_{\ell m}\equiv {\bf D}^E_sY_{\ell m}
     \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
                              \eth^s+\bar{\eth}^s\\
                              -i(\eth^s-\bar{\eth}^s)
                            \end{array}\right)Y_{\ell m}
     \equiv-\frac{1}{2}\left(\begin{array}{c}
                              _sY_{\ell m}+(-1)^s\,_{-s}Y_{\ell m}\\
                              -i(_sY_{\ell m}-(-1)^s\,_{-s}Y_{\ell m})
                            \end{array}\right) \\
    &_s{\bf Y}^B_{\ell m}\equiv {\bf D}^B_sY_{\ell m}
     \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
                              i(\eth^s-\bar{\eth}^s)\\
                              \eth^s+\bar{\eth}^s
                            \end{array}\right)Y_{\ell m}
     \equiv-\frac{1}{2}\left(\begin{array}{c}
                              i(_sY_{\ell m}-(-1)^s\,_{-s}Y_{\ell m})\\
                              _sY_{\ell m}+(-1)^s\,_{-s}Y_{\ell m}
                            \end{array}\right),
  \end{align}
  which also defines the differential operators ${\bf D}^{E,B}_s$. These functions, for  $s=0$ are simply ${\bf D}^{E}_0=(Y_{\ell m},0)$ and ${\bf D}^{B}_0=(0,Y_{\ell m})$.

  The matrix operator $\hat{\sf Y}^s_{\ell m}$ is then defined as having $_s{\bf Y}^{E,B}_{\ell m}$ as columns:
  \begin{equation}
    \hat{\sf Y}^s_{\ell m}\equiv\left(_s{\bf Y}^E_{\ell m},_s{\bf Y}^B_{\ell m},\right)
    \equiv-\frac{1}{2}
    \left(\begin{array}{cc}
            _sY_{\ell m}+(-1)^s\,_{-s}Y_{\ell m}     & i(_sY_{\ell m}-(-1)^s\,_{-s}Y_{\ell m})\\                
            -i(_sY_{\ell m}-(-1)^s\,_{-s}Y_{\ell m}) & _sY_{\ell m}+(-1)^s\,_{-s}Y_{\ell m}                
          \end{array}\right)
  \end{equation}

  The matrices $\hat{\sf Y}^s_{\ell m}$ satisfy the following relations:
  \begin{align}
    \hat{\sf Y}^{s\dag}_{\ell m}&=(-1)^{m+s}\hat{\sf Y}^{-s}_{\ell -m}\\
    \int d\nv \hat{\sf Y}^s_{\ell m}\hat{\sf Y}^{s\dag}_{\ell'm'}&=\hat{\sf 1}\delta_{\ell\ell'}\delta_{mm'}\\
    \hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2}&\equiv\int d\nv \left(\hat{\sf Y}^{s\dag}_{\bf l}(\nv)\,\hat{\sf Y}^{s}_{{\bf l}_1}(\nv)\right)\,\hat{\sf Y}^0_{{\bf l}_2}(\nv),\\
    &=(-1)^{s+m}\sqrt{\frac{(2\ell+1)(2\ell_1+1)(2\ell_2+1)}{4\pi}}\wtj{\ell}{\ell_1}{\ell_2}{-m}{m_1}{m_2}\wtj{\ell}{\ell_1}{\ell_2}{s}{-s}{0}\,\hat{\sf d}_{\ell+\ell_1+\ell_2},\\
    \hspace{12pt}&\hat{\sf d}^2_n=\frac{1}{2}\left(\begin{array}{cc}
                                       1+(-1)^n & -i[1-(-1)^n]\\
                                       i[1-(-1)^n] & 1+(-1)^n
                                     \end{array}\right),
  \end{align}
  where we have abbreviated the pair $(\ell,m)$ as ${\bf l}$.

  Finally, the following orthogonality relation for the Wigner 3$j$ symbols is useful:
  \begin{equation}
    \sum_{mm_1}\wtj{\ell}{\ell_1}{\ell_2}{m}{m_1}{m_2}\wtj{\ell}{\ell_1}{\ell_3}{m}{m_1}{m_3}=\frac{\delta_{\ell_2\ell_3}\delta_{m_2m_3}}{2\ell_2+1}
  \end{equation}

\subsection{$E$ and $B$ mode purification}
  We define a field ${\bf f}$ to be a $B$ ($E$) mode if $({\bf D}^{E(B)}_s)^\dag{\bf f}=0$. At the same time, and under the definition of the dot product:
  \begin{equation}
    ({\bf f},{\bf g})\equiv\int d\nv\,{\bf f}^\dag{\bf g},
  \end{equation}
  we define a {\sl pure} $B$ ($E$) mode as a field that is orthogonal to all $E$ ($B$) modes.

  Since ${\bf D}^{E\dag}_s{\bf D}^B_s=0$, one can always generate a $B$ ($E$) mode by applying ${\bf D}^{B(E)}_s$ to a scalar field. It is then possible to show that $E$ and $B$ modes thus defined are orthogonal in the full sky:
  \begin{equation}
    ({\bf D}^E_s\varphi,{\bf D}^B_s\psi)=\int d\nv\,({\bf D}^E_s\varphi)^\dag{\bf D}^B_s\psi=0
  \end{equation}
  This can be done by integrating by parts and noting that the celestial sphere has no boundaries. On a cut sky, however, and for $s=2$, this is only true if the fields satisfy Neumann and Dirichlet boundary conditions simultaneously (i.e. vanishing value and first derivative on the boundary of the cut sky region).

  Let $w(\nv)$ be a sky window function defining the sky region to be analyzed (and the weight to be applied in each pixel). The standard pseudo $B$-mode of a field ${\bf P}$ is then given by
  \begin{equation}
    \tilde{B}_{\ell m}=\int d\nv\,w(\nv)\left(_s{\bf Y}^{B}_{\ell m}(\nv)\right)^\dag{\bf P}=\int d\nv\,w(\nv)({\bf D}^B_sY_{\ell m})^\dag{\bf P}(\nv),
  \end{equation}
  Now, since ${\bf D}^B_sY_{\ell m}$ is a $B$-mode, in the absence of $w$ this expresion would correspond to a projection that filters out all the $E$-modes from ${\bf P}$. However, $w(\nv){\bf D}^B_sY_{\ell m}$ is not a $B$-mode, and therefore $\tilde{B}_{\ell m}$ receives contributions from ambiguous $E$ modes (which then propagate into the variance of the pseudo-$C_\ell$ estimator of the power spectrum).

  The idea behind $B$-mode purification is to move $w$ to the right of ${\bf D}^B_s$, defining the field:
  \begin{equation}
   B^p_{\ell m}=\int d\nv\left({\bf D}^B_2(w\,Y_{\ell m})\right)^\dag\,{\bf P}(\nv).
  \end{equation}
  Since ${\bf D}^B_2(wY_{\ell m})$ is a $B$-mode quantity, $B^p_{\ell m}$ should receive contributions only from $B$-modes.

  Expanding ${\bf D}^B_2(wY_{\ell m})$, we can write $B^p_{\ell m}$ as:
  \begin{equation}
    B^p_{\ell m}=
    \left(\tilde{P}_2\right)^B_{\ell m}+
    2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}}\left(\tilde{P}_1\right)^B_{\ell m}+\alpha_{\ell,2}\left(\tilde{P}_2\right)^B_{\ell m},
  \end{equation}
  where $(f)^B_{\ell m}$ stands for the $B$-mode of field $f$, and we have defined the fields $\tilde{P}_n=(\eth^{2-n}w)^*(Q+iU)$, where $Q$ and $U$ are the real and imaginary parts of the field $P$.

  Note that the derivatives of $w$ can be computed as:
  \begin{equation}
    w
    \rightarrow w_{\ell m}={\rm SHT}(w)
    \rightarrow\,\left\{_nw^E_{\ell m}=(-1)^Hw_{\ell m}/\alpha_{\ell,n},\,_nw^B_{\ell m}=0\right\}
    \rightarrow \eth^nw={\rm SHT}^{-1}(\{\,_nw^E_{\ell m},\,_nw^B_{\ell m}\})
  \end{equation}

\section{Contaminant cleaning}\label{sec:contaminant}
  Let ${\bf a}$ be a random field defined on the sphere, let $v(\nv)$ a mask for ${\bf a}$ and let ${\bf f}^i$ be a set of $N_a$ contaminants of ${\bf a}$ such that the observed version of ${\bf a}$ be:
  \begin{equation}
    {\bf d}_a(\nv)={\bf a}^v(\nv)+\sum_{i=1}^{N_a}\alpha_i{\bf f}^i(\nv),
  \end{equation}
  where ${\bf a}^v(\nv)=v(\nv)\,{\bf a}(\nv)$ (note that we have implicitly applied the same mask to ${\bf f}^i$). The best-fit value for the coefficients $\alpha_i$ assuming the same weights for all points in ${\bf d}_a$ can be found as
  \begin{equation}
    \tilde{\alpha}_i=\sum_jM_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf d}_a(\nv),\hspace{12pt}
    (\hat{\sf M}^{-1})_{ij}\equiv\int d\nv\,{\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv).
  \end{equation}
  Thus we can find a cleaned version of ${\bf a}$ as:
  \begin{align}\nonumber
    \tilde{\bf a}(\nv)&\equiv{\bf d}_a(\nv)-{\bf f}^i(\nv)\,M_{ij}\int d\nv'{\bf f}^{j\dag}(\nv'){\bf d}_a(\nv')\\\label{eq:deproject_real}
                      &={\bf a}^v(\nv)-{\bf f}^i(\nv)\,M_{ij}\int d\nv'{\bf f}^{j\dag}(\nv'){\bf a}^v(\nv'),
  \end{align}
  where there is an implicit summation sign over $i$ and $j$ (we will omit these from now on).

  The harmonic coefficients of the cleaned and masked field are:
  \begin{equation}
   \tilde{\bf a}_{\ell m}={\bf a}^v_{\ell m}-{\bf f}^i_{\ell m}M_{ij}\sum_{\ell' m'}{\bf f}^{j\dag}_{\ell' m'}{\bf a}^{v}_{\ell' m'}.
  \end{equation}
  From now on we will simplify the notation by abbreviating the pair $\ell m$ as ${\bf l}$, so that the previous equation reads:
  \begin{equation}\label{eq:deproject_harmonic}
   \tilde{\bf a}_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}M_{ij}\sum_{{\bf l}'}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^{v}_{{\bf l}'}.
  \end{equation}

  The harmonic coefficients for the masked field can be related to those of the unmasked one and the mask $v$ (understood as a spin-0 field) as:
  \begin{equation}
    {\bf a}^v_{\bf l}=\sum_{{\bf l}_1{\bf l}_2}\hat{\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}{\bf a}_{{\bf l}_1}v_{{\bf l}_2}.
  \end{equation}

\section{Pseudo-$C_\ell$ estimators with mode deprojection}  
  In what follows, for two fields ${\bf a}$ and ${\bf b}$ we will define their observed power spectrum as:
  \begin{equation}
    \tilde{\sf C}^{ab}_{\ell}\equiv\frac{1}{2\ell+1}\sum_m{\bf a}_{\ell m}{\bf b}^\dag_{\ell m}.
  \end{equation}
  This must not be confused with the true power spectrum defined as an ensemble average for isotropic fields:
  \begin{equation}
    \langle{\bf a}_{{\bf l}}{\bf b}^\dag_{{\bf l}'}\rangle\equiv\hat{\sf C}^{ab}_\ell\delta_{\ell\ell'}\delta_{mm'}.
  \end{equation}

  Now, let $\tilde{\bf a}$ and $\tilde{\bf b}$ be the contaminant-cleaned versions of two random fields ${\bf a}$ and ${\bf b}$ with contaminants ${\bf f}^i$ and ${\bf g}^j$ and masks $v$ and $w$ respectively, and let us define
  \begin{equation}
    (\hat{\sf N}^{-1})_{ij}\equiv\int d\nv\, {\bf g}^{i\dag}(\nv)\,{\bf g}^j(\nv).
  \end{equation}

  The observed power spectrum of the contaminant-cleaned maps can be written as:
  \begin{align}\nonumber
   \tilde{\sf C}^{\tilde{a}\tilde{b}}_\ell=&
   \frac{1}{2\ell+1}\sum_m{\bf a}^v_{\ell m}{\bf b}^{w\dag}_{\ell m}-\frac{N^*_{ij}}{2\ell+1}\sum_m\sum_{{\bf l}'}{\bf a}^v_{{\bf l}}{\bf b}^{w\dag}_{{\bf l}'}{\bf g}^{j}_{{\bf l}'}{\bf g}^{i\dag}_{\bf l}-\\
   &-\frac{M_{ij}}{2\ell+1}\sum_m\sum_{{\bf l}'}{\bf f}^{i}_{{\bf l}}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}{\bf b}^{w\dag}_{\bf l}+\frac{M_{ij}N^*_{pq}}{2\ell+1}\sum_m\sum_{{\bf l}'{\bf l}''}{\bf f}^{i}_{{\bf l}}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}{\bf b}^{w\dag}_{{\bf l}''}{\bf g}^{q}_{{\bf l}''}{\bf g}^{p\dag}_{\bf l}.
  \end{align}

  In order to compute the bias of $\tilde{\sf C}^{\tilde{a}\tilde{b}}_\ell$ with respect to $\hat{\sf C}^{ab}_\ell$, we need to compute the ensemble average of the former, which we will write as:
  \begin{equation}
    \left\langle\tilde{\sf C}^{\tilde{a}\tilde{b}}_\ell\right\rangle=\hat{\sf F}^1_\ell-\hat{\sf F}^2_\ell-\hat{\sf F}^3_\ell+\hat{\sf F}^4_\ell,
  \end{equation}
  where:
  \begin{align}
    \hat{\sf F}^1_\ell&\equiv\frac{1}{2\ell+1}\sum_m\langle{\bf a}^v_{\ell m}{\bf b}^{w\dag}_{\ell m}\rangle,\hspace{12pt}
    \hat{\sf F}^2_\ell\equiv\frac{N^*_{ij}}{2\ell+1}\sum_m\sum_{{\bf l}'}\langle{\bf a}^v_{{\bf l}}{\bf b}^{w\dag}_{{\bf l}'}{\bf g}^{j}_{{\bf l}'}{\bf g}^{i\dag}_{\bf l}\rangle\\
    \hat{\sf F}^3_\ell&\equiv\frac{M_{ij}}{2\ell+1}\sum_m\sum_{{\bf l}'}\langle{\bf f}^{i}_{{\bf l}}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}{\bf b}^{w\dag}_{\bf l}\rangle,\hspace{12pt}
    \hat{\sf F}^4_\ell\equiv\frac{M_{ij}N^*_{pq}}{2\ell+1}\sum_m\sum_{{\bf l}'{\bf l}''}\langle{\bf f}^{i}_{{\bf l}}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}{\bf b}^{w\dag}_{{\bf l}''}{\bf g}^{q}_{{\bf l}''}{\bf g}^{p\dag}_{\bf l}\rangle.
  \end{align}
  We will now compute the ensemble average of each of these terms.

  \subsection{$\hat{\sf F}^1_\ell$}
    \begin{align}\nonumber
      \hat{\sf F}^1_\ell&=\frac{1}{2\ell+1}\sum_{m{\bf l}_{1,2,3,4}}v_{{\bf l}_2}w^*_{{\bf l}_4}\hat{\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}\langle{\bf a}_{{\bf l}_1}{\bf b}^\dag_{{\bf l}_3}\rangle\hat{\sf D}^{s_b\dag}_{{\bf l}{\bf l}_3{\bf l}_4}\\\nonumber
      &=\frac{1}{2\ell+1}\sum_{m{\bf l}_{1,2,3}}v_{{\bf l}_2}w^*_{{\bf l}_3}\hat{\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}\hat{\sf C}^{ab}_{{\bf l}_1}\hat{\sf D}^{s_b\dag}_{{\bf l}{\bf l}_1{\bf l}_3}\\\nonumber
      &=\frac{1}{2\ell+1}\sum_{\ell_1{\bf l}_{2,3}}v_{{\bf l}_2}w^*_{{\bf l}_3}\frac{(2\ell+1)(2\ell_1+1)}{4\pi}\sqrt{(2\ell_2+1)(2\ell_3+1)}\wtj{\ell}{\ell_1}{\ell_2}{s_a}{-s_a}{0}\wtj{\ell}{\ell_1}{\ell_3}{s_b}{-s_b}{0}\\\nonumber
      &\hspace{50pt}\hat{\sf d}^{s_a}_{\ell+\ell_1+\ell_2}\hat{\sf C}^{ab}_{\ell_1}\hat{\sf d}^{s_b\dag}_{\ell+\ell_1+\ell_3}\sum_{mm_1}\wtj{\ell}{\ell_1}{\ell_2}{-m}{m_1}{m_2}\wtj{\ell}{\ell_1}{\ell_3}{-m}{m_1}{m_3}\\\label{eq:master_transformation}
      &=\sum_{\ell_1\ell_2}\frac{(2\ell_1+1)(2\ell_2+1)}{4\pi}\tilde{C}^{vw}_{\ell_2}\wtj{\ell}{\ell_1}{\ell_2}{s_a}{-s_a}{0}\wtj{\ell}{\ell_1}{\ell_2}{s_b}{-s_b}{0}\hat{\sf d}^{s_a}_{\ell+\ell_1+\ell_2}\hat{\sf C}^{ab}_{\ell_1}\hat{\sf d}^{s_b\dag}_{\ell+\ell_1+\ell_2}
    \end{align}

    For $v=w=1$ this reduces to $\tilde{C}^{vw}_\ell=4\pi\delta_{\ell0}$ and:
    \begin{align}\nonumber
      \hat{\sf F}^1_\ell
      &=\sum_{\ell_1\ell_2}(2\ell_1+1)(2\ell_2+1)\delta_{\ell_20}\wtj{\ell}{\ell_1}{\ell_2}{s_a}{-s_a}{0}\wtj{\ell}{\ell_1}{\ell_2}{s_b}{-s_b}{0}\hat{\sf d}^{s_a}_{\ell+\ell_1+\ell_2}\hat{\sf C}^{ab}_{\ell_1}\hat{\sf d}^{s_b\dag}_{\ell+\ell_1+\ell_2}\\\nonumber
      &=\sum_{\ell_1}(2\ell_1+1)\wtj{\ell}{\ell_1}{0}{s_a}{-s_a}{0}\wtj{\ell}{\ell_1}{0}{s_b}{-s_b}{0}\hat{\sf d}^{s_a}_{\ell+\ell_1}\hat{\sf C}^{ab}_{\ell_1}\hat{\sf d}^{s_b\dag}_{\ell+\ell_1}\\\nonumber
      &=\sum_{\ell_1}(2\ell_1+1)\delta_{\ell\ell_1}\frac{(-1)^{\ell-s_a}}{\sqrt{2\ell+1}}\delta_{\ell\ell_1}\frac{(-1)^{\ell-s_b}}{\sqrt{2\ell+1}}\hat{\sf d}^{s_a}_{\ell+\ell_1}\hat{\sf C}^{ab}_{\ell_1}\hat{\sf d}^{s_b\dag}_{\ell+\ell_1}\\\nonumber
      &=\hat{\sf d}^{s_a}_{2\ell}\hat{\sf C}^{ab}_{\ell}\hat{\sf d}^{s_b\dag}_{2\ell}\\\nonumber
      &=\hat{\sf C}^{ab}_{\ell}
    \end{align}

  \subsection{$\hat{\sf F}^2_\ell$}
    \begin{align}\nonumber
     \hat{\sf F}^2_\ell
     &=N^*_{ij}\int d\nv\,d\nv'\,\frac{\sum_{m{\bf l}'{\bf l}_{1,2,3,4}}}{2\ell+1}\hat{\sf Y}^{s_a\dag}_{\bf l}(\nv)\hat{\sf Y}^{s_a}_{{\bf l}_1}(\nv)\langle{\bf a}_{{\bf l}_1}{\bf b}^\dag_{{\bf l}_3}\rangle\hat{\sf Y}^{s_b\dag}_{{\bf l}_3}(\nv')\hat{\sf Y}^{s_b}_{{\bf l}'}(\nv'){\bf g}^j_{{\bf l}'}{\bf g}^{i\dag}_{\bf l}v_{{\bf l}_2}w^*_{{\bf l}_4}Y_{{\bf l}_2}(\nv)Y^*_{{\bf l}_4}(\nv')\\\nonumber
     &=N^*_{ij}\,\frac{\sum_m}{2\ell+1}\left\{\int d\nv\,v(\nv)\,\hat{\sf Y}^{s_a\dag}_{\bf l}(\nv)\left[\sum_{\ell_1m_1}\hat{\sf Y}^{s_a}_{{\bf l}_1}(\nv)\hat{\sf C}^{ab}_{\ell_1}\left(\int d\nv'\hat{\sf Y}^{s_b\dag}_{{\bf l}_1}(\nv'){\bf g}^j(\nv')w(\nv')\right)\right]{\bf g}^{i\dag}_{\bf l}\right\}\\\label{eq:bias_2}
     &=N^*_{ij}\,\frac{\sum_m}{2\ell+1}{\rm SHT}\left\{v(\nv){\rm SHT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm SHT}\left(w\,{\bf g}^j\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right\}^{s_a}_{\bf l}{\bf g}^{i\dag}_{\bf l}
    \end{align}

    For $v=w=1$ this reduces to:
    \begin{align}\nonumber
     \hat{\sf F}^2_\ell
     &=N^*_{ij}\,\frac{\sum_m}{2\ell+1}{\rm SHT}\left\{{\rm SHT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm SHT}\left({\bf g}^j\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right\}^{s_a}_{\bf l}{\bf g}^{i\dag}_{\bf l}\\\nonumber
     &=N^*_{ij}\,\hat{\sf C}^{ab}_{\ell}\frac{\sum_m{\bf g}^j_{\ell m}{\bf g}^{i\dag}_{\ell m}}{2\ell+1}\\\nonumber
     &=N^*_{ij}\,\hat{\sf C}^{ab}_{\ell}\tilde{\sf C}^{g^jg^i}_\ell
    \end{align}

  \subsection{$\hat{\sf F}^3_\ell$}
    \begin{align}\nonumber
     \hat{\sf F}^3_\ell
     &=M_{ij}\int d\nv\,d\nv'\,\frac{\sum_{m{\bf l}'{\bf l}_{1,2,3,4}}}{2\ell+1}{\bf f}^i_{\bf l}{\bf f}^{j\dag}_{{\bf l}'}\hat{\sf Y}^{s_a\dag}_{\bf l'}(\nv')\hat{\sf Y}^{s_a}_{{\bf l}_3}(\nv')\langle{\bf a}_{{\bf l}_3}{\bf b}^\dag_{{\bf l}_1}\rangle\hat{\sf Y}^{s_b\dag}_{{\bf l}_1}(\nv)\hat{\sf Y}^{s_b}_{{\bf l}}(\nv)v_{{\bf l}_4}w^*_{{\bf l}_2}Y_{{\bf l}_2}(\nv)Y^*_{{\bf l}_4}(\nv')\\\nonumber
     &=M_{ij}\frac{\sum_{m}}{2\ell+1}{\bf f}^i_{\bf l}\left\{\int d\nv\,w(\nv)\,\left[\sum_{\ell_1m_1}\left(\int d\nv'\,v(\nv'){\bf f}^{j\dag}(\nv')\hat{\sf Y}^{s_a}_{{\bf l}_1}(\nv')\right)\hat{\sf C}^{ab}_{\ell_1}\hat{\sf Y}^{s_b\dag}_{{\bf l}_1}(\nv)\right]\hat{\sf Y}^{s_b}_{{\bf l}}(\nv)\right\}\\\label{eq:bias_3}
     &=M_{ij}\frac{\sum_{m}}{2\ell+1}{\bf f}^i_{\bf l}\,{\rm SHT}\left\{w(\nv)\,{\rm SHT}^{-1}\left[\hat{\sf C}^{ab\dag}_{\ell_1}{\rm SHT}\left(v\,{\bf f}^j\right)^{s_a}_{{\bf l}_1}\right]^{s_b}_{\nv}\right\}^{s_b\dag}_{\bf l}
    \end{align}

    For $v=w=1$ this reduces to:
    \begin{align}\nonumber
     \hat{\sf F}^3_\ell
     &=M_{ij}\frac{\sum_{m}}{2\ell+1}{\bf f}^i_{\bf l}\,{\rm SHT}\left\{{\rm SHT}^{-1}\left[\hat{\sf C}^{ab\dag}_{\ell_1}{\rm SHT}\left({\bf f}^j\right)^{s_a}_{{\bf l}_1}\right]^{s_b}_{\nv}\right\}^{s_b\dag}_{\bf l}\\\nonumber
     &=M_{ij}\frac{\sum_{m}{\bf f}^i_{\ell m}{\bf f}^{j\dag}_{\ell m}}{2\ell+1}\hat{\sf C}^{ab}_{\ell}\\\nonumber
     &=M_{ij}\tilde{\sf C}^{f^if^j}_\ell\hat{\sf C}^{ab}_{\ell}
    \end{align}

  \subsection{$\hat{\sf F}^4_\ell$}
    \begin{align}\nonumber
     \hat{\sf F}^4_\ell
     &=\frac{M_{ij}N^*_{pq}}{2\ell+1}\int d\nv\,d\nv'\,\sum_{m{\bf l}'{\bf l}''{\bf l}_{1,2,3,4}}
     {\bf f}^i_{\bf l}{\bf f}^{j\dag}_{{\bf l}'}\hat{\sf Y}^{s_a\dag}_{\bf l'}(\nv)\hat{\sf Y}^{s_a}_{{\bf l}_1}(\nv)\langle{\bf a}_{{\bf l}_1}{\bf b}^\dag_{{\bf l}_3}\rangle\hat{\sf Y}^{s_b\dag}_{{\bf l}_3}(\nv')\hat{\sf Y}^{s_b}_{{\bf l}''}(\nv'){\bf g}^q_{{\bf l}''}{\bf g}^{p\dag}_{\bf l}v_{{\bf l}_2}w^*_{{\bf l}_4}Y_{{\bf l}_2}(\nv)Y^*_{{\bf l}_4}(\nv')\\\nonumber
     &=\frac{M_{ij}N^*_{pq}}{2\ell+1}\sum_{m}
     {\bf f}^i_{\bf l}\left\{\int d\nv\,v(\nv)\,{\bf f}^{j\dag}(\nv)\left[\sum_{\ell_1m_1}\hat{\sf Y}^{s_a}_{{\bf l}_1}(\nv)\hat{\sf C}^{ab}_{\ell_1}\left(\int d\nv'\,\hat{\sf Y}^{s_b\dag}_{{\bf l}_1}(\nv'){\bf g}^q(\nv')w(\nv')\right)\right]\right\}{\bf g}^{p\dag}_{\bf l}\\\label{eq:bias_4}
     &=M_{ij}N^*_{pq}
     \left\{\int d\nv\,v(\nv)\,{\bf f}^{j\dag}(\nv)\,{\rm SHT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm SHT}\left(w\,{\bf g}^q\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right\}\tilde{\sf C}^{f^ig^p}_\ell
    \end{align}

    For $v=w=1$ this reduces to:
    \begin{align}\nonumber
     \hat{\sf F}^4_\ell
     &=M_{ij}N^*_{pq}\left\{\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\rm SHT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm SHT}\left({\bf g}^q\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right\}\tilde{\sf C}^{f^ig^p}_\ell\\\nonumber
     &=M_{ij}N^*_{pq}\left\{\int d\nv\,\sum_{{\bf l}_2}{\bf f}^{j\dag}_{{\bf l}_2}\hat{\sf Y}^{s_a\dag}_{{\bf l}_2}(\nv)\,\sum_{{\bf l}_1}\hat{\sf Y}^{s_a}_{{\bf l}_1}(\nv)\hat{\sf C}^{ab}_{\ell_1}{\bf g}^q_{{\bf l}_1}\right\}\tilde{\sf C}^{f^ig^p}_\ell\\\nonumber
     &=M_{ij}N^*_{pq}{\bf f}^{j\dag}_{{\bf l}_1}\hat{\sf C}^{ab}_{\ell_1}{\bf g}^q_{{\bf l}_1}\tilde{\sf C}^{f^ig^p}_\ell\\\nonumber
     &=M_{ij}N^*_{pq}\left[\sum_{\ell_1}(2\ell_1+1){\rm Tr}\left(\hat{\sf C}^{ab}_{\ell_1}\tilde{\sf C}^{g^qf^j}_{\ell_1}\right)\right]\tilde{\sf C}^{f^ig^p}_\ell
    \end{align}

  \subsection{Final form of the estimator}
    Putting together the results from Equations \ref{eq:master_transformation}, \ref{eq:bias_2}, \ref{eq:bias_3} and \ref{eq:bias_4}, we can write down an unbiased estimator for the pseudo-$C_\ell$ of the cut-sky maps free from contamination from ${\bf f}$ and ${\bf g}$:
    \begin{align}\nonumber
      \tilde{\sf C}^{ab}_\ell=&\tilde{\sf C}^{\tilde{a}\tilde{b}}_\ell+N^*_{ij}\,\frac{\sum_m}{2\ell+1}{\rm SHT}\left\{v(\nv){\rm SHT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm SHT}\left(w\,{\bf g}^j\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right\}^{s_a}_{\bf l}{\bf g}^{i\dag}_{\bf l}+\\\nonumber
      &+M_{ij}\frac{\sum_{m}}{2\ell+1}{\bf f}^i_{\bf l}\,{\rm SHT}\left\{w(\nv)\,{\rm SHT}^{-1}\left[\hat{\sf C}^{ab\dag}_{\ell_1}{\rm SHT}\left(v\,{\bf f}^j\right)^{s_a}_{{\bf l}_1}\right]^{s_b}_{\nv}\right\}^{s_b\dag}_{\bf l}-\\\label{eq:unbiased_pcl}
      &-M_{ij}N^*_{pq}\left\{\int d\nv\,v(\nv)\,{\bf f}^{j\dag}(\nv)\,{\rm SHT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm SHT}\left(w\,{\bf g}^q\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right\}\tilde{\sf C}^{f^ig^p}_\ell
    \end{align}

    Once $\tilde{\sf C}^{ab}$ is calculated, it can be corrected for the effects of masking by inverting the linear transformation in Eq \ref{eq:master_transformation}. This transformation can be written explicitly by first transforming the power spectrum matrices into vectors $_v{\sf C}$. E.g. for $s_a=s_b=2$ we transform:
    \begin{equation}
      \hat{\sf C}^{ab}_\ell\equiv\left(
      \begin{array}{cc}
        C^{E_aE_b}_\ell & C^{E_aB_b}_\ell \\
        C^{B_aE_b}_\ell & C^{B_aB_b}_\ell
      \end{array}\right)
      \hspace{12pt}\text{into}\hspace{12pt}
      _v\hat{\sf C}^{ab}_\ell\equiv\left(
      \begin{array}{c}
        C^{E_aE_b}_\ell\\
        C^{E_aB_b}_\ell\\
        C^{B_aE_b}_\ell\\
        C^{B_aB_b}_\ell
      \end{array}\right).
    \end{equation}
    We can then write, in general:
    \begin{equation}\label{eq:cell_coupled}
      _v\tilde{\sf C}^{ab}_\ell=\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\cdot\,_v\hat{\sf C}^{ab}_{\ell'},
    \end{equation}
    where:
    \begin{align}
      &{\sf M}^{00}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}(2\ell''+1)C^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\\
      &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\hat{\sf 1},
      \hspace{12pt}
      M^{0+}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}(2\ell''+1)C^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}\\
      &{\sf M}^{22}_{\ell\ell'}=\left(
      \begin{array}{cccc}
        M^{++}_{\ell\ell'} &0 &0&M^{--}_{\ell\ell'}\\
        0&M^{++}_{\ell\ell'} &-M^{--}_{\ell\ell'}&0\\
        0&-M^{--}_{\ell\ell'} &M^{++}_{\ell\ell'}&0\\
        M^{--}_{\ell\ell'} &0 &0&M^{++}_{\ell\ell'}
      \end{array}\right)\\
      & \hspace{12pt}M^{\pm\pm}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}(2\ell''+1)C^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}^2\frac{1\pm(-1)^{\ell+\ell'+\ell''}}{2}
    \end{align}

    If either the $B$ or $E$ modes of a spin-2 field has been purified, the equations above must be modified by carrying out the following modification in the equations above:
    \begin{align}
      &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}\longrightarrow\\\nonumber
      &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}+2\sqrt{\frac{(\ell+1)!(\ell-2)!(\ell''+1)!}{(\ell-1)!(\ell+2)!(\ell''-1)!}}\wtj{\ell}{\ell'}{\ell''}{1}{-2}{1}+\sqrt{\frac{(\ell-2)!(\ell''+2)!}{(\ell+2)!(\ell''-2)!}}\wtj{\ell}{\ell'}{\ell''}{0}{-2}{2}.
    \end{align}
    This change must be applied to the corresponding factors of $\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}$.

\subsection{Beam}
  Adding the effect of a beam amounts to redefining:
  \begin{equation}
    \mathsf{M}^{s_as_b}_{\ell_1\ell_2}\rightarrow\mathsf{M}^{s_as_b}_{\ell_1\ell_2}\,b^{ab}_{\ell_2},
  \end{equation}
  where $b^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$.

\subsection{Computing $\langle\tilde{C}^N_l\rangle$}
Consider the case where the window function is just a flat top-hat multiplied by the inverse variance of the noise:
\begin{equation}
  W(\nv)\equiv\Theta(\nv)\frac{\bar{\sigma}_N^2}{\sigma_N^2(\nv)},
\end{equation}
where $\sigma_N^2(\nv)$ is the variance per sterad at each point and $\bar{\sigma}^2_N$ is its sky-averaged value.

For uncorrelated noise, we can write $N(\nv)$ as
\begin{equation}
 N(\nv)=u(\nv)\,\sigma_N(\nv)
\end{equation}
where $u(\nv)$ is a white GRF with power spectrum $C^u_l=1$. In this case, the noise pseudo-$C_l$ can be estimated theoretically:
\begin{equation}
 \langle\tilde{C}_l^N\rangle=\frac{1}{4\pi}\int d\Omega\,\sigma_N^2(\nv)W^2(\nv)=
 \frac{\bar{\sigma}_N^2}{4\pi}\int d\Omega\, \Theta(\nv)\frac{\bar{\sigma}_N^2}{\sigma_N^2(\nv)}=
 f_{\rm sky}\,\bar{\sigma}_N^2\overline{\left(\frac{\bar{\sigma}_N^2}{\sigma^2_N}\right)}
\end{equation}

\subsection{Binning into bandpowers}
  Given the loss of information implicit in masking the originally full-sky field, it is in general not possible to invert Eq. \ref{eq:cell_coupled} directly. De usual approach to doing so is by binning the pseudo-$C_\ell$ into bandpowers. A bandpower $b$ is defined by a set of $N_b$ multipoles $\vec{\ell}_b\equiv(\ell_b^1,...,\ell_b^{N_k})$ and a set of weights $\vec{w}_b\equiv(w_b^1,...,w_b^{N_b})$ normalized such that $\sum_{i=1}^{N_b}w_b^i=1$. The $b-$th bandpower for the coupled pseudo-$C_\ell$ is then defined as:
  \begin{equation}
    _v\tilde{\sf B}_b \equiv\sum_{i=1}^{N_b}w_b^i\,_v\tilde{\sf C}_{\ell_b^i}
    =\sum_{i=1}^{N_b}w_b^i\sum_{\ell'}{\sf M}^{s_as_b}_{\ell_b^i\ell'}\,_v\hat{\sf C}_\ell.
  \end{equation}
  One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: $_v\hat{\sf C}_\ell=\sum_b\,_v\hat{\sf B}_b\Theta(\ell\in\vec{\ell}_b)$ (where $\Theta$ is a binary step function). The previous equation then reads:
  \begin{equation}
   _v\tilde{\sf B}_b=\sum_{b'}\mathcal{M}_{bb'}\,_v\hat{\sf B}_{bb'}\equiv\sum_{b'}\left(\sum_{\ell\in\vec{\ell}_b}\sum_{\ell'\in\vec{\ell}_{b'}}w_b^\ell{\sf M}_{\ell\ell'}\right)\,_v\hat{\sf B}_{b'},
  \end{equation}
  which defines the binned coupling matrix $\mathcal{M}_{bb'}$. The decoupled bandpowers are then estimated by inverting $\mathcal{M}$:
  \begin{equation}
    _v\hat{\sf B}_b=\sum_{b'}\left(\mathcal{M}\right)^{-1}_{bb'}\,_v\tilde{\sf B}_{b'}.
  \end{equation}

  Note that, since this procedure is based on the assumption that the true power spectrum is step-wise constant, the bandpowers computed this way should be compared with the theoretical prediction subjected to the same type of transformation. I.e. the theoretical prediction for the bandpowers is:
  \begin{equation}
    _v\bar{\sf B}_b=\sum_{b'}\left(\mathcal{M}\right)^{-1}_{bb'}\sum_{\ell'\in\vec{\ell}_{b'}}w_{b'}^{\ell'}\sum_{\ell''}{\sf M}_{\ell'\ell''}\,_v\bar{\sf C}_{\ell''},
  \end{equation}
  where we the overline $\bar{\,}$ denotes theoretical predictions.

\section{Gaussian covariance matrices}
  For three scalar fields $a$, $b$, $c$ and $d$, with masks $w^a$, $w^b$, $w^c$ and $w^d$, the covariance matrix element between the coupled pseudo-$C_\ell$ for $(a,b)$ and $(c,d)$ is given by:
  \begin{equation}
    \left\langle\Delta\tilde{C}^{ab}_\ell\Delta\tilde{C}^{cd}_{\ell'}\right\rangle=\sum_{mm'}\sum_{{\bf l}_1{\bf l}_2}\left(C^{ac}_{\ell_1}C^{bd}_{\ell_2}W^a_{{\bf l}{\bf l}_1}W^b_{{\bf l}{\bf l}_2}W^c_{{\bf l}{\bf l}_1}W^d_{{\bf l}{\bf l}_2}+C^{ad}_{\ell_1}C^{bc}_{\ell_2}W^a_{{\bf l}{\bf l}_1}W^b_{{\bf l}{\bf l}_2}W^c_{{\bf l}{\bf l}_2}W^d_{{\bf l}{\bf l}_1}\right)
  \end{equation}
  Under the approximation $C^{ac}_{\ell_1}C^{bd}_{\ell_2}\rightarrow C^{ac}_{(\ell}C^{bd}_{\ell')}$\footnote{The approximation made by e.g. \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B} is $C^{ac}_{\ell_1}C^{bd}_{\ell_2}\rightarrow \sqrt{C^{ac}_{\ell}C^{ac}_{\ell'}C^{bd}_{\ell}C^{bd}_{\ell'}}$, however this is not possible for possibly-negative cross-correlations, which motivates the arithmetic rather than geometrical mean used here.}, where $X_{(a}Y_{b)}\equiv(X_aY_b+X_bY_a)/2$, this gets simplified to:
  \begin{equation}
    \left\langle\Delta\tilde{C}^{ab}_\ell\Delta\tilde{C}^{cd}_{\ell'}\right\rangle\simeq
    C^{ac}_{(\ell}C^{bd}_{\ell')}\,\Xi_{\ell\ell'}(w^aw^c,w^bw^d)+C^{ad}_{(\ell}C^{bc}_{\ell')}\,\Xi_{\ell\ell'}(w^aw^d,w^bw^c).
  \end{equation}
  Here:
  \begin{equation}
    \Xi_{\ell\ell'}(\phi,\psi)\equiv\sum_{\ell_1}\frac{2\ell_1+1}{4\pi}\wtj{\ell}{\ell'}{\ell_1}{0}{0}{0}^2C^{\phi\psi}_{\ell_1}.
  \end{equation}

  Finally, we can compute the covariance of the coupled bandpowers by averaging within each bandpower:
  \begin{equation}
    {\rm Cov}^{(ab),(cd)}_{kk'}\equiv\left\langle\Delta\,_v\tilde{\sf B}^{ab}_k\Delta\,_v\tilde{\sf B}^{cd}_{k'}\right\rangle=\sum_{\ell\in k}\sum_{\ell'\in k'}w_k^\ell w_{k'}^{\ell'}\left\langle\Delta\tilde{C}^{ab}_\ell\Delta\tilde{C}^{cd}_{\ell'}\right\rangle,
  \end{equation}
  and the covariance of the uncoupled bandpowers is given by the linear transformation:
  \begin{equation}
    \left\langle\left(\Delta\,_v\hat{\sf B}^{ab}\right)\left(\Delta\,_v\hat{\sf B}^{cd}\right)^T\right\rangle=({\cal M}_{ab})^{-1}{\rm Cov}^{(ab),(cd)}\left(({\cal M}_{cd})^{-1}\right)^T
  \end{equation}

\section{Flat-sky}
  \subsection{Fourier transforms}
    In the flat sky we will write the directional vector $\nv$ as ${\bf x}$. Let ${\bf a}({\bf x})$ be a spin-$s_a$ quantity. Under the approximation $\sin\theta\sim1$, $\Delta\theta\rightarrow -\Delta x$, $\Delta\varphi\rightarrow\Delta y$, the differential operator $\eth$ now takes the form:
    \begin{equation}
      \eth= (\partial_x-i\partial_y),\hspace{12pt}\bar{\eth}= (\partial_x+i\partial_y),
    \end{equation}
    and acts on a plane wave $e^{i{\bf k}{\bf x}}$ as:
    \begin{equation}
      \eth^s e^{i{\bf k}{\bf x}}=(ik)^se^{-i\,s\varphi_k}e^{i{\bf k}{\bf x}},\hspace{12pt}
      \bar{\eth}e^{i{\bf k}{\bf x}}=(ik)^se^{i\,s\varphi_k}e^{i{\bf k}{\bf x}}
    \end{equation}

    Let us define the basis functions:
    \begin{equation}
      \,_s{\cal Y}_{\bf k}({\bf x})\equiv k^{-s}\eth^se^{i{\bf k}{\bf x}}=i^s\,e^{-is\varphi_k}e^{i{\bf k}{\bf x}},\hspace{12pt}\,
      _{-s}{\cal Y}_{\bf k}({\bf x})\equiv (-k)^{-s}\bar{\eth}^se^{i{\bf k}{\bf x}}=(-i)^se^{is\varphi_k}e^{i{\bf k}{\bf x}},
    \end{equation}
    and the spin-$s$ Fourier coefficients:
    \begin{align}
      a({\bf x})  =\int\frac{d{\bf l}^2}{2\pi}\,_s{\cal Y}_{\bf l}({\bf x})_sa_{\bf l},\hspace{12pt}
      a^*({\bf x})=\int\frac{d{\bf l}^2}{2\pi}\,_{-s}{\cal Y}_{\bf l}({\bf x})_{-s}a_{\bf l},\\
      \,_sa_{\bf l}\equiv\int\frac{d{\bf x}^2}{2\pi}\,_s{\cal Y}^*_{\bf l}({\bf x})a({\bf x}),\hspace{12pt}
      \,_{-s}a_{\bf l}\equiv\int\frac{d{\bf x}^2}{2\pi}\,_{-s}{\cal Y}^*_{\bf l}({\bf x})a^*({\bf x}).
    \end{align}
    The $E$ and $B$-mode coefficients are then defined as:
    \begin{align}
      _sE_{\bf l}\equiv-\frac{1}{2}\left[_sa_{\bf l}+(-1)^s\,_{-s}a_{\bf l}\right],\hspace{12pt}
      i\,_sB_{\bf l}\equiv-\frac{1}{2}\left[_sa_{\bf l}-(-1)^s\,_{-s}a_{\bf l}\right].
    \end{align}
    Note the preceding $(-)$ sign. For scalar fields ($s\equiv0$) the $E$ and $B$ modes are defined omitting that sign.

    Let us now write $a$ as a vector such that in real space ${\bf a}({\bf x})\equiv({\rm Re}(a),{\rm Im}(a))$, and in Fourier space ${\bf a}_{\bf l}\equiv(\,_sE_{\bf l},\,_sB_{\bf l})$. We can rewrite the equations above in vectorial form:
    \begin{equation}
      {\bf a}({\bf x})\equiv\int\frac{d{\bf l}^2}{2\pi}\,_s{\sf E}_{\bf l}({\bf x}){\bf a}_{\bf l},\hspace{12pt}{\bf a}_{\bf l}\equiv\int\frac{d{\bf x}^2}{2\pi}\,_s{\sf E}^\dag_{\bf l}({\bf x}){\bf a}({\bf x}),
    \end{equation}
    where we have defined the matrix basis functions:
    \begin{align}
      \,_s{\sf E}_{\bf l}({\bf x})
      &\equiv-\frac{1}{2}
      \left(\begin{array}{cc}
                 \,_s{\cal Y}_{\bf l}+(-1)^s\,_{-s}{\cal Y}_{\bf l}  & i(\,_s{\cal Y}_{\bf l}-(-1)^s\,_{-s}{\cal Y}_{\bf l}) \\
              -i(\,_s{\cal Y}_{\bf l}-(-1)^s\,_{-s}{\cal Y}_{\bf l}) &   \,_s{\cal Y}_{\bf l}+(-1)^s\,_{-s}{\cal Y}_{\bf l}
            \end{array}\right)\\
      &=-\frac{1}{2l^s}
      \left(\begin{array}{cc}
                 \eth^s+\bar{\eth}^s & i(\eth^s-\bar{\eth}^s) \\
              -i(\eth^s-\bar{\eth}^s) &  \eth^s+\bar{\eth}^s
            \end{array}\right)e^{i{\bf l}{\bf x}}\\
      &=-i^s
      \left(\begin{array}{cc}
               \cos(s\varphi_l) & \sin(s\varphi_l) \\
              -\sin(s\varphi_l) & \cos(s\varphi_l)
            \end{array}\right)e^{i{\bf l}{\bf x}}\\
      &=-i^s{\sf R}^\dag(s\varphi_l)e^{i{\bf l}{\bf x}},
    \end{align}
    where ${\sf R}(\varphi)$ is a rotation matrix.

    Thus:
    \begin{align}
      \left(\begin{array}{c}
       Q({\bf x}) \\ U({\bf x})
      \end{array}\right)
      &=-i^s
      \int\frac{d{\bf l}^2}{2\pi}
      \left(\begin{array}{cc}
              \cos(s\varphi_l) & \sin(s\varphi_l) \\
             -\sin(s\varphi_l) & \cos(s\varphi_l) \\
            \end{array}\right)
      \left(\begin{array}{c}
       _sE_{\bf l} \\ _sB_{\bf l}
      \end{array}\right)e^{i{\bf l}{\bf x}}\\
      \left(\begin{array}{c}
       _sE_{\bf l} \\ _sB_{\bf l}
      \end{array}\right)
      &=-(-i)^s
      \left(\begin{array}{cc}
              \cos(s\varphi_l) & -\sin(s\varphi_l) \\
              \sin(s\varphi_l) &  \cos(s\varphi_l) \\
            \end{array}\right)
      \left(\begin{array}{c}
       Q_{\bf l} \\ U_{\bf l}
      \end{array}\right),
    \end{align}
    where $Q_{\bf l}$ and $U_{\bf l}$ are the standard Fourier transforms of $Q$ and $U$.

    The functions $\,_s{\sf E}_{\bf l}({\bf x})$ satisfy the following orthogonality and completeness relations:
    \begin{equation}
      \int \frac{d{\bf x}^2}{(2\pi)^2}\,_s{\sf E}_{\bf l}({\bf x})_s{\sf E}^\dag_{{\bf l}'}({\bf x})={\sf 1}\,\delta({\bf l}-{\bf l}'),\hspace{12pt}
      \int \frac{d{\bf l}^2}{(2\pi)^2}\,_s{\sf E}_{\bf l}({\bf x})_s{\sf E}^\dag_{{\bf l}}({\bf x}')={\sf 1}\,\delta({\bf x}-{\bf x}').
    \end{equation}

  \subsection{Pseudo-$C_\ell$ estimator - continuum limit}
    The Fourier coefficients of the masked field are:
    \begin{align}
      {\bf a}^v_{\bf l}&=\int\int\frac{d{\bf k}^2d{\bf q}^2}{2\pi}\left[\int\frac{d{\bf x}^2}{(2\pi)^2}\,_{s_a}{\sf E}^\dag_{\bf l}({\bf x})\,_{s_a}{\sf E}_{\bf k}({\bf x})\,_0E_{\bf q}({\bf x})\right]{\bf a}_{\bf k}\,v_{\bf q}\\
                       &=\int\frac{d{\bf k}^2}{2\pi}{\sf R}\left(s_a(\varphi_\ell-\varphi_k)\right)\,{\bf a}_{\bf k}v_{{\bf l}-{\bf k}},
    \end{align}
    which, for instance, in the case of a spin-$2$ field, read:
    \begin{equation}\label{eq:pseudo_flat_eb}
     \left(
     \begin{array}{c}
      E^v_{\bf l} \\ B^v_{\bf l}
     \end{array}\right)
     =\int\frac{d{\bf k}^2}{2\pi}v_{{\bf l}-{\bf k}}
     \left(
     \begin{array}{cc}
       \cos2\Delta\varphi & -\sin2\Delta\varphi \\
       \sin2\Delta\varphi &  \cos2\Delta\varphi
     \end{array}\right)\,\left(
     \begin{array}{c}
      E_{\bf k} \\ B_{\bf k}
     \end{array}\right),
    \end{equation}
    where $\Delta\varphi\equiv\varphi_\ell-\varphi_k$. Then, the covariance of the Fourier coefficients of two masked fields is:
    \begin{align}
      \left\langle{\bf a}^v_{\bf l}\,{\bf b}^{w\dag}_{\bf l}\right\rangle
      &=\int\int\frac{d{\bf k}^2d{\bf q}^2}{(2\pi)^2}{\sf R}(s_a(\varphi_\ell-\varphi_k))\left\langle{\bf a}_{\bf k}{\bf b}^\dag_{\bf q}\right\rangle{\sf R}^\dag(s_b(\varphi_\ell-\varphi_q))v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf q}}\\
      &=\int\frac{d{\bf k}^2}{(2\pi)^2}{\sf R}(s_a\Delta\varphi){\sf C}^{ab}_k{\sf R}^\dag(s_b\Delta\varphi)v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}.
    \end{align}

    The pseudo-$C_\ell$ is defined as the nomalized angular average of ${\bf a}_{\bf l}{\bf b}^\dag_{\bf l}$:
    \begin{align}
      \tilde{\sf C}^{ab}_\ell
      &\equiv\frac{(2\pi)^2}{S}\int\frac{d\varphi_\ell}{2\pi}\left\langle{\bf a}^v_{\bf l}\,{\bf b}^{w\dag}_{\bf l}\right\rangle\\
      &=\int\frac{kdk\,qdq}{(2\pi)^2}\left[\frac{(2\pi)^2}{S}\int\frac{d\varphi_q}{2\pi}v_{{\bf q}}w^*_{{\bf q}}
      \int d\varphi_\ell d\varphi_k{\sf R}(s_a\Delta\varphi){\sf C}^{ab}_k{\sf R}^\dag(s_b\Delta\varphi)\delta({\bf q}-{\bf l}+{\bf k})\right],
    \end{align}
    where $S$ is the observed sky area (in sterad), and where we have eliminated the dependence on ${\bf l}-{\bf k}$ by introducing an additional integral over $d{\bf q}^2\delta({\bf q}-{\bf l}+{\bf k})$.

    As shown in \cite{2016arXiv161204664A}, these expressions can be simplified through the following steps:
    \begin{enumerate}
     \item Substitute
       \begin{equation}
         \delta({\bf q}-{\bf l}+{\bf k})\rightarrow\int\frac{d{\bf r}^2}{(2\pi)^2}e^{i({\bf q}-{\bf l}+{\bf k}){\bf r}}
       \end{equation}
     \item Integrate over the angular parts of ${\bf r}$, ${\bf l}$ and ${\bf k}$ using the following relation:
       \begin{equation}
        \int_0^{2\pi} d\varphi\,e^{i\,x\,\cos\varphi}e^{i\,n\varphi}=2\pi\,i^n\,J_n(x),
       \end{equation}
       where $J_n$ is the cylindrical Bessel function of order $n$.
     \item Integrate over the angular part of ${\bf q}$, defining the pseudo-power spectrum of the masks:
       \begin{equation}
         \tilde{C}^{vw}_q\equiv\frac{(2\pi)^2}{S}\int d\varphi_qv_{\bf q}w^*_{\bf q}
       \end{equation}
     \item Solve the last isolated integral over the radial part of ${\bf r}$ by using the following relation:
       \begin{equation}
         \int_0^\infty dr\,r\,J_0(qr)\,J_n(kr)\,J_n(\ell r)=\frac{\cos n\theta}{\pi k\ell\sin\theta},
       \end{equation}
       where $\theta$ is the angle between sides $\ell$ and $k$ of the triangle formed by three sides of length $q$, $\ell$ and $k$.
     \item Make the change of variables $q^2(\ell,k,\theta)\equiv\ell^2+k^2-2k\ell\cos\theta$ to simplify the integral over the radial part of ${\bf q}$.
    \end{enumerate}

    This yields the following relation analogous to Eq. \ref{eq:cell_coupled}:
    \begin{equation}
      \,_v\tilde{\sf C}^{ab}_\ell=\int_0^{\infty} dk\, {\sf M}^{s_as_b}_{\ell k}\cdot\,_v\hat{\sf C}^{ab}_k,
    \end{equation}
    where:
    \begin{align}
      &{\sf M}^{00}_{\ell k}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\\
      &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\hat{\sf 1},
      \hspace{12pt}
      M^{0+}_{\ell k}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\cos2\theta\\
      &{\sf M}^{22}_{\ell\ell'}=\left(
      \begin{array}{cccc}
        M^{++}_{\ell\ell'} &0 &0&M^{--}_{\ell\ell'}\\
        0&M^{++}_{\ell\ell'} &-M^{--}_{\ell\ell'}&0\\
        0&-M^{--}_{\ell\ell'} &M^{++}_{\ell\ell'}&0\\
        M^{--}_{\ell\ell'} &0 &0&M^{++}_{\ell\ell'}
      \end{array}\right),
      \hspace{12pt}M^{\pm\pm}_{\ell\ell'}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\frac{1\pm\cos4\theta}{2}.
    \end{align}
    These can also be expressed as integrals over $q$:
    \begin{align}
      &M^{00}_{\ell k}=\frac{k}{2\pi}\int\frac{q\,dq}{2\pi}\tilde{C}^{vw}_qF(\ell,k,q)\\
      &M^{0+}_{\ell k}=\frac{k}{2\pi}\int\frac{q\,dq}{2\pi}\tilde{C}^{vw}_qF(\ell,k,q)\frac{\ell^4+k^4+q^4-2k^2q^2-2\ell^2q^2}{2k^2\ell^2}\\
      &M^{++}_{\ell k}=\frac{k}{2\pi}\int\frac{q\,dq}{2\pi}\tilde{C}^{vw}_qF(\ell,k,q)\left[\frac{\ell^4+k^4+q^4-2k^2q^2-2\ell^2q^2}{2k^2\ell^2}\right]^2\\
      &M^{--}_{\ell k}=\frac{k}{2\pi}\int\frac{q\,dq}{2\pi}\tilde{C}^{vw}_qF(\ell,k,q)\left(1-\left[\frac{\ell^4+k^4+q^4-2k^2q^2-2\ell^2q^2}{2k^2\ell^2}\right]^2\right),
    \end{align}
    where
    \begin{equation}
     F(\ell,k,q)\equiv\left\{
          \begin{array}{cl}
            4\,(2\ell^2k^2+2k^2q^2+2q^2\ell^2-\ell^4-k^4-q^4)^{-1/2} & {\rm if} |k-q|<\ell<k+q\\
            0 & {\rm otherwise}
          \end{array}\right.
    \end{equation}

    Since in the flat-sky limit $\ell$ is a continuous variable, we define bandpowers $_v{\sf B}_b$ as averages over a given interval in $\ell$, $[\ell_b^i,\ell_b^0]$:
    \begin{equation}
      \,_v\tilde{\sf B}_b\equiv N_b\int_{\ell_b^i}^{\ell_b^f} \frac{d\ell}{\ell_b^f-\ell_b^i}\,_v\tilde{\sf C}_\ell.
    \end{equation}
    The binned coupling matrix is therefore given by:
    \begin{equation}
      \mathcal{M}_{bb'}\equiv\int^{\ell_b^f}_{\ell_b^i} \frac{d\ell}{\ell_b^f-\ell_b^i}\int^{\ell_b'^f}_{\ell_b'^i} d\ell'\,{\sf M}_{\ell\ell'}
    \end{equation}

  \subsection{Pseudo-$C_\ell$ estimator - discrete formalism}
    The main complication in using the pseudo-$C_\ell$ formalism in the continuum limit is the need to compute the angle-averaged mask pseudo power spectrum $C^{vw}_\ell$. Flat fields are most easily analyzed when pixelized in a Cartesian grid, and under this setup the mask power spectrum is not a well defined quantity for infinitesitally small intervals of $\ell$. This leads to non-negligible biases and poor performance \cite{2016arXiv161204664A} due to the need to use highly resolved finite intervals to compute the integrals presented in the previous section. This motivates the discrete formalism described here, which connects directy with the data storage format of flat-sky maps.

    We will discuss the flat-sky pseudo-$C_\ell$ algorithm starting from a pixelized representation of the sky map.
    Let the patch of sky under inspection be contained by a rectangle of sides $L_x$ and $L_y$ (in units of radians),
    and let us discretize this rectangle by dividing it into an $N_x\times N_y$ grid with pixels of area
    $\Delta{\bf x}^2\equiv\Delta x\Delta y\equiv (L_x/N_x)(L_y/N_y)$. Each pixel in this grid is then labelled by a
    pair of integers ${\bf n}\equiv(n_x,n_y)$, and is assigned coordinates ${\bf x}_{\bf n}\equiv(n_x\Delta x,n_y\Delta y)$.
    The each component of the pixelized map ${\bf a}({\bf x})$ is therefore defined for
    $n_x\in[0,N_x-1],\,\,n_y\in[0,N_y-1]$.

    In this scenario, the spin Fourier transform of the pixelized field can be computed as its discrete Fourier transform:
    \begin{equation}
      {\bf a}_{\bf k}\equiv {\rm DFT}\left({\bf a}\right)^{s_a}_{\bf k}\equiv\sum_{\bf x}\frac{\Delta{\bf x}^2}{2\pi}\,_{s_a}{\sf E}_{\bf k}({\bf x}){\bf a}_{\bf x},
    \end{equation}
    where the wavenumber ${\bf k}$ is now discretized as ${\bf k}=(j_x\Delta k_x,j_y\Delta k_y)$, the integers
    $j_{(x,y)}$ run from $-N_{(x,y)}/2$ tof $N_{(x,y)}/2-1$\footnote{This is the valid domain when $N_{(x,y)}$ is even.
    For odd $N_{(x,y)}$ the interval becomes $[-(N_{(x,y)}-1)/2,(N_{(x,y)}-1)/2]$} and the pixel size is
    $\Delta k_{(x,y)}\equiv2\pi/L_{(x,y)}$.

    The following properties of the DFT are worth remembering:
    \begin{itemize}
     \item Periodicity: ${\bf a}_{(k_x,k_y)}={\bf a}_{(k_x+N_x\Delta k_x,k_y)}={\bf a}_{(k_x,k_y+N_y\Delta k_y)}={\bf a}_{(k_x+N_x\Delta k_x,k_y+N_y\Delta k_y)}$
     \item The power spectrum of the DFT is defined as:
           \begin{equation}
             \langle{\bf a}_{\bf l}{\bf b}^\dag_{\bf k}\rangle\equiv\frac{\delta_{{\bf l},{\bf k}}}{\Delta{\bf k}^2}{\sf C}^{ab}_{\bf l}
           \end{equation}
     \item For a real-valued scalar $a$, its DFT satisfies $a^*_{{\bf k}_{\bf j}}=a^{}_{{\bf k}_{{\bf N}-{\bf j}}}$ (where ${\bf N}\equiv(N_x,N_y)$).
     \item The orthogonality relation of the basis functions now takes the form:
           \begin{equation}
             \sum_{\bf x}\,_sE^\dag_{\bf l}({\bf x})\,_sE_{\bf k}({\bf x})=N_xN_y\delta_{{\bf l},{\bf k}}\hat{\sf 1}
           \end{equation}
    \end{itemize}

    Using the properties above one can prove that the Fourier coefficients of a masked field are:
    \begin{equation}
      {\bf a}^v_{\bf l}\equiv\left(v({\bf x}){\bf a}({\bf x})\right)_{\bf l}=\sum_{\bf k}\frac{\Delta{\bf k}^2}{2\pi}{\sf R}(s_a(\varphi_{\bf l}-\varphi_{\bf k})){\bf a}_{\bf k}v_{{\bf l}-{\bf k}},
    \end{equation}
    and we can easily find the following relation for the covariance of two masked fields at the same wavenumber:
    \begin{equation}
      \left\langle{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{\bf l}\right\rangle=\sum_{\bf k}\frac{\Delta{\bf k}^2}{(2\pi)^2}{\sf R}(s_a\Delta\varphi){\sf C}^{ab}_{\bf k}{\sf R}^\dag(s_b\Delta\varphi)\,v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}
    \end{equation}

    At this stage it is natural to connect directly with the final bandpowers (since there is no natural minimal binning as in the case of the full-sky multipoles). In this case we will define a bandpower ${\sf B}_b$, indexed by an integer $b$ as the average of the covariance above over a set of ${\bf l}$ values $S_b$:
    \begin{equation}
      \,_v{\sf B}^{ab}_b\equiv\frac{\Delta{\bf k}^2}{N_b}\sum_{{\bf l}\in S_b}{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{\bf l},
    \end{equation}
    where $N_b$ is the number of Fourier-space pixels covered by $S_b$. The bandpowers are related to the true power spectrum through:
    \begin{equation}
      \langle\,_v{\sf B}^{ab}_b\rangle=\sum_{{\bf l}\in S_b}N_b^{-1}\sum_{\bf k}{\sf M}^{s_a\,s_b}_{{\bf l}{\bf k}}\cdot\,_v\hat{\sf C}^{ab}_{\bf k},
    \end{equation}    
    where the un-binned mode-coupling matrix is:
    \begin{align}
      & M^{00}_{{\bf l}\,{\bf k}}\equiv\frac{(2\pi)^2}{L_x^2L_y^2}v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}},\\
      & {\sf M}^{02}_{{\bf l}\,{\bf k}}\equiv\frac{(2\pi)^2}{L_x^2L_y^2}v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}
      \left(
      \begin{array}{cc}
        \cos2\Delta\varphi & -\sin2\Delta\varphi \\
        \sin2\Delta\varphi &  \cos2\Delta\varphi
      \end{array}
      \right)\\
      & {\sf M}^{22}_{{\bf l}\,{\bf k}}\equiv\frac{(2\pi)^2}{L_x^2L_y^2}v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}
      \left(
      \begin{array}{cccc}
        \cos^22\Delta\varphi & -\cos2\Delta\varphi\sin2\Delta\varphi & -\cos2\Delta\varphi\sin2\Delta\varphi & \sin^22\Delta\varphi\\
        \cos2\Delta\varphi\sin2\Delta\varphi & \cos^22\Delta\varphi & -\sin^22\Delta\varphi & -\cos2\Delta\varphi\sin2\Delta\varphi\\
        \cos2\Delta\varphi\sin2\Delta\varphi & -\sin^22\Delta\varphi & \cos^22\Delta\varphi & -\cos2\Delta\varphi\sin2\Delta\varphi\\
        \sin^22\Delta\varphi &  \cos2\Delta\varphi\sin2\Delta\varphi & \cos2\Delta\varphi\sin2\Delta\varphi & \cos^22\Delta\varphi\\
      \end{array}
      \right)
    \end{align}
    The mode-coupling matrix for the bandpowers is therefore given by:
    \begin{equation}
      \mathcal{M}^{s_as_b}_{bb'}=\sum_{{\bf l}\in S_b}N_b^{-1}\sum_{{\bf k}\in S_{b'}}{\sf M}^{s_as_b}_{{\bf l}\,{\bf k}},
    \end{equation}

  \subsection{Contaminant cleaning}
    Using the same notation as in Section \ref{sec:contaminant}, the contaminant-cleaned version of ${\bf a}$ is (c.f. Eqs. \ref{eq:deproject_real} \ref{eq:deproject_harmonic}):
    \begin{equation}
      \tilde{\bf a}({\bf x})={\bf a}^v({\bf x})-{\bf f}^i({\bf x})M_{ij}\sum_{\bf x}\Delta{\bf x}^2\,{\bf f}^{j\dag}({\bf x}'){\bf a}^v({\bf x}'),\hspace{12pt}
      \tilde{\bf a}_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}M_{ij}\sum_{\bf k}\Delta{\bf k}^2\,{\bf f}^{j\dag}_{\bf k}{\bf a}^v_{\bf k},
    \end{equation}
    and the unbiased pseudo-$C_\ell$ estimator takes the form (c.f. Eq. \ref{eq:unbiased_pcl}):
    \begin{align}\nonumber
      \left\langle\tilde{\bf a}_{\bf l}\tilde{\bf b}_{\bf l}^\dag\right\rangle=&\left\langle{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{\bf l}\right\rangle+N^*_{ij}\,{\rm DFT}\left\{v({\bf x}){\rm DFT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm DFT}\left(w\,{\bf g}^j\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\bf x}\right\}^{s_a}_{\bf l}{\bf g}^{i\dag}_{\bf l}+\\\nonumber
      &+M_{ij}{\bf f}^i_{\bf l}\,{\rm DFT}\left\{w({\bf x})\,{\rm DFT}^{-1}\left[\hat{\sf C}^{ab\dag}_{\ell_1}{\rm DFT}\left(v\,{\bf f}^j\right)^{s_a}_{{\bf l}_1}\right]^{s_b}_{\bf x}\right\}^{s_b\dag}_{\bf l}-\\
      &-M_{ij}N^*_{pq}\left\{\sum_{\bf x} \Delta{\bf x}^D\,v({\bf x})\,{\bf f}^{j\dag}({\bf x})\,{\rm DFT}^{-1}\left[\hat{\sf C}^{ab}_{\ell_1}{\rm DFT}\left(w\,{\bf g}^q\right)^{s_b}_{{\bf l}_1}\right]^{s_a}_{\bf x}\right\}\,{\bf f}^i_{\bf l}{\bf g}^{p\dag}_{\bf l}
    \end{align}

  \subsection{$E$ and $B$ purification}
    The analogue equations for the pure $B$ mode component in the flat-sky approximation are:
    \begin{equation}
      B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2\ell^{-1}\left(\tilde{P}_1\right)^B_{\bf l}+\ell^{-2}\left(\tilde{P}_0\right)^B_{\bf l},
    \end{equation}
    (and a similar relation for the pure $E$ component), where $P_n=(\eth^{2-n}w)^*(Q+iU)$, and $w$ is the sky mask. The derivatives of $w$
    can be taken by using the following relation:
    \begin{align}
      &\eth^nw({\bf x})=\int\frac{d{\bf l}^2}{2\pi}\eth^ne^{i{\bf l}{\bf x}}w_{\bf l}=\int\frac{d{\bf l}^2}{2\pi}\,_n{\cal Y}_{\bf k}({\bf x})\,(l^nw_{\bf l})\hspace{6pt}
      \longrightarrow\hspace{6pt}\,_nw_{\bf l}=\ell^nw_{\bf l}\\
      &(\eth^nw({\bf x}))*=\bar{\eth}^nw({\bf x})=\int\frac{d{\bf l}^2}{2\pi}\,_{-n}{\cal Y}_{\bf k}({\bf x})\,((-l)^nw_{\bf l})\hspace{6pt}
      \longrightarrow\hspace{6pt}\,_{-n}w_{\bf l}=(-1)^n\ell^nw_{\bf l}\\
      &((\eth^nw({\bf x}))^E,(\eth^nw({\bf x}))^B)=(-\ell^nw_{\bf l},0)
    \end{align}

    Note that a mathematically simpler relation for the pure components is:
    \begin{equation}
      B^p_{\bf l}=\int\frac{d{\bf k}^2}{2\pi}B_{\bf k}w_{{\bf l}-{\bf k}}\frac{k^2}{\ell^2}
    \end{equation}
    (and similarly for $E^p_{\bf l}$). Comparing with Eq. \ref{eq:pseudo_flat_eb}, we see that the key to work out the expressions for the pure-$E$ and $B$ coupling matrices is to substitute all factors of $\cos2\Delta\varphi$ for $k^2/\ell^2$, and all factors of $\sin2\Delta\varphi$ for 0.

  \subsection{Gaussian covariance}
    Using approximations similar to those described in the full-sky case (see previous section), we can prove that the Gaussian covariance matrix for the coupled bandpowers of 4 scalar fields $a$, $b$, $c$ and $d$ is given by:
    \begin{equation}
      \left\langle \Delta{\sf B}^{ab}_q \Delta{\sf B}^{cd}_{q'}\right\rangle=
      {\sf B}^{ac}_{(q}{\sf B}^{bd}_{q')}\,\Xi_{qq'}(w^aw^c,w^bw^d)+{\sf B}^{ad}_{(q}{\sf B}^{bc}_{q')}\,\Xi_{qq'}(w^aw^d,w^bw^c),
    \end{equation}
    where
    \begin{equation}
      \Xi_{qq'}(\phi,\psi)=\sum_{{\bf l}\in S_q}N_q^{-1}\sum_{{\bf l}'\in S_{q'}}N_{q'}^{-1}\frac{(2\pi)^2}{L_x^2L_y^2}\phi_{{\bf l}-{\bf l}'}\psi^*_{{\bf l}-{\bf l}'}.
    \end{equation}

\begin{thebibliography}{}
 \bibitem{2002ApJ...567....2H} Hivon, E., G{\'o}rski, K.~M., Netterfield, C.~B., et al.\ 2002, \apj, 567, 2 
 \bibitem{2017MNRAS.465.1847E} Elsner, F., Leistedt, B., \& Peiris, H.~V.\ 2017, \mnras, 465, 1847 
 \bibitem{2003ApJS..148..161K} Kogut, A., Spergel, D.~N., Barnes, C., et al.\ 2003, \apjs, 148, 161 
 \bibitem{2016arXiv161204664A} Asgari, M., Taylor, A., Joachimi, B., \& Kitching, T.~D.\ 2016, arXiv:1612.04664 
 \bibitem{2004MNRAS.349..603E} Efstathiou, G.\ 2004, \mnras, 349, 603
 \bibitem{2005MNRAS.360.1262B} Brown, M.~L., Castro, P.~G., \& Taylor, A.~N.\ 2005, \mnras, 360, 1262 
\end{thebibliography}

\end{document}

\subsection{Bandpowers}
Consider the case where you want to compute the power spectrum in band-powers given by
\begin{equation}
  \mathbf{B}^{ab}_k\equiv\frac{1}{N_k}\sum_{\ell=\ell_k}^{\ell_k+N_k-1} f(\ell)\,\mathbf{C}^{ab}_\ell,
\end{equation}
then Eq. \ref{eq:pcl_gen} above becomes
\begin{equation}\label{eq:pcl_bin}
  \langle\tilde{\mathbf{B}}^{ab}_k\rangle=\sum_{k'}\mathsf{M}^{B,s_as_b}_{k k'}\cdot\mathbf{B}^{ab}_{k'}+
  \langle\tilde{\mathsf{N}}^{B,ab}_k\rangle
\end{equation}
where the binned coupling matrix $\mathsf{M}^{B,s_as_b}$ is
\begin{equation}
  \mathsf{M}^{B,s_as_b}_{k_1,k_2}\equiv\frac{1}{N_{k_1}}\sum_{\ell_1=\ell_{k_1}}^{\ell_{k_1}+N_{k_1}-1}\sum_{\ell_2=\ell_{k_2}}^{\ell_{k_2}+N_{k_2}-1}\frac{f(\ell_1)}{f(\ell_2)}
  \mathsf{M}^{s_as_b}_{\ell_1\ell_2}
\end{equation}

\section{Mode deprojection}
Let ${\bf t}$ be a set of contaminant templates such that we can write:
\begin{equation}
  {\bf d}={\bf s}+{\bf n}+\hat{\bf t}\cdot{\bf a}.
\end{equation}
A maximum-likelihood estimate for ${\bf a}$ can be found as
\begin{equation}
 \tilde{\bf a}=(\hat{\bf t}^T\hat{\bf C}^{-1}\hat{\bf t})^{-1}\hat{\bf t}\hat{\bf C}^{-1}{\bf d}
\end{equation}

\section{Minimum variance quadratic estimator}

Let ${\bf d}_1$ and ${\bf d}_2$ be the data, given as a pixelized full-sky maps,
with covariance matrix
\begin{equation}
 \langle{\bf d}_1\,{\bf d}^T_1\rangle\equiv C^{12}(\hat{\mu})\equiv
 S^{12}(\hat{\mu})+N^{12}(\hat{\mu})=
 \sum_\ell C^{12}_\ell\,P_\ell(\hat{\mu})+N^{12}(\hat{\mu}),
\end{equation}
where $\hat{\mu}={\bf n}\,{\bf n}^T$, ${\bf n}$ is the vector of unit vectors
containing the angular coordinates to each pixel and $P_\ell(x)\equiv (2\ell+1)
L_\ell(x)/(4\pi)$, where $L_\ell$ are the Legendre polynomials. Note that we can
expand $P_\ell$ as
\begin{equation}
  P_\ell(\hat{\mu})\equiv\sum_{m=-\ell}^\ell
  Y_{\ell m}({\bf n})\cdot Y^\dag_{\ell m}({\bf n}).
\end{equation}

We will parametrize the power spectrum with a set of step functions, such that
\begin{equation}
 C^{12}_\ell=\sum_b c_b\,\Theta_b(\ell),\hspace{12pt}\leftarrow\hspace{12pt}
 \Theta_b(\ell)=1\,{\rm if}\,\ell\in[\ell^b_{\rm min},\ell^b_{\rm max}),
 \,\,0\,{\rm otherwise},
\end{equation}
thus we can write
\begin{equation}
  C^{12}(\hat{\mu})=\sum_b\,c_b\,Q_b(\hat{\mu})+N^{12}(\hat{\mu}),
  \hspace{6pt}{\rm with}\hspace{6pt}
  Q_b(\mu)=\sum_\ell\Theta_b(\ell)P_\ell(\mu).
\end{equation}

Let us write the most general quadratic estimator for the coefficients $c_b$:
\begin{equation}
 \tilde{c}_b\equiv {\bf d}^T_1\hat{E}_b{\bf d}_2-B_b.
\end{equation}
The mean value of $\tilde{c}_b$ is:
\begin{equation}
 \langle\tilde{c}_b\rangle=\sum_a c_a\Tr(\hat{E}_b\,\hat{Q}_a)+
 \Tr(\hat{E_b}\hat{N}^{12})-B_b,
\end{equation}
and therefore we can remove the noise bias by defining
\begin{equation}
  B_b\equiv\Tr(\hat{E_b}\hat{N}^{12}).
\end{equation}

The covariance matrix of this estimator is:
\begin{equation}
 \left\langle(\tilde{c}_a-\langle\tilde{c}_a\rangle)\,
 (\tilde{c}_b-\langle\tilde{c}_b\rangle)\right\rangle=
 \Tr(\hat{C}^{11}\hat{E}_a\hat{C}^{22}\hat{E}_b),
\end{equation}
where we have approximated $C_\ell^{12}\ll[C_\ell^{11},C_\ell^{22}]$. Minimizing
the variance of $\tilde{c}_a$ under the constrain that the coupling matrix
$W_{ab}\equiv\Tr(\hat{E}_b\,\hat{Q}_a)$ have unit diagonal, we find the following
minimization problem:
\begin{equation}
  L=\Tr\left[\hat{E}_b\hat{C}^{11}\hat{E}_b\hat{C}^{22}-
  2(\lambda\hat{E}_b\hat{Q}_b-1)\right],
\end{equation}
where $\lambda$ is the lagrange multiplier related to the constraint. The solution
to this problem is:
\begin{equation}
 \hat{E}_b=\frac{\hat{C}_{11}^{-1}\hat{Q}_b\hat{C}_{22}^{-1}}
 {\Tr(\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}\hat{Q}_b)}.
\end{equation}

Thus, a decoupled, unbiased and minimum-variance estimator, $\hat{c}_b$, for
$c_b$ can be found as:
\begin{equation}
 {\bf d}^T_1\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}{\bf d}_2-
 \Tr(\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}\hat{N})
 =\sum_a \hat{c}_a\Tr(\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}\hat{Q}_a).
\end{equation}
Note that this estimator requires a guess for the covariance matrix of the data
$C^{ij}(\hat{\mu})$, and exact minimum variance will only be attained for the true
covariance (which itself depends on the power spectrum coefficients $c_b$).
The choice of prior covariance will define the type of estimator. Note that this
formalism immediately encompasses cut-skies by setting to zero all elements of the
full covariance matrix involving unobserved pixels (this would correspond to the
limit of infinite uncorrelated noise for those pixels).
\subsection{Strategies for the different terms}
\begin{itemize}
\item $\hat{C}^{-1}_{ii}\,{\bf d}_i$. This term can be computed by solving the
linear system:
\begin{equation}
  \hat{C}_{ii}{\bf z}_i={\bf d}_i
\end{equation}
via conjugate gradients. The action of $\hat{C}=\hat{S}+\hat{N}$ can be computed as
follows:
\begin{align}
 [\hat{S}{\bf z}](\nv_1)&\equiv\sum_{\nv_2}S(\nv_1\cdot\nv_2)z(\nv_2)\\
                        &=\frac{1}{\Delta\Omega}\int d\nv_2\sum_{\ell m}C_\ell
                          Y^*_{\ell m}(\nv_1)Y^*_{\ell m}(\nv_1)\,\tilde{z}(\nv_2)\\
                        &=\frac{1}{\Delta\Omega}\sum_{\ell m} Y^*_{\ell m}(\nv_1)
                        C_\ell \tilde{z}_{\ell m}\\
                        &=\frac{1}{\Delta\Omega}
{\rm SHT}^{-1}\left[C_\ell\,{\rm SHT}[\tilde{z}(\nv)]_{\ell m}\right],
\end{align}
where $\tilde{z}$ is the extension of $z$ to the whole sphere, with $\tilde{z}=0$ in
all unobserved pixels.

In most cases the noise power spectrum will be white, in which case $\hat{N}{\bf z}$
is just $\sum_{\nv}\sigma_N^2(\nv)z(\nv)$, with $\sigma^2(\nv)$ the per-pixel
variance.

Note that this procedure works for any $\hat{C}^{-1}{\bf v}$-type operation
(see below).

\item ${\bf d}^T_1\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}{\bf d}_2$. With
${\bf z}_i$ defined as above, this is just:
\begin{align}
 {\bf z}^T_i\hat{Q}_b{\bf z}_j=
 \frac{1}{(\Delta\Omega)^2}\sum_\ell\Theta_b(\ell)
 \sum_m(\tilde{z}^{i}_{\ell m})^*\tilde{z}^{i}_{\ell m}
\end{align}
\item $\Tr(\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}\hat{Q}_a)$. Let ${\bf v}$ be a random
vector with covariance $\langle {\bf v}{\bf v}^T\rangle=\hat{1}$, then one can
calculate traces by averaging over realizations of such
vectors:
 \begin{equation}
   \Tr{\hat{A}}=\left\langle {\bf v}^T\hat{A}{\bf v}\right\rangle.
 \end{equation}
 Then, the only thing to bear in mind in order to compute the trace above is that the action of the
 $\hat{Q}_a$ operator is:
 \begin{equation}
   \left(\hat{Q}_a{\bf v}\right)_{\nv}=\frac{1}{\Delta\Omega}{\rm SHT}^{-1}
   \left[\Theta_b(\ell){\rm SHT}\left[\tilde{v}\right]\right]_{\nv},
 \end{equation}
 where, as before $\tilde{v}$ is the extension of $v$ with zeros in all unobserved pixels.

 In order to minimize the number of operations needed to compute this trace we can compute
 it as:
 \begin{equation}
   \Tr(\hat{C}^{-1}_{11}\hat{Q}_b\hat{C}^{-1}_{22}\hat{Q}_a)=
   \left\langle(\hat{C}^{-1}_{11}\,{\bf v})^T\hat{Q}_b(\hat{C}^{-1}_{22}\hat{Q}_a\,{\bf v})\right\rangle
 \end{equation}

\end{itemize}

%\subsection{Computing $\langle\tilde{C}^N_l\rangle$}
%Consider the case where the window function is just a flat top-hat multiplied by the inverse variance of the noise:
%\begin{equation}
%  W(\nv)\equiv\Theta(\nv)\frac{\bar{\sigma}_N^2}{\sigma_N^2(\nv)},
%\end{equation}
%where $\sigma_N^2(\nv)$ is the variance per sterad at each point and $\bar{\sigma}^2_N$ is its sky-averaged value.
%
%For uncorrelated noise, we can write $N(\nv)$ as
%\begin{equation}
% N(\nv)=u(\nv)\,\sigma_N(\nv)
%\end{equation}
%where $u(\nv)$ is a white GRF with power spectrum $C^u_l=1$. In this case, the noise pseudo-$C_l$ can be estimated theoretically:
%\begin{equation}
% \langle\tilde{C}_l^N\rangle=\frac{1}{4\pi}\int d\Omega\,\sigma_N^2(\nv)W^2(\nv)=
% \frac{\bar{\sigma}_N^2}{4\pi}\int d\Omega\, \Theta(\nv)\frac{\bar{\sigma}_N^2}{\sigma_N^2(\nv)}=
% f_{\rm sky}\,\bar{\sigma}_N^2\overline{\left(\frac{\bar{\sigma}_N^2}{\sigma^2_N}\right)}
%\end{equation}

### 2018-07-29 doc/source/conf.py

@@ -1,0 +1,180 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+#
+# pymaster documentation build configuration file, created by
+# sphinx-quickstart on Fri Feb 17 04:34:56 2017.
+#
+# This file is execfile()d with the current directory set to its
+# containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+#
+import os
+import sys
+from mock import Mock as MagicMock
+sys.path.insert(0, os.path.abspath('../../'))
+sys.path.append(os.path.abspath('../../pymaster/'))
+
+autodoc_mock_imports = ['nmtlib','_nmtlib','numpy']
+
+class Mock(MagicMock):
+    @classmethod
+    def __getattr__(cls, name):
+            return MagicMock()
+
+MOCK_MODULES = ['nmtlib']
+sys.modules.update((mod_name, Mock()) for mod_name in MOCK_MODULES)
+# -- General configuration ------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+#
+# needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = ['sphinx.ext.autodoc',
+              'sphinx.ext.mathjax']
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix(es) of source filenames.
+# You can specify multiple suffix as a list of string:
+#
+# source_suffix = ['.rst', '.md']
+source_suffix = '.rst'
+
+# The master toctree document.
+master_doc = 'index'
+
+# General information about the project.
+project = 'pymaster'
+copyright = '2017, David Alonso'
+author = 'David Alonso'
+
+# The version info for the project you're documenting, acts as replacement for
+# |version| and |release|, also used in various other places throughout the
+# built documents.
+#
+# The short X.Y version.
+version = '0.1'
+# The full version, including alpha/beta/rc tags.
+release = '0.1'
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+#
[…]

### 2018-07-29 pymaster/__init__.py

@@ -1,0 +1,46 @@
+"""
+:mod:`pymaster` contains three basic classes:
+
+- :class:`pymaster.field.NmtField`
+- :class:`pymaster.bins.NmtBin`
+- :class:`pymaster.workspaces.NmtWorkspace`
+- :class:`pymaster.covariance.NmtCovarianceWorkspace`
+
+and a number of functions
+
+- :func:`pymaster.workspaces.deprojection_bias`
+- :func:`pymaster.workspaces.compute_coupled_cell`
+- :func:`pymaster.workspaces.compute_full_master`
+- :func:`pymaster.covariance.gaussian_covariance`
+- :func:`pymaster.utils.mask_apodization`
+- :func:`pymaster.utils.synfast_spherical`
+
+:mod:`pymaster` also comes with a flat-sky version with most of the same functionality:
+
+- :class:`pymaster.field.NmtFieldFlat`
+- :class:`pymaster.bins.NmtBinFlat`
+- :class:`pymaster.workspaces.NmtWorkspaceFlat`
+
+- :func:`pymaster.workspaces.deprojection_bias_flat`
+- :func:`pymaster.workspaces.compute_coupled_cell_flat`
+- :func:`pymaster.workspaces.compute_full_master_flat`
+- :func:`pymaster.covariance.gaussian_covariance_flat`
+- :func:`pymaster.utils.mask_apodization_flat`
+- :func:`pymaster.utils.synfast_flat`
+
+Many of these function accept or return sets of power spectra (arrays with one element per angular multipole) or bandpowers (binned versions of power spectra). In all cases, these are returned and provided as 2D arrays with shape [n_cls][nl], where n_cls is the number of power spectra and nl is either the number of multipoles or bandpowers. In all cases, n_cls should correspond with the spins of the two fields being correlated, and the ordering is as follows:
+
+- Two spin-0 fields: n_cls=1, [C_T1T2]
+- One spin-0 field and one spin-2 field: n_cls=2, [C_TE,C_TB]
+- Two spin-2 fields: n_cls=4, [C_E1E2,C_E1B2,C_E2B1,C_B1B2]
+
+All sky maps accepted and returned by these functions are in the form of HEALPix maps exclusively with RING ordering.
+
+"""
+from pymaster import nmtlib as lib
+import numpy as np
+from pymaster.utils import mask_apodization, mask_apodization_flat, synfast_spherical, synfast_flat
[…]

### 2018-07-29 pymaster/bins.py

@@ -1,0 +1,161 @@
+from pymaster import nmtlib as lib
+import numpy as np
+
+class NmtBin(object) :
+    """
+    An NmtBin object defines the set of bandpowers used in the computation of the pseudo-Cl estimator. The definition of bandpowers is described in Section 3.6 of the scientific documentation.
+
+    :param int nside: HEALPix nside resolution parameter of the maps you intend to correlate. The maximum multipole considered for bandpowers will be 3*nside-1.
+    :param array-like ells: array of integers corresponding to different multipoles
+    :param array-like bpws: array of integers that assign the multipoles in ells to different bandpowers
+    :param array-like weights: array of floats corresponding to the weights associated to each multipole in ells. The sum of weights within each bandpower is normalized to 1.
+    :param int nlb: integer value corresponding to a constant bandpower width. I.e. the bandpowers will be defined as consecutive sets of nlb multipoles from l=2 to l=lmax (see below) with equal weights. If this argument is provided, the values of ells, bpws and weights are ignored.
+    :param int lmax: integer value corresponding to the maximum multiple used by these bandpowers. If None, it will be set to 3*nside-1. In any case the actual maximum multipole will be chosen as the minimum of lmax, 3*nside-1 and the maximum element of ells.
+    """
+    def __init__(self,nside,bpws=None,ells=None,weights=None,nlb=None,lmax=None) :
+        if((bpws is None) and (ells is None) and (weights is None) and (nlb is None)) :
+            raise KeyError("Must supply bandpower arrays or constant bandpower width")
+
+        if(lmax is None) :
+            lmax_in=3*nside-1
+        else :
+            lmax_in=lmax
+
+        if(nlb is None) :
+            if((bpws is None) and (ells is None) and (weights is None)) :
+                raise KeyError("Must provide bpws, ells and weights")
+            ell_max=min(3*nside-1,np.amax(ells),lmax_in)
+            self.bin=lib.bins_create_py(bpws.astype(np.int32),ells.astype(np.int32),weights,ell_max)
+        else :
+            ell_max=min(3*nside-1,lmax_in)
+            self.bin=lib.bins_constant(nlb,ell_max)
+        self.lmax=ell_max
+
+    def __del__(self) :
+        lib.bins_free(self.bin)
+
+    def get_n_bands(self) :
+        """
+        Returns the number of bandpowers stored in this object
+
+        :return: number of bandpowers
+        """
[…]

### 2018-07-29 pymaster/covariance.py

@@ -1,0 +1,132 @@
+from pymaster import nmtlib as lib
+import numpy as np
+
+class NmtCovarianceWorkspace(object) :
+    """
+    NmtCovarianceWorkspace objects are used to compute and store the coupling coefficients needed to calculate the Gaussian covariance matrix under the Efstathiou approximation (astro-ph/0307515). When initialized, this object is practically empty. The information describing the coupling coefficients must be computed or read from a file afterwards.
+    """
+    def __init__(self) :
+        self.wsp=None
+    
+    def __del__(self) :
+        if(self.wsp is not None) :
+            lib.covar_workspace_free(self.wsp)
+
+    def read_from(self,fname) :
+        """
+        Reads the contents of an NmtCovarianceWorkspace object from a file (encoded using an internal binary format).
+
+        :param str fname: input file name
+        """
+        if self.wsp is not None :
+            lib.covar_workspace_free(self.wsp)
+        self.wsp=lib.covar_workspace_read(fname);
+
+    def compute_coupling_coefficients(self,wa,wb) :
+        """
+        Computes coupling coefficients of the Gaussian covariance between the power spectra computed using wa and wb (two NmtWorkspace objects).
+
+        :param NmtWorkspace wa,wb: workspaces used to compute the two power spectra whose covariance matrix you want to compute.
+        """
+        ns=wa.wsp.nside;
+        if(wa.wsp.nside!=wb.wsp.nside) :
+            raise ValueError("Everything should have the same resolution!")
+        if((wa.wsp.ncls!=1) or (wb.wsp.ncls!=1)) :
+            raise ValueError("Gaussian covariances only supported for spin-0 fields")
+        if self.wsp is not None :
+            lib.covar_workspace_free(self.wsp)
+        self.wsp=lib.covar_workspace_init(wa.wsp,wb.wsp)
+
+    def write_to(self,fname) :
+        """
+        Writes the contents of an NmtCovarianceWorkspace object to a file (encoded using an internal binary format).
+
+        :param str fname: output file name
+        """
+        if self.wsp is None :
+            raise KeyError("Must initialize workspace before writing")
+        lib.covar_workspace_write(self.wsp,fname)
+
+class NmtCovarianceWorkspaceFlat(object) :
+    """
[…]

### 2018-07-29 pymaster/field.py

@@ -1,0 +1,189 @@
+from pymaster import nmtlib as lib
+import numpy as np
+
+class NmtField(object) :
+    """
+    An NmtField object contains all the information describing the fields to correlate, including their observed maps, masks and contaminant templates.
+
+    :param mask: array containing a HEALPix map corresponding to the field's mask.
+    :param maps: 2D array containing the observed maps for this field. The first dimension corresponds to the number of maps, which should be 1 for a spin-0 field and 2 for a spin-2 field.
+    :param templates: 3D array containing a set of contaminant templates for this field. This array should have shape [ntemp][nmap][npix], where ntemp is the number of templates, nmap should be 1 for spin-0 fields and 2 for spin-2 fields, and npix is the number of pixels per map. The best-fit contribution from each contaminant is automatically removed from the maps unless templates=None
+    :param beam: spherical harmonic transform of the instrumental beam (assumed to be rotationally symmetric - i.e. no m dependence). If None, no beam will be corrected for. Otherwise, this array should have 3*nside elements, corresponding to multipoles from 0 to 3*nside-1.
+    :param purify_e: use pure E-modes?
+    :param purify_b: use pure B-modes?
+    :param n_iter_mask_purify: number of iterations used to compute an accurate SHT of the mask when using E/B purification
+    :param tol_pinv: when computing the pseudo-inverse of the contaminant covariance matrix, all eigenvalues below tol_pinv * max_eval will be treated as singular values, where max_eval is the largest eigenvalue. Only relevant if passing contaminant templates that are likely to be highly correlated.
+
+    """
+    def __init__(self,mask,maps,templates=None,beam=None,purify_e=False,purify_b=False,
+                 n_iter_mask_purify=3,tol_pinv=1E-10) :
+        pure_e=0
+        if(purify_e) :
+            pure_e=1
+        pure_b=0
+        if(purify_b) :
+            pure_b=1
+
+        nside=2
+        while(12*nside*nside!=len(mask)) :
+            nside*=2
+            if(nside>65536) :
+                raise KeyError("Something is wrong with your input arrays")
+
+        if((len(maps)!=1) and (len(maps)!=2)) :
+            raise KeyError("Must supply 1 or 2 maps per field")
+
+        if isinstance(templates,(list,tuple,np.ndarray)) :
[…]

### 2018-07-29 pymaster/nmtlib.py

@@ -1,0 +1,1083 @@
+# This file was automatically generated by SWIG (http://www.swig.org).
+# Version 3.0.8
+#
+# Do not make changes to this file unless you know what you are doing--modify
+# the SWIG interface file instead.
+
+
+
+
+
+from sys import version_info
+if version_info >= (2, 6, 0):
+    def swig_import_helper():
+        from os.path import dirname
+        import imp
+        fp = None
+        try:
+            fp, pathname, description = imp.find_module('_nmtlib', [dirname(__file__)])
+        except ImportError:
+            import _nmtlib
+            return _nmtlib
+        if fp is not None:
+            try:
+                _mod = imp.load_module('_nmtlib', fp, pathname, description)
+            finally:
+                fp.close()
+            return _mod
+    _nmtlib = swig_import_helper()
+    del swig_import_helper
+else:
+    import _nmtlib
+del version_info
+try:
+    _swig_property = property
+except NameError:
+    pass  # Python < 2.2 doesn't have 'property'.
+
+
+def _swig_setattr_nondynamic(self, class_type, name, value, static=1):
+    if (name == "thisown"):
+        return self.this.own(value)
+    if (name == "this"):
+        if type(value).__name__ == 'SwigPyObject':
+            self.__dict__[name] = value
+            return
+    method = class_type.__swig_setmethods__.get(name, None)
+    if method:
+        return method(self, value)
+    if (not static):
+        if _newclass:
+            object.__setattr__(self, name, value)
+        else:
+            self.__dict__[name] = value
+    else:
+        raise AttributeError("You cannot add attributes to %s" % self)
+
+
+def _swig_setattr(self, class_type, name, value):
+    return _swig_setattr_nondynamic(self, class_type, name, value, 0)
+
+
+def _swig_getattr_nondynamic(self, class_type, name, static=1):
+    if (name == "thisown"):
+        return self.this.own()
[…]

### 2018-07-29 pymaster/utils.py

@@ -1,0 +1,128 @@
+from pymaster import nmtlib as lib
+import numpy as np
+
+def mask_apodization(mask_in,aposize,apotype="C1") :
+    """
+    Apodizes a mask with an given apodization scale using different methods.
+
+    :param mask_in: input mask, provided as an array of floats corresponding to a HEALPix map in RING order.
+    :param aposize: apodization scale in degrees.
+    :param apotype: apodization type. Three methods implemented: "C1", "C2" and "Smooth". See the description of the C-function nmt_apodize_mask in the C API documentation for a full description of these methods.
+    :return: apodized mask as a HEALPix map
+    """
+    return lib.apomask(mask_in.astype('float64'),len(mask_in),aposize,apotype)
+
+def mask_apodization_flat(mask_in,lx,ly,aposize,apotype="C1") :
+    """
+    Apodizes a flat-sky mask with an given apodization scale using different methods.
+
+    :param mask_in: input mask, provided as a 2D array (ny,nx) of floats.
+    :param float lx: patch size in the x-axis (in radians)
+    :param float ly: patch size in the y-axis (in radians)
+    :param aposize: apodization scale in degrees.
+    :param apotype: apodization type. Three methods implemented: "C1", "C2" and "Smooth". See the description of the C-function nmt_apodize_mask in the C API documentation for a full description of these methods.
+    :return: apodized mask as a 2D array (ny,nx)
+    """
+    nx=len(mask_in[0])
+    ny=len(mask_in)
+    mask_apo_flat=lib.apomask_flat(nx,ny,lx,ly,mask_in.flatten().astype('float64'),nx*ny,aposize,apotype)
+    return mask_apo_flat.reshape([ny,nx])
+
+def synfast_spherical(nside,cls,pol=False,beam=None) :
+    """
+    Generates a full-sky Gaussian random field according to a given power spectrum. This function should produce outputs similar to healpy's synfast.
+
+    :param int nside: HEALpix resolution parameter
+    :param array-like cls: array containing power spectra. If pol=False, cls should be a 1D array. If pol=True it should be a 2D array with 4 (TT,EE,BB,TE) or 6 (TT,EE,BB,TE,EB,TB) power spectra.
+    :param boolean pol: Set to True if you want to generate T, Q and U
[…]

### 2018-07-29 pymaster/workspaces.py

@@ -1,0 +1,358 @@
+from pymaster import nmtlib as lib
+import numpy as np
+
+class NmtWorkspace(object) :
+    """
+    NmtWorkspace objects are used to compute and store the coupling matrix associated with an incomplete sky coverage, and used in the MASTER algorithm. When initialized, this object is practically empty. The information describing the coupling matrix must be computed or read from a file afterwards.
+    """
+    def __init__(self) :
+        self.wsp=None
+
+    def __del__(self) :
+        if(self.wsp is not None) :
+            lib.workspace_free(self.wsp)
+
+    def read_from(self,fname) :
+        """
+        Reads the contents of an NmtWorkspace object from a file (encoded using an internal binary format).
+
+        :param str fname: input file name
+        """
+        if self.wsp is not None :
+            lib.workspace_free(self.wsp)
+        self.wsp=lib.workspace_read(fname);
+        
+    def compute_coupling_matrix(self,fl1,fl2,bins) :
+        """
+        Computes coupling matrix associated with the cross-power spectrum of two NmtFields and an NmtBin binning scheme.
+
+        :param NmtField fl1,fl2: fields to correlate
+        :param NmtBin bin: binning scheme
+        """
+        if self.wsp is not None :
+            lib.workspace_free(self.wsp)
+        self.wsp=lib.compute_coupling_matrix(fl1.fl,fl2.fl,bins.bin)
+
+    def write_to(self,fname) :
+        """
+        Writes the contents of an NmtWorkspace object to a file (encoded using an internal binary format).
+
+        :param str fname: output file name
+        """
+        if self.wsp is None :
+            raise KeyError("Must initialize workspace before writing")
+        lib.workspace_write(self.wsp,fname)
+
+    def couple_cell(self,cl_in) :
+        """
+        Convolves a set of input power spectra with a coupling matrix (see Eq. 6 of the C API documentation).
+
+        :param cl_in: set of input power spectra. The number of power spectra must correspond to the spins of the two fields that this NmtWorkspace object was initialized with (i.e. 1 for two spin-0 fields, 2 for one spin-0 and one spin-2 field and 4 for two spin-2 fields).
+        :return: coupled power spectrum
+        """
[…]

### 2018-07-29 sandbox_validation/check_flat.py

@@ -1,0 +1,370 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import matplotlib.pyplot as plt
+import pymaster as nmt
+import os
+import sys
+import data.flatmaps as fm
+from matplotlib import rc
+import matplotlib
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+
+DTOR=np.pi/180
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--isim-ini', dest='isim_ini', default=1, type=int,
+                  help='Index of first simulation')
+parser.add_option('--isim-end', dest='isim_end', default=100, type=int,
+                  help='Index of last simulation')
+parser.add_option('--wo-mask', dest='wo_mask', default=False, action='store_true',
+                  help='Set if you don\'t want to use a mask')
+parser.add_option('--wo-contaminants', dest='wo_cont', default=False, action='store_true',
+                  help='Set if you don\'t want to use contaminants')
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--aposize', dest='aposize', default=0.0, type=float,
+                  help='Mask apodization (in degrees)')
+parser.add_option('--no-deproject',dest='no_deproject',default=False,action='store_true',
+                  help='Set if you will include contaminants but won\'t clean them')
+parser.add_option('--no-debias',dest='no_debias',default=False,action='store_true',
+                  help='Set if you will include contaminants, clean them but won\'t correct for the bias')
[…]

### 2018-07-29 sandbox_validation/check_flat_pure.py

@@ -1,0 +1,265 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import matplotlib.pyplot as plt
+import pymaster as nmt
+import os
+import sys
+import data.flatmaps as fm
+
+
+DTOR=np.pi/180
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--isim-ini', dest='isim_ini', default=1, type=int,
+                  help='Index of first simulation')
+parser.add_option('--isim-end', dest='isim_end', default=100, type=int,
+                  help='Index of last simulation')
+parser.add_option('--wo-contaminants', dest='wo_cont', default=False, action='store_true',
+                  help='Set if you don\'t want to use contaminants')
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--wo-pureb', dest='wo_pureb', default=False, action='store_true',
+                  help='Set if you don\'t want to purify B-modes')
+(o, args) = parser.parse_args()
+
+nsims=o.isim_end-o.isim_ini+1
+w_cont=not o.wo_cont
+w_pureb=not o.wo_pureb
+
+#Create output directory
+predir="tests_flat"
+os.system("mkdir -p "+predir)
+prefix=predir+"/run_pure0%d_cont%d"%(w_pureb,w_cont)
+
+#Read theory power spectra
+def read_cl_camb(fname) :
+    data=np.loadtxt(fname,unpack=True)
+    ll=np.arange(len(data[0])+2)+0.
+    fac=2*np.pi/(ll[2:]*(ll[2:]+1.))
+    cl_tt=np.zeros_like(ll); cl_tt[2:]=data[1,:]*fac
+    cl_ee=np.zeros_like(ll); cl_ee[2:]=data[2,:]*fac
[…]

### 2018-07-29 sandbox_validation/check_sph.py

@@ -1,0 +1,321 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+import os
+import sys
+
+DTOR=np.pi/180
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--nside', dest='nside_out', default=512, type=int,
+                  help='Resolution parameter')
+parser.add_option('--isim-ini', dest='isim_ini', default=1, type=int,
+                  help='Index of first simulation')
+parser.add_option('--isim-end', dest='isim_end', default=100, type=int,
+                  help='Index of last simulation')
+parser.add_option('--wo-mask', dest='wo_mask', default=False, action='store_true',
+                  help='Set if you don\'t want to use a mask')
+parser.add_option('--wo-contaminants', dest='wo_cont', default=False, action='store_true',
+                  help='Set if you don\'t want to use contaminants')
+parser.add_option('--wo-varnoise', dest='wo_nvar', default=False, action='store_true',
+                  help='Set if you don\'t want to use inhomogeneous noise')
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--aposize', dest='aposize', default=0.0, type=float,
+                  help='Mask apodization (in degrees)')
+(o, args) = parser.parse_args()
+
+nsims=o.isim_end-o.isim_ini+1
+w_mask=not o.wo_mask
+w_cont=not o.wo_cont
+w_nvar=not o.wo_nvar
+
+#Switch off contaminants and inhomogeneous noiseif there's no mask
+if not w_mask :
[…]

### 2018-07-29 sandbox_validation/check_sph_pure.py

@@ -1,0 +1,259 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+import os
+import sys
+
+DTOR=np.pi/180
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--nside', dest='nside_out', default=512, type=int,
+                  help='Resolution parameter')
+parser.add_option('--isim-ini', dest='isim_ini', default=1, type=int,
+                  help='Index of first simulation')
+parser.add_option('--isim-end', dest='isim_end', default=100, type=int,
+                  help='Index of last simulation')
+parser.add_option('--wo-contaminants', dest='wo_cont', default=False, action='store_true',
+                  help='Set if you don\'t want to use contaminants')
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--wo-pureb', dest='wo_pureb', default=False, action='store_true',
+                  help='Set if you don\'t want to purify B-modes')
+(o, args) = parser.parse_args()
+
+nsims=o.isim_end-o.isim_ini+1
+w_cont=not o.wo_cont
+w_pureb=not o.wo_pureb
+
+#Create output directory
+predir="tests_sph"
+os.system("mkdir -p "+predir)
+prefix=predir+"/run_pure0%d_ns%d_cont%d"%(w_pureb,o.nside_out,w_cont)
+
+#Read theory power spectra
+def read_cl_camb(fname) :
+    data=np.loadtxt(fname,unpack=True)
+    ll=np.arange(3*o.nside_out,dtype=float)
+    fac=2*np.pi/(ll[2:]*(ll[2:]+1.))
[…]

### 2018-07-29 sandbox_validation/data/flatmaps.py

@@ -1,0 +1,560 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import pymaster as nmt
+from astropy.io import fits
+from astropy.wcs import WCS
+from scipy.interpolate import interp1d
+
+class FlatMapInfo(object) :
+    def __init__(self,wcs,nx=None,ny=None,lx=None,ly=None) :
+        """
+        Creates a flat map
+        wcs : WCS object containing information about reference point and resolution
+        nx,ny : Number of pixels in the x/y axes. If None, dx/dy must be provided
+        lx,ly : Extent of themap in the x/y axes. If None, nx/ny must be provided
+        """
+        self.wcs=wcs.copy()
+
+        if nx is None and lx is None :
+            raise ValueError("Must provide either nx or lx")
+
+        if ny is None and ly is None :
+            raise ValueError("Must provide either ny or ly")
+
+        if nx is None :
+            self.lx=lx
+            self.nx=int(self.lx/np.abs(self.wcs.wcs.cdelt[0]))+1
+        else :
+            self.nx=nx
+            self.lx=np.fabs(nx*self.wcs.wcs.cdelt[0])
+        self.dx=self.lx/self.nx
+
+        if ny is None :
+            self.ly=ly
+            self.ny=int(self.ly/np.abs(self.wcs.wcs.cdelt[1]))+1
+        else :
+            self.ny=ny
+            self.ly=np.fabs(ny*self.wcs.wcs.cdelt[1])
+        self.dy=self.ly/self.ny
+
+        self.lx_rad=self.lx*np.pi/180
+        self.ly_rad=self.ly*np.pi/180
+
+        self.npix=self.nx*self.ny
+
+    def get_dims(self) :
+        """
+        Returns map size
+        """
+        return [self.ny,self.nx]
+        
+    def get_size(self) :
+        """
+        Returns map size
+        """
+        return self.npix
+
+    def pos2pix(self,ra,dec) :
+        """
+        Returns pixel indices for arrays of x and y coordinates.
+        Will return -1 if (x,y) lies outside the map
+        """
+        ra=np.asarray(ra)
+        scalar_input=False
+        if ra.ndim==0 :
+            ra=x[None]
+            scalar_input=True
+
+        dec=np.asarray(dec)
+        if dec.ndim==0 :
+            dec=dec[None]
+
+        if len(ra)!=len(dec) :
[…]

### 2018-07-29 sandbox_validation/data/get_cmb_contaminants.py

@@ -1,0 +1,72 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import os
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--nside', dest='nside_out', default=512, type=int,
+                  help='Resolution parameter')
+(o, args) = parser.parse_args()
+
+def read_cl_camb(fname) :
+    data=np.loadtxt(fname,unpack=True)
+    ll=np.arange(3*o.nside_out,dtype=float)
+    fac=2*np.pi/(ll[2:]*(ll[2:]+1.))
+    cl_tt=np.zeros_like(ll); cl_tt[2:]=data[1,:3*o.nside_out-2]*fac
+    cl_ee=np.zeros_like(ll); cl_ee[2:]=data[2,:3*o.nside_out-2]*fac
+    cl_bb=np.zeros_like(ll); cl_bb[2:]=data[3,:3*o.nside_out-2]*fac
+    cl_te=np.zeros_like(ll); cl_te[2:]=data[4,:3*o.nside_out-2]*fac
+
+    return ll,cl_tt,cl_ee,cl_bb,cl_te
+
+l,cltt,clee,clbb,clte=read_cl_camb("cls_cmb.txt")
+msk=hp.ud_grade(hp.read_map("mask_cmb_ns%d.fits"%o.nside_out,verbose=False),nside_out=o.nside_out)
+mz=np.zeros_like(msk)
+fsky=np.mean(msk)
+larr=np.arange(3*o.nside_out)
+
+if not os.path.isfile('cont_cmb_ns%d.fits'%o.nside_out) :
+    if o.nside_out>256 :
+        #Generate random homogeneous realization with the right power spectrum.
+        ccee=2E-5*(100./(larr+0.1))**2.5; ccee[:10]=ccee[10]
+        ccbb=9E-6*(100./(larr+0.1))**2.3; ccbb[:10]=ccbb[10]
+        ratio=np.sqrt(0.1*np.mean(clbb[40:60])/np.mean(ccbb[40:60]))
+        
+        czero=np.zeros_like(ccee)
[…]

### 2018-07-29 sandbox_validation/data/get_cmb_contaminants_flat.py

@@ -1,0 +1,63 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import flatmaps as fm
+import matplotlib.pyplot as plt
+import os
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+(o, args) = parser.parse_args()
+
+def read_cl_camb(fname) :
+    data=np.loadtxt(fname,unpack=True)
+    ll=np.arange(len(data[0])+2)+0.
+    fac=2*np.pi/(ll[2:]*(ll[2:]+1.))
+    cl_tt=np.zeros_like(ll); cl_tt[2:]=data[1,:]*fac
+    cl_ee=np.zeros_like(ll); cl_ee[2:]=data[2,:]*fac
+    cl_bb=np.zeros_like(ll); cl_bb[2:]=data[3,:]*fac
+    cl_te=np.zeros_like(ll); cl_te[2:]=data[4,:]*fac
+
+    return ll,cl_tt,cl_ee,cl_bb,cl_te
+
+l,cltt,clee,clbb,clte=read_cl_camb("cls_cmb.txt")
+fmi,msk=fm.read_flat_map("mask_cmb_flat.fits")
+
+#Generate random homogeneous realization with the right power spectrum.
+ccee=2E-5*(100./(l+0.1))**2.5; ccee[:10]=ccee[10]
+ccbb=9E-6*(100./(l+0.1))**2.3; ccbb[:10]=ccbb[10]
+czero=np.zeros_like(ccee)
+ratio=np.sqrt(0.2*np.mean(clbb[500:600])/np.mean(ccbb[500:600]))
+t,q,u=fmi.synfast(l,np.array([czero,ccee,ccbb,czero]))
+q*=ratio; u*=ratio;
+
+fmi.write_flat_map("cont_cmb_flat.fits",np.array([q,u]))
+
+if o.plot_stuff :
+    fdum,[cq,cu]=fm.read_flat_map("cont_cmb_flat.fits",i_map=-1)
+    dt,dq,du=fmi.synfast(l,np.array([cltt,clee,clbb,clte]))
+
[…]

### 2018-07-29 sandbox_validation/data/get_cmb_mask.py

@@ -1,0 +1,44 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import matplotlib.pyplot as plt
+import os
+import healpy as hp
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--nside', dest='nside_out', default=256, type=int,
+                  help='Resolution parameter')
+(o, args) = parser.parse_args()
+
+DTOR=np.pi/180
+
+def getmaskapoana(ns,aps,fsk=0.1,dec0=-50,ra0=0.) :
+    #This generates a correctly-apodized mask
+    v0=np.array([np.sin(DTOR*(90-dec0))*np.cos(DTOR*ra0),
+                 np.sin(DTOR*(90-dec0))*np.sin(DTOR*ra0),
+                 np.cos(DTOR*(90-dec0))])
+    vv=np.array(hp.pix2vec(ns,np.arange(hp.nside2npix(ns))))
+    cth=np.sum(v0[:,None]*vv,axis=0); th=np.arccos(cth); th0=np.arccos(1-2*fsk); th_apo=aps*DTOR
+    id0=np.where(th>=th0)[0]
+    id1=np.where(th<=th0-th_apo)[0]
+    idb=np.where((th>th0-th_apo) & (th<th0))[0]
+    x=np.sqrt((1-np.cos(th[idb]-th0))/(1-np.cos(th_apo)))
+    mask_apo=np.zeros(hp.nside2npix(ns))
+    mask_apo[id0]=0.
+    mask_apo[id1]=1.
+    mask_apo[idb]=x-np.sin(2*np.pi*x)/(2*np.pi)
+    return mask_apo
+
+if not os.path.isfile('mask_cmb_ns%d.fits'%o.nside_out) :
+    msk=getmaskapoana(o.nside_out,10.,0.1,ra0=250.)
+    hp.write_map("mask_cmb_ns%d.fits"%o.nside_out,msk,overwrite=True)
+
+if o.plot_stuff :
+    msk=hp.read_map("mask_cmb_ns%d.fits"%o.nside_out,verbose=False)
+    hp.mollview(msk);
+    plt.show()
+

### 2018-07-29 sandbox_validation/data/get_cmb_mask_flat.py

@@ -1,0 +1,65 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import matplotlib.pyplot as plt
+import os
+import healpy as hp
+from scipy.special import jv
+import pymaster as nmt
+from astropy.wcs import WCS
+import flatmaps as fm
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+(o, args) = parser.parse_args()
+
+DTOR=np.pi/180
+
+res_pix=2.*DTOR/60 #Pixels of about 2 arcmin
+taper_width=10. #20-deg edges
+area_rad=500.*DTOR*DTOR #Total area of ~500 sq-deg
+d_extra=1.2*taper_width*DTOR
+Ly=np.sqrt(area_rad)/1.25+d_extra
+Lx=np.sqrt(area_rad)*1.25+d_extra
+Nx=int(Lx/res_pix)
+Ny=int(Ly/res_pix)
+ioff=int(d_extra*0.5/res_pix)
+
+def plot_map(mp) :
+    plt.figure(); plt.imshow(mp,origin='lower',interpolation='nearest')
+
+def get_ellmod() :
+    ell1dx=np.fft.fftfreq(Nx)*2*np.pi*Nx/Lx
+    ell1dy=np.fft.fftfreq(Ny)*2*np.pi*Ny/Ly
+    ell2dx,ell2dy=np.meshgrid(ell1dx,ell1dy)
+    return np.sqrt(ell2dy**2+ell2dx**2)
+
+def smooth_map(mp,smooth_scale) :
+    ellmod=get_ellmod()
+    beam_taper=2*jv(1,ellmod*smooth_scale*DTOR)/(ellmod*smooth_scale*DTOR); beam_taper[0,0]=1;
+    mp=np.real(np.fft.ifft2(np.fft.fft2(mp+1j*0)*beam_taper)).flatten();
+    mp[mp<1E-3]=0;
+    mp=mp.reshape([Ny,Nx])
+    return mp
+
+#Make mask
+mask=np.ones([Ny,Nx])
[…]

### 2018-07-29 sandbox_validation/data/get_lss_cls.py

@@ -1,0 +1,48 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import matplotlib.pyplot as plt
+import pyccl as ccl
+import os
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+plot_stuff=True
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+(o, args) = parser.parse_args()
+
+if not os.path.isfile("cls_lss.txt") :
+    z,pz=np.loadtxt("nz.txt",unpack=True)
+    ndens=np.sum(pz)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
+
+    bz=np.ones_like(z)
+    cosmo = ccl.Cosmology(Omega_c=0.27, Omega_b=0.045, h=0.67, A_s=2.1e-9, n_s=0.96)
+    clust=ccl.ClTracerNumberCounts(cosmo,False,False,z=z,n=pz,bias=bz)
+    lens=ccl.ClTracerLensing(cosmo,False,z=z,n=pz)
+    ell=np.arange(30000)
+    cltt=ccl.angular_cl(cosmo,clust,clust,ell)
+    clte=ccl.angular_cl(cosmo,clust,lens,ell)
+    clee=ccl.angular_cl(cosmo,lens,lens,ell)
+    clbb=np.zeros_like(ell)
+    nltt=np.ones_like(ell)/ndens
+    nlte=np.zeros_like(ell)
+    nlee=0.28**2*np.ones_like(ell)/ndens
+    nlbb=nlee.copy()
+    np.savetxt("cls_lss.txt",np.transpose([ell,cltt,clee,clbb,clte,nltt,nlee,nlbb,nlte]))
+ell,cltt,clee,clbb,clte,nltt,nlee,nlbb,nlte=np.loadtxt("cls_lss.txt",unpack=True)
+
+if o.plot_stuff :
+    plt.figure()
+    plt.plot(ell,cltt,'r-',label='$\\delta_g-\\delta_g$')
+    plt.plot(ell,clte,'g-',label='$\\delta_g-\\gamma_E$')
+    plt.plot(ell,clee,'b-',label='$\\gamma_E-\\gamma_E$')
+    plt.plot(ell,nltt,'r--',label='$N_{gg}$')
[…]

### 2018-07-29 sandbox_validation/data/get_lss_contaminants.py

@@ -1,0 +1,218 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+from scipy.special import jv
+import os
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--nside', dest='nside_out', default=512, type=int,
+                  help='Resolution parameter')
+(o, args) = parser.parse_args()
+
+l,cltt,clee,clbb,clte,nltt,nlee,nlbb,nlte=np.loadtxt("cls_lss.txt",unpack=True)
+msk=hp.ud_grade(hp.read_map("mask_lss.fits",verbose=False),nside_out=o.nside_out); fsky=np.mean(msk)
+
+larr=np.arange(3*o.nside_out)
+
+if not os.path.isfile("cont_lss_nvar_ns%d.fits"%o.nside_out) :
+    #First, variations in \bar{n} due to stars
+    starmap=hp.read_map("star_template.fits",verbose=False)
+    starmap[starmap<=0]=1E-15
+    starmap_low=hp.smoothing(hp.ud_grade(hp.ud_grade(starmap,nside_out=64),nside_out=o.nside_out),fwhm=3.5*np.pi/180,verbose=False) #Smooth out the star map to 3.5deg FWHM
+    r_starmap=starmap_low/np.amax(starmap_low[msk>0]) #Normalize by maximum value
+    shotnoise_factor=(r_starmap-4)/(2*r_starmap-4) #Compute shot noise factor. This interpolates between 1 (no stars) and 1.5 (lots of stars)
+
+    #Now, impose additional depth variations on scales of 3.5 degrees
+    cl_extravar=np.exp(-larr*(larr+1)*(3.5*np.pi/180/2.355)**2)
+    norm=0.67E-3/np.sum(cl_extravar*(2*larr+1)/4/np.pi); cl_extravar*=norm; #This yields <10% variations
[…]

### 2018-07-29 sandbox_validation/data/get_lss_contaminants_flat.py

@@ -1,0 +1,192 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import flatmaps as fm
+import matplotlib.pyplot as plt
+from scipy.special import jv
+import os
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+(o, args) = parser.parse_args()
+
+l,cltt,clee,clbb,clte,nltt,nlee,nlbb,nlte=np.loadtxt("cls_lss.txt",unpack=True)
+fmi,msk=fm.read_flat_map("mask_lss_flat.fits")
+
+mpt,mpq,mpu=fmi.synfast(l,np.array([cltt,clee,clbb,clte]))
+
+#Star contaminant (basically something that affects mostly small scales)
+if not os.path.isfile("cont_lss_star_flat.fits") :
+    loff=500.; ltouch=5000; frac=0.2; plaw=1.2;
+    clstar=frac*cltt[ltouch]*((ltouch+loff)/(l+loff))**plaw; clstar[:2]=0;
+    mp_star=fmi.synfast(l,clstar)
+
+    if o.plot_stuff :
+        ls,cl_c=fmi.anafast(mp_star)
+        ls,cl_d=fmi.anafast(mpt)
+        ls,cl_t=fmi.anafast(mp_star+mpt)
+        plt.figure()
+        plt.plot(ls,fmi.clbin(l,cltt)[1],'k-')
+        plt.plot(ls,fmi.clbin(l,clstar)[1],'k--')
+        plt.plot(ls,cl_c,label='Star contaminant')
+        plt.plot(ls,cl_d,label='True $\\delta_g$')
+        plt.plot(ls,cl_t,label='True + contaminant')
+        plt.xlabel('$\\ell$',fontsize=18);
+        plt.ylabel('$C_\\ell$',fontsize=18);
+        plt.legend(); plt.yscale('log'); plt.xlim([17,20000])
+        
+        fmi.view_map(mpt*msk,title='True $\\delta_g$');
[…]

### 2018-07-29 sandbox_validation/data/get_lss_mask.py

@@ -1,0 +1,42 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import matplotlib.pyplot as plt
+import os
+import healpy as hp
+
+DTOR=np.pi/180
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--nside', dest='nside_out', default=512, type=int,
+                  help='Resolution parameter')
+parser.add_option('--nholes', dest='nholes', default=100, type=int,
+                  help='Number of holes to cut out')
+parser.add_option('--rholes', dest='rholes', default=1., type=float,
+                  help='Hole radius in degrees')
+(o, args) = parser.parse_args()
+
+if not os.path.isfile("mask_lss_ns%d.fits"%o.nside_out) :
+    mask0=hp.ud_grade(hp.read_map("mask_lss.fits",verbose=False),nside_out=o.nside_out)
+    dec0=0; decf=-57; cth0=np.cos(DTOR*(90-decf)); cthf=np.cos(DTOR*(90-dec0));
+    th_r_c=np.arccos(cth0+(cthf-cth0)*np.random.rand(o.nholes))
+    phi_r_c=2*np.pi*np.random.rand(o.nholes)
+    r=hp.Rotator(coord=['C','G'])
+    th_r_g,phi_r_g=r(th_r_c,phi_r_c)
+    vs=np.transpose(np.array([np.sin(th_r_g)*np.cos(phi_r_g),
+                              np.sin(th_r_g)*np.sin(phi_r_g),
+                              np.cos(th_r_g)]))
+    mskb=np.ones_like(mask0);
+    for v in vs :
+        mskb[hp.query_disc(o.nside_out,v,o.rholes*np.pi/180)]=0
+
+    hp.write_map("mask_lss_ns%d.fits"%o.nside_out,mskb*mask0,overwrite=True)
+
+if o.plot_stuff :
+    mask=hp.read_map("mask_lss_ns%d.fits"%o.nside_out,verbose=False)
+    hp.mollview(mask,coord=['G','C']);
+    plt.show();

### 2018-07-29 sandbox_validation/data/get_validation_data.sh

@@ -1,0 +1,49 @@
+#!/bin/bash
+
+nside_lss=1024
+nside_cmb=256
+
+if [ ! -f lambda_sfd_ebv.fits ] ; then
+    echo "Downloading reddening map"
+    wget https://lambda.gsfc.nasa.gov/data/foregrounds/SFD/lambda_sfd_ebv.fits
+fi
+
+if [ ! -f mask_lss.fits ] ; then
+    echo "Missing mask!"
+    exit
+fi
+
+if [ ! -f star_template.fits ] ; then
+    echo "Missing star template!"
+    exit
+fi
+
+if [ ! -f nz.txt ] ; then
+    echo "Missing N(z)"
+    exit
+fi
+
+echo "Generating LSS power spectra"
+python get_lss_cls.py --plot
+
+echo "Generating LSS mask"
+python get_lss_mask.py --plot --nside ${nside_lss} --nholes 100 --rholes 1.
+
+echo "Generating LSS contaminant templates"
+python get_lss_contaminants.py --plot --nside ${nside_lss}
+
+echo "Generating CMB mask"
+python get_cmb_mask.py --plot --nside ${nside_cmb}
+
+echo "Generating CMB contaminant templates"
+python get_cmb_contaminants.py --plot --nside ${nside_cmb}
+
+echo "Generating flat-sky LSS contaminant templates"
+python get_lss_contaminants_flat.py --plot
+
+echo "Generating flat-sky CMB mask"
+python get_cmb_mask_flat.py --plot
+
+echo "Generating flat-sky CMB contaminant templates"
+python get_cmb_contaminants_flat.py --plot
+exit

### 2018-07-29 sandbox_validation/data/plot_mock_data.py

@@ -1,0 +1,262 @@
+from __future__ import print_function
+from optparse import OptionParser
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import flatmaps as fm
+import os
+from matplotlib import rc
+import matplotlib
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+def opt_callback(option, opt, value, parser):
+    setattr(parser.values, option.dest, value.split(','))
+parser = OptionParser()
+parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true',
+                  help='Set if you want to produce plots')
+parser.add_option('--figure',dest='whichfig',default=-1, type=int,
+                  help='Figure number. Pass -1 for all figures')
+(o, args) = parser.parse_args()
+
+nside_lss=1024
+nside_cmb=256
+
+def tickfs(ax,x=True,y=True) :
+    if x :
+        for tick in ax.xaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+    if y :
+        for tick in ax.yaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+
+if (o.whichfig==3) or (o.whichfig==-1) :
+    #Plotting flat-sky LSS maps
+    l,cltt,clee,clbb,clte,nltt,nlee,nlbb,nlte=np.loadtxt("cls_lss.txt",unpack=True)
+    fmi,msk_f=fm.read_flat_map("mask_lss_flat.fits")
+    mpt_f,mpq_f,mpu_f=fmi.synfast(l,np.array([cltt,clee,clbb,clte]))
+    fdum,[cwp_q_f,cwp_u_f]=fm.read_flat_map("cont_wl_psf_flat.fits",i_map=-1)
+    fdum,[cws_q_f,cws_u_f]=fm.read_flat_map("cont_wl_ss_flat.fits",i_map=-1)
+    fdum,cld_f=fm.read_flat_map("cont_lss_dust_flat.fits")
[…]

### 2018-07-29 sandbox_validation/plot_flat.py

@@ -1,0 +1,224 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.stats as st
+from matplotlib import rc
+import matplotlib
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+nsims=1000
+prefix_clean="tests_flat/run_mask1_cont1_apo0.00"
+prefix_dirty="tests_flat/run_mask1_cont1_apo0.00_no_deproj"
+
+def tickfs(ax,x=True,y=True) :
+    if x :
+        for tick in ax.xaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+    if y :
+        for tick in ax.yaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+
+def read_cls(fname) :
+    l,ctt,cte,ctb,cee,ceb,cbe,cbb=np.loadtxt(fname,unpack=True);
+    return l,ctt,cte,ctb,cee,ceb,cbe,cbb
+l_th,clTT_th,clTE_th,clTB_th,clEE_th,clEB_th,clBE_th,clBB_th=read_cls(prefix_clean+"_cl_th.txt")
+ndof=len(l_th)
+
+print "Reading"
+clTT_clean=[]; clTE_clean=[]; clTB_clean=[]; clEE_clean=[]; clEB_clean=[]; clBB_clean=[];
+clTT_dirty=[]; clTE_dirty=[]; clTB_dirty=[]; clEE_dirty=[]; clEB_dirty=[]; clBB_dirty=[];
+for i in np.arange(nsims) :
+    ll,cctt,ccte,cctb,ccee,cceb,ccbe,ccbb=read_cls(prefix_clean+"_cl_%04d.txt"%(i+1))
+    clTT_clean.append(cctt); clTE_clean.append(ccte); clTB_clean.append(cctb);
+    clEE_clean.append(ccee); clEB_clean.append(cceb); clBB_clean.append(ccbb);
+    ll,cctt,ccte,cctb,ccee,cceb,ccbe,ccbb=read_cls(prefix_dirty+"_cl_%04d.txt"%(i+1))
[…]

### 2018-07-29 sandbox_validation/plot_flat_pure.py

@@ -1,0 +1,147 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.stats as st
+from matplotlib import rc
+import matplotlib
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+nsims=1000
+prefix_clean="tests_flat/run_pure01_cont0"
+
+def tickfs(ax,x=True,y=True) :
+    if x :
+        for tick in ax.xaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+    if y :
+        for tick in ax.yaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+
+def read_cls(fname) :
+    l,cee,ceb,cbe,cbb=np.loadtxt(fname,unpack=True);
+    return l,cee,ceb,cbe,cbb
+l_th,clEE_th,clEB_th,clBE_th,clBB_th=read_cls(prefix_clean+"_cl_th.txt")
+ndof=len(l_th)
+
+print "Reading"
+clEE_clean=[]; clEB_clean=[]; clBB_clean=[];
+for i in np.arange(nsims) :
+    ll,ccee,cceb,ccbe,ccbb=read_cls(prefix_clean+"_cl_%04d.txt"%(i+1))
+    clEE_clean.append(ccee); clEB_clean.append(cceb); clBB_clean.append(ccbb);
+clEE_clean=np.array(clEE_clean); clEB_clean=np.array(clEB_clean); clBB_clean=np.array(clBB_clean); 
+
+print "Computing statistics"
+hartfac=(nsims-ndof-2.)/(nsims-1.)
+def compute_stats(y,y_th) :
+    mean=np.mean(y,axis=0)
+    cov=np.mean(y[:,:,None]*y[:,None,:],axis=0)-mean[:,None]*mean[None,:]
+    icov=hartfac*np.linalg.inv(cov)
+    chi2_red=np.sum((mean-y_th)**2/np.diag(cov))*nsims
+    chi2_all=np.sum(((y-y_th)**2)/np.diag(cov)[None,:],axis=1)
+
+    return mean,cov,icov,chi2_red,chi2_all
+
[…]

### 2018-07-29 sandbox_validation/plot_sph_pure.py

@@ -1,0 +1,240 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.stats as st
+from matplotlib import rc
+import matplotlib
+import pymaster as nmt
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+nside=256
+nsims=1000
+prefix_pure="tests_sph/run_pure01_ns%d_cont1"%nside
+prefix_nopu="tests_sph/run_pure00_ns%d_cont1"%nside
+prefix_noco="tests_sph/run_pure01_ns%d_cont0"%nside
+
+def tickfs(ax,x=True,y=True) :
+    if x :
+        for tick in ax.xaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+    if y :
+        for tick in ax.yaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+
+def read_cls(fname) :
+    l,cee,ceb,cbe,cbb=np.loadtxt(fname,unpack=True);
+    id_good=np.where(l<2*nside)[0]
+    return l[id_good],cee[id_good],ceb[id_good],cbe[id_good],cbb[id_good]
+l_th,clEE_th,clEB_th,clBE_th,clBB_th=read_cls(prefix_pure+"_cl_th.txt")
+ndof=len(l_th)
+
+print "Reading"
+clEE_pure=[]; clEB_pure=[]; clBB_pure=[];
+clEE_nopu=[]; clEB_nopu=[]; clBB_nopu=[];
+clEE_noco=[]; clEB_noco=[]; clBB_noco=[];
+for i in np.arange(nsims) :
+    ll,ccee,cceb,ccbe,ccbb=read_cls(prefix_pure+"_cl_%04d.txt"%(i+1))
+    clEE_pure.append(ccee); clEB_pure.append(cceb); clBB_pure.append(ccbb);
+    ll,ccee,cceb,ccbe,ccbb=read_cls(prefix_nopu+"_cl_%04d.txt"%(i+1))
+    clEE_nopu.append(ccee); clEB_nopu.append(cceb); clBB_nopu.append(ccbb);
[…]

### 2018-07-29 sandbox_validation/run_stress_tests.sh

@@ -1,0 +1,164 @@
+#!/bin/bash
+
+#exc="addqueue -q cmb -s -n 1x12 -m 2.0 /usr/local/shared/python/2.7.6-gcc/bin/python"
+exc="python"
+nsims=1000
+which_partition="regular"
+timelim_hou=00
+timelim_min=15
+timelim_sec=00
+run_in_nodes=1
+
+mkdir -p tests_sph
+for nside in 512
+do
+    if [ ${nside} -eq 64 ]
+    then	
+	timelim_hou=00
+	timelim_min=05
+	timelim_sec=00
+    elif [ ${nside} -eq 256 ]
+    then
+	timelim_hou=00
+	timelim_min=15
+	timelim_sec=00
+    elif [ ${nside} -eq 512 ]
+    then
+	timelim_hou=00
+	timelim_min=30
+	timelim_sec=00
+    fi
+    for msk in 1
+    do
+	for cont in 0 1
+	do
+	    if [ ${cont} -eq 0 ] #Run pure B
+	    then
+		for aposize in 1. 2. 5. 10. 20.
+		do
+		    for pureB in 0 1
+		    do
+			command="${exc} check_sph_pure.py ${nside} ${cont} ${nsims} 0 ${aposize} C1 0 ${pureB}"
+			if [ ${run_in_nodes} -eq 1 ]
+			then
+			    runfile=tests_sph/run_batch_ns${nside}_mask${msk}_apo${aposize}_cont${cont}_pure${pureB}.sh
+			    cat > ${runfile} <<EOF
+#!/bin/bash -l
+#SBATCH --partition ${which_partition}
+##SBATCH --qos premium
+#SBATCH --nodes 1
+#SBATCH --time=${timelim_hou}:${timelim_min}:${timelim_sec}
+#SBATCH --job-name=sph_${nside}_${aposize}_${cont}_${pureB}
+#SBATCH --account=m1727
+#SBATCH -C haswell
+module load python/2.7-anaconda
+srun -n 1 ${command}
+EOF
+			    
+			    cat ${runfile}
+			    #sbatch ${runfile}
+			else
+			    echo ${command}
+			    #${command}
+			fi
+			echo " "
+		    done
+		done
+	    fi
+	    for aposize in 1.
+	    do
+		command="${exc} check_sph.py ${nside} ${msk} ${cont} ${nsims} 0 ${aposize}"
+		if [ ${run_in_nodes} -eq 1 ]
+		then
[…]

### 2018-07-29 setup.py

@@ -1,0 +1,37 @@
+#!/usr/bin/env python
+
+from distutils.core import *
+from distutils import sysconfig
+import os.path
+
+# Get numpy include directory (works across versions)
+import numpy
+try:
+    numpy_include = numpy.get_include()
+except AttributeError:
+    numpy_include = numpy.get_numpy_include()
+
+
+use_icc=False #Set to True if you compiled libsharp with icc
+if use_icc :
+    libs=['nmt','fftw3','fftw3_omp','sharp','fftpack','c_utils','chealpix','cfitsio','gsl','gslcblas','m','gomp','iomp5']
+    extra=['-openmp',]
+else :
+    libs=['nmt','fftw3','fftw3_omp','sharp','fftpack','c_utils','chealpix','cfitsio','gsl','gslcblas','m','gomp']
+    extra=['-O4', '-fopenmp',]
+
+
+_nmtlib = Extension("_nmtlib",
+                    ["pymaster/namaster_wrap.c"],
+                    libraries = libs,
+                    include_dirs = [numpy_include, "../src/"],
+                    extra_compile_args=extra,
+                    )
+
+setup(name = "pymaster",
+      description = "Library for pseudo-Cl computation",
+      author = "David Alonso",
+      version = "0.1",
+      packages = ['pymaster'],
+      ext_modules = [_nmtlib],
+      )

### 2018-07-29 src/flatsky_utils.c

@@ -1,0 +1,583 @@
+#include "utils.h"
+
+void *dftw_malloc(size_t n)
+{
+#ifdef _SPREC
+  void *p=fftwf_malloc(n);
+#else //_SPREC
+  void *p=fftw_malloc(n);
+#endif //_SPREC
+  if(p==NULL)
+    report_error(1,"Ran out of memory\n");
+  return p;
+}
+
+void dftw_free(void *p)
+{
+#ifdef _SPREC
+  fftwf_free(p);
+#else //_SPREC
+  fftw_free(p);
+#endif //_SPREC
+}
+
+void fs_map_product(nmt_flatsky_info *fs,flouble *mp1,flouble *mp2,flouble *mp_out)
+{
+#pragma omp parallel default(none)		\
+  shared(fs,mp1,mp2,mp_out)
+  {
+    long ip;
+
+#pragma omp for
+    for(ip=0;ip<fs->npix;ip++) {
+      mp_out[ip]=mp1[ip]*mp2[ip];
+    } //end omp for
+  } //end omp parallel
+}
+
+flouble fs_map_dot(nmt_flatsky_info *fs,flouble *mp1,flouble *mp2)
+{
+  double sum=0;
+
+#pragma omp parallel default(none)		\
+  shared(mp1,mp2,sum,fs)
+  {
+    long ip;
+    double sum_thr=0;
+    
+#pragma omp for
+    for(ip=0;ip<fs->npix;ip++) {
+      sum_thr+=mp1[ip]*mp2[ip];
+    } //end omp for
+
+#pragma omp critical
+    {
+      sum+=sum_thr;
+    } //end omp critical
+  } //end omp parallel
+
+  return (flouble)(sum*fs->pixsize);
+}  
+
+static void qu2eb(nmt_flatsky_info *fs,int spin,fcomplex **alm)
+{
+  int sig_overall=-1;
+  if(spin==0)
+    sig_overall=1;
+
+#pragma omp parallel default(none)		\
+  shared(fs,spin,alm,sig_overall)
+  {
+    int iy;
+    fcomplex sig=sig_overall*cpow(I,spin);
+    flouble dkx=2*M_PI/fs->lx;
+    flouble dky=2*M_PI/fs->ly;
+
+#pragma omp for
[…]

### 2018-07-29 src/healpix_extra.c

@@ -1,0 +1,1386 @@
+#include "utils.h"
+#include <fitsio.h>
+#include <chealpix.h>
+#include <sharp_almhelpers.h>
+#include <sharp_geomhelpers.h>
+#include <sharp.h>
+
+//HE_IO
+void he_write_healpix_map(flouble **tmap,int nfields,long nside,char *fname)
+{
+  fitsfile *fptr;
+  int ii,status=0;
+  char **ttype,**tform,**tunit;
+  float *map_dum=my_malloc(he_nside2npix(nside)*sizeof(float));
+
+  ttype=my_malloc(nfields*sizeof(char *));
+  tform=my_malloc(nfields*sizeof(char *));
+  tunit=my_malloc(nfields*sizeof(char *));
+  for(ii=0;ii<nfields;ii++) {
+    ttype[ii]=my_malloc(256);
+    tform[ii]=my_malloc(256);
+    tunit[ii]=my_malloc(256);
+    sprintf(ttype[ii],"map %d",ii+1);
+    sprintf(tform[ii],"1E");
+    sprintf(tunit[ii],"uK");
+  }
+
+  fits_create_file(&fptr,fname,&status);
+  fits_create_tbl(fptr,BINARY_TBL,0,nfields,ttype,tform,
+		  tunit,"BINTABLE",&status);
+  fits_write_key(fptr,TSTRING,"PIXTYPE","HEALPIX","HEALPIX Pixelisation",
+		 &status);
+
+  fits_write_key(fptr,TSTRING,"ORDERING","RING",
+		 "Pixel ordering scheme, either RING or NESTED",&status);
+  fits_write_key(fptr,TLONG,"NSIDE",&nside,
+		 "Resolution parameter for HEALPIX",&status);
+  fits_write_key(fptr,TSTRING,"COORDSYS","G",
+		 "Pixelisation coordinate system",&status);
+  fits_write_comment(fptr,
+		     "G = Galactic, E = ecliptic, C = celestial = equatorial",
+		     &status);
+  for(ii=0;ii<nfields;ii++) {
+    long ip;
+    for(ip=0;ip<he_nside2npix(nside);ip++)
+      map_dum[ip]=(float)(tmap[ii][ip]);
+    fits_write_col(fptr,TFLOAT,ii+1,1,1,he_nside2npix(nside),map_dum,&status);
+  }
+  fits_close_file(fptr, &status);
+
+  for(ii=0;ii<nfields;ii++) {
+    free(ttype[ii]);
[…]

### 2018-07-29 src/namaster.h

@@ -1,0 +1,289 @@
+#ifndef _NAMASTER_H_
+#define _NAMASTER_H_
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdarg.h>
+#include <string.h>
+#include <unistd.h>
+#include <math.h>
+#include <time.h>
+#include <complex.h>
+#include <omp.h>
+#include <gsl/gsl_vector.h>
+#include <gsl/gsl_matrix.h>
+#include <gsl/gsl_linalg.h>
+#include <gsl/gsl_blas.h>
+#include <gsl/gsl_eigen.h>
+#include <gsl/gsl_rng.h>
+#include <gsl/gsl_randist.h>
+#include <gsl/gsl_spline.h>
+#include <gsl/gsl_integration.h>
+#include <fftw3.h>
+
+#define NMT_MAX(a,b)  (((a)>(b)) ? (a) : (b)) // maximum
+#define NMT_MIN(a,b)  (((a)<(b)) ? (a) : (b)) // minimum
+
+#ifdef _SPREC
+typedef float flouble;
+typedef float complex fcomplex;
+#else //_SPREC
+typedef double flouble;
+typedef double complex fcomplex;
+#endif //_SPREC
+
+//Defined in bins_flat.c
+typedef struct {
+  int n_bands;
+  flouble *ell_0_list;
+  flouble *ell_f_list;
+} nmt_binning_scheme_flat;
+nmt_binning_scheme_flat *nmt_bins_flat_constant(int nlb,flouble lmax);
+nmt_binning_scheme_flat *nmt_bins_flat_create(int nell,flouble *l0,flouble *lf);
+void nmt_bins_flat_free(nmt_binning_scheme_flat *bin);
+void nmt_bin_cls_flat(nmt_binning_scheme_flat *bin,int nl,flouble *larr,flouble **cls_in,
+		      flouble **cls_out,int ncls);
+void nmt_unbin_cls_flat(nmt_binning_scheme_flat *bin,flouble **cls_in,
+			int nl,flouble *larr,flouble **cls_out,int ncls);
+void nmt_ell_eff_flat(nmt_binning_scheme_flat *bin,flouble *larr);
+int nmt_bins_flat_search_fast(nmt_binning_scheme_flat *bin,flouble l,int il);
+
+//Defined in bins.c
+typedef struct {
+  int n_bands;
+  int *nell_list;
+  int **ell_list;
+  flouble **w_list;
+  int ell_max;
+} nmt_binning_scheme;
[…]

### 2018-07-29 src/nmt_bins.c

@@ -1,0 +1,179 @@
+#include "utils.h"
+
+void nmt_bins_free(nmt_binning_scheme *bins)
+{
+  int ii;
+  free(bins->nell_list);
+  for(ii=0;ii<bins->n_bands;ii++) {
+    free(bins->ell_list[ii]);
+    free(bins->w_list[ii]);
+  }
+  free(bins->ell_list);
+  free(bins->w_list);
+  free(bins);
+}
+
+nmt_binning_scheme *nmt_bins_constant(int nlb,int lmax)
+{
+  int ii;
+  int nband_max=(lmax-1)/nlb;
+  flouble w0=1./nlb;
+
+  nmt_binning_scheme *bins=my_malloc(sizeof(nmt_binning_scheme));
+  bins->ell_max=lmax;
+  bins->n_bands=nband_max;
+  bins->nell_list=my_calloc(nband_max,sizeof(int));
+  bins->ell_list=my_malloc(nband_max*sizeof(int *));
+  bins->w_list=my_malloc(nband_max*sizeof(flouble *));
+
+  for(ii=0;ii<nband_max;ii++) {
+    int jj;
+    bins->nell_list[ii]=nlb;
+    bins->ell_list[ii]=my_malloc(nlb*sizeof(int));
+    bins->w_list[ii]=my_malloc(nlb*sizeof(flouble));
+    for(jj=0;jj<nlb;jj++) {
+      bins->ell_list[ii][jj]=2+ii*nlb+jj;
+      bins->w_list[ii][jj]=w0;
+    }
+  }
+
+  return bins;
+}
+
+nmt_binning_scheme *nmt_bins_create(int nell,int *bpws,int *ells,flouble *weights,int lmax)
+{
+  nmt_binning_scheme *bins;
+  int ii,nband_max=0;
+
+  for(ii=0;ii<nell;ii++) {
+    if(ells[ii]<=lmax) {
+      if(bpws[ii]>nband_max)
+	nband_max=bpws[ii];
+    }
+  }
+  nband_max++;
+
+  bins=my_malloc(sizeof(nmt_binning_scheme));
+  bins->ell_max=lmax;
+  bins->n_bands=nband_max;
+  bins->nell_list=my_calloc(nband_max,sizeof(int));
+  bins->ell_list=my_malloc(nband_max*sizeof(int *));
+  bins->w_list=my_malloc(nband_max*sizeof(flouble *));
+
[…]

### 2018-07-29 src/nmt_bins_flat.c

@@ -1,0 +1,107 @@
+#include "utils.h"
+
+int nmt_bins_flat_search_fast(nmt_binning_scheme_flat *bin,flouble l,int il)
+{
+  if(l<bin->ell_0_list[il]) {
+    int ilback=il-1;
+    while(ilback>=0) {
+      if(l<bin->ell_0_list[ilback])
+	ilback--;
+      else
+	return ilback;
+    }
+    return -1;
+  }
+  else if(l>=bin->ell_f_list[il]) {
+    int ilback=il+1;
+    while(ilback<bin->n_bands) {
+      if(l>bin->ell_f_list[ilback])
+	ilback++;
+      else
+	return ilback;
+    }
+    return -1;
+  }
+  else {
+    return il;
+  }
+}
+
+void nmt_bins_flat_free(nmt_binning_scheme_flat *bin)
+{
+  free(bin->ell_0_list);
+  free(bin->ell_f_list);
+  free(bin);
+}
+
+nmt_binning_scheme_flat *nmt_bins_flat_constant(int nlb,flouble lmax)
+{
+  int ii;
+  int nband_max=(int)((lmax-1)/nlb);
+
+  nmt_binning_scheme_flat *bin=my_malloc(sizeof(nmt_binning_scheme_flat));
+  bin->n_bands=nband_max;
+  bin->ell_0_list=my_calloc(nband_max,sizeof(flouble));
+  bin->ell_f_list=my_calloc(nband_max,sizeof(flouble));
+
+  for(ii=0;ii<nband_max;ii++) {
+    bin->ell_0_list[ii]=2+nlb*ii;
+    bin->ell_f_list[ii]=2+nlb*(ii+1)-1;
+  }
+
+  return bin;
+}
+
+nmt_binning_scheme_flat *nmt_bins_flat_create(int nell,flouble *l0,flouble *lf)
+{
+  nmt_binning_scheme_flat *bin=my_malloc(sizeof(nmt_binning_scheme_flat));
+  bin->n_bands=nell;
+  bin->ell_0_list=my_calloc(nell,sizeof(flouble));
+  bin->ell_f_list=my_calloc(nell,sizeof(flouble));
+  memcpy(bin->ell_0_list,l0,nell*sizeof(flouble));
+  memcpy(bin->ell_f_list,lf,nell*sizeof(flouble));
+
+  return bin;
+}
+
[…]

### 2018-07-29 src/nmt_covar.c

@@ -1,0 +1,293 @@
+#include "utils.h"
+
+static nmt_binning_scheme *nmt_bins_copy(nmt_binning_scheme *b_or)
+{
+  int ii;
+  nmt_binning_scheme *b=my_malloc(sizeof(nmt_binning_scheme));
+  b->n_bands=b_or->n_bands;
+  b->nell_list=my_malloc(b->n_bands*sizeof(int));
+  memcpy(b->nell_list,b_or->nell_list,b->n_bands*sizeof(int));
+  b->ell_list=my_malloc(b->n_bands*sizeof(int *));
+  b->w_list=my_malloc(b->n_bands*sizeof(flouble *));
+  for(ii=0;ii<b->n_bands;ii++) {
+    b->ell_list[ii]=my_malloc(b->nell_list[ii]*sizeof(int));
+    b->w_list[ii]=my_malloc(b->nell_list[ii]*sizeof(flouble));
+    memcpy(b->ell_list[ii],b_or->ell_list[ii],b->nell_list[ii]*sizeof(int));
+    memcpy(b->w_list[ii],b_or->w_list[ii],b->nell_list[ii]*sizeof(flouble));
+  }
+  
+  return b;
+}
+
+nmt_covar_workspace *nmt_covar_workspace_init(nmt_workspace *wa,nmt_workspace *wb)
+{
+  if((wa->nside!=wb->nside) || (wa->lmax!=wb->lmax))
+    report_error(1,"Can't compute covariance for fields with different resolutions\n");
+  if((wa->ncls!=1) || (wb->ncls!=1))
+    report_error(1,"Gaussian covariance only implemented for spin-0 fields\n");
+
+  nmt_covar_workspace *cw=my_malloc(sizeof(nmt_covar_workspace));
+  int ii;
+  int nside=wa->nside;
+  int npix=he_nside2npix(nside);
+  flouble *mask_a1b1=my_malloc(npix*sizeof(flouble));
+  flouble *mask_a1b2=my_malloc(npix*sizeof(flouble));
+  flouble *mask_a2b1=my_malloc(npix*sizeof(flouble));
[…]

### 2018-07-29 src/nmt_covar_flat.c

@@ -1,0 +1,366 @@
+#include "utils.h"
+
+static fcomplex *product_and_transform(nmt_flatsky_info *fs,flouble *m1,flouble *m2)
+{
+  flouble  *m12=dftw_malloc(fs->npix*sizeof(flouble));
+  fs_map_product(fs,m1,m2,m12);
+  fcomplex *cm12=dftw_malloc(fs->ny*(fs->nx/2+1)*sizeof(fcomplex));
+  fs_map2alm(fs,1,0,&m12,&cm12);
+
+  dftw_free(m12);
+  return cm12;
+}
+  
+static nmt_binning_scheme_flat *nmt_bins_copy(nmt_binning_scheme_flat *b_or)
+{
+  nmt_binning_scheme_flat *b=my_malloc(sizeof(nmt_binning_scheme_flat));
+  b->n_bands=b_or->n_bands;
+  b->ell_0_list=my_malloc(b->n_bands*sizeof(flouble));
+  memcpy(b->ell_0_list,b_or->ell_0_list,b->n_bands*sizeof(flouble));
+  b->ell_f_list=my_malloc(b->n_bands*sizeof(flouble));
+  memcpy(b->ell_f_list,b_or->ell_f_list,b->n_bands*sizeof(flouble));
+  
+  return b;
+}
+
+nmt_covar_workspace_flat *nmt_covar_workspace_flat_init(nmt_workspace_flat *wa,nmt_workspace_flat *wb)
+{
+  int ii;
+  
+  if((wa->fs->nx!=wb->fs->nx) || (wa->fs->ny!=wa->fs->ny))
+    report_error(1,"Can't compute covariance for fields with different resolutions\n");
+  nmt_flatsky_info *fs=wa->fs;
+  if((wa->ncls!=1) || (wb->ncls!=1))
+    report_error(1,"Gaussian covariance only implemented for spin-0 fields\n");
+  if(wa->bin->n_bands!=wb->bin->n_bands)
+    report_error(1,"Can't compute covariance for different binning schemes");
+
+  nmt_covar_workspace_flat *cw=my_malloc(sizeof(nmt_covar_workspace_flat));
+
+  cw->ncls_a=wa->ncls;
+  cw->ncls_b=wb->ncls;
[…]

### 2018-07-29 src/nmt_field.c

@@ -1,0 +1,429 @@
+#include "utils.h"
+
+void nmt_field_free(nmt_field *fl)
+{
+  int imap,itemp;
+  free(fl->beam);
+
+  for(imap=0;imap<fl->nmaps;imap++) {
+    free(fl->maps[imap]);
+    free(fl->alms[imap]);
+  }
+  free(fl->mask);
+  if(fl->ntemp>0) {
+    for(itemp=0;itemp<fl->ntemp;itemp++) {
+      for(imap=0;imap<fl->nmaps;imap++) {
+	free(fl->temp[itemp][imap]);
+	free(fl->a_temp[itemp][imap]);
+      }
+    }
+  }
+
+  free(fl->alms);
+  free(fl->maps);
+  if(fl->ntemp>0) {
+    for(itemp=0;itemp<fl->ntemp;itemp++) {
+      free(fl->temp[itemp]);
+      free(fl->a_temp[itemp]);
+    }
+    free(fl->temp);
+    free(fl->a_temp);
+    gsl_matrix_free(fl->matrix_M);
+  }
+  if(fl->a_mask!=NULL) {
+    for(imap=0;imap<fl->nmaps;imap++)
+      free(fl->a_mask[imap]);
+    free(fl->a_mask);
+  }
+  free(fl);
+}
+
+void nmt_purify(nmt_field *fl,flouble *mask,fcomplex **walm0,
+		flouble **maps_in,flouble **maps_out,fcomplex **alms)
+{
+  long ip;
+  int imap,mm,ll;
+  int purify[2]={0,0};
+  flouble *f_l=my_malloc((fl->lmax+1)*sizeof(flouble));
+  flouble  **pmap=my_malloc(fl->nmaps*sizeof(flouble *));
+  flouble  **wmap=my_malloc(fl->nmaps*sizeof(flouble *));
+  fcomplex **walm=my_malloc(fl->nmaps*sizeof(fcomplex *));
+  fcomplex **palm=my_malloc(fl->nmaps*sizeof(fcomplex *));
+  fcomplex **alm_out=my_malloc(fl->nmaps*sizeof(fcomplex *));
+  for(imap=0;imap<fl->nmaps;imap++) {
+    walm[imap]=my_calloc(he_nalms(fl->lmax),sizeof(fcomplex));
+    palm[imap]=my_calloc(he_nalms(fl->lmax),sizeof(fcomplex));
+    pmap[imap]=my_calloc(fl->npix,sizeof(flouble));
[…]

### 2018-07-29 src/nmt_field_flat.c

@@ -1,0 +1,503 @@
+#include "utils.h"
+
+nmt_k_function *nmt_k_function_alloc(int nk,flouble *karr,flouble *farr,flouble y0,flouble yf,int is_const)
+{
+  nmt_k_function *f=my_malloc(sizeof(nmt_k_function));
+  f->is_const=is_const;
+  f->y0=y0;
+  if(!is_const) {
+    f->x0=karr[0];
+    f->xf=karr[nk-1];
+    f->yf=yf;
+    //    f->spl=gsl_spline_alloc(gsl_interp_cspline,nk);
+    f->spl=gsl_spline_alloc(gsl_interp_linear,nk);
+    gsl_spline_init(f->spl,karr,farr,nk);
+  }
+  return f;
+}
+
+void nmt_k_function_free(nmt_k_function *f)
+{
+  if(!(f->is_const))
+    gsl_spline_free(f->spl);
+  free(f);
+}
+
+flouble nmt_k_function_eval(nmt_k_function *f,flouble k,gsl_interp_accel *intacc)
+{
+  if((f->is_const) || (k<=f->x0))
+    return f->y0;
+  else if(k>=f->xf)
+    return f->yf;
+  else
+    return gsl_spline_eval(f->spl,k,intacc);
+}
+
+#define N_DELL 1
+nmt_flatsky_info *nmt_flatsky_info_alloc(int nx,int ny,flouble lx,flouble ly)
+{
+  nmt_flatsky_info *fs=my_malloc(sizeof(nmt_flatsky_info));
+  fs->nx=nx;
+  fs->ny=ny;
+  fs->npix=nx*ny;
+  fs->lx=lx;
+  fs->ly=ly;
+  fs->pixsize=lx*ly/(nx*ny);
+
+  int ii;
+  flouble dkx=2*M_PI/lx;
+  flouble dky=2*M_PI/ly;
+  flouble kmax_x=dkx*(nx/2);
+  flouble kmax_y=dky*(ny/2);
+  double dk=NMT_MIN(dkx,dky);
+  double kmax=sqrt(kmax_y*kmax_y+kmax_x*kmax_x);
+  fs->dell=N_DELL*dk;
+  fs->i_dell=1./fs->dell;
+  fs->n_ell=0;
+  while((fs->n_ell+1)*fs->dell<=kmax)
+    fs->n_ell++;
+  fs->ell_min=my_malloc(fs->n_ell*sizeof(flouble));
[…]

### 2018-07-29 src/nmt_main.c

@@ -1,0 +1,277 @@
+#include "utils.h"
+
+void run_master(nmt_field *fl1,nmt_field *fl2,
+		char *fname_cl_noise,
+		char *fname_cl_proposal,
+		char *fname_coupling,
+		char *fname_out,
+		char *fname_bins,
+		int n_lbin)
+{
+  FILE *fi;
+  int ii;
+  int lmax=fl1->lmax;
+  int nspec=fl1->nmaps*fl2->nmaps;
+  flouble **cl_noise,**cl_proposal,**cl_out,**cl_bias,**cl_data;
+
+  if(fl1->nside!=fl2->nside)
+    report_error(1,"Can't correlate fields with different resolution\n");
+
+  //Binning
+  nmt_binning_scheme *bin;
+  if(!strcmp(fname_bins,"none"))
+    bin=nmt_bins_constant(n_lbin,fl1->lmax);
+  else
+    bin=nmt_bins_read(fname_bins,fl1->lmax);
+
+  //Allocate cl
+  cl_noise=my_malloc(nspec*sizeof(flouble *));
+  cl_proposal=my_malloc(nspec*sizeof(flouble *));
+  cl_bias=my_malloc(nspec*sizeof(flouble *));
+  cl_data=my_malloc(nspec*sizeof(flouble *));
+  cl_out=my_malloc(nspec*sizeof(flouble *));
+  for(ii=0;ii<nspec;ii++) {
+    cl_noise[ii]=my_calloc((lmax+1),sizeof(flouble));
+    cl_proposal[ii]=my_calloc((lmax+1),sizeof(flouble));
+    cl_bias[ii]=my_calloc((lmax+1),sizeof(flouble));
+    cl_data[ii]=my_calloc((lmax+1),sizeof(flouble));
+    cl_out[ii]=my_calloc(bin->n_bands,sizeof(flouble));
+  }
+
+  printf("Reading noise pseudo-cl\n");
+  if(strcmp(fname_cl_noise,"none")) {
+    fi=my_fopen(fname_cl_noise,"r");
+    int nlin=my_linecount(fi); rewind(fi);
+    if(nlin!=lmax+1)
+      report_error(1,"Wrong number of multipoles for noise p.spec.\n");
+    for(ii=0;ii<lmax+1;ii++) {
+      int status,jj;
+      flouble l;
+      status=fscanf(fi,"%lf",&l);
+      if(status!=1)
[…]

### 2018-07-29 src/nmt_mask.c

@@ -1,0 +1,142 @@
+#include "utils.h"
+
+static void apodize_mask_CX(long nside,flouble *mask_in,flouble *mask_out,flouble aposize,char *apotype)
+{
+  long npix=he_nside2npix(nside);
+  double aporad=aposize*M_PI/180;
+  double x2_thr=1-cos(aporad);
+  double inv_x2_thr=1./x2_thr;
+  flouble *vec=my_malloc(3*npix*sizeof(flouble));
+  flouble *cthv=my_malloc(npix*sizeof(flouble));
+  flouble *phiv=my_malloc(npix*sizeof(flouble));
+  int apotyp=0;
+  if(!strcmp(apotype,"C1"))
+    apotyp=0;
+  else if(!strcmp(apotype,"C2"))
+    apotyp=1;
+  else
+    report_error(1,"Unknown apodization type %s\n",apotype);
+
+  if(mask_out!=mask_in)
+    memcpy(mask_out,mask_in,npix*sizeof(flouble));
+
+  //Get coords for each pixel
+#pragma omp parallel default(none)		\
+  shared(vec,npix,nside,cthv,phiv)
+  {
+    long ip;
+#pragma omp for
+    for(ip=0;ip<npix;ip++) {
+      flouble *v=vec+3*ip;
+      he_pix2vec_ring(nside,ip,v);
+      cthv[ip]=v[2];
+      phiv[ip]=atan2(v[1],v[0]);
+      if(phiv[ip]<0)
+	phiv[ip]+=2*M_PI;
+    } //end omp for
+  }//end omp parallel
+
+#pragma omp parallel default(none)			\
+  shared(vec,npix,x2_thr,inv_x2_thr,mask_in,mask_out)	\
+  shared(nside,cthv,phiv,aporad,apotyp)
+  {
+    long ip;
+    int lenlist0=(int)(4*npix*(1-cos(1.2*aporad)));
+    int *listpix=my_malloc(lenlist0*sizeof(int));
+
+#pragma omp for schedule(dynamic)
+    for(ip=0;ip<npix;ip++) {
+      if(mask_in[ip]>0) {
+	int j;
+	int lenlist_half=lenlist0/2;
+	flouble *v0=vec+3*ip;
+	flouble x2dist=1000;
[…]

### 2018-07-29 src/nmt_mask_flat.c

@@ -1,0 +1,160 @@
+#include "utils.h"
+
+static void apodize_mask_CX(nmt_flatsky_info *fs,flouble *mask_in,flouble *mask_out,
+			    flouble aposize,char *apotype)
+{
+  double aporad=aposize*M_PI/180;
+  int apotyp=0;
+  if(!strcmp(apotype,"C1"))
+    apotyp=0;
+  else if(!strcmp(apotype,"C2"))
+    apotyp=1;
+  else
+    report_error(1,"Unknown apodization type %s\n",apotype);
+
+  if(mask_out!=mask_in)
+    memcpy(mask_out,mask_in,fs->nx*fs->ny*sizeof(flouble));
+
+#pragma omp parallel default(none)			\
+  shared(fs,mask_in,mask_out,aporad,apotyp)
+  {
+    int iy;
+    flouble **rarr;
+    double x_thr=aporad;
+    double inv_x_thr=1./x_thr;
+    flouble dx=fs->lx/fs->nx;
+    flouble dy=fs->lx/fs->nx;
+    int nx_patch=(int)(1.2*aporad/dx);
+    int ny_patch=(int)(1.2*aporad/dy);
+    rarr=my_malloc((2*ny_patch+1)*sizeof(flouble *));
+    for(iy=0;iy<=2*ny_patch;iy++) {
+      int ix;
+      flouble y=(iy-ny_patch)*dy;
+      rarr[iy]=my_malloc((2*nx_patch+1)*sizeof(flouble));
+      for(ix=0;ix<=2*nx_patch;ix++) {
+	flouble x=(ix-nx_patch)*dx;
+	rarr[iy][ix]=sqrt(x*x+y*y);
+      }
+    }
+
+#pragma omp for schedule(dynamic)
+    for(iy=0;iy<fs->ny;iy++) {
+      int ix;
+      for(ix=0;ix<fs->nx;ix++) {
+	int index=ix+fs->nx*iy;
+	flouble rmin=100000;
+	if(mask_in[index]>0) {
+	  int iyy;
+	  for(iyy=0;iyy<=2*ny_patch;iyy++) {
+	    int ixx;
+	    if(iy+iyy-ny_patch<0) continue;
+	    if(iy+iyy-ny_patch>=fs->ny) break;
+	    for(ixx=0;ixx<=2*nx_patch;ixx++) {
+	      if(ix+ixx-nx_patch<0) continue;
[…]

### 2018-07-29 src/nmt_master.c

@@ -1,0 +1,682 @@
+#include "utils.h"
+
+static void purify_generic(nmt_field *fl,flouble *mask,fcomplex **walm0,flouble **maps_in,fcomplex **alms_out)
+{
+  if(fl->pure_b || fl->pure_e) {
+    nmt_purify(fl,mask,walm0,maps_in,maps_in,alms_out);
+  }
+  else {
+    int im1;
+    for(im1=0;im1<fl->nmaps;im1++)
+      he_map_product(fl->nside,maps_in[im1],mask,maps_in[im1]);
+    he_map2alm(fl->nside,fl->lmax,1,2*fl->pol,maps_in,alms_out,HE_NITER_DEFAULT);
+  }
+}
+
+static flouble weigh_l(int l)
+{
+#ifdef _WEIGH_L2
+  return (flouble)(l*(l+1))/(2*M_PI);
+#else //_WEIGH_L2
+  return 1.;
+#endif //_WEIGH_L2
+}
+
+static nmt_workspace *nmt_workspace_new(int nside,int ncls,nmt_binning_scheme *bin)
+{
+  int ii;
+  nmt_workspace *w=my_malloc(sizeof(nmt_workspace));
+  w->lmax=bin->ell_max;
+  w->ncls=ncls;
+
+  w->nside=nside;
+  w->mask1=my_malloc(he_nside2npix(w->nside)*sizeof(flouble));
+  w->mask2=my_malloc(he_nside2npix(w->nside)*sizeof(flouble));
+  w->pcl_masks=my_malloc((w->lmax+1)*sizeof(flouble));
+
+  w->coupling_matrix_unbinned=my_malloc(w->ncls*(w->lmax+1)*sizeof(flouble *));
+  for(ii=0;ii<w->ncls*(w->lmax+1);ii++)
+    w->coupling_matrix_unbinned[ii]=my_calloc(w->ncls*(w->lmax+1),sizeof(flouble));
+
+  w->bin=my_malloc(sizeof(nmt_binning_scheme));
+  w->bin->n_bands=bin->n_bands;
+  w->bin->nell_list=my_malloc(w->bin->n_bands*sizeof(int));
+  memcpy(w->bin->nell_list,bin->nell_list,w->bin->n_bands*sizeof(int));
+  w->bin->ell_list=my_malloc(w->bin->n_bands*sizeof(int *));
[…]

### 2018-07-29 src/nmt_master_flat.c

@@ -1,0 +1,1175 @@
+#include "utils.h"
+
+static void purify_generic_flat(nmt_field_flat *fl,flouble *mask,fcomplex **walm0,flouble **maps_in,fcomplex **alms_out)
+{
+  if(fl->pure_b || fl->pure_e) {
+    nmt_purify_flat(fl,mask,walm0,maps_in,maps_in,alms_out);
+  }
+  else {
+    int im1;
+    for(im1=0;im1<fl->nmaps;im1++)
+      fs_map_product(fl->fs,maps_in[im1],mask,maps_in[im1]);
+    fs_map2alm(fl->fs,1,2*fl->pol,maps_in,alms_out);
+  }
+}
+
+void nmt_workspace_flat_free(nmt_workspace_flat *w)
+{
+  int ii;
+  gsl_permutation_free(w->coupling_matrix_perm);
+  gsl_matrix_free(w->coupling_matrix_binned_gsl);
+  for(ii=0;ii<w->ncls*w->bin->n_bands;ii++)
+    free(w->coupling_matrix_unbinned[ii]);
+  free(w->coupling_matrix_unbinned);
+  for(ii=0;ii<w->ncls*w->bin->n_bands;ii++)
+    free(w->coupling_matrix_binned[ii]);
+  free(w->coupling_matrix_binned);
+  free(w->n_cells);
+  nmt_bins_flat_free(w->bin);
+  nmt_flatsky_info_free(w->fs);
+  free(w->mask1);
+  free(w->mask2);
+#ifdef _ENABLE_FLAT_THEORY_ACCURATE
+  free(w->maskprod);
+#endif //_ENABLE_FLAT_THEORY_ACCURATE
+  free(w);
+}
+
+static nmt_workspace_flat *nmt_workspace_flat_new(int ncls,nmt_flatsky_info *fs,
+						  nmt_binning_scheme_flat *bin,
+						  flouble lmn_x,flouble lmx_x,flouble lmn_y,flouble lmx_y)
+{
+  int ii,ib=0;
+  nmt_workspace_flat *w=my_malloc(sizeof(nmt_workspace_flat));
+  w->ncls=ncls;
+
+  w->ellcut_x[0]=lmn_x;
+  w->ellcut_x[1]=lmx_x;
+  w->ellcut_y[0]=lmn_y;
[…]

### 2018-07-29 src/utils.c

@@ -1,0 +1,357 @@
+#include "utils.h"
+
+int my_linecount(FILE *f)
+{
+  int i0=0;
+  char ch[1000];
+  while((fgets(ch,sizeof(ch),f))!=NULL) {
+    i0++;
+  }
+  return i0;
+}
+
+void report_error(int level,char *fmt,...)
+{
+  va_list args;
+  char msg[256];
+
+  va_start(args,fmt);
+  vsprintf(msg,fmt,args);
+  va_end(args);
+  
+  if(level) {
+    fprintf(stderr," Fatal error: %s",msg);
+    exit(level);
+  }
+  else
+    fprintf(stderr," Warning: %s",msg);
+}
+
+void *my_malloc(size_t size)
+{
+  void *outptr=malloc(size);
+  if(outptr==NULL) report_error(1,"Out of memory\n");
+
+  return outptr;
+}
+
+void *my_calloc(size_t nmemb,size_t size)
+{
+  void *outptr=calloc(nmemb,size);
+  if(outptr==NULL)
+    report_error(1,"Out of memory\n");
+
+  return outptr;
+}
+
+FILE *my_fopen(const char *path,const char *mode)
+{
+  FILE *fout=fopen(path,mode);
+  if(fout==NULL)
+    report_error(1,"Couldn't open file %s\n",path);
+
+  return fout;
+}
+
+size_t my_fwrite(const void *ptr, size_t size, size_t nmemb,FILE *stream)
+{
+  if(fwrite(ptr,size,nmemb,stream)!=nmemb)
+    report_error(1,"Error fwriting\n");
+
+  return nmemb;
+}
+
+size_t my_fread(void *ptr,size_t size,size_t count,FILE *stream)
+{
+  if(fread(ptr,size,count,stream)!=count)
+    report_error(1,"Error freading\n");
+
+  return count;
+}
+
+gsl_rng *init_rng(unsigned int seed)
+{
+  gsl_rng *rng=gsl_rng_alloc(gsl_rng_mt19937);
+  gsl_rng_set(rng,seed);
+
+  return rng;
+}
+
+double rng_01(gsl_rng *rng)
+{
[…]

### 2018-07-29 src/utils.h

@@ -1,0 +1,110 @@
+#ifndef _NM_UTILS_
+#define _NM_UTILS_
+#include "namaster.h"
+
+//Defined in utils.c
+gsl_rng *init_rng(unsigned int seed);
+double rng_01(gsl_rng *rng);
+int rng_poisson(double lambda,gsl_rng *rng);
+void rng_delta_gauss(double *module,double *phase,
+		     gsl_rng *rng,double sigma2);
+void rng_gauss(gsl_rng *rng,double *r1,double *r2);
+void end_rng(gsl_rng *rng);
+int my_linecount(FILE *f);
+void report_error(int level,char *fmt,...);
+void *my_malloc(size_t size);
+void *my_calloc(size_t nmemb,size_t size);
+FILE *my_fopen(const char *path,const char *mode);
+size_t my_fwrite(const void *ptr, size_t size, size_t nmemb,FILE *stream);
+size_t my_fread(void *ptr,size_t size,size_t count,FILE *stream);
+//Returns all non-zero wigner-3j symbols
+// il2 (in) : l2
+// il3 (in) : l3
+// im2 (in) : m2
+// im3 (in) : m3
+// l1min_out (out) : min value for l1
+// l1max_out (out) : max value for l1
+// thrcof (out) : array with the values of the wigner-3j
+// size (in) : size allocated for thrcof
+int drc3jj(int il2,int il3,int im2, int im3,int *l1min_out,
+	   int *l1max_out,double *thrcof,int size);
+// Moore-penrose pseudo-inverse
+void moore_penrose_pinv(gsl_matrix *M,double threshold);
+
+//Defined in flatsky_utils.c
+void *dftw_malloc(size_t n);
+void dftw_free(void *p);
+void fs_map_product(nmt_flatsky_info *fs,flouble *mp1,flouble *mp2,flouble *mp_out);
+flouble fs_map_dot(nmt_flatsky_info *fs,flouble *mp1,flouble *mp2);
+void fs_map2alm(nmt_flatsky_info *fs,int ntrans,int spin,flouble **map,fcomplex **alm);
+void fs_alm2map(nmt_flatsky_info *fs,int ntrans,int spin,flouble **map,fcomplex **alm);
+nmt_k_function *fs_generate_beam_window(double fwhm_amin);
+void fs_zero_alm(nmt_flatsky_info *fs,fcomplex *alm);
[…]

### 2018-07-29 test/check.py

@@ -1,0 +1,44 @@
+import numpy as np
+import healpy as hp
+import pymaster as nmt
+import matplotlib.pyplot as plt
+
+nside=256
+
+mp_t,mp_q,mp_u=hp.read_map("maps.fits",field=[0,1,2],verbose=False)
+mask=np.ones(hp.nside2npix(nside))
+aft_tt,aft_ee,aft_bb,aft_te,aft_eb,aft_tb=hp.anafast([mp_t,mp_q,mp_u],pol=True)
+
+b=nmt.NmtBin(nside,nlb=1)
+larr=b.get_effective_ells()
+
+f0=nmt.NmtField(mask,[mp_t])
+f2=nmt.NmtField(mask,[mp_q,mp_u])
+w00=nmt.NmtWorkspace(); w00.compute_coupling_matrix(f0,f0,b);
+w02=nmt.NmtWorkspace(); w02.compute_coupling_matrix(f0,f2,b);
+w22=nmt.NmtWorkspace(); w22.compute_coupling_matrix(f2,f2,b);
+
+def compute_master(f_a,f_b,wsp) :
+    cl_coupled=nmt.compute_coupled_cell(f_a,f_b)
+    cl_decoupled=wsp.decouple_cell(cl_coupled)
+
+    return cl_decoupled
+
+nmt_tt=compute_master(f0,f0,w00)[0]
+nmt_te,nmt_tb=compute_master(f0,f2,w02)
+nmt_ee,nmt_eb,nmt_be,nmt_bb=compute_master(f2,f2,w22)
+
+def compare_cl(aft_cl,nmt_cl,title) :
+    print(np.sqrt(np.sum((aft_cl[2:2*nside]-nmt_cl[:2*nside-2])**2)/(len(nmt_cl)-2)))
+    plt.figure()
+    plt.title(title)
+    plt.plot(larr[:2*nside-2],aft_cl[2:2*nside]-nmt_cl[:2*nside-2])
+
+compare_cl(aft_tt,nmt_tt,'TT')
+compare_cl(aft_te,nmt_te,'TE')
+compare_cl(aft_tb,nmt_tb,'TB')
+compare_cl(aft_ee,nmt_ee,'EE')
+compare_cl(aft_eb,nmt_eb,'EB')
+compare_cl(aft_eb,nmt_be,'BE')
+compare_cl(aft_bb,nmt_bb,'BB')
+plt.show()

### 2018-07-29 test/ctest.h

@@ -1,0 +1,532 @@
+/* Copyright 2011-2016 Bas van den Berg
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
+#ifndef CTEST_H
+#define CTEST_H
+
+#if defined _WIN32 || defined __CYGWIN__
+#ifndef WIN32
+#define WIN32
+#endif
+#endif
+
+#ifndef WIN32
+#define WEAK __attribute__ ((weak))
+#else
+#define WEAK
+#endif
+
+#include <inttypes.h> /* intmax_t, uintmax_t, PRI* */
+#include <stddef.h> /* size_t */
+
+typedef void (*SetupFunc)(void*);
+typedef void (*TearDownFunc)(void*);
+
+struct ctest {
+    const char* ssname;  // suite name
+    const char* ttname;  // test name
+    void (*run)();
+    int skip;
+
+    void* data;
+    SetupFunc setup;
+    TearDownFunc teardown;
+
+    unsigned int magic;
+};
+
+#define __FNAME(sname, tname) __ctest_##sname##_##tname##_run
+#define __TNAME(sname, tname) __ctest_##sname##_##tname
+
+#define __CTEST_MAGIC (0xdeadbeef)
+#ifdef __APPLE__
+#define __Test_Section __attribute__ ((used, section ("__DATA, .ctest")))
+#else
+#define __Test_Section __attribute__ ((used, section (".ctest")))
+#endif
+
+#define __CTEST_STRUCT(sname, tname, _skip, __data, __setup, __teardown) \
+    static struct ctest __TNAME(sname, tname) __Test_Section = { \
+        .ssname=#sname, \
+        .ttname=#tname, \
+        .run = __FNAME(sname, tname),		\
+        .skip = _skip, \
+        .data = __data, \
[…]

### 2018-07-29 test/nmt_checks.c

@@ -1,0 +1,278 @@
+#include <stdio.h>
+
+#define CTEST_MAIN
+#include <fitsio.h>
+#include "namaster.h"
+#include "ctest.h"
+
+#define NSIDE_TESTS 128
+
+
+double *he_read_healpix_map(char *fname,long *nside,int nfield)
+{
+  //////
+  // Reads a healpix map from file fname. The map will be
+  // read from column #nfield. It also returns the map's nside.
+  int status=0,hdutype,nfound,anynul,ncols;
+  long naxes,*naxis,npix;
+  fitsfile *fptr;
+  double *map,nulval;
+  char order_in_file[32];
+  int nested_in_file=0;
+
+  fits_open_file(&fptr,fname,READONLY,&status);
+  fits_movabs_hdu(fptr,2,&hdutype,&status);
+  fits_read_key_lng(fptr,"NAXIS",&naxes,NULL,&status);
+  naxis=malloc(naxes*sizeof(long));
+  fits_read_keys_lng(fptr,"NAXIS",1,naxes,naxis,&nfound,&status);
+  fits_read_key_lng(fptr,"NSIDE",nside,NULL,&status);
+  npix=12*(*nside)*(*nside);
+  ASSERT_EQUAL(npix%naxis[1],0);
+
+  if (fits_read_key(fptr, TSTRING, "ORDERING", order_in_file, NULL, &status))
+    printf("WARNING: Could not find %s keyword in in file %s\n","ORDERING",fname);
+  if(!strncmp(order_in_file,"NEST",4))
+    nested_in_file=1;
+
+  if(nested_in_file) {
+    printf("Map should be in ring ordering\n");
+    ASSERT_EQUAL(0,1);
+  }
+
+  map=malloc(npix*sizeof(double));
+  fits_get_num_cols(fptr,&ncols,&status);
+  if(nfield>=ncols) {
+    printf("Not enough columns in FITS file\n");
+    ASSERT_EQUAL(0,1);
+  }
+
+#ifdef _SPREC
+  fits_read_col(fptr,TFLOAT,nfield+1,1,1,npix,&nulval,map,&anynul,&status);
+#else //_SPREC
[…]

### 2018-07-29 test/sample.c

@@ -1,0 +1,49 @@
+#include <stdlib.h>
+#include <stdio.h>
+#include <namaster.h>
+#include <math.h>
+
+int main(int argc,char **argv)
+{
+  long i,nside;
+  if(argc!=4) {
+    fprintf(stderr,"Usage: ./sample <fname_map> <fname_mask> <fname_out>\n");
+    exit(1);
+  }
+
+  //Create spin-0 field
+  nmt_field *fl1=nmt_field_read(argv[2],argv[1],"none","none",0,0,0,3);
+
+  //Create a binning scheme (20 multipoles per bandpower)
+  nmt_binning_scheme *bin=nmt_bins_constant(20,fl1->lmax);
+  //Compute array of effective multipoles
+  double *ell_eff=calloc(bin->n_bands,sizeof(double));
+  nmt_ell_eff(bin,ell_eff);
+
+  //Allocate memory for power spectrum
+  double *cl_out=malloc(bin->n_bands*sizeof(double));
+
+  //Dummy array to be passed as proposal and noise power spectrum
+  //These are not needed here, because we don't want to remove noise bias
+  //and we have assumed the field is not contaminated
+  double *cl_dum=calloc((fl1->lmax+1),sizeof(double));
+
+  //Compute pseudo-Cl estimator
+  nmt_workspace *w=nmt_compute_power_spectra(fl1,fl1,bin,NULL,&cl_dum,&cl_dum,&cl_out);
+
+  //Write output
+  FILE *fo=fopen(argv[3],"w");
+  for(i=0;i<bin->n_bands;i++)
+    fprintf(fo,"%.2lE %lE\n",ell_eff[i],cl_out[i]);
+  fclose(fo);
+
+  //Free stuff up
+  nmt_workspace_free(w);
+  free(cl_dum);
+  free(cl_out);
+  free(ell_eff);
+  nmt_bins_free(bin);
+  nmt_field_free(fl1);
+
+  return 0;
+}

### 2018-07-29 test/sample_bins.py

@@ -1,0 +1,56 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+
+#This script showcases the use of the NmtBin structure to define bandpowers.
+
+#HEALPix map resolution
+nside=256
+
+#Initialize binning scheme with bandpowers of constant width
+#(4 multipoles per bin)
+bin1=nmt.NmtBin(nside,nlb=4)
+
+#Initialize binning scheme with custom-made bandpowers.
+#In this case we simply manually choose these bandpowers to also have
+#4 multipoles per bin.
+ells=np.arange(3*nside,dtype='int32') #Array of multipoles
+weights=0.25*np.ones_like(ells) #Array of weights
+bpws=-1+np.zeros_like(ells) #Array of bandpower indices
+i=0;
+while 4*(i+1)+2<3*nside :
+    bpws[4*i+2:4*(i+1)+2]=i
+    i+=1
+bin2=nmt.NmtBin(nside,bpws=bpws,ells=ells,weights=weights)
+
+#At this stage bin1 and bin2 should be identical
+print np.sum(bin1.get_effective_ells()-bin2.get_effective_ells())
+
+#Array with effective multipole per bandpower
+ell_eff=bin1.get_effective_ells()
+
+#Bandpower info:
+print "Bandpower info:"
+print " %d bandpowers"%(bin1.get_n_bands())
+print "The columns in the following table are:"
+print " [1]-band index, [2]-list of multipoles, [3]-list of weights, [4]=effective multipole"
+for i in np.arange(bin1.get_n_bands()) :
+    print i, bin1.get_ell_list(i), bin1.get_weight_list(i), ell_eff[i]
+print ""
+
+#Binning a power spectrum
+#Read the TT power spectrum
+data=np.loadtxt("cls.txt",unpack=True);
+ell_arr=data[0]; cl_tt=data[1]
+#Bin the power spectrum into bandpowers
+cl_tt_binned=bin1.bin_cell([cl_tt])
+#Unbin bandpowers
+cl_tt_binned_unbinned=bin1.unbin_cell(cl_tt_binned)
+#Plot all to see differences
+plt.plot(ell_arr,cl_tt                   ,'r-',label='Original $C_\\ell$')
+plt.plot(ell_eff,cl_tt_binned[0]         ,'g-',label='Binned $C_\\ell$')
[…]

### 2018-07-29 test/sample_covariance.py

@@ -1,0 +1,73 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+
+#This script showcases the ability of namaster to compute Gaussian
+#estimates of the covariance matrix. This is currently only
+#supported for spin-0 fields
+#A similar example for flat-sky fields can be found in test/sample_covariance_flat.py
+
+#HEALPix map resolution
+nside=256
+
+#We start by creating some synthetic masks and maps with contaminants.
+#Here we will focus on the auto-correlation of a spin-1 field.
+#a) Read and apodize mask
+mask=nmt.mask_apodization(hp.read_map("mask.fits",verbose=False),1.,apotype="Smooth")
+
+#Let's now create a fictitious theoretical power spectrum to generate
+#Gaussian realizations:
+larr=np.arange(3*nside)
+clarr=((larr+1.)/80.)**(-1.1)+1.
+
+#This routine generates a scalar Gaussian random field based on this
+#power spectrum
+def get_sample_field() :
+    mp=hp.synfast(clarr,nside,verbose=False)
+    return nmt.NmtField(mask,[mp])
+
+#We also copy this function from sample_workspaces.py. It computes
+#power spectra given a pair of fields and a workspace.
+def compute_master(f_a,f_b,wsp) :
+    cl_coupled=nmt.compute_coupled_cell(f_a,f_b)
+    cl_decoupled=wsp.decouple_cell(cl_coupled)
+
+    return cl_decoupled
+
+#Let's generate one particular sample and its power spectrum.
+print "Field"
+f0=get_sample_field()
+b=nmt.NmtBin(nside,nlb=20) #We will use 20 multipoles per bandpower.
+print "Workspace"
+w=nmt.NmtWorkspace()
+w.compute_coupling_matrix(f0,f0,b)
+cl_0=compute_master(f0,f0,w)[0]
+
+#Let's now compute the Gaussian estimate of the covariance!
+print "Covariance"
+#First we generate a NmtCovarianceWorkspace object to precompute
+#and store the necessary coupling coefficients
+cw=nmt.NmtCovarianceWorkspace()
+cw.compute_coupling_coefficients(w,w) #<- This is the time-consuming operation
+covar=nmt.gaussian_covariance(cw,clarr,clarr,clarr,clarr)
+
+#Let's now compute the sample covariance
[…]

### 2018-07-29 test/sample_covariance_flat.py

@@ -1,0 +1,106 @@
+import numpy as np
+#import healpy as hp
+import pymaster as nmt
+import matplotlib.pyplot as plt
+import os
+
+#This script describes the functionality of the flat-sky version of pymaster
+
+#We start by defining the flat-sky field
+Lx=72.*np.pi/180; Ly=48.*np.pi/180;
+Nx=602; Ny=410;
+#Nx=301; Ny=205;
+
+#Let's now create a mask:
+mask=np.ones(Nx*Ny)
+xarr=np.ones(Ny)[:,None]*np.arange(Nx)[None,:]*Lx/Nx
+yarr=np.ones(Nx)[None,:]*np.arange(Ny)[:,None]*Ly/Ny
+def dig_hole(x,y,r) :
+    rad=(np.sqrt((xarr-x)**2+(yarr-y)**2)).flatten()
+    return np.where(rad<r)[0]
+mask[dig_hole(0.3*Lx,0.6*Ly ,0.05*np.sqrt(Lx*Ly))]=0.
+mask[dig_hole(0.7*Lx,0.12*Ly,0.07*np.sqrt(Lx*Ly))]=0.
+mask[dig_hole(0.7*Lx,0.8*Ly ,0.03*np.sqrt(Lx*Ly))]=0.
+mask[np.where(xarr.flatten()<Lx/16.)]=0; mask[np.where(xarr.flatten()>15*Lx/16.)]=0;
+mask[np.where(yarr.flatten()<Ly/16.)]=0; mask[np.where(yarr.flatten()>15*Ly/16.)]=0;
+mask=mask.reshape([Ny,Nx])
+mask=nmt.mask_apodization_flat(mask,Lx,Ly,aposize=2.,apotype="C1");
+plt.figure(); plt.imshow(mask,interpolation='nearest',origin='lower'); plt.colorbar()
+
+#Binning scheme
+l0_bins=np.arange(Nx/8)*8*np.pi/Lx
+lf_bins=(np.arange(Nx/8)+1)*8*np.pi/Lx
+b=nmt.NmtBinFlat(l0_bins,lf_bins)
+ells_uncoupled=b.get_effective_ells()
+
+#Let's create a fictitious theoretical power spectrum to generate
+#Gaussian simulations:
+larr=np.arange(3000.)
+clarr=((larr+50.)/300.)**(-1.1)+0.5
+
+#This function will generate random fields
[…]

### 2018-07-29 test/sample_fields.py

@@ -1,0 +1,55 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+
+#This script showcases the use of the NmtField structure to store information
+#about the fields to be correlated an define bandpowers.
+
+#HEALPix map resolution
+nside=256
+
+#### Read input maps
+#a) Read and apodize mask
+mask=nmt.mask_apodization(hp.read_map("mask.fits",verbose=False),1.,apotype="C1")
+#b) Read maps
+mp_t,mp_q,mp_u=hp.read_map("maps.fits",field=[0,1,2],verbose=False)
+#c) Read contaminants maps
+tm_t,tm_q,tm_u=hp.read_map("temp.fits",field=[0,1,2],verbose=False)
+
+#### Create fields
+#Create spin-0 field with no contaminants
+f0_clean=nmt.NmtField(mask,[mp_t])
+#Create spin-2 field with no contaminants
+f2_clean=nmt.NmtField(mask,[mp_q,mp_u])
+#Create contaminated spin-0 field
+f0_cont=nmt.NmtField(mask,[mp_t+tm_t],templates=[[tm_t]])
+#Create contaminated spin-2 field
+f2_cont=nmt.NmtField(mask,[mp_q+tm_q,mp_u+tm_u],templates=[[tm_q,tm_u]])
+
+#Note: when passing "templates", the constructor cleans the maps by finding
+#the best-fit linear coefficient that fits the contaminant templates.
+#I.e. the maps returned by, e.g. f0_cont.get_maps(), are already cleaned.
+# - f0_clean and f2_clean now hold masked copies of the original maps.
+# - f0_cont and f2_cont now hold masked and ***template-cleaned*** copies of
+#   the original maps
+
+#We can verify this by plotting them:
+#Spin-0
+hp.mollview(mp_t                  ,title='Original map T'        ,coord=['G','C'])
+hp.mollview(mp_t+tm_t             ,title='Contaminated map T'    ,coord=['G','C'])
+hp.mollview(f0_clean.get_maps()[0],title='Masked original map T' ,coord=['G','C'])
+hp.mollview(f0_cont.get_maps()[0] ,title='Masked & cleaned map T',coord=['G','C'])
+plt.show()
+#Spin-2, Q 
+hp.mollview(mp_q                  ,title='Original map Q'        ,coord=['G','C'])
[…]

### 2018-07-29 test/sample_flat.py

@@ -1,0 +1,98 @@
+import numpy as np
+import healpy as hp
+import pymaster as nmt
+import matplotlib.pyplot as plt
+
+
+#This script describes the functionality of the flat-sky version of pymaster
+
+#Dimensions:
+#First, a flat-sky field is defined by four quantities:
+# - Lx and Ly: the size of the patch in the x and y dimensions (in radians)
+Lx=72.*np.pi/180; Ly=48.*np.pi/180;
+# - Nx and Ny: the number of pixels in the x and y dimensions
+Nx=602; Ny=410;
+
+#Gaussian simulations:
+#pymaster allows you to generate random realizations of both spherical and
+#flat fields given a power spectrum. These are returned as 2D arrays with
+#shape (Ny,Nx)
+l,cl_tt,cl_ee,cl_bb,cl_te=np.loadtxt('cls.txt',unpack=True);
+beam=np.exp(-(0.25*np.pi/180*l)**2);
+cl_tt*=beam; cl_ee*=beam; cl_bb*=beam; cl_te*=beam; 
+mpt,mpq,mpu=nmt.synfast_flat(Nx,Ny,Lx,Ly,[cl_tt,cl_ee,cl_bb,cl_te],pol=True)
+
+#You can have a look at the maps using matplotlib's imshow:
+plt.figure(); plt.imshow(mpt,interpolation='nearest',origin='lower'); plt.colorbar()
+plt.figure(); plt.imshow(mpq,interpolation='nearest',origin='lower'); plt.colorbar()
+plt.figure(); plt.imshow(mpu,interpolation='nearest',origin='lower'); plt.colorbar()
+plt.show()
+
+#Masks:
+#Let's now create a mask:
+mask=np.ones_like(mpt).flatten()
+xarr=np.ones(Ny)[:,None]*np.arange(Nx)[None,:]*Lx/Nx
+yarr=np.ones(Nx)[None,:]*np.arange(Ny)[:,None]*Ly/Ny
+#First we dig a couple of holes
+def dig_hole(x,y,r) :
+    rad=(np.sqrt((xarr-x)**2+(yarr-y)**2)).flatten()
+    return np.where(rad<r)[0]
+mask[dig_hole(0.3*Lx,0.6*Ly ,0.05*np.sqrt(Lx*Ly))]=0.
[…]

### 2018-07-29 test/sample_masks.py

@@ -1,0 +1,40 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+
+#This script showcases the apodization routine included with pymaster
+#and the three apodization modes supported.
+
+
+#Read input binary mask
+mask_raw=hp.read_map("mask.fits",verbose=False)
+
+#The following function calls create apodized versions of the raw mask
+#with an apodization scale of 2.5 degrees using three different methods
+
+#Apodization scale in degrees
+aposcale=2.5
+
+#C1 and C2: in these cases, pixels are multiplied by a factor f
+#           (with 0<=f<=1) based on their distance to the nearest fully
+#           masked pixel. The choices of f in each case are documented in
+#           Section 3.4 of the C API documentation. All pixels separated
+#           from any masked pixel by more than the apodization scale are
+#           left untouched.
+mask_C1=nmt.mask_apodization(mask_raw,aposcale,apotype="C1")
+mask_C2=nmt.mask_apodization(mask_raw,aposcale,apotype="C2")
+
+#Smooth: in this case, all pixels closer to a masked pixel than 2.5 times
+#        the apodization scale are initially set to zero. The resulting
+#        map is then smoothed with a Gaussian kernel with standard
+#        deviation given by the apodization scale. Finally, all pixels
+#        originally masked are forced back to zero.
+mask_Sm=nmt.mask_apodization(mask_raw,aposcale,apotype="Smooth")
+
+#Let's plot the results
+hp.mollview(mask_raw,title='Binary mask'       ,coord=['G','C'])
+hp.mollview(mask_C1 ,title='C1 apodization'    ,coord=['G','C'])
+hp.mollview(mask_C2 ,title='C2 apodization'    ,coord=['G','C'])
+hp.mollview(mask_Sm ,title='Smooth apodization',coord=['G','C'])
+plt.show()

### 2018-07-29 test/sample_pureB.py

@@ -1,0 +1,75 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+import os
+
+#This script describes the computation of polarized power spectra using the
+#pure-E and B approach
+
+#We'll run this many simulations
+nsim=10
+#HEALPix map resolution
+nside=256
+
+#Let us first create a square mask:
+msk=np.zeros(hp.nside2npix(nside))
+th,ph=hp.pix2ang(nside,np.arange(hp.nside2npix(nside)))
+ph[np.where(ph>np.pi)[0]]-=2*np.pi
+msk[np.where((th<2.63) & (th>1.86) & (ph>-np.pi/4) & (ph<np.pi/4))[0]]=1.
+
+#Now we apodize the mask. The pure-B formalism requires the mask to be differentiable
+#along the edges. The 'C1' and 'C2' apodization types supported by mask_apodization
+#achieve this.
+msk_apo=nmt.mask_apodization(msk,10.0,apotype='C1')
+
+#Select a binning scheme
+b=nmt.NmtBin(nside,nlb=16)
+leff=b.get_effective_ells()
+
+#Read power spectrum and provide function to generate simulated skies
+l,cltt,clee,clbb,clte=np.loadtxt('cls.txt',unpack=True);
+def get_fields() :
+    mp_t,mp_q,mp_u=hp.synfast([cltt,clee,clbb,clte],nside=nside,new=True,verbose=False)
+    #This creates a spin-2 field without purifying either E or B
+    f2_np=nmt.NmtField(msk_apo,[mp_q,mp_u])
+    #This creates a spin-2 field with both pure E and B.
+    f2_yp=nmt.NmtField(msk_apo,[mp_q,mp_u],purify_e=True,purify_b=True)
+    #Note that generally it's not a good idea to purify both, since you'll lose sensitivity on E
+    return f2_np,f2_yp
+
+#We initialize two workspaces for the non-pure and pure fields:
+f2np0,f2yp0=get_fields()
+w_np=nmt.NmtWorkspace(); w_np.compute_coupling_matrix(f2np0,f2np0,b)
+w_yp=nmt.NmtWorkspace(); w_yp.compute_coupling_matrix(f2yp0,f2yp0,b)
+
+#This wraps up the two steps needed to compute the power spectrum
[…]

### 2018-07-29 test/sample_simple.py

@@ -1,0 +1,42 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+
+#Import the NaMaster python wrapper
+import pymaster as nmt
+
+# Simple example showcasing the use of NaMaster to compute the pseudo-Cl
+# estimator of the angular cross-power spectrum of a spin-0 field and a
+# spin-2 field
+
+#HEALPix resolution parameter used here
+nside=256
+
+#Read mask and apodize it on a scale of ~1deg
+mask=nmt.mask_apodization(hp.read_map("mask.fits",verbose=False),1.,apotype="Smooth")
+hp.mollview(mask,coord=['G','C'],title='Apodized mask'); plt.show()
+
+#Read healpix maps and initialize a spin-0 and spin-2 field
+f_0=nmt.NmtField(mask,[hp.read_map("maps.fits",field=0  ,verbose=False)])
+f_2=nmt.NmtField(mask,hp.read_map("maps.fits",field=[1,2],verbose=False))
+
+#Initialize binning scheme with 4 ells per bandpower
+b=nmt.NmtBin(nside,nlb=4)
+
+#Compute MASTER estimator
+#spin-0 x spin-0
+cl_00=nmt.compute_full_master(f_0,f_0,b)
+#spin-0 x spin-2
+cl_02=nmt.compute_full_master(f_0,f_2,b)
+#spin-2 x spin-2
+cl_22=nmt.compute_full_master(f_2,f_2,b)
+
+#Plot results
+ell_arr=b.get_effective_ells()
+plt.plot(ell_arr,cl_00[0],'r-',label='TT')
+plt.plot(ell_arr,np.fabs(cl_02[0]),'g-',label='TE')
+plt.plot(ell_arr,cl_22[0],'b-',label='EE')
+plt.plot(ell_arr,cl_22[3],'y-',label='BB')
+plt.loglog(); plt.xlabel('$\\ell$',fontsize=16); plt.ylabel('$C_\\ell$',fontsize=16);
+plt.legend(loc='upper right',ncol=2,labelspacing=0.1)
+plt.show()

### 2018-07-29 test/sample_workspaces.py

@@ -1,0 +1,99 @@
+import numpy as np
+import healpy as hp
+import matplotlib.pyplot as plt
+import pymaster as nmt
+
+#This script showcases the use of NmtWorkspace objects to speed up the
+#computation of power spectra for many pairs of fields with the same masks.
+
+#HEALPix map resolution
+nside=256
+
+#We start by creating some synthetic masks and maps with contaminants.
+#Here we will focus on the cross-correlation of a spin-2 and a spin-1 field.
+#a) Read and apodize mask
+mask=nmt.mask_apodization(hp.read_map("mask.fits",verbose=False),1.,apotype="Smooth")
+#b) Read maps
+mp_t,mp_q,mp_u=hp.read_map("maps.fits",field=[0,1,2],verbose=False)
+#c) Read contaminants maps
+tm_t,tm_q,tm_u=hp.read_map("temp.fits",field=[0,1,2],verbose=False)
+#d) Create contaminated fields
+#   Spin-0
+f0=nmt.NmtField(mask,[mp_t+tm_t],templates=[[tm_t]])
+#   Spin-2
+f2=nmt.NmtField(mask,[mp_q+tm_q,mp_u+tm_u],templates=[[tm_q,tm_u]])
+#e) Create binning scheme. We will use 20 multipoles per bandpower.
+b=nmt.NmtBin(nside,nlb=20)
+#f) Finally, we read our best guess for the true power spectrum. We will
+#   use this to:
+#   i)  Compute the bias to the power spectrum from contaminant cleaning
+#   ii) Generate random realizations of our fields to compute the errors
+l,cltt,clee,clbb,clte=np.loadtxt("cls.txt",unpack=True)
+cl_02_th=np.array([clte,np.zeros_like(clte)])
+
+#We then generate an NmtWorkspace object that we use to compute and store
+#the mode coupling matrix. Note that this matrix depends only on the masks
+#of the two fields to correlate, but not on the maps themselves (in this
+#case both maps are the same.
+w=nmt.NmtWorkspace()
+w.compute_coupling_matrix(f0,f2,b)
+
+#Since we suspect that our maps are contaminated (that's why we passed the
+#contaminant templates as arguments to the NmtField constructor), we also
+#need to compute the bias to the power spectrum caused by contaminant
+#cleaning (deprojection bias).
+cl_bias=nmt.deprojection_bias(f0,f2,cl_02_th)
+
[…]

### 2018-07-30 doc/Paper/main.tex
wrote abstract

\RequirePackage{docswitch}
% \flag is set by the user, through the makefile:
%    make note
%    make apj
% etc.
\setjournal{\flag}

\documentclass[\docopts]{\docclass}

% You could also define the document class directly
%\documentclass[]{emulateapj}

% Custom commands from LSST DESC, see texmf/styles/lsstdesc_macros.sty
\usepackage{lsstdesc_macros}

\usepackage{graphicx}
\graphicspath{{./}{./figures/}}
\bibliographystyle{apj}

% Add your own macros here:
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\lstset{ %
language=python,                % choose the language of the code
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=false,           % adds a frame around the code
tabsize=2,          % sets default tabsize to 2 spaces
captionpos=b,           % sets the caption-position to bottom
breaklines=true,        % sets automatic line breaking
breakatwhitespace=false,    % sets if automatic breaks should only happen at whitespace
escapeinside={\%*}{*)}          % if you want to add a comment within your code
}
\newcommand{\nv}{\hat{\bf n}}
\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
\newcommand{\todo}[1]{{\bf TODO: #1}}
\newcommand{\nmt}{{\tt NaMaster}}
\newcommand{\va}{{\mathbf{a}}}
\newcommand{\vb}{{\mathbf{b}}}
\newcommand{\mC}{{\sf C}}
\newcommand{\mM}{{\sf M}}

\newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
\newcommand{\as}[1]{{\textcolor{red}{[\textbf{AS}: #1]}}}
\newcommand{\dam}[1]{{\textcolor{green}{[\textbf{DA}: #1]}}}
\newcommand{\js}[1]
{{\textcolor{orange}
{[\textbf{JS}: #1]}}}

% ======================================================================

\begin{document}

\title{A unified pseudo-$C_\ell$ framework}

\maketitlepre

\begin{abstract}
  The pseudo-$C_\ell$ provides a fast and simpler alternative to the optimal quadratic estimator to compute angular power spectra. The algorithm can be extended to deal with contaminant deprojection and $E/B$ purification, and can therefore be applied in a wide variety of scenarios of interest for current and future cosmological observations. This paper presents \nmt, a public, validated, robust, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both in the sphere or in the flat-sky approximation. We describe the mathematical background of the estimator, including all the features above, and its software implementation in \nmt. We construct a validation suite that aims to resemble the types of observations that next-generation large-scale structure and ground-based CMB experiments will face, and use it to show that the code is able to recover the input power spectra in the most complex scenarios with no detectable bias. \nmt~can be found at \url{add_url}, and is provided with comprehensive documentation and a number of code examples to maximize its usefulness in the community.
\end{abstract}

% Keywords are ignored in the LSST DESC Note style:
\dockeys{methods: data analysis -- methods: numerical}

\maketitlepost

% ----------------------------------------------------------------------
% {% raw %}

  \section{Introduction}\label{sec:intro}
    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citet{1995PhRvL..74.4369B,1997PhRvD..55.5895T} \todo{More cites?}. Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier/harmonic-space power spectra $C_\ell$ \todo{cite one of the angular xcorr papers}. Although correlation functions are often easier to estimate from complex observations~\citet{1993ApJ...412...64L}, a far higher level of data compression and decorrelation can be obtained through $C_\ell$s. This, together with other advantages associated with power spectra (e.g. clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citet{2008PhRvD..77j3013H} \todo{cite Sellentin} has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators.

    While maximum-likelihood~\citet{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citet{1997PhRvD..55.5895T,2000MNRAS.317L..23H} can recover the power spectrum with virtually no loss of information, their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ and $\ell_{\rm max}^4$ respectively with the maximum multipole $\ell_{\rm max}$. For this reason, pseudo-$C_\ell$ algorithms \citet{1973ApJ...185..413P,2002ApJ...567....2H}\todo{cite more everywhere} have become a popular alternative that preserves a the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation). This method can be used on fields with arbitrary spin \citet{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citet{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $B$-mode power spectrum \citet{2009PhRvD..79l3515G}. Unfortunately, in spite of the advantages of this method, no public, validated and well-maintained codes exist supporting all of these capabilities. This paper presents a public software package, \nmt, that unifies and extends all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.

    The paper is structured as follows: Section \ref{sec:maths} introduces the pseudo-$C_\ell$ estimator in the curved sky, including mode deprojection and $E/B$ purification. In Section \ref{sec:structure} we briefly describe the implementation of these methods on \nmt, as well as its structure and usage. Section \ref{sec:validation} presents the stress tests used to validate the code. We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicatiblity in typical use-cases scenarios in both fields. This section also studies the speed and scaling. We conclude in Section \ref{sec:discussion}.

  \section{The pseudo-\texorpdfstring{$C_\ell$}{Cl} estimator}\label{sec:maths}
    Here we introduce the pseudo-$C_\ell$ estimator and its relevant extensions. Many of the results presented here can be found in the literature (e.g. \cite{2002ApJ...567....2H,2003ApJS..148..161K,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2013A&A...554A.112R,2017MNRAS.465.1847E} and references therein), although some others are new as far as we are aware. In what follows we will use boldface symbols for vector fields (e.g. ${\bf a}$) and sans-serif symbols for matrices (e.g. ${\sf C}$).

    \subsection{Spherical harmonic transform and pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.pcl101}
      Let ${\bf a}(\nv)$ be a spin-$s_a$ quantity defined on the sphere, where $\nv$ is the unit vector pointing in a particular direction $(\theta,\varphi)$. We define its spherical harmonic coefficients through a spherical harmonic transform (SHT) as:
      \begin{align}\label{eq:sht1}
        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,_{s_a}{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
        {\bf a}(\nv)&=\mathcal{S}^{-1}\left[{\bf a}_{\ell m}\right]^{s_a}_{\nv}\equiv\sum_{\ell m}\,{\sf Y}^{s_a}_{\ell m}(\nv)\,{\bf a}_{\ell m}.
      \end{align}
      Note that, in general we will use a vector notation such that, for a complex spin-$s_a$ field $a$, we form the vector ${\bf a}\equiv({\rm Re}(a),{\rm Im}(a))$. The harmonic coefficients above are decomposed in a similar manner into $E$ and $B$ modes: ${\bf a}_{\ell m}\equiv(a^E_{\ell m},a^B_{\ell m})$. The matrices ${\sf Y}^s_{\ell m}$ are defined in terms of the spin-weighed spherical harmonics, and are described in detail in Appendix \ref{app:shts}. To simplify the notation, we will also often abbreviate the pair $(\ell,m)$ as ${\bf l}$.

      For two isotropic fields ${\bf a}$ and ${\bf b}$, the power spectrum is given by their covariance matrix:
      \begin{equation}\label{eq:cldef}
        \left\langle{\bf a}_{\bf l}{\bf b}_{{\bf l}'}^\dag\right\rangle\equiv{\sf C}^{ab}_\ell\delta_{\ell\ell'}\delta_{mm'}.
      \end{equation}
      Note that for a general non-zero spin, vectors $\va$ and $\vb$ have two-components, each of which is a complex number. Nevertheless the rotational invariance requires that elements of the $2\times2$ matrix $\mC$ be real.

      \subsubsection{Power-spectrum estimation}
        The problem we are attempting to solve is how to estimate ${\sf C}^{ab}_\ell$ from a single realization of ${\bf a}$ and ${\bf b}$ measured on a cut sky. In general, the input data for measuring the power spectrum come pixelated on a sphere. Instead of dealing with an infinitely sampled points, we have a finite number of measurements
        \begin{equation}
          \va_i = \va_i^{\rm true} + {\bf n}_i,
        \end{equation}
        where $\va_i^{\rm true}$ is is the true underlying field whose power spectrum we would like to estimate and ${\bf n}_i$ is the noise vector, which in general can be non-white and non-homogeneous. We will denote the pixel covariance of the noise component as ${\sf N}$. Note that for the sake of power-spectrum estimation the there is no conceptual difference between inhomogeneous noise and a survey mask: masked areas can be simply though of as areas of infinite noise. Conversely, when we talk about ``multiplying by the mask'' we really mean a more general process of dividing by the expected variance field.

        The underlying power spectrum in Eq.~\ref{eq:cldef} means that there is a non-trivial true underlying covariance matrix $\mC^{\rm true} = \left<\va^{\rm true} \va^{{\rm true}\dagger}\right>$, which in addition to the noise component leads to the actual data covariance $\mC = \left<\va \va^\dagger\right>=\mC^{\rm true} + {\sf N}$.

        In the standard incarnation of the optimal quadratic estimator~\citet{1998PhRvD..57.2117B}\todo{Cite others?}, the measurement is produced on inverse covariance weighted data, namely
        \begin{equation}
          {\sf F_{ij}} B_j = \va^\dagger \mC^{-1}
          \left(\frac{\partial \mC}{\partial B_i} \right)
          \mC^{-1} \va - \rm{Tr}\left[\left(\frac{\partial \mC}{\partial B_i}
          \right) {\sf N} \right],
        \end{equation}
        where ${\sf F}$ is the Fisher matrix,the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
        \begin{equation}
          \mC = \sum_i \left(\frac{\partial \mC}{\partial B_i}\right) + {\sf N},
        \end{equation}
        and the last term correspond to bias of convolved estimates by the presence of noise. Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices $\partial {\sf C}/\partial B_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix. The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particuarly in order to estimate the Fisher matrix. This motivates the pseudo-$C_\ell$ estimator, which replaces these computationally very intensive steps with numerically more efficient ones at the expense of optimality.

        The main improvement is gained by replacing the inverse covariance by its diagonal, assuming that the data is uncorrelated between pixels. The response matrices are also replaced by simple spherical harmonic transforms and multiplication of the corresponding harmonic coefficients, followed by an average over bandpowers. Both of these steps reduce to exact optimal estimator in the limit of uncorrelated data. The resulting estimates are still biased and require a mode-decoupling matrix, which can be calculated as we show in Section \ref{sssec:maths.pcl101.mcoup}.

        The pseudo-$C_\ell$ estimates will be optimal when these steps are close to what an optimal quadratic estimator would do. First, this requires that $\mC$ be close to diagonal, which is true when either the noise is large, or when the underlying power spectrum is close to white. Multiplication by the mask mixes modes, and replacement of multiplication by the response matrix with band-power averaging, only makes sense, when the mask is ``well-behaved'', i.e. compact with no high-frequency structure in it. If these conditions are not met, the pseudo-$C_\ell$ estimator will still be unbiased (by construction), but it will get progressively less optimal. In practical surveys, however, the loss of optimality rarely exceeds 10-20\% and is typically smaller~\citet{2013MNRAS.435.1857L}.

        In what follows we will work with two-fields $\va$ and $\vb$ and spell out in detail how to calculate the cross-power spectrum. The auto-power spectrum case is the same (with $\vb=\va$), but one must additionally subtract the noise-contribution to the estimate. So far, the discussion has been completely general, but from now onwards, we will focus on the power estimation of auto, and cross-power spectra for the spin-0 and spin-2 fields only, as these are the most commonly used in cosmology.

      \subsubsection{Mode coupling}\label{sssec:maths.pcl101.mcoup}
        Let $v(\nv)$ be a sky mask or weights map for ${\bf a}$, and let us define ${\bf a}^v\equiv v(\nv){\bf a}(\nv)$. As discussed above we start by considering a na\"ive estimator for ${\sf C}^{ab}_\ell$ from the masked fields:
        \begin{equation}\label{eq:pcldef}
          {\rm PCL}_\ell\left({\bf a}^v,{\bf b}^w\right)\equiv\frac{1}{2\ell+1}\sum_{m=-\ell}^\ell{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{\bf l}.
        \end{equation}
        The incomplete sky coverage couples different $\ell$ modes, and makes this estimator biased. The pseudo-$C_\ell$ method is based on computing an analytical prediction for this bias and correcting for it. This is straightforward to do using the results in Appendix \ref{app:shts}. The harmonic coefficients of a masked field are given by
        \begin{equation}
          {\bf a}^v_{\bf l}=\sum_{{\bf l}_1{\bf l}_2}={\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}{\bf a}_{{\bf l}_1}v_{{\bf l}_2},
        \end{equation}
        where ${\sf D}$ is defined in Eq. \ref{eq:dmat}. Using the statistical isotropy of the unmasked field (Eq. \ref{eq:cldef}) and the orthogonality relation Eq. \ref{eq:3jorth}, we obtain:
        \begin{equation}\label{eq:modecoup}
          {\rm vec}\left[\left\langle{\rm PCL}_\ell\left({\bf a}^v_{\bf l},{\bf b}^w_{\bf l}\right)\right\rangle\right]=\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\cdot{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
        \end{equation}
        where we have defined the vectorization operator \citet{2008PhRvD..77j3013H}
        \begin{equation}
          {\rm vec}\left[\left(\begin{array}{cc}
                          C^{EE} & C^{EB}\\
                          C^{BE} & C^{BB}
                         \end{array}\right)\right]\equiv
          \left(
          \begin{array}{c}
            C^{EE} \\
            C^{EB} \\
            C^{BE} \\
            C^{BB} \\
          \end{array}
          \right).
        \end{equation}

        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq. \ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by:
        \begin{align}\label{eq:m00}
          &{\sf M}^{00}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\,{\sf 1},\\\label{eq:m02}
          &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,{\sf 1},\\\nonumber
          &M^{0+}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0},\\\label{eq:m22}
          &{\sf M}^{22}=\left(
                        \begin{array}{cccc}
                          M^{+}_{\ell\ell'} & 0 & 0 & M^{-}_{\ell\ell'} \\
                          0 & M^{+}_{\ell\ell'} & -M^{-}_{\ell\ell'} & 0 \\
                          0 & -M^{-}_{\ell\ell'} & M^{+}_{\ell\ell'} & 0 \\
                          M^{-}_{\ell\ell'} & 0 & 0 & M^{+}_{\ell\ell'}
                        \end{array}\right),\\\nonumber
          &M^{\pm}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}^2\frac{1\pm(-1)^{\ell+\ell'+\ell''}}{2},
        \end{align}
        where ${\sf 1}$ is the unit matrix, and $P^{vw}_\ell\equiv(2\ell+1)\,{\rm PCL}_\ell(v,w)$.

      \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
        Given the loss of information implicit in masking the originally full-sky field, it is in general not possible to invert Eq. \ref{eq:modecoup} directly. One solution to this problem is to convolve the theory prediction with the mode-coupling matrix at the likelihood level. A more usual approach is instead to bin the coupled pseudo-$C_\ell$ into bandpowers. A bandpower $q$ is defined as a set of $N_q$ multipoles $\vec{\ell}_q\equiv(\ell_q^1,...,\ell_q^{N_q})$ and a set of weights defined at those multipoles $w^\ell_q$, and normalized such that $\sum_{\ell\in\vec{\ell}_q}w_q^\ell=1$. The $q$-th bandpower for the coupled pseudo-$C_\ell$ is then defined as
        \begin{align}\label{eq:bpws}
          {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
          &=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right].
        \end{align}
        One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function). Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
        \begin{equation}\label{eq:decoup}
          {\rm vec}\left[{\sf B}^{ab}_q\right]=\sum_{q'}\left({\cal M}^{s_as_b}\right)^{-1}_{qq'}\,{\rm vec}\left[\tilde{\sf B}^{ab}_{q'}\right],
        \end{equation}
        where the binned coupling matrix ${\cal M}^{s_as_b}$ is
        \begin{equation}
          {\cal M}^{s_as_b}_{qq'}=\sum_{\ell\in\vec{\ell}_q}\sum_{\ell'\in\vec{\ell}_{q'}}w_q^\ell\,{\sf M}^{s_as_b}_{\ell\ell'}.
        \end{equation}

        Since, in general, the true power spectrum is not a step-wise function, the theoretical prediction must be corrected for the bandpower binning before making any comparison with the data:
        \begin{equation}\label{eq:theory_correct}
          {\rm vec}\left[{\sf B}^{ab,{\rm th}}_q\right]=\sum_{\ell}{\cal F}^{s_as_b}_{q\ell}\cdot{\rm vec}\left[{\sf C}^{ab,{\rm th}}_\ell\right],
        \end{equation}
        where the filter ${\cal F}^{s_as_b}_{q\ell}$ is given by
        \begin{equation}\label{eq:bpresponse}
          {\cal F}^{s_as_b}_{q\ell}=\sum_{q'}({\cal M}^{s_as_b})^{-1}_{qq'}\sum_{\ell'\in\vec{\ell}_{q'}}w_{q'}^{\ell'}{\sf M}^{s_as_b}_{\ell'\ell}.
        \end{equation}
        The action of the filter ${\cal F}^{s_as_b}$ is therefore given by a sequence of three operations:
        \begin{enumerate}
          \item Coupling the different multipoles (Eq. \ref{eq:modecoup}).
          \item Binning into bandpowers (Eq. \ref{eq:bpws}).
          \item Decoupling the bandpowers (Eq. \ref{eq:decoup}).
        \end{enumerate}
        Depending on the structure of the mode-coupling matrix, ${\cal F}^{s_as_b}$ can be significantly different from a na\"ive binning operator.

    \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
      In a typical data-analysis case, we deal with contaminants that pollute the measured signal. Usually, we know something about these contaminants. As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which we might want to fit and marginalise out. But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening. Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales \dam{sounds like a very strong statement that we should either prove or give references for! I think I would rather say that you have either reduced contamination enough that a first-order Taylor term should account for the rest or that, in general, you can use a Taylor expansion where each term is a contaminant map.}\js{I agree with David, I'd go with the Taylor expansion motivation}. Thus, we generically deal with a linear model of the kind:
      \begin{equation}\label{eq:aobs}
        {\bf a}^{\rm obs}(\nv)={\bf a}^v+\sum_{i=1}^{N_a}\alpha_i{\bf f}^i(\nv),
      \end{equation}
      where the observed map ${\bf a}^{\rm obs}$ is composed of a masked signal map that receives contributions from $N_a$ known contaminants ${\bf f}^i$ and where the contamination coefficients $\alpha_i$ are unknown (note that we have implicitly applied the same mask to the contaminant templates ${\bf f}^i$). A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citet{2013MNRAS.435.1857L}. Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
      \begin{align}\label{eq:deproj_coeff}
        &\hat{\alpha}_i=\sum_iF_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv).
      \end{align}
      The clean map is then given by
      \begin{align}\label{eq:clean}
        {\bf a}^c(\nv)
        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')\\\nonumber
        &={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv'),
      \end{align}
      where there is an implicit summation over $i$ and $j$ (which we will omit from now on).
      The harmonic coefficients of ${\bf a}^c$ are therefore given by
      \begin{equation}\label{eq:clean_f}
        {\bf a}^c_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}\,F_{ij}\sum_{{\bf l}'}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}
      \end{equation}

      Due to the residual statistical contamination in Eq. \ref{eq:clean}, a direct computation of the pseudo-$C_\ell$ of the cleaned map would yield a biased estimator beyond the mode coupling induced by the mask $v$. This bias can however be estimated analytically and corrected for. This was done in \cite{2017MNRAS.465.1847E} for scalar quantities, and we generalize their result here to fields of arbitrary spin.

      Considering a second map ${\bf b}$ with mask $w$, $N_b$ contaminants ${\bf g}^i$ and correlation matrix $G_{ij}$, the mode-coupled pseudo-$C_\ell$ of ${\bf a}^c$ and ${\bf b}^c$ is given by:
      \begin{align}\nonumber
        {\rm PCL}_\ell({\bf a}^c,{\bf b}^c)=&{\rm PCL}_\ell({\bf a}^v,{\bf b}^w)\\\nonumber
        &-\frac{G_{ij}}{2\ell+1}\sum_m\sum_{{\bf l}'}{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{{\bf l}'}{\bf g}^j_{{\bf l}'}{\bf g}^{i\dag}_{\bf l}\\\nonumber
        &-\frac{F_{ij}}{2\ell+1}\sum_m\sum_{{\bf l}'}{\bf f}^i_{\bf l}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}{\bf b}^{w\dag}_{\bf l}\\\label{eq:dpbias0}
        &+\frac{F_{ij}G_{rs}}{2\ell+1}\sum_m\sum_{{\bf l}',{\bf l}''}{\bf f}^i_{\bf l}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}{\bf b}^{w\dag}_{{\bf l}''}{\bf g}^r_{{\bf l}''}{\bf g}^{s\dag}_{\bf l}.
      \end{align}
      Taking the ensemble average of the equation above, we find that the first term is the usual pseudo-$C_\ell$ estimator. The combination of the last three terms is what we will refer to as the \emph{deprojection bias}, and can be computed as\footnote{These equations can be easily derived by expanding ${\bf a}^v$ and ${\bf b}^v$ in terms of the spherical harmonics of ${\bf a}$, ${\bf b}$, $v$ and $w$, making use of Eq. \ref{eq:cldef} and repeatedly employing the definitions of the spherical harmonic transform (Eqs. \ref{eq:sht1} and \ref{eq:sht2}).}:
      \begin{align}\nonumber
        \Delta {\sf C}^{ab}_\ell=&-G_{ij}\,{\rm PCL}_\ell\left(\tilde{\bf g}^j_{\bf l},{\bf g}^i_{\bf l}\right)-F_{ij}\,{\rm PCL}_\ell\left({\bf f}^i_{\bf l},\tilde{\bf f}^j_{\bf l}\right)\\\label{eq:dpbias}
        &+F_{ij}G_{rs}\left[\int d\nv\,{\bf f}^{j\dag}(\nv)\tilde{\bf g}^r(\nv)\right]{\rm PCL}_\ell\left({\bf f}^{i},{\bf g}^s\right),
      \end{align}
      where
      \begin{align}\label{eq:dp.g.aux}
        \tilde{\bf g}^i_{\bf l}&\equiv{\cal S}\left[v(\nv){\cal S}^{-1}\left[{\sf C}^{ab}_{\ell_1}{\cal S}\left[w\,{\bf g}^i\right]^{s_b}_{{\bf l}_1}\right]^{s_a}_{\nv}\right]^{s_a}_{\bf l},\\\label{eq:dp.f.aux}
        \tilde{\bf f}^i_{\bf l}&\equiv{\cal S}\left[w(\nv){\cal S}^{-1}\left[{\sf C}^{ab\dag}_{\ell_1}{\cal S}\left[v\,{\bf f}^i\right]^{s_a}_{{\bf l}_1}\right]^{s_b}_{\nv}\right]^{s_b}_{\bf l}.
      \end{align}

      \as{I would move this procedure together with full procedure to Section 3.}
      The procedure to obtain an unbiased estimator of the power spectrum in the presence of contaminants can therefore be summarized as follows:
      \begin{enumerate}
        \item Produce clean versions of both maps (i.e. estimate the best-fit coefficients in Eq. \ref{eq:deproj_coeff} and subtract the contamination as in Eq. \ref{eq:clean}).
        \item Estimate the pseudo-$C_\ell$ of the clean maps (Eq. \ref{eq:pcldef}).
        \item Estimate the deprojection bias in Eq. \ref{eq:dpbias} and subtract it from the pseudo-$C_\ell$ above.
        \item Use the methods described in Sections \ref{sssec:maths.pcl101.mcoup} and  \ref{sssec:maths.pcl101.bpws} to account for the mode-coupling matrix.        
      \end{enumerate}

      Two further points must be addressed before moving on. First, the computation of the deprojection bias requires an estimate of the true power spectrum ${\sf C}^{ab}_\ell$ (see Eqs. \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}). In practice, this can be done through an iterative approach, starting from an initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the estimated power spectrum in the previous iteration. A sufficiently good guess will likely be enough to remove the bias in a single iteration. Alternatively, the ensemble average of Eq. \ref{eq:dpbias0} can be seen as a convolution of ${\sf C}^{ab}_\ell$, which could be treated by computing the convolution matrix as a modified version of the standard mode-coupling matrix that accounts for mode deprojection. Estimating this convolution matrix in this case becomes significantly more computationally intensive, and therefore an iterative approach is preferred. \as{I think this is not quite true. This latter way of looking at it, in effect replaces the true power spectrum with its observed quantity, which is fine in practice, but conceptually not quite right, for the same raeason e.g. error on power spectrum should be based on real rather than measured power as illustrated by low CMB quadrupole. I would just remove text after ``Alternatively...''} \dam{we can do this, but I'd like to understand what the problem with this is. Since you can understand a mask as a set of deprojected contaminants (one per masked pixel), the pseudo-$C_\ell$ mode-coupling matrix is basically a flavour of the equations above for one particular set of templates. In this sense then writing the equation above as a matrix-vector product and inverting that matrix (after binning etc.) seems like the pseudo-$C_\ell$ way to go. Are you saying that you'll get more optimal results by putting in a guess for the power spectrum because that's closer to what a QML estimator would do?}

      Secondly, estimating the deprojection coefficients (Eq. \ref{eq:deproj_coeff}) involves inverting the correlation matrix of all the contaminant templates. This will not be in general possible (e.g. if a set of templates are linearly related). The problem can however be solved self-consistently by using the Moore-Penrose pseudo-inverse. This is trivial for a symmetric and positive-definite matrix such as ${\sf F}$, and corresponds to setting the inverse of all of its zero eigenvalues (or all those below a given threshold) to zero.

    \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
      As described in \cite{2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa. The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB. In this case, removing the leakage at the power spectrum level (i.e. the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes. This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above. In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e. purifying $E$-modes from $B$-mode contamination is conceptually exactly the same. Moreover. our discussion is also specific for spin-2 fields.

      We define a field ${\bf f}$ to be a $B$ mode if $({\bf D}^{E}_s)^\dag{\bf f}=0$, where ${\bf D}^{E}_s$ is the differential operator defined in Eq.~\ref{eq:ederiv}. At the same time, and under the definition of the dot product:
      \begin{equation}
        ({\bf f},{\bf g})\equiv\int d\nv\,{\bf f}^\dag(\nv){\bf g}(\nv),
      \end{equation}
      we define a \emph{pure} $B$ mode as a field that is orthogonal to all $E$ modes.

      Since ${\bf D}^{E\dag}_s{\bf D}^B_s=0$, one can always generate a $B$ mode by applying ${\bf D}^{B}_s$ to a scalar field. It is then possible to show that $E$ and $B$ modes thus defined are orthogonal in the full sky:
      \begin{equation}
        ({\bf D}^E_s\varphi,{\bf D}^B_s\psi)=\int d\nv\,({\bf D}^E_s\varphi)^\dag{\bf D}^B_s\psi=0
      \end{equation}
      This can be done by integrating by parts and noting that the celestial sphere has no boundaries. On a cut sky, however, and for $s=2$, this is only true if the fields satisfy Neumann and Dirichlet boundary conditions simultaneously (i.e. vanishing value and first derivative on the boundary of the cut sky region).

      Let $w(\nv)$ be a sky window function defining the sky region to be analyzed (and the weight to be applied in each pixel). The standard pseudo $B$-mode of a field ${\bf P}$ is then given by
      \begin{align}\nonumber
        \tilde{B}_{\bf l}&\equiv\int d\nv\,w(\nv)\left(_s{\bf Y}^{B}_{\bf l}(\nv)\right)^\dag{\bf P}\\
                         &=\int d\nv\,w(\nv)({\bf D}^B_sY_{\bf l})^\dag{\bf P}(\nv),
      \end{align}
      Since ${\bf D}^B_sY_{\ell m}$ is a $B$-mode, in the absence of $w$ this expression would correspond to a projection that filters out all the $E$-modes from ${\bf P}$. However, $w(\nv){\bf D}^B_sY_{\ell m}$ is not a $B$-mode, and therefore $\tilde{B}_{\ell m}$ receives contributions from ambiguous $E$ modes (which then propagate into the variance of the pseudo-$C_\ell$ estimator of the power spectrum). The idea behind $B$-mode purification is to move $w$ to the right of ${\bf D}^B_s$, defining the pure $B$ component:
      \begin{equation}
        B^p_{\bf l}=\int d\nv\left({\bf D}^B_2(w\,Y_{\bf l})\right)^\dag\,{\bf P}(\nv).
      \end{equation}
      Since ${\bf D}^B_2(wY_{\bf l})$ is a $B$-mode quantity, $B^p_{\bf l}$ should receive contributions only from $B$-modes.

      Expanding ${\bf D}^B_2(wY_{\bf l})$, we can write $B^p_{\bf l}$ as:
      \begin{equation}\label{eq:pureb}
        B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}}\left(\tilde{P}_1\right)^B_{\bf l}+\alpha_{\ell,2}\left(\tilde{P}_0\right)^B_{\bf l},
      \end{equation}
      where $(a)^B_{\bf l}$ stands for the $B$-mode component of field ${\bf a}$, and we have defined the fields $\tilde{P}_n=(\eth^{2-n}w)^*(Q+iU)$, where $Q$ and $U$ are the real and imaginary parts of the field $P$ (see Appendix \ref{app:shts} for the definitions of $\eth$ and $\alpha_{\ell,s}$).

      The $B(E)$-purification of a given field is simply achieved by applying Eq. \ref{eq:pureb} to the $B(E)$-mode component of the field. Note that doing so requires the computation of the first and second-order derivatives of the weights map $w$, and therefore purification methods require masks for which this quantities are well defined. This usually involves tapering the mask boundaries to avoid sharp edges. A thorough discussion of these methods can be found in \cite{2009PhRvD..79l3515G}. Note that the derivatives of $w$ can be computed analytically in harmonic space:
      \begin{align}
        \eth^nw={\cal S}^{-1}\left[\left(-\frac{w_{\bf l}}{\alpha_{\ell,n}},0\right)\right],
      \end{align}

      At this stage, the pseudo-$C_\ell$ method proceeds as usual, applying Eq. \ref{eq:pcldef} to the purified field. The purification process however requires a slight modification of the analytical form of the mode-coupling matrix. Specifically, if either $E$ or $B$ modes of a spin-2 field have been purified Eqs. \ref{eq:m02} and \ref{eq:m22} must be modified by exchanging one factor of
      \begin{equation}\nonumber
        \wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}
      \end{equation}
      for
      \begin{align}\nonumber
        &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}+2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}\alpha_{\ell'',1}}\wtj{\ell}{\ell'}{\ell''}{1}{-2}{1}\\
        &\,\,\,+\frac{\alpha_{\ell,2}}{\alpha_{\ell'',2}}\wtj{\ell}{\ell'}{\ell''}{0}{-2}{2}.
      \end{align}
      \emph{per purified field}. \todo{check this}.

      Finally, it is also possible to combine purification and contaminant deprojection. Since $E/B$ purifications involves destroying part of the signal (e.g. projecting out all leaked $E$ modes from a $B$-mode component), it is more optimal to use the non-pure maps to compute the contamination coefficients ($\alpha_i$ in Eq. \ref{eq:deproj_coeff}). Therefore we will assume here that purification takes place between the steps (i) and (ii) described at the end of Section \ref{ssec:maths.deproj}, and not before step (i). In this case, although no additional modifications are needed in the mode-coupling matrix, beyond those we have just described, the expression for the deprojection bias (Eq. \ref{eq:dpbias}) must be slightly altered. In particular, although the fields $\tilde{\bf g}^i$ and $\tilde{\bf f}^i$ in Equations \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux} are still computed from the non-pure versions of ${\bf f}^i$ and ${\bf g}^i$, all the maps entering the three instances of ${\rm PCL}_\ell$ in Eq. \ref{eq:dpbias} must be consistently $E/B$-purified before computing those pseudo-$C_\ell$s.

      \subsection{Beams and noise}\label{ssec:maths.beam_noise}
        In any practical scenario, the observed sky map will contain a mixture of signal ${\bf s}$ (i.e. the component whose power spectrum we are really interested in) and noise ${\bf n}$ (i.e. a stochastic contaminant that we do not have a template for). Furthermore, the signal component will often be smoothed on the smallest scales by an instrumental beam. We discuss how to deal with both of these complications here.

        \subsubsection{Beam deconvolution}\label{sssec:maths.beam_noise.beam}
          Including the effect of an spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
          \begin{equation}
            {\sf M}^{s_as_b}_{\ell_1\ell_2}\rightarrow{\sf M}^{s_as_b}_{\ell_1\ell_2}W_{\ell_2}^{ab},
          \end{equation}
          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$. Note that it is impossible to have an anisotropic beam that is uniform over the entire sphere. However, it is possible that the beam is both anisotropic and spatially varying. In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam. Correcting for the beam in such cases can also be done analytically, but we leave it for future work. \as{Anything to add here?} \dam{maybe be a bit more precise.}

        \subsubsection{Noise bias}
         In auto-correlation, the two-point correlation of the noise component ${\bf n}$ will contaminate our estimate of the power spectrum of ${\bf s}$, and we must correct for this noise bias. Depending on the statistical properties of ${\bf n}$, different strategies can be used to do this:
         \paragraph*{Homogeneous noise.}
           If the noise is homogeneous, and can be described by a known noise power spectrum ${\sf N}_\ell$, the procedure is simple: first, convolve the known noise power spectrum with the pseudo-$C_\ell$ mode-coupling matrix. Then, estimate the contribution from contaminant deprojection by using ${\sf N}_\ell$ in lieu of ${\sf C}^{ab}_\ell$ in Eqs. \ref{eq:dpbias}, \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}. The total noise bias is then the sum of both contributions, corrected for the pseudo-$C_\ell$ convolution using Eq. \ref{eq:theory_correct}.

         \paragraph*{Uncorrelated noise.}
           If the noise is homogeneous but uncorrelated (i.e. $\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle={\sf 1}\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
           \begin{itemize}
             \item It is not difficult to prove that the pseudo-$C_\ell$ of the noise contribution is given by
                   \begin{align}
                     {\rm PCL}_\ell({\bf n},{\bf n})&={\sf 1}\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
                                                    &={\sf 1}\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
                   \end{align}
                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=\alpha\sigma_n^{-2}(\nv)$.
             \item The additional bias from contaminant deprojection can also be computed analytically, starting from Eq. \ref{eq:dpbias0}, as:
                   \begin{align}\nonumber
                     \Delta{\sf C}^{nn}_\ell=&-2F_{ij}{\rm PCL}_\ell\left({\bf f}^i,v^2\sigma_n^2{\bf f}^j\right)\\
                     &+F_{ij}F_{rs}\left[\int d\nv\, v^2\sigma_n^2\,{\bf f}^{j\dag}{\bf f}^r\right]\,{\rm PCL}_\ell\left({\bf f}^i,{\bf f}^s\right).
                   \end{align}
           \end{itemize}

        \paragraph*{General case.} In general, if the noise is both non-white and inhomogeneous, the total noise bias can be estimated by averaging the result of applying the pseudo-$C_\ell$ estimator to a large number of noise realizations with the same noise properties as the data, assuming that those properties are sufficiently well characterised to produce those simulations. If that is not the case, a common approach to avoiding the noise bias altogether is to use only cross-correlations of data splits with independent noise contributions.

    \subsection{Flat-sky pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.flat}
      Using the full-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
      \begin{enumerate}
        \item Full-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch. This can lead to an inefficient use of memory and storage space, because the standard numerical implementations of spherical transforms operate on full-sky data.
        \item Spherical harmonic transforms are notoriously slower than discrete Fourier transforms (DFTs), both due to the scaling of both algorithms (${\cal O}(N_{\rm pix}^{3/2})$ for SHTs vs. ${\cal O}(N_{\rm pix}\log(N_{\rm pix}))$ for DFTs) and because by SHTs involve operations over the full sphere.
        \item Fourier transform over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
      \end{enumerate}
      For this reason, \nmt\, also supports the computation of power spectra from flat-sky maps. Appendix \ref{app:flat} describes the dictionary between spin-$s$ fields and their Fourier/harmonic coefficients defined in flat and curved skies, and we present the flat-sky versions of the pseudo-$C_\ell$ methods described in the previous sections here.

      \subsubsection{Standard pseudo-$C_\ell$}\label{sssec:maths.flat.pcl101}
        Using the definitions presented in Appendix \ref{app:flat}, it is easy to show that the Fourier coefficients of a masked field are given by
        \begin{equation}\label{eq:pcl_flat_masked}
          {\bf a}^v_{\bf l}\equiv{\cal D}(v\,{\bf a})^{s_a}_{\bf l}=\sum_{\bf k}\frac{\Delta {\bf k}^2}{2\pi}{\sf R}(s_a(\varphi_{\bf l}-\varphi_{\bf k}))\,{\bf a}_{\bf k}\,v_{{\bf l}-{\bf k}}.
        \end{equation}
        Note that, in the context of flat-sky fields, the vector ${\bf l}$ is a 2D wave vector, and must not be confused with the abbreviation ${\bf l}\equiv(\ell,m)$ used in previous sections.
        It is then straightforward to estimate the covariance of two masked field at the same wavenumber:
        \begin{equation}\label{eq:flat_coup}
          \left\langle{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{\bf l}\right\rangle=\sum_{\bf k}\frac{\Delta{\bf k}^2}{(2\pi)^2}{\sf R}(s_a\Delta\varphi){\sf C}^{ab}_{\bf k}{\sf R}^\dag(s_b\Delta\varphi)\,v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}},
        \end{equation}
        where $\Delta\varphi\equiv\varphi_{\bf l}-\varphi_{\bf k}$.

        At this stage it is natural to connect directly with the final bandpowers since, unlike in the full-sky case, there is no natural minimal binning ($\Delta\ell=1$) for the un-masked power spectrum. In this case we will define a bandpower ${\sf B}_q$, indexed by an integer $q$ as the average of the covariance above over a set of ${\bf l}$ values $S_q$:
        \begin{equation}
          {\rm vec}\left[{\sf B}^{ab}_q\right]\equiv\frac{\Delta{\bf k}^2}{N_q}\sum_{{\bf l}\in S_q}{\rm vec}\left[{\bf a}^v_{\bf l}{\bf b}^{w\dag}_{\bf l}\right],
        \end{equation}
        where $N_q$ is the number of Fourier-space pixels covered by $S_q$. Typically, $S_q$ will be the set of pixels within an annulus defined by an interval $\left(|{\bf k}|_{\rm min},|{\bf k}|_{\rm max}\right)$.  Taking the expectation value of the equation above, we can relate this estimator to the true power spectrum as:
        \begin{equation}\label{eq:predflat}
          \langle{\rm vec}\left[{\sf B}^{ab}_q\right]\rangle=\sum_{{\bf l}\in S_b}N_q^{-1}\sum_{\bf k}{\sf M}^{s_a\,s_b}_{{\bf l}{\bf k}}\cdot\,{\rm vec}\left[\hat{\sf C}^{ab}_{\bf k}\right],
        \end{equation}
        where the un-binned mode-coupling matrix is:
        \begin{align}\label{eq:m00_flat}
          {\sf M}^{00}_{{\bf l}\,{\bf k}}&\equiv\frac{(2\pi)^2}{L_x^2L_y^2}v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}},\\\label{eq:m02_flat}
          {\sf M}^{02}_{{\bf l}\,{\bf k}}&\equiv\frac{(2\pi)^2}{L_x^2L_y^2}v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}
          \left(
          \begin{array}{cc}
           {\rm c} & -{\rm s} \\
           {\rm s} &  {\rm c}
          \end{array}\right)\\\label{eq:m22_flat}
          {\sf M}^{22}_{{\bf l}\,{\bf k}}&\equiv\frac{(2\pi)^2}{L_x^2L_y^2}v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}\left(
          \begin{array}{cccc}
           {\rm c}^2 & -{\rm c\,}{\rm s} & -{\rm c}\,{\rm s} & {\rm s}^2 \\
           {\rm c}\,{\rm s} & {\rm c}^2 & -{\rm s}^2 & -{\rm c}\,{\rm s} \\
           {\rm c}\,{\rm s} & -{\rm s}^2 & {\rm c}^2 & -{\rm c}\,{\rm s} \\
           {\rm s}^2 &  {\rm c}\,{\rm s} &  {\rm c}\,{\rm s} & {\rm c}^2
          \end{array}\right),
        \end{align}
        where $\rm{c} \equiv \cos2\Delta\varphi$ and $\rm{s}\equiv \sin2\Delta\varphi$. The mode-coupling matrix for the bandpowers is therefore given by:
        \begin{equation}\label{eq:modcoup_flat}
          \mathcal{M}^{s_as_b}_{qq'}=\sum_{{\bf l}\in S_q}N_q^{-1}\sum_{{\bf k}\in S_{q'}}{\sf M}^{s_as_b}_{{\bf l}\,{\bf k}}.
        \end{equation}

        At this point it is important to note that, with this procedure, computing the coupling matrix becomes an ${\cal O}(N_{\rm pix}^2)$ problem (or at best ${\cal O}(\ell_{\rm max}^2N_{\rm pix})$), which is worse than the ${\cal O}(\ell_{\rm max}^3$) pseudo-$C_\ell$ algorithm we described in the curved sky case. The key in that case is the orthogonality relation Eq. \ref{eq:3jorth}, which significantly simplifies the expressions \ref{eq:m00}-\ref{eq:m22}. In a flat sky, and in the presence of a discrete and finite Cartesian grid, it is not possible however to average over azimuthal Fourier degrees of freedom at constant $l$ in order to obtain a numerically stable version of the mask power spectrum for arbitrary weight maps $v$ and $w$. This would not be a problem in the continuum limit ($(\Delta x,\Delta y)\rightarrow0$), where similar simplifying relations can be found to vastly improve the computational efficiency of the method. Taking the continuum approximation is unfortunately not accurate enough in practice, as described in \cite{2016arXiv161204664A}, and a fully discrete approach is usually necessary. This slower performance is normally compensated by the smaller pixel numbers and faster DFTs that can be used in the flat sky, so that the flat-sky approximation is still preferable over a full-sky treatment for small sky patches. Appendix \ref{app:flat.pcl} discusses these issues in detail, providing the continuum-limit expressions mentioned above.

      \subsubsection{Contaminant cleaning}\label{sssec:maths.flat.deproj}
        Using the same notation as in Section \ref{ssec:maths.deproj}, the contaminant-cleaned version of ${\bf a}$ is simply given by(c.f. Eqs. \ref{eq:clean} and \ref{eq:clean_f}):
        \begin{align}
          &{\bf a}^c({\bf x})={\bf a}^v({\bf x})-{\bf f}^i({\bf x})F_{ij}\sum_{\bf x}\Delta{\bf x}^2\,{\bf f}^{j\dag}({\bf x}'){\bf a}^v({\bf x}'),\\
          &{\bf a}^c_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}F_{ij}\sum_{\bf k}\Delta{\bf k}^2\,{\bf f}^{j\dag}_{\bf k}{\bf a}^v_{\bf k},
        \end{align}
        where $({\sf F}^{-1})_{ij}=\sum_{\bf x}\Delta{\bf x}^2{\bf f}^{i\dag}({\bf x}){\bf f}^j({\bf x})$.

        The deprojection bias also takes a similar form (c.f. Eq. \ref{eq:dpbias}):
        \begin{align}\nonumber
          \Delta{\sf C}^{ab}_{\bf l}=&-G_{ij}\,\tilde{\bf g}^j_{\bf l}\,{\bf g}^{i\dag}_{\bf l}-F_{ij}\,{\bf f}^i_{\bf l}\,\tilde{\bf f}^{j\dag}_{\bf l}\\\label{eq:dpbias_flat}
          &+F_{ij}G_{rs}\left[\sum_{\bf x}\Delta{\bf x}^2{\bf f}^{j\dag}({\bf x})\tilde{\bf g}^r({\bf x})\right]\,{\bf f}^i_{\bf l}{\bf g}^{s\dag}_{\bf l},
        \end{align}
        where
        \begin{align}\label{eq:dp.g.aux_flat}
          &\tilde{\bf g}^i_{\bf l}\equiv{\cal D}\left[v({\bf x}){\cal D}^{-1}\left[{\sf C}^{ab}_{{\bf l}_1}{\cal D}\left[w{\bf g}^i\right]^{s_b}_{{\bf l}_1}\right]^{s_a}_{\bf x}\right]^{s_a}_{\bf l},\\\label{eq:dp.f.aux_flat}
          &\tilde{\bf f}^i_{\bf l}\equiv{\cal D}\left[w({\bf x}){\cal D}^{-1}\left[{\sf C}^{ab\dag}_{{\bf l}_1}{\cal D}\left[v{\bf f}^i\right]^{s_a}_{{\bf l}_1}\right]^{s_b}_{\bf x}\right]^{s_b}_{\bf l}.
        \end{align}

      \subsubsection{$E$ and $B$ purification}\label{sssec:maths.flat.pureb}
        The logic behind $E/B$ purification is the same in flat and curved skies, and we will not repeat it here. The analogue of a pure $B$-mode (Eq. \ref{eq:pureb}) in the flat-sky approximation is:
        \begin{equation}
          B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2l^{-1}\left(\tilde{P}_1\right)^B_{\bf l}+l^{-2}\left(\tilde{P}_0\right)^B_{\bf l},
        \end{equation}
        (and a similar relation for the pure $E$ component), where $P_n=(\eth^{2-n}w)^*(Q+iU)$, and $w$ is the sky mask. The derivatives of $w$ can be taken by using the following relation:
        \begin{equation}
          \eth^nw={\cal D}^{-1}\left[\left(-l^n\,w_{\bf l},0\right)\right],\\
        \end{equation}

        Note that, in the flat sky, a mathematically (but not computationally) simpler relation for the pure component can be found, given by:
        \begin{equation}
          B^p_{\bf l}=\int\frac{d{\bf k}^2}{2\pi}B_{\bf k}w_{{\bf l}-{\bf k}}\frac{k^2}{\ell^2},
        \end{equation}
        and similarly for $E$ modes. Comparing with Eq. \ref{eq:flat_coup} (where  ${\sf R}$ is defined in Eq. \ref{eq:flat_sht}), it is easy to see that the key to work out the expressions for the pure-$E$ and $B$ coupling matrices is simply to replace all factors of ${\rm c}$ and ${\rm s}$ with $k^2/\ell^2$ and 0 respectively in Equations \ref{eq:m02_flat} and \ref{eq:m22_flat}. Again, under the assumption that purification takes place after contaminant deprojection, the only modification to the expression for the deprojection bias is to make sure that all functions of ${\bf l}$ in Eq. \ref{eq:dpbias_flat} are consistently purified (but not any of the other fields appearing there or in Eqs. \ref{eq:dp.g.aux_flat} and \ref{eq:dp.f.aux_flat}).

% ----------------------------------------------------------------------

  \section{Code structure}\label{sec:structure}
    All of the methods described above have been implemented in a software package called \nmt\footnote{The code source is hosted in \url{https://github.com/damonge/NaMaster}, and its documentation can be found at \url{http://namaster.readthedocs.io/}.}. We briefly describe the code structure here.

    The code is written in C and wrapped into {\tt python} to facilitate its use and its combination with other software libraries for astronomy. \nmt~itself is based on a number of these libraries:
    \begin{itemize}
      \item {\tt HEALPix} \citet{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations. Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citet{1999ASPC..172..487P}.
      \item The basic full-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citet{2013A&A...554A.112R}.
      \item In flat sky, SHTs are replaced by 2D DFTs, for which \nmt~uses the {\tt FFTW} library~\citet{FFTW05}.
    \end{itemize}

    As described in the previous sections, some of the operations (e.g. SHTs, deprojection, purification) take place in each field individually, while others depend on the correlation of a pair of fields (e.g. PCL, mode-coupling matrix, bias deprojection). It is therefore inefficient to carry out the per-field operations on every field every time a power spectrum is computed, especially if a large number of fields are being cross-correlated. For this reason, at the \texttt{python} level, \nmt~is structured around two main classes that incorporate the two different types of operations we just described. Each of these classes are associated to a counterpart structure in C:

    \paragraph*{NmtField.} This class stores all the necessary information about one individual observd spin-0 or spin-2 field. A {\tt NmtField} object is defined by a sky mask or weights map (e.g. $v$ in Eq. \ref{eq:aobs}), 1 or 2 sky maps (depending on the spin), corresponding to an observation of the field on the sky (${\bf a}^{\rm obs}$ in Eq. \ref{eq:aobs}), a set of contaminant templates (${\bf f}_i$ in Eq. \ref{eq:aobs}), and the choice to purify or not the $E$ or $B$ mode component of the field. Once initialized, this class carries out the following operations:
    \begin{itemize}
      \item SHT or DFT of the field (Eqs. \ref{eq:sht1} or \ref{eq:dft}).
      \item Contaminant deprojection (Eq. \ref{eq:clean}).
      \item $E$-mode and/or $B$-mode purification (Eq. \ref{eq:pureb}).
    \end{itemize}
    \todo{check if any other functionality is worth reporting here.}

    \paragraph*{NmtWorkspace.} This class stores the information necessary to compute an unbiased estimate of the power spectrum of two {\tt NmtField}s. The main objective of these objects is to compute mode-coupling matrices. {\tt NmtWorkspace}s are equipped with read/write methods to avoid redoing these calculations when estimating the power spectra of several fields with a common set of masks (e.g. for a large number of tomographic bin or simulations).

    \nmt~also includes other convenience classes to handle bandpowers, covariance matrices etc. as well as routines to carry out useful operations (e.g. mask apodization, Gaussian simulations) and to wrap up commonly used sequences of operations (e.g. pseudo-$C_\ell$ computation followed by binning and deconvolution).

    The workflow for a typical pseudo-$C_\ell$ run would be:
    \begin{lstlisting}
from pymaster import *

...

f2=NmtField(mask,[map_q,map_u],
            templates=[[dust_q,dust_u]],
            purify_b=True,beam=beam_ell)
f0=NmtField(mask,[map_t])

b=NmtBin(nside,nlb=10)

wsp=NmtWorkspace()
wsp.compute_coupling_matrix(f0,f2,b)

cl_bias=deprojection_bias(f0,f2,cl_theory)

cl_coupled=compute_coupled_cell(f0,f2)

cl_decoupled=wsp.decouple_cell(cl_coupled,
                               cl_bias)
    \end{lstlisting}
    Briefly:
    \begin{enumerate}
      \item In line 1 we import \nmt's python module.
      \item All preliminary I/O operations (reading maps, masks etc.) take place implicitly in line 3.
      \item In line 5 we generate a field with spin 2. The field is defined in terms of a mask and maps of the $(Q,U)$ Stokes parameters. We also provide a set of contaminants to deproject, an instrumental beam and request for the field's $B$-modes to be purified. Deprojection and purification will take place at this stage in that order, as explained in Section \ref{ssec:maths.pureb}. The field's SHT is also estimated upon initialization.
      \item We define a second field in line 8. This one is a simpler spin-0 field with no contaminants or beam.
      \item We define the output bandpower structure in line 10. In this case we will use bins of 10 multipoles, although more general schemes are supported.
      \item In lines 12 and 13 we use an {\tt NmtWorkspace} object to compute the mode-coupling matrix.
      \item The deprojection bias is computed in line 15 using a best-guess for the true underlying power spectrum.
      \item The pseudo-$C_\ell$ is computed in line 17.
      \item The deconvolved bandpowers are finally estimated using the workspace in line 19.
    \end{enumerate}

    Further details regarding the different features of \nmt~can be found in the online documentation \url{http://namaster.readthedocs.io/}.

  \section{Code performance and validation} \label{sec:validation}
    The core of our code-validation suite is based on two science examples. These examples are not meant to be realistic mocks of actual data, however, they do contain the dynamic range and noise-levels of a typical experimental set-up that we expect in the next decade, while providing us with an exact, known true power spectrum to compare against. This section presents these two examples and the results of this validation in terms of accuracy and computational performance. Results will be presented for both full-sky and flat-sky realizations.

    \subsection{Validation suite} \label{ssec:validation.suite}
      We base our validation suite on two science cases of relevance for the most relevant next-generation cosmology experiments. We describe these here.

      \subsubsection{Galaxy clustering and weak lensing} \label{sssec:validation.suite.lss}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_lss}
          \caption{Signal and noise power spectra used in the LSS validation set. Red, green and blue curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode. All other cross-correlations are zero. The dashed lines show the noise power spectra, associated to shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
        \end{figure}
        \begin{figure*}
          \centering
          \includegraphics[width=0.9\textwidth]{maps_lss_sph}
          \caption{Inputs for the large-scale structure, full-sky validation set. From top to bottom and left to right: sky mask (1), Gaussian realization of the galaxy overdensity $\delta$ (2) and cosmic shear $\gamma_i$ (3,4), dust contaminant for $\delta$ (5), star contaminant for $\delta$ (6), PSF contaminant for $\gamma$ (7, $Q$ component) and small-scale contaminant for $\gamma$ (8, $Q$ component).}
          \label{fig:lss_mocks_sph}
        \end{figure*}
        \begin{figure*}
          \centering
          \includegraphics[width=0.9\textwidth]{maps_lss_flat}
          \caption{Same as Fig. \ref{fig:lss_mocks_sph} for the large-scale structure, flat-sky validation set.}
          \label{fig:lss_mocks_flat}
        \end{figure*}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_cont_lss}
          \caption{{\sl Top:} signal power spectrum (red) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (green) and stars (blue). {\sl Middle:} $E$-mode lensing power spectrum (red) and the spectra of the PSF and small-scale contaminants used in our validation set (green and blue respectively). {\sl Bottom:} same as middle panel for the $B$-mode power spectrum. Note that in this case the input signal power spectrum is $0$. In all cases the yellow line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
        \end{figure}
        The first validation suite exemplifies the use of pseudo-$C_\ell$ algorithms in the analysis of large-scale structure data. Current and next-generation imaging surveys such as DES~\citet{2016MNRAS.460.1270D}, KiDS~\citet{2017A&A...604A.134D}, HSC~\citet{2018PASJ...70S...4A}, LSST~\citet{2008arXiv0805.2366I} and Euclid~\citet{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing. These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \todo{cite 3x2pt}.

        Large-scale structure data are often characterized\js{brit: characterised} by inhomogeneous sky coverages and complex mask structures, associated with the presence of bright stars and observational artifacts. They are also affected by many sky contaminants, associated to Galactic sources (dust absorption, star density) and observing conditions (seeing, airmass etc.)~\citet{2016ApJS..226...24L}. The typically flatter power spectra for both galaxy clustering and lensing (e.g. when compared with the damping tail of CMB temperature fluctuations), makes the pseudo-$C_\ell$ approach ideal for this kind of analysis~\citet{2013MNRAS.435.1857L,2016ApJS..226...24L}. This science case therefore allows us to validate several aspects of the code simultaneously:
        \begin{itemize}
          \item The unbiased reconstruction of the 6 possible correlations between spin-0 and spin-2 fields.
          \item The inclusion of inhomogeneous noise and the correction of the associated noise bias.
          \item The effects of deprojection for several contaminant templates in different-spin fields.
          \item The code's ability to handle complex masks and high-resolution pixelization.
        \end{itemize}

        For this validation suite we therefore use a mock dataset mimicking a typical observation of an LSST-like survey in a particular redshift bin. Slightly different set-ups are used for the full-sky and flat-sky validation suites. The characteristics of these data are as follows:
        \begin{itemize}
          \item {\bf Signal:} we generate maps of the galaxy overdensity $\delta$ and of the shear field $(\gamma_1,\gamma_2)$ for a redshift bin centred at $z\simeq1$ with a width $\Delta z\sim0.1$ and Gaussian photo-$z$ tails with width $\sigma_z=0.06$. For $\delta_g$ we assume unit bias $b(z)=1$ and no contributions from redshift-space distortions or magnification. For $\gamma$, we assume no contribution from intrinsic alignments. The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library \todo{cite if paper is out} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}. These power spectra are shown in Fig. \ref{fig:lss_cls}. Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the full-sky and flat-sky cases respectively.
          \item {\bf Mask:} for the full-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}. To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask. We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citet{2018PASJ...70S...4A}. The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales. As can be seen in the top-left panel of Figure~\ref{fig:lss_mocks_flat}, the mask is not binary, and its value at each pixel correspond to the pixel's fractional masked area.
          \item {\bf Contaminants:} we contaminate the signal maps described above with several different types of residuals. Maps of these are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig. \ref{fig:lss_cont_cls} for the full-sky case.
          \begin{itemize}
            \item {\sl Dust.} For the full-sky suite, we include a linear contaminant for $\delta_g$ proportional to the fluctuations around the mean of the dust reddening map of~\citet{1998wfsc.conf..297S}. This contamination affects mostly the largest scales, and simulates the effect of an imperfect correction of the dust absorption on the galaxy number density. The amplitude of the dust fluctuations is chosen so that their power spectrum within the unmasked region be $\sim20\%$ of the galaxy power spectrum at $\ell\sim120$. For the flat-sky suite, we generate a flat-sky Gaussian realization of a field with a power spectrum $C_\ell\propto\ell^{-2.4}$, corresponding to the large-$\ell$ behaviour of the full-sky dust map. The amplitude was chosen such that the contaminant power spectrum be $\sim10\%$ of the signal at $\ell\sim400$.
            \item {\sl Stars.} We add a second contaminant to $\delta_g$ associated with star contamination. For the full-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citet{2014SPIE.9150E..14C}. We modify this contaminant slightly by adding a white-noise component that dominates the small-scale spectrum beyond $\ell\simeq400$. We do this in order to simulate the noise-like distribution of stars on small scales as well as to include a contaminant source that, unlike dust, can surpass the signal on small scales. The amplitude of this contaminant is chosen such that its power spectrum is about 3\% of the signal at $\ell\sim400$.
            \item {\sl PSF fluctuations.} We add a large-scale contaminant for weak lensing in the form of fluctuations in the image point-spread function (PSF) with a characteristic scale corresponding to the telescope field of view. We generate this contaminant as a spin-2 field with equal $E$ and $B$-mode amplitude and a power spectrum
            \begin{equation}
              C^{EE}_\ell=C^{BB}_\ell\propto\left[2\frac{J_1(\ell\theta_{\rm FoV})}{\ell\theta_{\rm FoV}}\right]^2,
            \end{equation}
            where $J_1(x)$ is the order-1 Bessel function, and $2J_1(\ell\theta)/\ell\theta$ is the Fourier transform of a circular aperture with radius $\theta$. For this we assume $\theta_{\rm FoV}=1.75^\circ$. The amplitude of the power spectrum is chosen so that the contaminant amounts to $30\%$ of the signal power spectrum at $\ell\sim50$. We use the same prescription for both the full-sky and the flat-sky cases. 
            \item {\sl Small-scale contamination.} Finally we include an additional small-scale contaminant to weak lensing with a flat power spectrum, unit $E$-$B$ ratio, and an amplitude fixed so the contaminant's power spectrum is $20\%$ of the signal at $\ell=500$. Although we do not identify this contaminant with any specific source, it could arise from e.g. the impact of stars on shape measurement \todo{does this make sense?} \js{I guess that if you throw out sources that are blended above certain threshold and consider that most blends are going to be purely random, this will have a shot-noise-like white power-spectrum? So, I guess this makes sense.}.
          \end{itemize}
          \item {\bf Noise:} for the full-sky case we generate inhomogeneous noise realizations. To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is then intrinsic ellipticity scatter per component. These noise realizations are then locally rescaled by a factor $f(\nv)=f_{\rm star}(\nv)\,f_{\rm depth}(\nv)$ made up of two components:
          \begin{itemize}
            \item $f_{\rm star}(\nv)$ is a function of the local star density modeled with the same star map used to generate the galaxy clustering contaminant described above. The function interpolates between $f_{\rm star}=1$ for low star densities and $f_{\rm star}=1.5$ for high star densities. The specific functional form used is
            \begin{equation}
              f_{\rm star}=\frac{r_{\rm star}-4}{2r_{\rm star}-4},
            \end{equation}
            where $r_{\rm star}=N_{\rm star}/{\rm max}(N_{\rm star})$ and $N_{\rm star}$ is the star density map.
            \item $f_{\rm depth}(\nv)$ is a map of stochastic depth variations generated as a Gaussian random field with mean 1 and fluctuations with a typical amplitude of 3-4\% and a characteristic scale of $\theta_{\rm FoV}=1.75^\circ$.
          \end{itemize}
          The logic behind these simulated noise maps is that both the presence of stars and fluctuations in the survey depth will modulate the number density of observed objects, and with it the noise in both $\delta$ and $\gamma$. A more optimal estimate of the power spectrum could therefore be achieved by using an inverse-variance weighting scheme in which the mask is inversely proportional to $f(\nv)$.

          For simplicity, the flat-sky simulations use only the homogeneous part of these noise realizations.

          \item {\bf Pixelization: \js{brit: pixellation}} the main suite of full-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$. This resolution allows us to validate the code on a large number of realizations without facing strong computational challenges on the main scales of interest ($\ell\lesssim2000$).  As noted in Appendix~\ref{app:full_sky_nside}, however, the code has also been validated with a smaller suite of high-resolution simulations with $N_{\rm side}=4096$, and we have been able to run it successfully up to a maximum $N_{\rm side}=8192$ ($\theta_{\rm pix}\sim0.4'$). The flat-sky simulations use square pixels of $0.6$ arcmin a side.
        \end{itemize}

        We should note that, unlike the simulated maps used here, the true galaxy distribution is markedly non-Gaussian. However, none of the expressions presented in Section \ref{sec:maths} for the pseudo-$C_\ell$ method assume Gaussianity, and therefore the method will deliver unbiased estimates of the harmonic-space two-point function for non-Gaussian fields too.

      \subsubsection{CMB $B$-modes from the ground} \label{sssec:validation.suite.cmb}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_cmb}
          \caption{$E$-mode and $B$-mode power spectra (blue and yellow respectively) for the cosmological CMB signal (solid lines), for the full-sky foreground contaminants (dashed lines) and for the flat-sky foreground contaminants (dot-dashed lines). The noise power spectra used for the full- and flat-sky simulations are shown as black dashed and dot-dashed lines respectively. These noise curves are intended to depict two different types of experiments: a small-aperture telescope (S.A., dashed lines) targetting primordial tensor modes, and a large-aperture telescope (L.A., dot-dashed lines) targetting the small-scale lensing $B$-modes.} \label{fig:cmb_cls}
        \end{figure}
        \begin{figure*}
          \centering
          \includegraphics[width=0.9\textwidth]{maps_cmb}
          \caption{{\sl Left panels:} mask (top), and $Q$ and $U$ components (middle, bottom) of the foreground contaminant used in the CMB full-sky validation set. {\sl Right panels:} same as left panels for the flat-sky validation set.}
          \label{fig:cmb_mocks}
        \end{figure*}
        Our second example corresponds to the measurement of $B$-modes in the polarized CMB emission, closely associated with the presence of primordial gravitational waves. This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citet{2016ApJ...833..228B}, Advanced ACTPol~\citet{2018JLTP..tmp..133K} or the Simons Array~\citet{2016JLTP..184..805S}\todo{Check this citation}, as well as next-generation facilities such as the Simons Observatory\todo{cite} \js{I can't find a reference for this. Maybe just cite website?} or CMB Stage-4~\citet{2016arXiv161002743A}. Potential future space missions, such as LiteBIRD~\citet{2018JLTP..tmp..124S} will be able to make low-resolution, full-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods. Ground-based experiments, on the other hand, will be systematics-limited at low $\ell$, and will be able to make higher-resolution maps, for which pseudo-$C_\ell$ methods are better suited~\citet{2017MNRAS.465.1847E}.

        The measurement of CMB $B$ modes is affected by two challenges: the presence of foreground residuals left in the data by any component separation method, and the much larger amplitude of the $E$-mode signal, which can dominate the uncertainties in the $B$-mode power spectrum due to $E$-to-$B$ leakage. This science case therefore allows us to validate the code when both contaminant deprojection and $B$-mode purification are required, as well as in the presence of a non-negligible instrumental beam.

        Our first validation suite is therefore inspired on a Southern-Hemisphere, high-sensitivity polarization experiment with the following characteristics:
        \begin{itemize}
          \item {\bf Signal:} we generate $Q$ and $U$ maps of the CMB as Gaussian realization of a lensed power spectrum corresponding to the best-fit parameters from Planck~\citet{2016A&A...594A..13P}. For simplicity we assume no delensing and zero primordial $B$-modes, so that the $B$-mode power spectrum is entirely due to the effects of CMB lensing. The signal power spectra are shown in Fig. \ref{fig:cmb_cls}.
          \item {\bf Mask:} for the full-sky case we assume a survey targeting a 4000 deg$^2$ footprint. The corresponding mask is analytically apodized \js{I have seen apodised too, should we go brit?} using the ``C1'' apodization method of \todo{cite}\js{can't find any C1 apodization but I found coefficients for an apodization function: \url{http://atmos.difa.unibo.it/LabFA/M/MATERIALE/TECNICHE_INVERSIONE/01_Norton-Beer_Apodization_functions_JOSA_1976.pdf}} with an apodization scale of 10 degrees. The observed area encompasses one of the lowest-foreground regions observable from the Southern Hemisphere. For the flat-sky case we assume a smaller sky fraction of 500 deg$^2$ with a similar apodization. The assumed masks are shown in the top panels of Figure~\ref{fig:cmb_mocks}.
          \item {\bf Noise:} for the full-sky and flat-sky case we assume white noise levels of 1 and 0.5$\mu K\,{\rm arcmin}$ in intensity respectively. We also include a $1/f$-like component associated with atmospheric noise with a characteristic scale $\ell_{\rm knee}$ and tilt $\alpha_{\rm knee}$ such that the noise power spectrum can be written as
          \begin{equation}
            N_\ell=\sigma_N^2\left[1+\left(\frac{\ell_{\rm knee}}{\ell}\right)^{\alpha_{\rm knee}}\right],
          \end{equation}
          where $\sigma_N^2$ is the white-noise variance in one steradian. We use $\alpha_{\rm knee}=2.4$ and $\ell_{\rm knee}=10$ and 300 for the full-sky and flat-sky cases respectively. The beam-deconvolved noise power spectra in both cases are shown in Fig. \ref{fig:cmb_cls}.
          \item {\bf Beam:} we assume a Gaussian beam with full-width at half-maximum of $\theta_{\rm FWHM}=20$ and 1.4 arcmin for the full-sky and flat-sky cases respectively. This resolution corresponds to what could be achieved at 150 GHz with with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case. The choice of $\ell_{\rm knee}$ described above for the noise curves is also consistent with the atmospheric noise levels achievable with those types of telescopes \todo{cite Simons Observatory paper}. As shown in Fig. \ref{fig:cmb_cls}, the combination of beam size and $\ell_{\rm knee}$ makes the two experiments sensitive to the $B$-mode power spectrum in different ranges of scales, corresponding to the primordial $B$-mode signal around the recombination bump (S.A.) and to the high-ell lensing signal (L.A.).
          \item {\bf Contaminants:} we assume a dust-like residual for both the full-sky an flat-sky cases. For the full-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citet{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$. For the flat-sky signal we generate a Gaussian realization of a dust-like component with power spectra
          \begin{equation}
            C_\ell^{EE}\propto\ell^{-2.5},\hspace{12pt}
            C_\ell^{BB}\propto\ell^{-2.3},
          \end{equation}
          where the spectral tilts were estimated from the high-$\ell$ behaviour of the {\tt PySM} maps in the low-foreground footprint. The amplitudes of these power spectra were fixed by requiring that the $B$-mode power be about 20\% of the lensing $B$-mode power spectrum at $\ell=550$ and so that $C_\ell^{EE}/C_\ell^{BB}\simeq2$, as found in \citet{2016A&A...594A..13P}. The foreground residual maps we use are shown in the bottom panels of Figure~\ref{fig:cmb_mocks}, and their power spectra are compared with the signal and noise spectra in Fig. \ref{fig:cmb_cls}.
          \item {\bf Pixelization:} the full-sky case, targetting the $B$-mode recombination bump at $\ell\sim100$, uses a HEALPix resolution parameter $N_{\rm side}=256$, corresponding to $14$-arcmin pixels and sufficient to measure power spectra to multipoles $\ell\sim500$. The flat-sky case, targetting the lensing $B$-modes, uses square 2-arcmin pixels. These correspond to scales $\ell\lesssim5000$, well within the signal-dominated regime given the noise level and beam sizes quoted above.
        \end{itemize}

    \subsection{Full-sky large-scale structure validation}\label{ssec:validation.lss_full}
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph}
        \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
        \caption{{\sl Left panel:} average over all simulations in our full-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines). The analytical prediction for the input signal power spectra are shown as dashed lines, and lie mostly below the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$). The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero. The bottom subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero and their fluctuations are well within those expected under the assumption of Gaussian statistics given 1000 independent realizations (see Fig. \ref{fig:lss_val_chi2}). {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
      \end{figure*}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our full-sky LSS validation suite (blue histograms). Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot. The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=XXX$ degrees of freedom (denoted by the dashed vertical lines). The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see lower half of the left panel in Fig. \ref{fig:lss_val_cls}). The associated PTE values are $\gtrsim0.1$, and thus no statistically significant bias is found by our validation suite.} \label{fig:lss_val_chi2}
      \end{figure*}
      We generate 1000 realizations of contaminated maps of the galaxy overdensity $\delta$ and cosmic shear $\gamma$ using the models described in Section \ref{sssec:validation.suite.lss} with resolution $N_{\rm side}=1024$. In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations. For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E.B}$), incuding map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum. We use band-powers defined by averaging over groups of $\Delta\ell=XX$ multipoles.

      The upper half in the left panel of Fig. \ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra. The lower half of the same figure shows the average residual with respect to the analytical prediction normalize by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$). We see that the latter are accurately recovered by \nmt~in all cases, including all null correlations ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$), and that the normalized residuals fluctuate around zero with typical variations of $\Delta C_\ell/\sigma_\ell\lesssim3$, corresponding to expected $\lesssim3\sigma$ fluctuations.

      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_cmb_sph}
        \includegraphics[width=0.99\columnwidth]{val_cmb_bpw}
        \caption{{\sl Left:} average over all simulations in our full-sky CMB validation suite of the three spin-2 power spectra ($EE$, $EB$ and $BB$, solid lines). The analytical prediction for the input signal power spectra are shown as dashed lines, and line mostly below the solid lines for the non-zero cases ($EE$ and $BB$). The null power spectrum $EB$ is compatible with zero. The bottom panel of this figure shows the mean residuals with respect to the input power spectra normalized by the $1\sigma$ error on the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero with random fluctuations of up to $\sim3\sigma$. {\sl Right:} exact bandpower window functions for the fully accounting for the effects of mode-coupling (solid black) compared with the input top-hat bandpower weights (red). Using the correct window functions is in general important in order to avoid parameter biases. Results are shown for the $BB$ power spectrum.} \label{fig:cmb_val_cls}
      \end{figure*}
      \begin{figure}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_sigmas_cmb_sph}
        \caption{1$\sigma$ errors on the $BB$, $EB$ and $EE$ power spectra (solid, dashed and dot-dashed lines respectively) for the standard pseudo-$C_\ell$ estimator (blue) and for the estimator with purified $B$ modes (red). $B$-mode purification is able to produce significant gains in sensitivity for the $B$-mode power spectrum, particularly at low $\ell$.}\label{fig:cmb_val_sigma}
      \end{figure}
      To quantify the presence of any residual bias in the estimator we start by computing, for each simulation and cross-power spectrum in the validation suite, the quantity:
      \begin{equation}
        \chi^2_i=({\bf s}_i-{\bf t})^T\,{\sf Cov}^{-1}\,({\bf s}_i-{\bf t}),
      \end{equation}
      where ${\bf s}_i$ is the vector of $N_{\rm dof}$ power spectrum values estimated in the $i$-th simulation, ${\bf t}$ is the analytical prediction for the input power spectrum, and ${\sf Cov}$ is the covariance matrix of the estimator, which we compute by averaging over all simulations:
      \begin{equation}
        {\sf Cov}=\frac{1}{N_{\rm sim}}\sum_{i=1}^{N_{\rm sim}}({\bf s}_i-\bar{\bf s})({\bf s}_i-\bar{\bf s})^T,\hspace{6pt} \bar{\bf s}\equiv\frac{1}{N_{\rm sim}}\sum_{i=1}^{N_{\rm sim}}{\bf s}_i.
      \end{equation}
      The full covariance matrix is very close to diagonal, as shown in the right panel of Fig. \ref{fig:lss_val_cls}. Under the assumption that the estimator is unbiased ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$ should follow a $\chi^2$ distribution with a $N_{\rm dof}$ degrees of freedom. Figure \ref{fig:lss_val_chi2} shows the distribution of $\chi^2_i$ across simulations (blue histograms), compared with the predicted $\chi^2$ distribution for $N_{\rm dof}=XXX$ (marked by the vertical dashed lines) for the six different power spectrum combinations. Both distributions are found to be compatible. Moreover, the dot-dashed lines show the values of $\chi^2$ found for the mean residual with respect to the input power spectra (lower panel of Fig. \ref{fig:lss_val_cls}). The probability to exceed (PTE) associated with these $\chi^2$ values are higher than $10\%$ in all cases, and therefore our validation suite does not show any statistically significant bias in the estimator. This bias, if any, must be at least $\sqrt{1000}\sim30$ times smaller than the statistical uncertainties associated with our validation suite, which are representative of the statistical power of next-generation galaxy surveys.

      \todo{List of more research-y things we could include:
      \begin{itemize}
        \item Present the statistical significance of the bias if we don't deproject.
        \item Present the statistical significance of the bias if we deproject but don't correct the deprojection bias.
        \item Show size of error bars using different masks (binary vs. inverse-noise-variance weights).
      \end{itemize}}

    \subsection{Full-sky CMB \texorpdfstring{$B$}{B}-mode validation}\label{ssec:validation.cmb_full}
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_covar_cmb_sph_nocont}
        \includegraphics[width=0.99\columnwidth]{val_covar_cmb_sph}
        \caption{{\sl Left:} covariance matrix for power spectra computed in the absence of foreground contaminants computed from the 1000 simulations of our full-sky CMB validation suite. The incomplete sky coverage causes a noticeable anti-correlation of neighboring bandpowers at low $\ell$ for the three power spectra ($EE$, $EB$ and $BB$), but beyond that, the covariance matrix is close to diagonal. {\sl Right:} the same as the left panel but this time for simulations containing foreground contaminants. The combination of $B$-mode purification and contaminant deprojection produces noticeable correlations at low $\ell$, especially in the $BB$ spectrum, that must be taken into account.} \label{fig:cmb_val_cov}
      \end{figure*}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_cmb_sph}
        \caption{Distribution of $\chi^2$ values found for the 1000 simulations in our full-sky CMB validation suite (blue histograms). Results are shown for the three polarized spectra ($EE$, $EB$ and $BB$), labeled in the top right corner of each panel. The simulated distributions agree well with the expected $\chi^2$ distribution (solid line) assuming Gaussian statistics for $N_{\rm dof}=43$ degrees of freedom (denoted by the dashed vertical lines). The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see lower half of the left panel in Fig. \ref{fig:cmb_val_cls}). The associated PTE values are $\gtrsim0.4$, and thus no statistically significant bias is found by our validation suite.} \label{fig:cmb_val_chi2}
      \end{figure*}
      We have carried out a similar set of tests on an ensemble of 1000 simulations from the CMB validation suite described in Section \ref{sssec:validation.suite.cmb}. In our fiducial run we both deprojected the foreground contaminants and subsequently used $B$-mode purification to decouple $E$ and $B$-modes at the map level, since this is the main feature of the code we want to validate. The left panel of Figure \ref{fig:cmb_val_cls} shows the mean of the $EE$, $EB$ and $BB$ power spectra computed for all simulations (solid lines) compared with the expectation from the input spectra for $EE$ and $BB$ (dashed lines, mostly lying below the solid ones). The average residuals normalized over the error on the mean, shown in the lower panel of the same Figure, fluctuate around zero with an amplitude of up to $3\sigma$, as would be expected in the absence of a bias in the estimator.

      Figure \ref{fig:cmb_val_sigma} illustrates the impact of $B$-mode purification on the variance of the pseudo-$C_\ell$ estimator. The Figure shows the diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra. Although there is no practical improvement in the error bars for the $EE$ spectrum (there is even a slight increase due to the loss of the ambiguous $E$ modes to purification), and the improvement is only mild for the $EB$ case, very significant gains can be obtained for the $BB$ power spectrum, reducing uncertainties by up to a factor of $\sim10$ on large scales.
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_flat}
        \includegraphics[width=0.99\columnwidth]{val_cl_cmb_flat}
        \caption{{\sl Left and Right:} the analogues of the left panels of Figures \ref{fig:lss_val_cls} and \ref{fig:cmb_cls} for the flat-sky LSS and CMB validation sets respectively.} \label{fig:val_cls_flat}
      \end{figure*}

      At this point we can use the validation suite to illustrate the relation between the input power spectrum and the theoretical prediction for the de-convolved bandpowers. As described by Equations \ref{eq:theory_correct} and \ref{eq:bpresponse}, the mode-coupling asociated with the sky mask and the subsequent de-coupling into bandpowers performed by inverting the binned mode-coupling matrix implies that the window function ${\cal F}_{q\ell}$ in Eq. \ref{eq:bpresponse} differs in general from a simple binning operator, even if the bandpower weights in Eq. \ref{eq:bpws} are chosen to be top-hat functions with a fixed width $\Delta\ell$. This is illustrated in the right panel of Figure \ref{fig:cmb_val_cls}, which shows, in black, the exact window function for different bandpowers (solid black), in comparison with the input top-hat bandpower weights (solid red). Results are shown for the $BB$ power spectrum. Depending on the specific problem, and on the range of scales under study using the simpler top-hat windows or even simply evaluating the theory power spectra at the bandpower centres may lead to e.g. significant biases in subsequent parameter inference stages. \nmt~provides simple and fast functions to apply the exact window functions to theory power spectra.

      As in the case of the LSS validation set, we also study the distribution of $\chi^2$ values from each simulation in order to quantify the presence of any possible bias. Unlike in the LSS suite, we find that in this case the covariance matrix is noticeably non-diagonal, and cannot be replaced by the diagonal errors when computing the $\chi^2$. The non-zero off-diagonal elements are caused by two effects. Firstly, the limited footprint area, map apodization and choice of bandpowers leads to a noticeable level of anti-correlation between neighboring bandpowers at low $\ell$. This can be seen in the left panel of Figure \ref{fig:cmb_val_cov}, which shows the covariance matrix of simulations without foreground contaminants or deprojection. Then, introducing contaminant deprojection produces additional correlations that affect the lowest bandpowers and also involves a level of cross-talk between the three different cross-spectra.

      The distribution of $\chi^2$ values found from the simulations is shown in Fig. \ref{fig:cmb_val_chi2} for the three power spectra. As before, the vertical dashed lines show the number of degrees of freedom ($N_{\rm dof}=43$), while the  dot-dashed line marks the $\chi^2$ value of the mean residuals (bottom panel of Fig. \ref{fig:cmb_cls}). The associated PTE values are all above $40\%$ and therefore, as in the case of the LSS validation set, we do not detect any significant bias in the estimator when including both contaminant deprojection and $B$-mode purification.

    \subsection{Flat-sky validation}\label{ssec:validation.flat}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_lss_flat}
        \includegraphics[width=0.80\textwidth]{val_chi2_cmb_flat}
        \caption{{\sl Top and bottom:} the analogues of the Figures \ref{fig:lss_val_chi2} and \ref{fig:cmb_val_chi2} for the flat-sky LSS and CMB validation sets respectively.} \label{fig:val_chi2_flat}
      \end{figure*}

      In order to validate the flat-sky features of \nmt, we follow the same procedure used in Sections \ref{ssec:validation.lss_full} and \ref{ssec:validation.cmb_full} for the LSS and CMB validation suites respectively.

      Before we discuss these results, a technical point, specific to the flat-sky case, must first be addressed. The theory prediction for the pseudo-$C_\ell$ method is given by Eq. \ref{eq:predflat} (before multiplying by the inverse binned mode-coupling matrix, Eq. \ref{eq:modcoup_flat}). This involves a convolution of the power spectra interpolated into the discretized Fourier plane with the un-binned mode-coupling matrix. This is an $\mathcal{O}(N_{\rm pix}^2)$ operation that would be too expensive to perform at every step of a Monte-Carlo chain, and therefore \nmt~uses a different approximate approach. This consists of first interpolating the theory power spectrum into the discrete Fourier plane, followed by an averaging of the interpolated power spectrum into a set of radial rings with a width given by the minimum wavenumber probed (given by ${\rm min}(2\pi/L_x,2\pi/L_y)$), and spanning all possible values of $|{\bf l}|$. This binned power spectrum, an object of size $\sim\sqrt{N_{\rm pix}}$, is then convolved with a high-resolution version of the mode-coupling matrix. This high-resolution matrix is given by a variation of Eq. \ref{eq:modcoup_flat} in which $S_{q'}$ represents all the values of the wavenumber ${\bf k}$ in one of the narrow rings described above (instead of the final wide bandpowers). The convolved power spectrum is then multiplied by the inverse binned coupling matrix (given by Eq. \ref{eq:modcoup_flat} exactly) to produce the final theoretical prediction. As we show below, this procedure, which is $\mathcal{O}(N_{\rm pix})$, is able to reproduce the exact theoretical prediction with a negligible error in all cases explored here.

      Figure \ref{fig:val_cls_flat} shows, with solid lines, the mean over all simulations of the 6 LSS power spectra and the 3 polarized CMB power spectra in the left and right panels respectively. The analytical predictions for the non-zero input power spectra are shown as dashed lines in both cases, and agree with the simulation mean almost perfectly. The mean residuals normalized by the 1$\sigma$ error on the mean are shown in the lower half of both figures, and feature fluctuations with amplitudes smaller than $\sim3$-$4\sigma$.

      Again, we quantify the presence of a bias in the estimator by studying the distribution of the $\chi^2$ values across simulations. In analogy with Figures \ref{fig:lss_val_chi2} and \ref{fig:cmb_val_chi2}, these are shown in the upper and lower panels of figure \ref{fig:val_chi2_flat} for the LSS and CMB suites respectively. In all cases the distributions (blue histograms) agree well with the expected $\chi^2$ distribution (solid lines) for the appropriate number of degrees of freedom (vertical dashed lines). The $\chi^2$ values of the mean residuals shown in the lower panels of Fig. \ref{fig:val_cls_flat} and marked by vertical dot-dashed lines in Fig. \ref{fig:val_chi2_flat} have PTE values above $10\%$. We therefore do not observe any significant bias in the estimator, proving \nmt's usefulness in the analysis of current and next-generation experiments in situations requiring the use of the flat-sky approximation.

  \section{Discussion}\label{sec:discussion}
    The pseudo-$C_\ell$ power spectrum estimator is one particular example of quadratic minimum-variance estimators as presented by \todo{cite tegmark} in which the proposal covariance matrix of the data is assumed to be diagonal in real space. This allows us to significantly reduce the computational complexity of the problem, from ${\cal O}(N_{\rm pix}^3)$ to ${\cal O}(N_{\rm pix}^{3/2})$ through both analytical and numerical simplifications, as described in Section \ref{sec:maths}. A number of techniques that are naturally included in the standard quadratic estimator, such as mode deprojection and $B$-mode purification, can also be implemented in the standard pseudo-$C_\ell$ algorithm without altering the computational complexity of the method. Although this speed boost is achieved at the expense of optimality (in terms of the estimator's variance), the degradation in sensitivity will be small or negligible as long as the data does not deviate significantly from the diagonal assumption. This implies that the pseudo-$C_\ell$ approach is a good choice for non-steep power spectra for which the typical correlation length is smaller than the mask structure, and in general on small scales, where the impact of the power spectrum shape drops as the number of available modes increases \todo{cite stuff}. As we have argued, this will be the case for a wide range of relevant science cases pursued by next-generation cosmological observations.

    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method including all its generalizations (flat-sky, deprojection and purification). We have described the steps taken to validate the code, showing that it is able to provide unbiased results in a wide variety of scenarios. In particular, we have constructed a realistic validation suite that resembles the type of data, masks and contaminants that will affect future wide-area optical galaxy surveys and ground-based CMB polarization experiments. As shown in Section \ref{sec:validation}, we are able to recover the input power spectrum in all cases (which include varying degrees of contamination, purification, different spins etc.) with no detectable bias down to $\lesssim1/30$-th of the statistical uncertainties.

    Although \nmt~is intended to make power spectrum estimation as easy as possibe, some tasks are too dependent upon the properties of the data to be analyized to automatize them in a general way, and users are therefore in charge of them. We list the most important of these here:
    \begin{itemize}
      \item \nmt~operates directly on {\sl maps}. Galaxy survey data, however, are normally provided in the form of object catalogs, and the conversion those into the desired maps (e.g. galaxy overdensity or shear) must be done externally. 
      \item $E/B$ purification can only be carried out in the context of pseudo-$C_\ell$ estimators when the mask is differentiable up to its second derivatives, which can usually be achieved with an appropriate apodization. Although \nmt~provides tools to carry our this type of operations, some of the properties of the input mask (e.g. sharp corners) will in general spoil differentiability even after apodization. Users must therefore take care to inspect the mask derivatives before making use of \nmt's purification feature.
      \item A crucial part of any quadratic power spectrum estimation is the removal of noise bias in the auto-correlation of a given dataset. Although we have provided analytic formulas to estimate this bias in certain simple scenarios (see Section \ref{ssec:maths.beam_noise}), these are often not applicable. Users must therefore make sure that a sufficiently accurate estimate of the noise bias can be achieved (e.g. through simulations) or to use only cross-correlation data.
      \item Contaminant deprojection will be able to provide an accurate estimate of the power spectrum as long as two conditions are met. First, users must be able to provide a comprehensive list of possible contaminants as would be observed in their data (e.g. beam-smoothed, mean-subtracted etc.). Second, a good estimate of the underlying signal power spectra must be provided in order to calculate and subtract the deprojection bias. We have outlined a possible iterative process to do so in Section \ref{ssec:maths.deproj}.
    \end{itemize}

    \nmt~is open-source software and publicly available at \url{add_url}. To maximize the usefulness of this software for the community, it is accompanied by an extensive documentation and example code\footnote{See \url{add_url}.}. \nmt~is written in {\tt C} and {\tt OpenMP}-parallelized to maximize its performance, although we encourage its use through the companion {\tt python} wrapper, the fastest-growing language of choice in the analysis of astronomical data. The code is in constant development, and we will strive to implement further functionality to e.g. estimate covariance matrices, extend the range of supported sky pixelizations and enhance its computational performance.

\subsection*{Acknowledgements}
We would like to thank Rupert Allison, Jo Dunkley, Franz Elsner, Boris Leistedt, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya Peiris, Naomi Robertson and Ben Thorne for useful comments and discussions. DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.

\input{contributions}

\input{desc-tex/ack/standard}

\bibliography{main}

\appendix

  \section{Generalities and Spherical Harmonic Transforms}\label{app:shts}
    Let $\eth$ and $\bar{\eth}$ be the following complex differential operators defined on the sphere when acting on a spin-$s$ quantity $f_s$:
    \begin{equation}
      \begin{split}
        \eth f_s\equiv-(\sin\theta)^s\left(\partial_\theta+i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^{-s}\,f_s(\theta,\varphi), \\
        \bar{\eth} f_s\equiv-(\sin\theta)^{-s}\left(\partial_\theta-i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^{s}\,f_s(\theta,\varphi).
      \end{split}
    \end{equation}
    The following properties can be easily derived for the action of these operators, and are useful to derive some of the formulas presented here:
    \begin{itemize}
      \item If $f_s$ is a spin-$s$ quantity, $(f_s)^*$ is a spin-$(-s)$ quantity.
      \item $\eth f_s$ is a spin-$(s+1)$ quantity, and $\bar{\eth} f_s$ is a spin-$(s-1)$ quantity.
      \item $(\eth^nf_s)^*=\bar{\eth}^n(f_s)^*$
      \item $\eth(f\,g)=f\eth g+g\eth f$
      \item $\eth^2(f\,g)=f\eth^2g+g\eth^2f+\eth f\eth g$
    \end{itemize}

    We start by defining the spin-weighed spherical harmonics with spin $s\geq0$:
    \begin{equation}
      _sY_{\ell m}\equiv \alpha_{\ell,s} \eth^s Y_{\ell m},\hspace{6pt}
      _{-s}Y_{\ell m}\equiv \alpha_{\ell,s} (-1)^s\bar{\eth}^s Y_{\ell m},
    \end{equation}
    where $\alpha_{\ell,s}\equiv\sqrt{(\ell-s)!/(\ell+s)!}$ and $Y_{\ell m}$ are the standard spherical harmonics. These functions satisfy the property: $(_sY_{\ell m})^*=(-1)^{s+m}\,_{-s}Y_{\ell-m}$.

    We can then define the $E$-mode and $B$-mode spherical harmonic vectors as:
    \begin{align}\nonumber
      _s{\bf Y}^E_{\ell m}&\equiv {\bf D}^E_sY_{\ell m}
      \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
                              \eth^s+\bar{\eth}^s\\
                              -i(\eth^s-\bar{\eth}^s)
                            \end{array}\right)Y_{\ell m}\\\label{eq:ederiv}
        &=-\frac{1}{2}\left(\begin{array}{c}
                               _sY_{\ell m}+(-1)^s\,_{-s}Y_{\ell m}\\
                            -i(_sY_{\ell m}-(-1)^s\,_{-s}Y_{\ell m})
                            \end{array}\right) \\\nonumber
          _s{\bf Y}^B_{\ell m}&\equiv {\bf D}^B_sY_{\ell m}
          \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
                              i(\eth^s-\bar{\eth}^s)\\
                              \eth^s+\bar{\eth}^s
                            \end{array}\right)Y_{\ell m}\\\label{eq:bderiv}
        &=-\frac{1}{2}\left(\begin{array}{c}
                             i(_sY_{\ell m}-(-1)^s\,_{-s}Y_{\ell m})\\
                               _sY_{\ell m}+(-1)^s\,_{-s}Y_{\ell m}
                            \end{array}\right),
      \end{align}
      which also defines the differential operators ${\bf D}^{E,B}_s$. For $s=0$, these functions are simply $_0{\bf Y}^{E}_{\ell m}=(Y_{\ell m},0)$ and $_0{\bf Y}^{B}_{\ell m}=(0,Y_{\ell m})$.

      Finally we define the matrix operator $\hat{\sf Y}^s_{\ell m}$ to have $_s{\bf Y}^{E,B}_{\ell m}$ as columns:
      \begin{equation}
        \hat{\sf Y}^s_{\ell m}\equiv\left(_s{\bf Y}^E_{\ell m},_s{\bf Y}^B_{\ell m},\right).
      \end{equation}
      These matrices satisfy the following properties:
      \begin{align}
        &\hat{\sf Y}^{s\dag}_{\ell m}=(-1)^{m+s}\hat{\sf Y}^{-s}_{\ell -m}\\
        &\int d\nv \hat{\sf Y}^s_{\ell m}\hat{\sf Y}^{s\dag}_{\ell'm'}=\hat{\sf 1}\delta_{\ell\ell'}\delta_{mm'},\\
        &\int d\nv \left(\hat{\sf Y}^{s\dag}_{\bf l}(\nv)\hat{\sf Y}^{s}_{{\bf l}_1}(\nv)\right)\,\hat{\sf Y}^0_{{\bf l}_2}(\nv)\equiv\hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2},
      \end{align}
      where we have abbreviated the pair $(\ell,m)$ as ${\bf l}$, and
      \begin{align}\label{eq:dmat}
        \hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2}&=(-1)^{s+m}\sqrt{\frac{(2\ell+1)(2\ell_1+1)(2\ell_2+1)}{4\pi}}\\\nonumber
        &\hspace{10pt}\wtj{\ell}{\ell_1}{\ell_2}{-m}{m_1}{m_2}
        \wtj{\ell}{\ell_1}{\ell_2}{s}{-s}{0}\,\hat{\sf d}_{\ell+\ell_1+\ell_2}.
      \end{align}
      Here
      \begin{equation}
        \hat{\sf d}_n= \frac{1}{2}\left(\begin{array}{cc}
                                            1+(-1)^n & -i[1-(-1)^n]\\
                                            i[1-(-1)^n] & 1+(-1)^n
                                          \end{array}\right),
      \end{equation}
      and the Wigner 3$j$ symbols satisfy the orthogonality relation
      \begin{equation}\label{eq:3jorth}
        \sum_{mm_1}\wtj{\ell}{\ell_1}{\ell_2}{m}{m_1}{m_2}\wtj{\ell}{\ell_1}{\ell_3}{m}{m_1}{m_3}=\frac{\delta_{\ell_2\ell_3}\delta_{m_2m_3}}{2\ell_2+1}
      \end{equation}

  \section{Flat skies and Fourier transforms}\label{app:flat}
    \subsection{Spin-\texorpdfstring{$s$}{s} fields in flat sky}\label{app:flat.fields}
      In the flat sky we will label a position in the plane by two
      coordinates $(x,y)\equiv{\bf x}$. These coordinates can be
      directly related to increments in the spherical coordinates
      $(\theta,\varphi)$ by considering a sufficiently small map
      centered around the equator (i.e. $\sin\theta\sim 1$). In
      this case we identify the $x$ coordinate with a latitude shift
      ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
      shifts ($\delta\varphi=\delta y$) \todo{make a plot of
        this?}. \as{no, but isn't it other way round? Latitute is $y$
        in my brain? :)} Using this dictionary, the differential operator $\eth$ now takes the form:
      \begin{equation}
        \eth= (\partial_x-i\partial_y),\hspace{12pt}\bar{\eth}= (\partial_x+i\partial_y).
      \end{equation}
      The analog of the standard spherical harmonics for flat skies are plane waves $e^{i{\bf k}{\bf x}}$, and the action of $\eth$ on these is:
      \begin{equation}
        \eth^s e^{i{\bf k}{\bf x}}=(ik)^se^{-i\,s\varphi_k}e^{i{\bf k}{\bf x}},\hspace{12pt}
        \bar{\eth}e^{i{\bf k}{\bf x}}=(ik)^se^{i\,s\varphi_k}e^{i{\bf k}{\bf x}}
      \end{equation}

      In analogy with the full sky case, we start by defining the basis functions:
      \begin{align}
        \,_s{\cal Y}_{\bf k}({\bf x})\equiv k^{-s}\eth^se^{i{\bf k}{\bf x}}=i^s\,e^{-is\varphi_k}e^{i{\bf k}{\bf x}},\\
        _{-s}{\cal Y}_{\bf k}({\bf x})\equiv (-k)^{-s}\bar{\eth}^se^{i{\bf k}{\bf x}}=(-i)^se^{is\varphi_k}e^{i{\bf k}{\bf x}},
      \end{align}
      where $\varphi_k$ is the polar angle of ${\bf k}$. We then define the Fourier coefficients of a spin-$s$ complex field $a({\bf x})$ as:
      \begin{align}\nonumber
        &\,_sa_{\bf l}\equiv\int\frac{d{\bf x}^2}{2\pi}\,_s{\cal Y}^*_{\bf l}({\bf x})a({\bf x}),\hspace{6pt}
        \,_{-s}a_{\bf l}\equiv\int\frac{d{\bf x}^2}{2\pi}\,_{-s}{\cal Y}^*_{\bf l}({\bf x})a^*({\bf x}).\\\label{eq:fft0}
        &a({\bf x})  =\int\frac{d{\bf l}^2}{2\pi}\,_s{\cal Y}_{\bf l}({\bf x})_sa_{\bf l},\hspace{6pt}
        a^*({\bf x})=\int\frac{d{\bf l}^2}{2\pi}\,_{-s}{\cal Y}_{\bf l}({\bf x})_{-s}a_{\bf l}.
      \end{align}
      These are then related to the $E$ and $B$-mode coefficients as:
      \begin{align}
        _sE_{\bf l}&\equiv-\frac{1}{2}\left[_sa_{\bf l}+(-1)^s\,_{-s}a_{\bf l}\right],\\
        i\,_sB_{\bf l}&\equiv-\frac{1}{2}\left[_sa_{\bf l}-(-1)^s\,_{-s}a_{\bf l}\right].
      \end{align}
      Note the preceding $(-)$ sign. For scalar fields ($s\equiv0$) the $E$ and $B$ modes are defined omitting that sign.

      In analogy with our full-sky nomenclature, let us now write $a$ as a vector such that in real space ${\bf a}({\bf x})\equiv({\rm Re}(a),{\rm Im}(a))$, and in Fourier space ${\bf a}_{\bf l}\equiv(\,_sE_{\bf l},\,_sB_{\bf l})$. We can rewrite the Eq. \ref{eq:fft0} in vectorial form:
      \begin{equation}
        {\bf a}({\bf x})\equiv\int\frac{d{\bf l}^2}{2\pi}\,_s{\sf E}_{\bf l}({\bf x}){\bf a}_{\bf l},\hspace{12pt}{\bf a}_{\bf l}\equiv\int\frac{d{\bf x}^2}{2\pi}\,_s{\sf E}^\dag_{\bf l}({\bf x}){\bf a}({\bf x}),
      \end{equation}
      where, in analogy with Eqs. \ref{eq:ederiv} and \ref{eq:bderiv},  we have defined the matrix basis functions:
      \begin{align}\nonumber
        \,_s{\sf E}_{\bf l}
        &\equiv-\frac{1}{2}
        \left[\begin{array}{cc}
                 \,_s{\cal Y}_{\bf l}+(-1)^s\,_{-s}{\cal Y}_{\bf l}  & i(\,_s{\cal Y}_{\bf l}-(-1)^s\,_{-s}{\cal Y}_{\bf l}) \\
              -i(\,_s{\cal Y}_{\bf l}-(-1)^s\,_{-s}{\cal Y}_{\bf l}) &   \,_s{\cal Y}_{\bf l}+(-1)^s\,_{-s}{\cal Y}_{\bf l}
            \end{array}\right]\\\nonumber
        &=-\frac{1}{2l^s}
        \left(\begin{array}{cc}
                 \eth^s+\bar{\eth}^s & i(\eth^s-\bar{\eth}^s) \\
              -i(\eth^s-\bar{\eth}^s) &  \eth^s+\bar{\eth}^s
            \end{array}\right)e^{i{\bf l}{\bf x}}\\\nonumber
        &=-i^s
        \left(\begin{array}{cc}
               \cos(s\varphi_l) & \sin(s\varphi_l) \\
              -\sin(s\varphi_l) & \cos(s\varphi_l)
            \end{array}\right)e^{i{\bf l}{\bf x}}\\\label{eq:flat_sht}
        &=-i^s{\sf R}^\dag(s\varphi_l)e^{i{\bf l}{\bf x}}.
      \end{align}
      Here ${\sf R}(\varphi)$ is a rotation matrix.

    \subsection{Discrete description and DFTs}\label{app:flat.dft}
      As we discuss in \todo{where}, it is more convenient to describe power spectrum estimation methods in a discretized flat sky. Let the patch of the sky under inspection be contained by a rectangle of sides $L_x$ and $L_y$ (in units of radians), and let us discretize this rectangle by dividing it into an $N_x\times N_y$ grid with pixels of area $\Delta{\bf x}^2\equiv\Delta x\Delta y=(L_x/N_x)(L_y/N_y)$. Each pixel in this grid is then labeled by a pair of integers ${\bf n}\equiv(n_x,n_y)$, and is assigned coordinates ${\bf x}_{\bf n}\equiv(n_x\Delta x,n_y\Delta y)$. Each field in the pixelized map ${\bf a}({\bf x})$ is therefore defined for $n_x\in[0,N_x-1]$, $n_y\in[0,N_y-1]$.

      In this case, the spin Fourier transform (Eq. \ref{eq:flat_sht}) of the pixelized field can be computed as its discrete Fourier transform (DFT):
      \begin{equation}\label{eq:dft}
        {\bf a}_{\bf k}\equiv {\cal D}\left({\bf a}\right)^{s_a}_{\bf k}\equiv\sum_{\bf x}\frac{\Delta{\bf x}^2}{2\pi}\,_{s_a}{\sf E}^\dag_{\bf k}({\bf x}){\bf a}_{\bf x},
      \end{equation}
      where the wavenumber ${\bf k}$ is now discretized as ${\bf k}=(j_x\Delta k_x,j_y\Delta k_y)$, the integers $j_{(x,y)}$ run from $-N_{(x,y)}/2$ to $N_{(x,y)}/2-1$\footnote{This is the valid domain when $N_{(x,y)}$ is even. For odd $N_{(x,y)}$ the interval becomes $[-(N_{(x,y)}-1)/2,(N_{(x,y)}-1)/2]$.} and the pixel size is $\Delta k_{(x,y)}\equiv2\pi/L_{(x,y)}$.

      The following properties of the DFT are worth recalling:
      \begin{itemize}
       \item Periodicity:
             \begin{align}\nonumber
               {\bf a}_{(k_x,k_y)}&={\bf a}_{(k_x+N_x\Delta k_x,k_y)}\\\nonumber&={\bf a}_{(k_x,k_y+N_y\Delta k_y)}\\&={\bf a}_{(k_x+N_x\Delta k_x,k_y+N_y\Delta k_y)}.
             \end{align}
       \item For a real-valued scalar $a$, its DFT satisfies $a^*_{{\bf k}_{\bf j}}=a^{}_{{\bf k}_{{\bf N}-{\bf j}}}$ (where ${\bf N}\equiv(N_x,N_y)$).
       \item The orthogonality relation of the basis functions now takes the form:
             \begin{equation}
               \sum_{\bf x}\,_sE^\dag_{\bf l}({\bf x})\,_sE_{\bf k}({\bf x})=N_xN_y\delta_{{\bf l},{\bf k}}\hat{\sf 1}.
             \end{equation}
       \item The power spectrum of a pixelized field is defined as:
             \begin{equation}
               \langle{\bf a}_{\bf l}{\bf b}^\dag_{\bf k}\rangle\equiv\frac{\delta_{{\bf l},{\bf k}}}{\Delta{\bf k}^2}{\sf C}^{ab}_{\bf l}.
             \end{equation}
      \end{itemize}

  \section{Flat-sky pseudo-\texorpdfstring{$C_\ell$}{Cl}s in the continuum limit}\label{app:flat.pcl}
    In the continuum limit, the Fourier coefficients of the masked field (see Eq. \ref{eq:pcl_flat_masked}) are:
    \begin{align}\nonumber
      {\bf a}^v_{\bf l}&=\int\int\frac{d{\bf k}^2d{\bf q}^2}{2\pi}\left[\int\frac{d{\bf x}^2}{(2\pi)^2}\,_{s_a}{\sf E}^\dag_{\bf l}({\bf x})\,_{s_a}{\sf E}_{\bf k}({\bf x})\,_0E_{\bf q}({\bf x})\right]{\bf a}_{\bf k}\,v_{\bf q}\\
                       &=\int\frac{d{\bf k}^2}{2\pi}{\sf R}\left(s_a\Delta\varphi\right)\,{\bf a}_{\bf k}v_{{\bf l}-{\bf k}},
    \end{align}
    The covariance of the Fourier coefficients of two masked fields is then given by:
    \begin{equation}
      \left\langle{\bf a}^v_{\bf l}\,{\bf b}^{w\dag}_{\bf l}\right\rangle=\int\frac{d{\bf k}^2}{(2\pi)^2}{\sf R}(s_a\Delta\varphi){\sf C}^{ab}_k{\sf R}^\dag(s_b\Delta\varphi)v_{{\bf l}-{\bf k}}w^*_{{\bf l}-{\bf k}}.
    \end{equation}
    Defining the pseudo-$C_\ell$ as the nomalized angular average of ${\bf a}_{\bf l}{\bf b}^\dag_{\bf l}$ we obtain:
    \begin{align}
      \left\langle{\rm PCL}_\ell({\bf a},{\bf b})\right\rangle
      &\equiv\frac{(2\pi)^2}{S}\int\frac{d\varphi_l}{2\pi}\left\langle{\bf a}^v_{\bf l}\,{\bf b}^{w\dag}_{\bf l}\right\rangle\\\nonumber
      &=\int\frac{kdk\,qdq}{(2\pi)^2}\left[\frac{(2\pi)^2}{S}\int\frac{d\varphi_qd\varphi_kd\varphi_l}{2\pi}v_{{\bf q}}w^*_{{\bf q}}\right.\\\nonumber
      &\hspace{12pt}\left.{\sf R}(s_a\Delta\varphi){\sf C}^{ab}_k{\sf R}^\dag(s_b\Delta\varphi)\delta({\bf q}-{\bf l}+{\bf k})\right],
    \end{align}
    where $S$ is the observed sky area (in steradian), and where we have eliminated the dependence on ${\bf l}-{\bf k}$ by introducing an additional integral over $d{\bf q}^2\delta({\bf q}-{\bf l}+{\bf k})$.

    As shown in \cite{2016arXiv161204664A}, these expressions can be simplified through the following steps:
    \begin{enumerate}
     \item Substitute
       \begin{equation}
         \delta({\bf q}-{\bf l}+{\bf k})\rightarrow\int\frac{d{\bf r}^2}{(2\pi)^2}e^{i({\bf q}-{\bf l}+{\bf k}){\bf r}}
       \end{equation}
     \item Integrate over the angular parts of ${\bf r}$, ${\bf l}$ and ${\bf k}$ using the following relation:
       \begin{equation}
        \int_0^{2\pi} d\varphi\,e^{i\,x\,\cos\varphi}e^{i\,n\varphi}=2\pi\,i^n\,J_n(x),
       \end{equation}
       where $J_n$ is the cylindrical Bessel function of order $n$.
     \item Integrate over the angular part of ${\bf q}$, defining the power spectrum of the masks as:
       \begin{equation}
         \tilde{C}^{vw}_q\equiv\frac{(2\pi)^2}{S}\int d\varphi_qv_{\bf q}w^*_{\bf q}
       \end{equation}
     \item Solve the last isolated integral over the radial part of ${\bf r}$ by using the following relation:
       \begin{equation}
         \int_0^\infty dr\,r\,J_0(qr)\,J_n(kr)\,J_n(\ell r)=\frac{\cos n\theta}{\pi k\ell\sin\theta},
       \end{equation}
       where $\theta$ is the angle between sides $\ell$ and $k$ of the triangle formed by three sides of length $q$, $\ell$ and $k$.
     \item Make the change of variables $q^2(\ell,k,\theta)\equiv\ell^2+k^2-2k\ell\cos\theta$ to simplify the integral over the radial part of ${\bf q}$.
    \end{enumerate}

    This yields the following relation analogous to Eq. \ref{eq:modecoup}:
    \begin{equation}
      {\rm vec}\left[{\sf C}^{ab}_l\right]=\int_0^{\infty} dk\, {\sf M}^{s_as_b}_{\ell k}\cdot\,{\rm vec}\left[{\sf C}^{ab}_k\right],
    \end{equation}
    where:
    \begin{align}
      &{\sf M}^{00}_{\ell k}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\\
      &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\hat{\sf 1},
      \hspace{12pt}
      M^{0+}_{\ell k}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\cos2\theta\\
      &{\sf M}^{22}_{\ell\ell'}=\left(
      \begin{array}{cccc}
        M^{++}_{\ell\ell'} &0 &0&M^{--}_{\ell\ell'}\\
        0&M^{++}_{\ell\ell'} &-M^{--}_{\ell\ell'}&0\\
        0&-M^{--}_{\ell\ell'} &M^{++}_{\ell\ell'}&0\\
        M^{--}_{\ell\ell'} &0 &0&M^{++}_{\ell\ell'}
      \end{array}\right)\\
      &M^{\pm\pm}_{\ell\ell'}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\frac{1\pm\cos4\theta}{2}.
    \end{align}
    These can also be expressed as integrals over $q$:
    \begin{equation}
      M^{XY}_{\ell k}=\frac{k}{2\pi}\int\frac{q\,dq}{2\pi}\tilde{C}^{vw}_qF(\ell,k,q)G^{XY}(\ell,k,q),
    \end{equation}
    where $F(\ell,k,q)=0$ if $\ell$, $k$ and $q$ do not form a triangle, and
    \begin{equation}\nonumber
      F(\ell,k,q)\equiv\frac{4}{\sqrt{2\ell^2k^2+2k^2q^2+2q^2\ell^2-\ell^4-k^4-q^4}}
    \end{equation}
    otherwise. The functions $G^{XY}$ are given by
    \begin{align}
      &G^{00}\equiv1,\\
      &G^{0+}(\ell,k,q)\equiv\frac{\ell^4+k^4+q^4-2k^2q^2-2\ell^2q^2}{2k^2\ell^2},\\
      &G^{++}(\ell,k,q)\equiv\left[\frac{\ell^4+k^4+q^4-2k^2q^2-2\ell^2q^2}{2k^2\ell^2}\right]^2,\\
      &G^{--}(\ell,k,q)\equiv1-G^{++}(\ell,k,q).
    \end{align}

    Since in the flat-sky limit $\ell$ is a continuous variable, bandpowers ${\sf B}^{ab}_q$ can be defined as averages over a given interval in $\ell$, $[\ell_q^{\rm min},\ell_q^{\rm max}]$:
    \begin{equation}
      {\sf B}^{ab}_q\equiv \int_{\ell_q^{\rm min}}^{\ell_q^{\rm max}} \frac{d\ell}{\ell_q^{\rm max}-\ell_q^{\rm min}}\,{\sf C}^{ab}_\ell.
    \end{equation}
    The binned coupling matrix is therefore given by:
    \begin{equation}
      \mathcal{M}^{s_as_b}_{qq'}\equiv\int^{\ell_q^{\rm max}}_{\ell_q^{\rm min}} \frac{d\ell}{\ell_q^{\rm max}-\ell_q^{\rm min}}\int^{\ell_{q'}^{\rm max}}_{\ell_{q'}^{\rm min}} d\ell'\,{\sf M}^{s_as_b}_{\ell\ell'}.
    \end{equation}

\todo{link this paragraph}    The main complication in using the pseudo-$C_\ell$ formalism in the continuum limit is the need to compute the angle-averaged mask pseudo power spectrum $C^{vw}_\ell$. Flat fields are most easily analyzed when pixellated in a Cartesian grid, and under this setup the mask power spectrum is not a well defined quantity for infinitesimally small intervals of $\ell$. This leads to non-negligible biases and poor performance \cite{2016arXiv161204664A} due to the need to use highly resolved finite intervals to compute the integrals presented in the previous section. This motivates the discrete formalism described here, which connects directly with the data storage format of flat-sky maps.

\section{Full-sky validation for different pixel scales}
\label{app:full_sky_nside}
We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of $N_{side}$. We show here some examples for 100 realisations of $N_{side}=128$ and $N_{side}=4096$.

% Don't change these lines
\bsp	% typesetting comment

\end{document}

% ======================================================================

### 2018-07-30 doc/Paper/main.tex
added missing references

@@
   \section{Introduction}\label{sec:intro}
-    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citet{1995PhRvL..74.4369B,1997PhRvD..55.5895T} \todo{More cites?}.
-Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier/harmonic-space power spectra $C_\ell$ \todo{cite one of the angular xcorr papers}.
-Although correlation functions are often easier to estimate from complex observations~\citet{1993ApJ...412...64L}, a far higher level of data compression and decorrelation can be obtained through $C_\ell$s.
+    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}.
+Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier/harmonic-space power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.
+Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a far higher level of data compression and decorrelation can be obtained through $C_\ell$s.
 This, together with other advantages associated with power spectra (e.g.
-clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citet{2008PhRvD..77j3013H} \todo{cite Sellentin} has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators.
-
-    While maximum-likelihood~\citet{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citet{1997PhRvD..55.5895T,2000MNRAS.317L..23H} can recover the power spectrum with virtually no loss of information, their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ and $\ell_{\rm max}^4$ respectively with the maximum multipole $\ell_{\rm max}$.
-For this reason, pseudo-$C_\ell$ algorithms \citet{1973ApJ...185..413P,2002ApJ...567....2H}\todo{cite more everywhere} have become a popular alternative that preserves a the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
-This method can be used on fields with arbitrary spin \citet{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citet{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $B$-mode power spectrum \citet{2009PhRvD..79l3515G}.
+clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}) has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators.
+
+    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information, their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ and $\ell_{\rm max}^4$ respectively with the maximum multipole $\ell_{\rm max}$.
+For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves a the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
+This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
 Unfortunately, in spite of the advantages of this method, no public, validated and well-maintained codes exist supporting all of these capabilities.
@@
     The paper is structured as follows: Section \ref{sec:maths} introduces the pseudo-$C_\ell$ estimator in the curved sky, including mode deprojection and $E/B$ purification.
+This is then generalized in the flat-sky approximation.
 In Section \ref{sec:structure} we briefly describe the implementation of these methods on \nmt, as well as its structure and usage.
 Section \ref{sec:validation} presents the stress tests used to validate the code.
-We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicatiblity in typical use-cases scenarios in both fields.
-This section also studies the speed and scaling.
+We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicatiblity in typical user-case scenarios in both fields.
 We conclude in Section \ref{sec:discussion}.
+The appendices provide further details about the mathematical background used throughout the paper, and present some additional validation tests.
 
@@
 
-        In the standard incarnation of the optimal quadratic estimator~\citet{1998PhRvD..57.2117B}\todo{Cite others?}, the measurement is produced on inverse covariance weighted data, namely
+        In the standard incarnation of the optimal quadratic estimator~\citep{1998PhRvD..57.2117B}, the measurement is produced on inverse covariance weighted data, namely
         \begin{equation}
@@
 If these conditions are not met, the pseudo-$C_\ell$ estimator will still be unbiased (by construction), but it will get progressively less optimal.
-In practical surveys, however, the loss of optimality rarely exceeds 10-20\% and is typically smaller~\citet{2013MNRAS.435.1857L}.
+In practical surveys, however, the loss of optimality rarely exceeds 10-20\% and is typically smaller~\citep{2013MNRAS.435.1857L}.
 
@@
         \end{equation}
-        where we have defined the vectorization operator \citet{2008PhRvD..77j3013H}
+        where we have defined the vectorization operator \citep{2008PhRvD..77j3013H}
         \begin{equation}
@@
 Usually, we know something about these contaminants.
-As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which we might want to fit and marginalise out.
+As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which can be marginalized out.
 But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening.
@@
       where the observed map ${\bf a}^{\rm obs}$ is composed of a masked signal map that receives contributions from $N_a$ known contaminants ${\bf f}^i$ and where the contamination coefficients $\alpha_i$ are unknown (note that we have implicitly applied the same mask to the contaminant templates ${\bf f}^i$).
-A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citet{2013MNRAS.435.1857L}.
+The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}.
+A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}.
 Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
@@
 
-      \as{I would move this procedure together with full procedure to Section 3.}
       The procedure to obtain an unbiased estimator of the power spectrum in the presence of contaminants can therefore be summarized as follows:
@@
     \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
-      As described in \cite{2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa.
+      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa.
 The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB.
@@
   \section{Code structure}\label{sec:structure}
-    All of the methods described above have been implemented in a software package called \nmt\footnote{The code source is hosted in \url{https://github.com/damonge/NaMaster}, and its documentation can be found at \url{http://namaster.readthedocs.io/}.}.
+    All of the methods described above have been implemented in a software package called \nmt\footnote{The code source is hosted in \url{add_url}, and its documentation can be found at \url{http://namaster.readthedocs.io/}.}.
 We briefly describe the code structure here.
@@
     \begin{itemize}
-      \item {\tt HEALPix} \citet{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations.
-Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citet{1999ASPC..172..487P}.
-      \item The basic full-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citet{2013A&A...554A.112R}.
-      \item In flat sky, SHTs are replaced by 2D DFTs, for which \nmt~uses the {\tt FFTW} library~\citet{FFTW05}.
+      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations.
+Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citep{1999ASPC..172..487P}.
+      \item The basic full-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citep{2013A&A...554A.112R}.
+      \item In flat sky, SHTs are replaced by 2D DFTs, for which \nmt~uses the {\tt FFTW} library~\citep{FFTW05}.
     \end{itemize}
@@
     \end{itemize}
-    \todo{check if any other functionality is worth reporting here.}
 
@@
 Current and next-generation imaging surveys such as DES~\citet{2016MNRAS.460.1270D}, KiDS~\citet{2017A&A...604A.134D}, HSC~\citet{2018PASJ...70S...4A}, LSST~\citet{2008arXiv0805.2366I} and Euclid~\citet{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing.
-These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \todo{cite 3x2pt}.
-
-        Large-scale structure data are often characterized\js{brit: characterised} by inhomogeneous sky coverages and complex mask structures, associated with the presence of bright stars and observational artifacts.
+These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
+
+        Large-scale structure data are often characterized by inhomogeneous sky coverages and complex mask structures, associated with the presence of bright stars and observational artifacts.
 They are also affected by many sky contaminants, associated to Galactic sources (dust absorption, star density) and observing conditions (seeing, airmass etc.)~\citet{2016ApJS..226...24L}.
@@
 Although we do not identify this contaminant with any specific source, it could arise from e.g.
-the impact of stars on shape measurement \todo{does this make sense?} \js{I guess that if you throw out sources that are blended above certain threshold and consider that most blends are going to be purely random, this will have a shot-noise-like white power-spectrum?
-So, I guess this makes sense.}.
+the impact of stars or blending on shape measurement.
           \end{itemize}
@@
 
-          \item {\bf Pixelization: \js{brit: pixellation}} the main suite of full-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$.
+          \item {\bf Pixelization:} the main suite of full-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$.
 This resolution allows us to validate the code on a large number of realizations without facing strong computational challenges on the main scales of interest ($\ell\lesssim2000$).
@@
         Our second example corresponds to the measurement of $B$-modes in the polarized CMB emission, closely associated with the presence of primordial gravitational waves.
-This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citet{2016ApJ...833..228B}, Advanced ACTPol~\citet{2018JLTP..tmp..133K} or the Simons Array~\citet{2016JLTP..184..805S}\todo{Check this citation}, as well as next-generation facilities such as the Simons Observatory\todo{cite} \js{I can't find a reference for this.
-Maybe just cite website?} or CMB Stage-4~\citet{2016arXiv161002743A}.
-Potential future space missions, such as LiteBIRD~\citet{2018JLTP..tmp..124S} will be able to make low-resolution, full-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
-Ground-based experiments, on the other hand, will be systematics-limited at low $\ell$, and will be able to make higher-resolution maps, for which pseudo-$C_\ell$ methods are better suited~\citet{2017MNRAS.465.1847E}.
+This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citep{2016PhRvL.116c1302B}, Advanced ACTPol~\citep{2016SPIE.9910E..14D,2017JCAP...06..031L} or the Simons Array~\citep{2016JLTP..184..805S,2017ApJ...848..121P}, as well as next-generation facilities such as the Simons Observatory\todo{cite if out} or CMB Stage-4~\citep{2016arXiv161002743A}.
+Potential future space missions, such as LiteBIRD~\citep{2018JLTP..tmp..124S} will be able to make low-resolution, full-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
+Ground-based experiments, on the other hand, will be systematics-limited at low $\ell$, and will be able to make higher-resolution maps, for which pseudo-$C_\ell$ methods are better suited.
 
@@
         \begin{itemize}
-          \item {\bf Signal:} we generate $Q$ and $U$ maps of the CMB as Gaussian realization of a lensed power spectrum corresponding to the best-fit parameters from Planck~\citet{2016A&A...594A..13P}.
+          \item {\bf Signal:} we generate $Q$ and $U$ maps of the CMB as Gaussian realization of a lensed power spectrum corresponding to the best-fit parameters from Planck~\citep{2016A&A...594A..13P}.
 For simplicity we assume no delensing and zero primordial $B$-modes, so that the $B$-mode power spectrum is entirely due to the effects of CMB lensing.
@@
           \item {\bf Mask:} for the full-sky case we assume a survey targeting a 4000 deg$^2$ footprint.
-The corresponding mask is analytically apodized \js{I have seen apodised too, should we go brit?} using the ``C1'' apodization method of \todo{cite}\js{can't find any C1 apodization but I found coefficients for an apodization function: \url{http://atmos.difa.unibo.it/LabFA/M/MATERIALE/TECNICHE_INVERSIONE/01_Norton-Beer_Apodization_functions_JOSA_1976.pdf}} with an apodization scale of 10 degrees.
+The corresponding mask is analytically apodized using the ``$C^2$'' apodization method of \cite{2009PhRvD..79l3515G} with an apodization scale of 10 degrees.
 The observed area encompasses one of the lowest-foreground regions observable from the Southern Hemisphere.
@@
 This resolution corresponds to what could be achieved at 150 GHz with with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case.
-The choice of $\ell_{\rm knee}$ described above for the noise curves is also consistent with the atmospheric noise levels achievable with those types of telescopes \todo{cite Simons Observatory paper}.
+The choice of $\ell_{\rm knee}$ described above for the noise curves is also consistent with the atmospheric noise levels achievable with those types of telescopes \todo{cite Simons Observatory paper if out}.
 As shown in Fig.
@@
   \section{Discussion}\label{sec:discussion}
-    The pseudo-$C_\ell$ power spectrum estimator is one particular example of quadratic minimum-variance estimators as presented by \todo{cite tegmark} in which the proposal covariance matrix of the data is assumed to be diagonal in real space.
+    The pseudo-$C_\ell$ power spectrum estimator is one particular example of quadratic minimum-variance estimators as presented by \cite{1997PhRvD..55.5895T} in which the proposal covariance matrix of the data is assumed to be diagonal in real space.
 This allows us to significantly reduce the computational complexity of the problem, from ${\cal O}(N_{\rm pix}^3)$ to ${\cal O}(N_{\rm pix}^{3/2})$ through both analytical and numerical simplifications, as described in Section \ref{sec:maths}.
@@
 Although this speed boost is achieved at the expense of optimality (in terms of the estimator's variance), the degradation in sensitivity will be small or negligible as long as the data does not deviate significantly from the diagonal assumption.
-This implies that the pseudo-$C_\ell$ approach is a good choice for non-steep power spectra for which the typical correlation length is smaller than the mask structure, and in general on small scales, where the impact of the power spectrum shape drops as the number of available modes increases \todo{cite stuff}.
+This implies that the pseudo-$C_\ell$ approach is a good choice for non-steep power spectra for which the typical correlation length is smaller than the mask structure, and in general on small scales, where the impact of the power spectrum shape drops as the number of available modes increases \citep{2004MNRAS.349..603E,2006MNRAS.370..343E}.
 As we have argued, this will be the case for a wide range of relevant science cases pursued by next-generation cosmological observations.

### 2018-07-30 doc/Paper/main.tex
link styles

@@
 \RequirePackage{docswitch}
-% \flag is set by the user, through the makefile:
-%    make note
-%    make apj
-% etc.
 \setjournal{\flag}
 
-\documentclass[\docopts]{\docclass}
-
-% You could also define the document class directly
-%\documentclass[]{emulateapj}
+\documentclass[usenatbib]{mnrasb}
 
@@
 \usepackage{lsstdesc_macros}
-
 \usepackage{graphicx}
-\graphicspath{{./}{./figures/}}
-\bibliographystyle{apj}
 
@@
 }
+%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
 \newcommand{\nv}{\hat{\bf n}}
@@
 
-% ======================================================================
-
 \begin{document}
@@
 
-% Keywords are ignored in the LSST DESC Note style:
 \dockeys{methods: data analysis -- methods: numerical}
@@
 \maketitlepost
-
-% ----------------------------------------------------------------------
-% {% raw %}
 
@@
 \ref{eq:dp.g.aux_flat} and \ref{eq:dp.f.aux_flat}).
-
-% ----------------------------------------------------------------------
 
@@
             \item {\sl Stars.} We add a second contaminant to $\delta_g$ associated with star contamination.
-For the full-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citet{2014SPIE.9150E..14C}.
+For the full-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citep{2014SPIE.9150E..14C}.
 We modify this contaminant slightly by adding a white-noise component that dominates the small-scale spectrum beyond $\ell\simeq400$.
@@
 
+\bibliographystyle{mnrasb}
 \bibliography{main}
@@
 
-% Don't change these lines
-\bsp	% typesetting comment
+\bsp
 
 \end{document}
-
-% ======================================================================

### 2018-07-30 sandbox_validation/check_sph_pure.py
show effects of deprojection bias

@@ -24,6 +24,10 @@ parser.add_option('--plot', dest='plot_stuff', default=False, action='store_true
                   help='Set if you want to produce plots')
 parser.add_option('--wo-pureb', dest='wo_pureb', default=False, action='store_true',
                   help='Set if you don\'t want to purify B-modes')
+parser.add_option('--no-deproject',dest='no_deproject',default=False,action='store_true',
+                  help='Set if you will include contaminants but won\'t clean them')
+parser.add_option('--no-debias',dest='no_debias',default=False,action='store_true',
+                  help='Set if you will include contaminants, clean them but won\'t correct for the bias')
 (o, args) = parser.parse_args()
 
 nsims=o.isim_end-o.isim_ini+1
@@ -34,6 +38,10 @@ w_pureb=not o.wo_pureb
 predir="tests_sph"
 os.system("mkdir -p "+predir)
 prefix=predir+"/run_pure0%d_ns%d_cont%d"%(w_pureb,o.nside_out,w_cont)
+if o.no_deproject :
+    prefix+="_no_deproj"
+if o.no_debias :
+    prefix+="_no_debias"
 
 #Read theory power spectra
 def read_cl_camb(fname) :
@@ -89,8 +97,12 @@ def get_fields() :
                         o.nside_out,new=True,verbose=False,pol=True)
     if w_cont :
         sq+=np.sum(fgp,axis=0)[0,:]; su+=np.sum(fgp,axis=0)[1,:]
-        ff2=nmt.NmtField(mask,[sq,su],templates=fgp,beam=beam,
-                         purify_e=False,purify_b=w_pureb,n_iter_mask_purify=10)
+        if o.no_deproject :
+            ff2=nmt.NmtField(mask,[sq,su],beam=beam,
+                             purify_e=False,purify_b=w_pureb,n_iter_mask_purify=10)
+        else :
+            ff2=nmt.NmtField(mask,[sq,su],templates=fgp,beam=beam,
+                             purify_e=False,purify_b=w_pureb,n_iter_mask_purify=10)
     else :
         ff2=nmt.NmtField(mask,[sq,su],beam=beam,
                          purify_e=False,purify_b=w_pureb,n_iter_mask_purify=10)
@@ -118,7 +130,7 @@ if not os.path.isfile(prefix+"_clb22.npy") :
     #Compute noise bias
[…]

### 2018-07-30 sandbox_validation/plot_sph_pure.py

@@ -12,6 +12,7 @@ nsims=1000
 prefix_pure="tests_sph/run_pure01_ns%d_cont1"%nside
 prefix_nopu="tests_sph/run_pure00_ns%d_cont1"%nside
 prefix_noco="tests_sph/run_pure01_ns%d_cont0"%nside
+prefix_nodb="tests_sph/run_pure01_ns%d_cont1_no_debias"%nside
 
 def tickfs(ax,x=True,y=True) :
     if x :
@@ -32,6 +33,7 @@ print "Reading"
 clEE_pure=[]; clEB_pure=[]; clBB_pure=[];
 clEE_nopu=[]; clEB_nopu=[]; clBB_nopu=[];
 clEE_noco=[]; clEB_noco=[]; clBB_noco=[];
+clEE_nodb=[]; clEB_nodb=[]; clBB_nodb=[];
 for i in np.arange(nsims) :
     ll,ccee,cceb,ccbe,ccbb=read_cls(prefix_pure+"_cl_%04d.txt"%(i+1))
     clEE_pure.append(ccee); clEB_pure.append(cceb); clBB_pure.append(ccbb);
@@ -39,9 +41,12 @@ for i in np.arange(nsims) :
     clEE_nopu.append(ccee); clEB_nopu.append(cceb); clBB_nopu.append(ccbb);
     ll,ccee,cceb,ccbe,ccbb=read_cls(prefix_noco+"_cl_%04d.txt"%(i+1))
     clEE_noco.append(ccee); clEB_noco.append(cceb); clBB_noco.append(ccbb);
+    ll,ccee,cceb,ccbe,ccbb=read_cls(prefix_nodb+"_cl_%04d.txt"%(i+1))
+    clEE_nodb.append(ccee); clEB_nodb.append(cceb); clBB_nodb.append(ccbb);
 clEE_pure=np.array(clEE_pure); clEB_pure=np.array(clEB_pure); clBB_pure=np.array(clBB_pure); 
 clEE_nopu=np.array(clEE_nopu); clEB_nopu=np.array(clEB_nopu); clBB_nopu=np.array(clBB_nopu);
 clEE_noco=np.array(clEE_noco); clEB_noco=np.array(clEB_noco); clBB_noco=np.array(clBB_noco);
+clEE_nodb=np.array(clEE_nodb); clEB_nodb=np.array(clEB_nodb); clBB_nodb=np.array(clBB_nodb);
 
[…]

### 2018-07-31 doc/Paper/main.tex
show effects of deprojection bias
Added 4096 validation, corrected typos

@@
 Section \ref{sec:validation} presents the stress tests used to validate the code.
-We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicatiblity in typical user-case scenarios in both fields.
+We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicability in typical user-case scenarios in both fields.
 We conclude in Section \ref{sec:discussion}.
@@
 Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices $\partial {\sf C}/\partial B_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
-The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particuarly in order to estimate the Fisher matrix.
+The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix.
 This motivates the pseudo-$C_\ell$ estimator, which replaces these computationally very intensive steps with numerically more efficient ones at the expense of optimality.
@@
 \as{I think this is not quite true.
-This latter way of looking at it, in effect replaces the true power spectrum with its observed quantity, which is fine in practice, but conceptually not quite right, for the same raeason e.g.
+This latter way of looking at it, in effect replaces the true power spectrum with its observed quantity, which is fine in practice, but conceptually not quite right, for the same reason e.g.
 error on power spectrum should be based on real rather than measured power as illustrated by low CMB quadrupole.
@@
     \begin{itemize}
-      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations.
+      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixellation scheme supported for full-sky calculations.
 Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citep{1999ASPC..172..487P}.
@@
 
-    \paragraph*{NmtField.} This class stores all the necessary information about one individual observd spin-0 or spin-2 field.
+    \paragraph*{NmtField.} This class stores all the necessary information about one individual observed spin-0 or spin-2 field.
 A {\tt NmtField} object is defined by a sky mask or weights map (e.g.
@@
           \item The effects of deprojection for several contaminant templates in different-spin fields.
-          \item The code's ability to handle complex masks and high-resolution pixelization.
+          \item The code's ability to handle complex masks and high-resolution pixellation.
         \end{itemize}
@@
 Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot.
-The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=XXX$ degrees of freedom (denoted by the dashed vertical lines).
+The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=681$ degrees of freedom (denoted by the dashed vertical lines).
 The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see lower half of the left panel in Fig.
@@
 In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations.
-For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E.B}$), incuding map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum.
-We use band-powers defined by averaging over groups of $\Delta\ell=XX$ multipoles.
+For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E.B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum.
+We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles.
 
@@
 The null power spectrum $EB$ is compatible with zero.
+For comparison, the red dot-dashed line shows the result of not correcting for the mode-deprojection bias when estimating the $B$-mode power spectrum, which leads to a strong bias on small $\ell$.
 The bottom panel of this figure shows the mean residuals with respect to the input power spectra normalized by the $1\sigma$ error on the mean (i.e.
@@
 Under the assumption that the estimator is unbiased ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$ should follow a $\chi^2$ distribution with a $N_{\rm dof}$ degrees of freedom.
-Figure \ref{fig:lss_val_chi2} shows the distribution of $\chi^2_i$ across simulations (blue histograms), compared with the predicted $\chi^2$ distribution for $N_{\rm dof}=XXX$ (marked by the vertical dashed lines) for the six different power spectrum combinations.
+Figure \ref{fig:lss_val_chi2} shows the distribution of $\chi^2_i$ across simulations (blue histograms), compared with the predicted $\chi^2$ distribution for $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the six different power spectrum combinations.
 Both distributions are found to be compatible.
@@
 
-    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method including all its generalizations (flat-sky, deprojection and purification).
+    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalisations (flat-sky, deprojection and purification).
 We have described the steps taken to validate the code, showing that it is able to provide unbiased results in a wide variety of scenarios.
@@
 
-    Although \nmt~is intended to make power spectrum estimation as easy as possibe, some tasks are too dependent upon the properties of the data to be analyized to automatize them in a general way, and users are therefore in charge of them.
+    Although \nmt~is intended to make power spectrum estimation as easy as possible, some tasks are too dependent upon the properties of the data to be analyzed to automatize them in a general way, and users are therefore in charge of them.
 We list the most important of these here:
@@
 \label{app:full_sky_nside}
-We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of $N_{side}$. We show here some examples for 100 realisations of $N_{side}=128$ and $N_{side}=4096$.
-
+We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of $N_{side}$.
+In this appendix we show that we recover the correct power spectra using the inputs and foregrounds from Section \ref{sssec:validation.suite.lss}.
+In particular we tested $N_{side}$ ranging from 8 to 8192.
+For the latter, the computational resources needed make very difficult to generate and analyze a large enough number of simulations to have a good statistical measurement.
+However, we generate 100 realisations of $N_{side}=4096$ and show the results in Figure~\ref{fig:validation.fullsky.lss.4096}.
+We can see that the power spectra is recovered without any noticeable biases in the expected range of validity $\ell \leq 2 N_{side}$.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\columnwidth]{val_cl_lss_sph_4096}
+\caption{Input (continuous darker lines) and measured (broken semi-transparent lines) power spectra using maps with \texttt{HEALPix} $N_{side}=4096$ using the inputs and foregrounds presented in Section \ref{sssec:validation.suite.lss}.}
+\label{fig:validation.fullsky.lss.4096}
+\end{figure}
 \bsp

### 2018-07-31 doc/Paper/main.tex
minor changes, added new results, homogenized to American English

@@
         The problem we are attempting to solve is how to estimate ${\sf C}^{ab}_\ell$ from a single realization of ${\bf a}$ and ${\bf b}$ measured on a cut sky.
-In general, the input data for measuring the power spectrum come pixelated on a sphere.
+In general, the input data for measuring the power spectrum come pixelized on a sphere.
 Instead of dealing with an infinitely sampled points, we have a finite number of measurements
@@
     \begin{itemize}
-      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixellation scheme supported for full-sky calculations.
+      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations.
 Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citep{1999ASPC..172..487P}.
@@
           \item The effects of deprojection for several contaminant templates in different-spin fields.
-          \item The code's ability to handle complex masks and high-resolution pixellation.
+          \item The code's ability to handle complex masks and high-resolution pixelization.
         \end{itemize}
@@
 \ref{fig:lss_val_chi2}).
+For comparison, the \todo{solid lines} show the mean relative difference for the power spectra of contaminated maps withouth contaminant deprojection.
 {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
@@
 In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations.
-For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E.B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum.
+For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E,B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum.
 We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles.
@@
 We see that the latter are accurately recovered by \nmt~in all cases, including all null correlations ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$), and that the normalized residuals fluctuate around zero with typical variations of $\Delta C_\ell/\sigma_\ell\lesssim3$, corresponding to expected $\lesssim3\sigma$ fluctuations.
+For comparison, the solid lines in this part of the figure show the mean relative difference for power spectra of contaminated maps without deprojection \todo{comment on this}.
 
@@
 
-      \todo{List of more research-y things we could include:
-      \begin{itemize}
-        \item Present the statistical significance of the bias if we don't deproject.
-        \item Present the statistical significance of the bias if we deproject but don't correct the deprojection bias.
-        \item Show size of error bars using different masks (binary vs. inverse-noise-variance weights).
-      \end{itemize}}
+      Our validation suite also allows us to explore the impact of the sub-optimal inverse-variance weighting of the pseudo-$C_\ell$ estimator.
+The comparison with the uncertainties associated with optimal quadratic estimators has been made before in the literature \citep{2013MNRAS.435.1857L}, and we will not explore this here.
+Instead, we can study the impact on the final uncertainties of using a mask that implements exact inverse-noise weighting (i.e.
+the mask shown in the top left corner of Figure \ref{fig:lss_mocks_sph}, inversely proportional to the number density variations induced by stars and depth fluctuations) against a top-hat mask that weights all pixels equally.
+Repeating our analysis for the latter mask, we observe a negligible increase of $\sim0.5-1\%$ in the diagonal error bars.
+Thus, except in the case of large variations in noise properties (as will be the case for inhomogeneous scanning strategies in CMB experiments, for instance), the exact weighting scheme used with the pseudo-$C_\ell$ estimator is unlikely to produce significantly better uncertainties.
 
@@
 
-    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalisations (flat-sky, deprojection and purification).
+    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalizations (flat-sky, deprojection and purification).
 We have described the steps taken to validate the code, showing that it is able to provide unbiased results in a wide variety of scenarios.
@@
         this?}. \as{no, but isn't it other way round? Latitute is $y$
-        in my brain? :)} Using this dictionary, the differential operator $\eth$ now takes the form:
+        in my brain? :)} \dam{it isn't, have a look here: \url{https://healpix.jpl.nasa.gov/html/intronode12.htm}.} Using this dictionary, the differential operator $\eth$ now takes the form:
       \begin{equation}
@@
 
-\todo{link this paragraph}    The main complication in using the pseudo-$C_\ell$ formalism in the continuum limit is the need to compute the angle-averaged mask pseudo power spectrum $C^{vw}_\ell$.
-Flat fields are most easily analyzed when pixellated in a Cartesian grid, and under this setup the mask power spectrum is not a well defined quantity for infinitesimally small intervals of $\ell$.
+    The main complication in using the pseudo-$C_\ell$ formalism in the continuum limit is the need to compute the angle-averaged mask pseudo power spectrum $C^{vw}_\ell$.
+Flat fields are most easily analyzed when pixelized in a Cartesian grid, and under this setup the mask power spectrum is not a well defined quantity for infinitesimally small intervals of $\ell$.
 This leads to non-negligible biases and poor performance \cite{2016arXiv161204664A} due to the need to use highly resolved finite intervals to compute the integrals presented in the previous section.
-This motivates the discrete formalism described here, which connects directly with the data storage format of flat-sky maps.
-
-\section{Full-sky validation for different pixel scales}
-\label{app:full_sky_nside}
-We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of $N_{side}$.
-In this appendix we show that we recover the correct power spectra using the inputs and foregrounds from Section \ref{sssec:validation.suite.lss}.
-In particular we tested $N_{side}$ ranging from 8 to 8192.
-For the latter, the computational resources needed make very difficult to generate and analyze a large enough number of simulations to have a good statistical measurement.
-However, we generate 100 realisations of $N_{side}=4096$ and show the results in Figure~\ref{fig:validation.fullsky.lss.4096}.
-We can see that the power spectra is recovered without any noticeable biases in the expected range of validity $\ell \leq 2 N_{side}$.
+This motivates the discrete formalism used throughout this paper, which connects directly with the storage format of flat-sky maps.
+
+  \section{Full-sky validation for different pixel resolutions}\label{app:full_sky_nside}
+    We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of the HEALPix resolution parameter $N_{\rm side}$.
+In this appendix we show that we recover the correct power spectra using the inputs from our full-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).
+In particular we tested $N_{\rm side}$ ranging from 8 to 8192.
+For the latter, the computational resources needed make it very challenging to generate and analyze a large enough number of simulations to have a good statistical power.
+However, we generate 100 simulations with $N_{\rm side}=4096$ and show the results in Figure~\ref{fig:validation.fullsky.lss.4096}.
+We can see that the power spectra are recovered without any noticeable biases in the expected range of validity $\ell \lesssim 2 N_{\rm side}$.
 
@@
       \centering
-\includegraphics[width=0.9\columnwidth]{val_cl_lss_sph_4096}
-\caption{Input (continuous darker lines) and measured (broken semi-transparent lines) power spectra using maps with \texttt{HEALPix} $N_{side}=4096$ using the inputs and foregrounds presented in Section \ref{sssec:validation.suite.lss}.}
+      \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
+      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the full-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
       \label{fig:validation.fullsky.lss.4096}
     \end{figure}
+
 \bsp

### 2018-08-01 doc/Paper/main.tex
Modified text to be according to new Figure 7 and added plotting routines to sandbox_validation

@@
         \caption{{\sl Left panel:} average over all simulations in our full-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
-The analytical prediction for the input signal power spectra are shown as dashed lines, and lie mostly below the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
+The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly below the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
 The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero.
-The bottom subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e.
+We include, as well, the average over all simulations of the power spectra when no deprojection is performed shown as the lighter dash-dotted lines.
+The middle subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e.
 the scatter over simulations divided by $\sqrt{N_{\rm sim}}$).
@@
 \ref{fig:lss_val_chi2}).
-For comparison, the \todo{solid lines} show the mean relative difference for the power spectra of contaminated maps withouth contaminant deprojection.
+For comparison, the lower panel show the mean relative difference for the power spectra of contaminated maps without contaminant deprojection.
 {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
@@
 The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=681$ degrees of freedom (denoted by the dashed vertical lines).
-The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see lower half of the left panel in Fig.
+The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see middle panel of the left panel in Fig.
 \ref{fig:lss_val_cls}).
@@
 
-      The upper half in the left panel of Fig.
-\ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra.
-The lower half of the same figure shows the average residual with respect to the analytical prediction normalize by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).
+      The upper plot in the left panel of Fig.
+\ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra (darker dashed lines), and when no deprojection is performed (lighter dash-dotted lines).
+The middle panel of the same figure shows the average residual with respect to the analytical prediction normalize by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).
 We see that the latter are accurately recovered by \nmt~in all cases, including all null correlations ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$), and that the normalized residuals fluctuate around zero with typical variations of $\Delta C_\ell/\sigma_\ell\lesssim3$, corresponding to expected $\lesssim3\sigma$ fluctuations.
-For comparison, the solid lines in this part of the figure show the mean relative difference for power spectra of contaminated maps without deprojection \todo{comment on this}.
+For comparison, bottom panel of the figure show the mean relative difference for power spectra of contaminated maps without deprojection, where we can appreciate that the residuals, $\Delta C'_{\ell}/\sigma'_{\ell}$, appear to not be centered around zero in some cases, i.e., the estimated power-spectra are biased.
 
@@
 Both distributions are found to be compatible.
-Moreover, the dot-dashed lines show the values of $\chi^2$ found for the mean residual with respect to the input power spectra (lower panel of Fig.
+Moreover, the dot-dashed lines show the values of $\chi^2$ found for the mean residual with respect to the input power spectra (middle panel of Fig.
 \ref{fig:lss_val_cls}).

### 2018-08-01 sandbox_validation/plot_sph.py

@@ -1,0 +1,359 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.stats as st
+from matplotlib import rc
+import matplotlib
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+nsims=1000
+rebin_step=-1
+prefix_clean="/global/cscratch1/sd/jsanch87/tests_sph_july_v2/run_ns1024_mask1_cont1_nvar1_apo0.00"
+#prefix_dirty="/global/cscratch1/sd/jsanch87/tests_sph_v2_binary/run_ns1024_mask1_cont1_nvar1_apo0.00"
+prefix_dirty="/global/cscratch1/sd/jsanch87/tests_sph_v2_no_dep_no_binary/run_ns1024_mask1_cont1_nvar1_apo0.00"
+def subsample(a,n):
+   return np.array([a[i:n+i].mean() for i in range(0,len(a),n)])
+
+def tickfs(ax,x=True,y=True) :
+    if x :
+        for tick in ax.xaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+    if y :
+        for tick in ax.yaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+
+def read_cls(fname) :
+    l,ctt,cte,ctb,cee,ceb,cbe,cbb=np.loadtxt(fname,unpack=True);
+    mask = l <= 2048
+    return l[mask],ctt[mask],cte[mask],ctb[mask],cee[mask],ceb[mask],cbe[mask],cbb[mask]
+
+l_th,clTT_th,clTE_th,clTB_th,clEE_th,clEB_th,clBE_th,clBB_th=read_cls(prefix_clean+"_cl_th.txt")
+if rebin_step>0:
+    l_th = subsample(l_th,rebin_step)
+    clTT_th = subsample(clTT_th,rebin_step); clTE_th = subsample(clTE_th,rebin_step); clTB_th = subsample(clTB_th,rebin_step);
[…]

### 2018-08-01 sandbox_validation/plot_sph_4096.py

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import rc
+import os
+import astropy.table
+
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+path = '/global/cscratch1/sd/jsanch87/tests_sph_july2018/'
+
+def tickfs(ax,x=True,y=True) :
+    if x :
+        for tick in ax.xaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+    if y :
+        for tick in ax.yaxis.get_major_ticks():
+            tick.label.set_fontsize(12)
+
+def plot_results(nside,apo,mask,cont,nvar,nrange=1000):
+    files = [os.path.join(path,'run_ns%d_mask%d_cont%d_nvar%d_apo%.2f_cl_%04d.txt' % (nside,mask,cont,nvar,apo,run)) for run in range(0,nrange)]
+    file_th = os.path.join(path,'run_ns%d_mask%d_cont%d_nvar%d_apo%.2f_cl_th.txt' % (nside,mask,cont,nvar,apo))
+    cl00_all = []
+    cl02_all = []
+    cl22_all = []
+    for fname in files:
+        tab = astropy.table.Table.read(fname,format='ascii')
+        cl00_all.append(tab['col2'])
+        cl02_all.append([tab['col3'],tab['col4']])
+        cl22_all.append([tab['col5'],tab['col5'],tab['col6'],tab['col7'],tab['col8']])
+    cl00_all = np.array(cl00_all)
+    cl02_all = np.array(cl02_all)
+    cl22_all = np.array(cl22_all)
+    cols=plt.cm.rainbow(np.linspace(0,1,6))
+    tab_th = astropy.table.Table.read(file_th,format='ascii')
+    cl00_th = tab_th['col2']
+    cl02_th = [tab_th['col3'],tab_th['col4']]
+    cl22_th = [tab_th['col5'],tab_th['col6'],tab_th['col7'],tab_th['col8']]
+    #f1, ax = plt.subplots(1,1,figsize=(11,3))
+    f1 = plt.figure()
+    ax = plt.gca()
[…]

### 2018-08-01 doc/Paper/main.tex
updated fig7 and text

@@
 The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero.
-We include, as well, the average over all simulations of the power spectra when no deprojection is performed shown as the lighter dash-dotted lines.
 The middle subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e.
@@
       The upper plot in the left panel of Fig.
-\ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra (darker dashed lines), and when no deprojection is performed (lighter dash-dotted lines).
+\ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra (darker dashed lines).
 The middle panel of the same figure shows the average residual with respect to the analytical prediction normalize by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).

### 2018-08-01 sandbox_validation/plot_sph.py

@@ -107,44 +107,45 @@ plt.colorbar(im)
 plt.savefig("plots_paper/val_covar_lss_sph.pdf",bbox_inches='tight')
 
 #Plot residuals
-cols=plt.cm.rainbow(np.linspace(0,1,60))
+cols=plt.cm.rainbow(np.linspace(0,1,6))
 plot_dirty=True
 fig=plt.figure()
 ax=fig.add_axes((0.12,0.4,0.78,0.55))
 ic=0
 
-ax.plot(l_th,clTT_clean_mean,label='$\\delta\\times\\delta$',c=cols[ic],alpha=0.7); ic+=4
-if plot_dirty : 
-   ax.plot(l_th,clTT_dirty_mean,'-.',c=cols[ic],alpha=0.3);
+ax.plot(l_th,clTT_clean_mean,label='$\\delta\\times\\delta$',c=cols[ic],alpha=0.45);
+#if plot_dirty : 
+#   ax.plot(l_th,clTT_dirty_mean,'-.',c=cols[ic],alpha=0.3);
 ax.plot(l_th,clTT_th,'--',c=cols[ic]); 
-ic+=6
-ax.plot(l_th,clTE_clean_mean,label='$\\delta\\times\\gamma_E$',c=cols[ic],alpha=0.7); ic+=4
-if plot_dirty :
-    ax.plot(l_th,clTE_dirty_mean,'-.',c=cols[ic],alpha=0.3); 
+ic+=1
+ax.plot(l_th,clTE_clean_mean,label='$\\delta\\times\\gamma_E$',c=cols[ic],alpha=0.45);
+#if plot_dirty :
+#    ax.plot(l_th,clTE_dirty_mean,'-.',c=cols[ic],alpha=0.3); 
 ax.plot(l_th,clTE_th,'--',c=cols[ic]);
-ic+=6
-ax.plot(l_th,clTB_clean_mean,label='$\\delta\\times\\gamma_B$',c=cols[ic],alpha=0.7); ic+=4
+ic+=1
+ax.plot(l_th,clTB_clean_mean,label='$\\delta\\times\\gamma_B$',c=cols[ic],alpha=0.45); 
 if plot_dirty :
-    ax.plot(l_th,clTB_dirty_mean,'-.',c=cols[ic],alpha=0.3);
+    pass
[…]

### 2018-08-02 doc/Paper/main.tex
Added timing section

@@
 
+ \subsection{Performance and scalability}
+ \label{ssec:scalability}
+     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
+In this section we will show that \nmt follows this same scaling with $\ell_{max}$, and we will also study how the code behaves when the number of parallel threads is changed.
+
+     We first focus on checking the dependence of the execution time with $\ell_{max}$.
+We compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no foreground contamination.
+We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
+We use one interactive Cori-Haswell node at the National Energy Research Scientific Computing Center (NERSC).
+This kind of nodes have two 16-core Intel Xeon (E5-2698 v3) processors at 2.3 GHz, and 128 GB DDR4 of RAM.
+Each core supports two hyper-threads, so we end up executing the code using 64 \texttt{OpenMP} threads.
+In Figure~\ref{fig:scaling_lmax} we can see the results for this test.
+We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 20).
+In all cases when we compare with the reference line $\propto \ell_{max}^{3}$ we can see that the code scales as expected.
+We fixed the proportionality constant for the reference line using the timing of the auto-power of the spin-0 field at $\ell_{max}=96$.
+We can also see that for very small $N_{side}$ data handling time starts to be comparable to the pure power-spectra computation time, increasing the expected execution time.
+
+     \begin{figure}
+     \centering
+     \includegraphics[width=0.9\columnwidth]{time_scaling}
+     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) foreground-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
+We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}
+     \label{fig:scaling_lmax}
+     \end{figure}
+
   \section{Discussion}\label{sec:discussion}
@@
 DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
+This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
+Department of Energy Office of Science User Facility operated under Contract No.
+DE-AC02-05CH11231.
 

### 2018-08-02 doc/Paper/main.tex
Latest version of the tex
added timing with number of threads

@@
      One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
-In this section we will show that \nmt follows this same scaling with $\ell_{max}$, and we will also study how the code behaves when the number of parallel threads is changed.
-
-     We first focus on checking the dependence of the execution time with $\ell_{max}$.
-We compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no foreground contamination.
+In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
 We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
@@
 This kind of nodes have two 16-core Intel Xeon (E5-2698 v3) processors at 2.3 GHz, and 128 GB DDR4 of RAM.
-Each core supports two hyper-threads, so we end up executing the code using 64 \texttt{OpenMP} threads.
+We executed the code using 32 \texttt{OpenMP} threads.
 In Figure~\ref{fig:scaling_lmax} we can see the results for this test.
@@
      \includegraphics[width=0.9\columnwidth]{time_scaling}
-     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) foreground-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
+     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
 We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}
@@
 
+    In addition, we also check the parallel performance of \nmt.
+We proceed again, as in the previous case, to compute the auto- and cross-power spectra of spin-0 and spin-2 fields with no contaminants for maps with $N_{side}=256$ and $N_{side}=1024$ and we vary the number of \texttt{OpenMP} threads from 1 to 32 in powers of 2.
+The results can be seen in Figure~\ref{fig:scaling_nthreads}.
+We can see that the operations involving spin-0 fields seem to be more affected by the number of threads used, however, the spin-2 auto-power spectra computing time seems to be very stable for any number of threads larger than 2.
+\todo{Discuss this further?
+Take the second part out?
+Ideas?}
+
+    \begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{time_scaling_threads}
+     \caption{Computing time for auto-power spectra of spin-0 (red lines) and spin-2 (cyan lines) contaminant-free maps, and their cross-power spectra (blue lines) as a function of the number of \texttt{OpenMP} threads used for the computation.
+We check maps with $N_{side}=256$ (solid lines) and $N_{side}=1024$ (broken lines).}
+     \label{fig:scaling_nthreads}
+     \end{figure}
   \section{Discussion}\label{sec:discussion}

### 2018-08-06 src/utils.c
all unit tests in utils.c

@@ -1,5 +1,10 @@
 #include "utils.h"
 
+#include <setjmp.h>
+jmp_buf exception_buffer;
+int exception_status;
+int error_policy=EXIT_ON_ERROR;
+
 int my_linecount(FILE *f)
 {
   int i0=0;
@@ -8,6 +13,11 @@ int my_linecount(FILE *f)
     i0++;
   }
   return i0;
+}
+
+void set_error_policy(int i)
+{
+  error_policy=i;
 }
 
 void report_error(int level,char *fmt,...)
@@ -21,7 +31,10 @@ void report_error(int level,char *fmt,...)
   
   if(level) {
     fprintf(stderr," Fatal error: %s",msg);
-    exit(level);
+    if(error_policy==EXIT_ON_ERROR)
+      exit(level);
+    else
+      throw(level);
   }
   else
     fprintf(stderr," Warning: %s",msg);

### 2018-08-06 src/utils.h

@@ -2,6 +2,19 @@
 #define _NM_UTILS_
 #include "namaster.h"
 
+#include <setjmp.h>
+
+#define EXIT_ON_ERROR 0
+#define THROW_ON_ERROR 1
+extern jmp_buf exception_buffer;
+extern int exception_status;
+extern int error_policy;
+
+#define try if ((exception_status = setjmp(exception_buffer)) == 0)
+#define catch(val) else if (exception_status == val)
+#define throw(val) longjmp(exception_buffer,val)
+#define finally else
+
 //Defined in utils.c
 gsl_rng *init_rng(unsigned int seed);
 double rng_01(gsl_rng *rng);
@@ -11,6 +24,7 @@ void rng_delta_gauss(double *module,double *phase,
 void rng_gauss(gsl_rng *rng,double *r1,double *r2);
 void end_rng(gsl_rng *rng);
 int my_linecount(FILE *f);
+void set_error_policy(int i);
 void report_error(int level,char *fmt,...);
 void *my_malloc(size_t size);
 void *my_calloc(size_t nmemb,size_t size);

### 2018-08-06 test/nmt_test_utils.c

@@ -34,18 +34,96 @@ void test_compare_arrays(int n,double *y,int narr,int iarr,char *fname)
   fclose(fi);
 }
 
-CTEST(nmt,my_linecount) {
-  FILE *f=my_fopen("test/cls.txt","r");
-  int cnt=my_linecount(f);
-  ASSERT_EQUAL(cnt,768);
-  fclose(f);
+CTEST(nmt,drc3jj) {
+  int sizew=1000;
+  double *w3=my_malloc(sizew*sizeof(double));
+  int r,l1min,l1max;
+
+  r=drc3jj(2,3,0,0,&l1min,&l1max,w3,sizew);
+  ASSERT_EQUAL(r,0);
+  ASSERT_EQUAL(l1max,2+3);
+  ASSERT_EQUAL(l1min,1);
+  ASSERT_DBL_NEAR_TOL(-sqrt(3./35.),w3[0],1E-10);
+  ASSERT_DBL_NEAR_TOL(0,w3[1],1E-10);
+  ASSERT_DBL_NEAR_TOL(2/sqrt(105.),w3[2],1E-10);
+  ASSERT_DBL_NEAR_TOL(0,w3[3],1E-10);
+  ASSERT_DBL_NEAR_TOL(-sqrt(10./231.),w3[4],1E-10);
+
+  r=drc3jj(100,200,2,-2,&l1min,&l1max,w3,sizew);
+  ASSERT_EQUAL(r,0);
+  ASSERT_EQUAL(l1max,100+200);
+  ASSERT_EQUAL(l1min,100);
+  ASSERT_DBL_NEAR_TOL(0.0139534,w3[0],1E-5);
+  ASSERT_DBL_NEAR_TOL(-0.00192083,w3[30],1E-5);
+  ASSERT_DBL_NEAR_TOL(0.000639717,w3[54],1E-5);
+  ASSERT_DBL_NEAR_TOL(0.000648742,w3[131],1E-5);
+
+  set_error_policy(THROW_ON_ERROR);
+  printf("\nError messages expected: \n");
+  try { r=drc3jj(100,200,2,-2,&l1min,&l1max,w3,100); }
+  catch(1) {}
+  ASSERT_EQUAL(exception_status,1);
+  try { r=drc3jj(2,3,3,0,&l1min,&l1max,w3,sizew); }
+  catch(1) {}
[…]

### 2018-08-06 doc/Paper/main.tex
performance commented
fixes

@@
 % Add your own macros here:
-\usepackage{lipsum}
 \usepackage{amsmath}
@@
 \usepackage{listings}
+\usepackage{dsfont}
+
 \lstset{ %
@@
 \newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
-\newcommand{\todo}[1]{{\bf TODO: #1}}
 \newcommand{\nmt}{{\tt NaMaster}}
@@
 \newcommand{\mM}{{\sf M}}
-
+\newcommand{\mI}{{\mathds{1}}}
+
+\usepackage{lipsum}
+\usepackage{lineno}
+\linenumbers
 \newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
@@
 {[\textbf{JS}: #1]}}}
+\newcommand{\todo}[1]{{\bf TODO: #1}}
 
@@
 \begin{abstract}
-  The pseudo-$C_\ell$ provides a fast and simpler alternative to the optimal quadratic estimator to compute angular power spectra.
-The algorithm can be extended to deal with contaminant deprojection and $E/B$ purification, and can therefore be applied in a wide variety of scenarios of interest for current and future cosmological observations.
-This paper presents \nmt, a public, validated, robust, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both in the sphere or in the flat-sky approximation.
-We describe the mathematical background of the estimator, including all the features above, and its software implementation in \nmt.
-We construct a validation suite that aims to resemble the types of observations that next-generation large-scale structure and ground-based CMB experiments will face, and use it to show that the code is able to recover the input power spectra in the most complex scenarios with no detectable bias.
-\nmt~can be found at \url{add_url}, and is provided with comprehensive documentation and a number of code examples to maximize its usefulness in the community.
+  The pseudo-$C_\ell$ is an algorithm for estimation of the angular
+  power and cross-power spectra that is very fast and in realistic
+  cases also nearly optimal. The algorithm can be extended to deal
+  with contaminant deprojection and $E/B$ purification, and can
+  therefore be applied in a wide variety of scenarios of interest for
+  current and future cosmological observations. This paper presents
+  \nmt, a public, validated, robust, accurate and easy-to-use software
+  package that, for the first time, provides a unified framework to
+  compute angular cross-power spectra of any pair of spin-0 or spin-2
+  fields, contaminated by an arbitrary number of linear systematics
+  and requiring $B$- or $E$-mode purification, both in the sphere or
+  in the flat-sky approximation. We describe the mathematical
+  background of the estimator, including all the features above, and
+  its software implementation in \nmt. We construct a validation suite
+  that aims to resemble the types of observations that next-generation
+  large-scale structure and ground-based CMB experiments will face,
+  and use it to show that the code is able to recover the input power
+  spectra in the most complex scenarios with no detectable
+  bias. \nmt~can be found at \url{add_url}, and is provided with
+  comprehensive documentation and a number of code examples to
+  maximize its usefulness in the community.
 \end{abstract}
@@
   \section{Introduction}\label{sec:intro}
-    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}.
-Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier/harmonic-space power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.
-Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a far higher level of data compression and decorrelation can be obtained through $C_\ell$s.
-This, together with other advantages associated with power spectra (e.g.
-clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}) has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators.
+  Two-point functions have proven to be the most useful summary
+  statistic for cosmology in terms of data compression, due to the
+  near-Gaussian nature of the modes of the cosmic density field that
+  can be easily analyzed to extract cosmological
+  information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since
+  the only true observables in any astronomical dataset are the
+  intensity and polarization as a function of frequency $\nu$ and sky
+  position $\nv$, angular 2-point correlators play a particularly
+  central role in most cosmological analyses, both in the form of
+  configuration-space two-point correlation functions $\xi(\theta)$
+  and as Fourier-space\footnote{In this paper we use the term
+    ``Fourier'' to describe both the flat space Fourier expansion as
+    well as harmonic-space expansion on the sphere. We will also use
+    the term ``power spectrum'' to describe both the auto power
+    spectrum of a field and cross-power spectra between two fields. }
+  power spectra $C_\ell$
+  \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.  Although
+  correlation functions are often easier to estimate from complex
+  observations~\citep{1993ApJ...412...64L}, a Fourier space analysis
+  is advantageous for many other reasons. The most important is that
+  Fourier space is a natural space for working with statistics of
+  translation-invariant fields. Therefore, Fourier space analysis has
+  a clearer separation between linear and non-linear scales, as well
+  as Gaussian and non-Gaussian
+  modes~\citep{2008PhRvD..77j3013H}. Moreover, individual band-power
+  measurements are typically significantly less correlated compared to
+  configuration-space analysis which makes covariance matrix
+  estimation easier. This has motivated a large body of literature
+  aimed at designing optimal and efficient power spectrum estimators.
 
@@
         \begin{align}\label{eq:m00}
-          &{\sf M}^{00}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\,{\sf 1},\\\label{eq:m02}
-          &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,{\sf 1},\\\nonumber
+          &{\sf M}^{00}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\,\mI,\\\label{eq:m02}
+          &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\mI,\\\nonumber
           &M^{0+}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0},\\\label{eq:m22}
@@
         \end{align}
-        where ${\sf 1}$ is the unit matrix, and $P^{vw}_\ell\equiv(2\ell+1)\,{\rm PCL}_\ell(v,w)$.
+        where $\mI$ is the unit matrix, and $P^{vw}_\ell\equiv(2\ell+1)\,{\rm PCL}_\ell(v,w)$.
 
@@
     \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
-      In a typical data-analysis case, we deal with contaminants that pollute the measured signal.
-Usually, we know something about these contaminants.
-As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which can be marginalized out.
-But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening.
-Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales \dam{sounds like a very strong statement that we should either prove or give references for!
-I think I would rather say that you have either reduced contamination enough that a first-order Taylor term should account for the rest or that, in general, you can use a Taylor expansion where each term is a contaminant map.}\js{I agree with David, I'd go with the Taylor expansion motivation}.
-Thus, we generically deal with a linear model of the kind:
+    In a typical data-analysis case, we deal with contaminants that
+    pollute the measured signal. Usually, we know something about
+    these contaminants. As an example, in case of CMB analysis, we
+    might have have a map of a given foreground at a different
+    frequency which can be marginalized out. But even when we do not,
+    we often have a tracer map of a given effect, for example, the map
+    of dust might tell us something about systematic errors related to
+    imperfect correction to dust reddening. Even if the dust map is
+    non-linearly and perhaps stochastically related to the amount of
+    reddening, as long as relation is local in space, we get a linear
+    contamination on large scales, very much like complex astrophysics
+    of galaxy formation nevertheless reduces to linear biasing on very
+    large scales (see Appendix \ref{app:lincont}). Thus, we generically deal with a
+    linear model of the kind:
       \begin{equation}\label{eq:aobs}
@@
 
-      Two further points must be addressed before moving on.
-First, the computation of the deprojection bias requires an estimate of the true power spectrum ${\sf C}^{ab}_\ell$ (see Eqs.
-\ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}).
-In practice, this can be done through an iterative approach, starting from an initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the estimated power spectrum in the previous iteration.
-A sufficiently good guess will likely be enough to remove the bias in a single iteration.
-Alternatively, the ensemble average of Eq.
-\ref{eq:dpbias0} can be seen as a convolution of ${\sf C}^{ab}_\ell$, which could be treated by computing the convolution matrix as a modified version of the standard mode-coupling matrix that accounts for mode deprojection.
-Estimating this convolution matrix in this case becomes significantly more computationally intensive, and therefore an iterative approach is preferred.
-\as{I think this is not quite true.
-This latter way of looking at it, in effect replaces the true power spectrum with its observed quantity, which is fine in practice, but conceptually not quite right, for the same reason e.g.
-error on power spectrum should be based on real rather than measured power as illustrated by low CMB quadrupole.
-I would just remove text after ``Alternatively...''} \dam{we can do this, but I'd like to understand what the problem with this is.
-Since you can understand a mask as a set of deprojected contaminants (one per masked pixel), the pseudo-$C_\ell$ mode-coupling matrix is basically a flavour of the equations above for one particular set of templates.
-In this sense then writing the equation above as a matrix-vector product and inverting that matrix (after binning etc.) seems like the pseudo-$C_\ell$ way to go.
-Are you saying that you'll get more optimal results by putting in a guess for the power spectrum because that's closer to what a QML estimator would do?}
+      Two further points must be addressed before moving on. First,
+      the computation of the deprojection bias requires an estimate of
+      the true power spectrum ${\sf C}^{ab}_\ell$ (see
+      Eqs. \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}). In practice, this
+      can be done through an iterative approach, starting from an
+      initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the
+      estimated power spectrum in the previous iteration. A
+      sufficiently good guess will likely be enough to remove the bias
+      in a single iteration.  Alternatively, the ensemble average of
+      Eq. \ref{eq:dpbias0} can be seen as a convolution of
+      ${\sf C}^{ab}_\ell$, which could be treated by computing the
+      convolution matrix as a modified version of the standard
+      mode-coupling matrix that accounts for mode
+      deprojection. Estimating this convolution matrix in this case
+      becomes significantly more computationally intensive, and
+      therefore an iterative approach is preferred.
 
@@
     \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
-      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa.
-The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB.
-In this case, removing the leakage at the power spectrum level (i.e.
-the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes.
-This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above.
-In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e.
-purifying $E$-modes from $B$-mode contamination is conceptually exactly the same.
-Moreover.
-our discussion is also specific for spin-2 fields.
+    As described in
+    \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B},
+    $B$-mode purification refers to the map-level removal of the
+    contamination from $E$-modes in a the $B$-mode component of a
+    given map caused by an incomplete sky coverage and vice-versa. The
+    procedure is particularly useful in situations in which the
+    $E$-mode component of the signal is significantly larger than the
+    $B$ modes, as is the case in CMB. In this case, removing the
+    leakage at the power spectrum level (i.e. the standard
+    pseudo-$C_\ell$ approach) produces a sub-optimal estimator in
+    which the variance in the $B$-mode power spectrum is dominated by
+    the variance of the leaked $E$-modes. This section presents the
+    purification method and the associated modifications to the
+    pseudo-$C_\ell$ algorithm as described in the references above. In
+    this section we describe the algorithm in terms of purifying
+    $B$-modes from $E$-mode contamination, but we note that the
+    reverse, i.e. purifying $E$-modes from $B$-mode contamination is
+    conceptually exactly the same. Our discussion is also
+    specific for the spin-2 fields.
 
@@
           \end{equation}
-          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$.
-Note that it is impossible to have an anisotropic beam that is uniform over the entire sphere.
-However, it is possible that the beam is both anisotropic and spatially varying.
-In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam.
-Correcting for the beam in such cases can also be done analytically, but we leave it for future work.
-\as{Anything to add here?} \dam{maybe be a bit more precise.}
+          where $W^{ab}_\ell$ is the product of the harmonic transform
+          of the beams for maps $a$ and $b$. Note that due to hairy
+          ball theorem, it is impossible to have an anisotropic beam
+          that is uniform over the entire sphere.  However, it is
+          possible that the beam is both anisotropic and spatially
+          varying. In this case, the pseudo-$C_\ell$ power spetrum
+          estimation will produce biased estimates that will be
+          suppressed by an \emph{effective} beam. Correcting for the
+          beam in such cases can also be done analytically, but we
+          leave it for future work.  In practice, if required, the
+          suppression factor can be found by applying the PCL
+          estimator to a a set of mock input maps with the same
+          underlying signal but  with and without application of the
+          beam smoothing (to both data maps and contaminant maps) and taking
+          the ration. \as{How does this interact with E/B purification?}
 
@@
            If the noise is homogeneous but uncorrelated (i.e.
-$\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle={\sf 1}\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
+$\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle=\mI\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
            \begin{itemize}
@@
                    \begin{align}
-                     {\rm PCL}_\ell({\bf n},{\bf n})&={\sf 1}\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
-                                                    &={\sf 1}\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
+                     {\rm PCL}_\ell({\bf n},{\bf n})&=\mI\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
+                                                    &=\mI\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
                    \end{align}
@@
 ${\cal O}(N_{\rm pix}\log(N_{\rm pix}))$ for DFTs) and because by SHTs involve operations over the full sphere.
-        \item Fourier transform over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
+        \item Fourier transforms over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
       \end{enumerate}
@@
         \end{figure}
-        The first validation suite exemplifies the use of pseudo-$C_\ell$ algorithms in the analysis of large-scale structure data.
-Current and next-generation imaging surveys such as DES~\citet{2016MNRAS.460.1270D}, KiDS~\citet{2017A&A...604A.134D}, HSC~\citet{2018PASJ...70S...4A}, LSST~\citet{2008arXiv0805.2366I} and Euclid~\citet{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing.
-These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
-
-        Large-scale structure data are often characterized by inhomogeneous sky coverages and complex mask structures, associated with the presence of bright stars and observational artifacts.
-They are also affected by many sky contaminants, associated to Galactic sources (dust absorption, star density) and observing conditions (seeing, airmass etc.)~\citet{2016ApJS..226...24L}.
-The typically flatter power spectra for both galaxy clustering and lensing (e.g.
-when compared with the damping tail of CMB temperature fluctuations), makes the pseudo-$C_\ell$ approach ideal for this kind of analysis~\citet{2013MNRAS.435.1857L,2016ApJS..226...24L}.
-This science case therefore allows us to validate several aspects of the code simultaneously:
+        The first validation suite exemplifies the use of
+        pseudo-$C_\ell$ algorithms in the analysis of large-scale
+        structure data. Current and next-generation imaging surveys
+        such as DES~\citep{2016MNRAS.460.1270D},
+        KiDS~\citep{2017A&A...604A.134D},
+        HSC~\citep{2018PASJ...70S...4A},
+        LSST~\citep{2008arXiv0805.2366I} and
+        Euclid~\citep{2011arXiv1110.3193L} will constrain the growth
+        and geometry of structure through cosmic time by using the
+        overdensity of galaxies as a proxy for the matter
+        inhomogeneities (a spin-0 field) and their correlated shape
+        distortions (a spin-2 field at first order), mostly caused by
+        weak lensing. These measurements will mostly follow a
+        tomographic approach, using the angular auto- and cross-power
+        spectra between these observables across different bins of
+        photometric redshift
+        \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
+
+        Large-scale structure data are often characterized by
+        inhomogeneous sky coverages and complex mask structures,
+        associated with the presence of bright stars and observational
+        artifacts. They are also affected by many sky contaminants,
+        associated to Galactic sources (dust absorption, star density)
+        and observing conditions (seeing, airmass
+        etc.) as discussed in \citet{2016ApJS..226...24L}. The typically flatter power
+        spectra for both galaxy clustering and lensing (e.g. when
+        compared with the damping tail of CMB temperature
+        fluctuations), makes the pseudo-$C_\ell$ approach ideal for
+        this kind of analysis \citep{2013MNRAS.435.1857L,2016ApJS..226...24L}. This science case therefore allows us to validate several aspects of the code simultaneously:
         \begin{itemize}
@@
 Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the full-sky and flat-sky cases respectively.
-          \item {\bf Mask:} for the full-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}.
-To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask.
-We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig.
-\ref{fig:lss_mocks_sph}).
-For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citet{2018PASJ...70S...4A}.
+          \item {\bf Mask:} for the full-sky data, we use a mask built
+            as a combination of the sky coverage output by the LSST
+            {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a
+            more conservative galactic cut based on the reddening map
+            of \citet{1998wfsc.conf..297S}. To add an extra layer of
+            complexity, we drill 100 additional 1-degree holes on the
+            resulting combined mask. We also explore the improvement
+            stemming from the down-weighting of higher-noise in an
+            inverse-variance way in the presence of inhomogeneous
+            noise (see description below), which is the reason why the
+            default version of this mask is not binary, but tracers
+            the inverse of the noise variance (see top left panel of
+            Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we
+            use a high-resolution mask constructed from the
+            bright-object mask used for the first public data release
+            from the HSC collaboration~\citep{2018PASJ...70S...4A}.
 The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales.
@@
           \item {\bf Contaminants:} we assume a dust-like residual for both the full-sky an flat-sky cases.
-For the full-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citet{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$.
+For the full-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citep{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$.
 For the flat-sky signal we generate a Gaussian realization of a dust-like component with power spectra
@@
         \caption{{\sl Left panel:} average over all simulations in our full-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
-The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly below the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
+The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
 The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero.
@@
         \includegraphics[width=0.99\columnwidth]{val_cmb_bpw}
-        \caption{{\sl Left:} average over all simulations in our full-sky CMB validation suite of the three spin-2 power spectra ($EE$, $EB$ and $BB$, solid lines).
-The analytical prediction for the input signal power spectra are shown as dashed lines, and line mostly below the solid lines for the non-zero cases ($EE$ and $BB$).
+        \caption{{\sl Left:} average over all simulations in our
+          full-sky CMB validation suite of the three spin-2 power
+          spectra ($EE$, $EB$ and $BB$, solid lines) using mode
+          deprojection and E/B mode purification.
+The analytical prediction for the input signal power spectra are shown as dashed lines, and line mostly under the solid lines for the non-zero cases ($EE$ and $BB$).
 The null power spectrum $EB$ is compatible with zero.
@@
       \end{equation}
-      The full covariance matrix is very close to diagonal, as shown in the right panel of Fig.
-\ref{fig:lss_val_cls}.
-Under the assumption that the estimator is unbiased ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$ should follow a $\chi^2$ distribution with a $N_{\rm dof}$ degrees of freedom.
-Figure \ref{fig:lss_val_chi2} shows the distribution of $\chi^2_i$ across simulations (blue histograms), compared with the predicted $\chi^2$ distribution for $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the six different power spectrum combinations.
-Both distributions are found to be compatible.
-Moreover, the dot-dashed lines show the values of $\chi^2$ found for the mean residual with respect to the input power spectra (middle panel of Fig.
-\ref{fig:lss_val_cls}).
-The probability to exceed (PTE) associated with these $\chi^2$ values are higher than $10\%$ in all cases, and therefore our validation suite does not show any statistically significant bias in the estimator.
-This bias, if any, must be at least $\sqrt{1000}\sim30$ times smaller than the statistical uncertainties associated with our validation suite, which are representative of the statistical power of next-generation galaxy surveys.
+      The full covariance matrix is very close to diagonal, as shown
+      in the right panel of Fig. \ref{fig:lss_val_cls}. Under the
+      assumption that the estimator is unbiased
+      ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$
+      should follow a $\chi^2$ distribution with a $N_{\rm dof}$
+      degrees of freedom. Figure \ref{fig:lss_val_chi2} shows the
+      distribution of $\chi^2_i$ across simulations approximating
+      covariance matrix as diagonal (blue histograms),
+      compared with the predicted $\chi^2$ distribution for
+      $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the
+      six different power spectrum combinations. Both distributions
+      are found to be compatible. Moreover, the dot-dashed lines show
+      the values of $\chi^2$ found for the mean residual with respect
+      to the input power spectra (middle panel of
+      Fig. \ref{fig:lss_val_cls}). The probability to exceed (PTE)
+      associated with these $\chi^2$ values are higher than $10\%$ in
+      all cases, and therefore our validation suite does not show any
+      statistically significant bias in the estimator. This bias, if
+      any, must be at least $\sqrt{1000}\sim30$ times smaller than the
+      statistical uncertainties associated with our validation suite,
+      which are representative of the statistical power of
+      next-generation galaxy surveys.
 
@@
       \end{figure*}
-      We have carried out a similar set of tests on an ensemble of 1000 simulations from the CMB validation suite described in Section \ref{sssec:validation.suite.cmb}.
-In our fiducial run we both deprojected the foreground contaminants and subsequently used $B$-mode purification to decouple $E$ and $B$-modes at the map level, since this is the main feature of the code we want to validate.
-The left panel of Figure \ref{fig:cmb_val_cls} shows the mean of the $EE$, $EB$ and $BB$ power spectra computed for all simulations (solid lines) compared with the expectation from the input spectra for $EE$ and $BB$ (dashed lines, mostly lying below the solid ones).
-The average residuals normalized over the error on the mean, shown in the lower panel of the same Figure, fluctuate around zero with an amplitude of up to $3\sigma$, as would be expected in the absence of a bias in the estimator.
-
-      Figure \ref{fig:cmb_val_sigma} illustrates the impact of $B$-mode purification on the variance of the pseudo-$C_\ell$ estimator.
-The Figure shows the diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra.
+      We have carried out a similar set of tests on an ensemble of
+      1000 simulations from the CMB validation suite described in
+      Section \ref{sssec:validation.suite.cmb}. In our fiducial run we
+      both deprojected the foreground contaminants and subsequently
+      used $B$-mode purification to decouple $E$ and $B$-modes at the
+      map level, since this is the main feature of the code we want to
+      validate. The left panel of Figure \ref{fig:cmb_val_cls} shows
+      the mean of the $EE$, $EB$ and $BB$ power spectra computed for
+      all simulations (solid lines) compared with the expectation from
+      the input spectra for $EE$ and $BB$ (dashed lines, mostly lying
+      under the solid ones). The average residuals normalized over the
+      error on the mean, shown in the lower panel of the same Figure,
+      fluctuate around zero with an amplitude of up to $3\sigma$, as
+      would be expected in the absence of a bias in the
+      estimator.
+
+      The above figure was created using E/B mode
+      purification. Disabling this option has no impact on the bias of
+      the estimates, but increases its variance as we show in the
+      Figure \ref{fig:cmb_val_sigma}.
+We plot diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra.
 Although there is no practical improvement in the error bars for the $EE$ spectrum (there is even a slight increase due to the loss of the ambiguous $E$ modes to purification), and the improvement is only mild for the $EB$ case, very significant gains can be obtained for the $BB$ power spectrum, reducing uncertainties by up to a factor of $\sim10$ on large scales.
@@
 The non-zero off-diagonal elements are caused by two effects.
-Firstly, the limited footprint area, map apodization and choice of bandpowers leads to a noticeable level of anti-correlation between neighboring bandpowers at low $\ell$.
+First, the limited footprint area, map apodization and choice of bandpowers leads to a noticeable level of anti-correlation between neighboring bandpowers at low $\ell$.
 This can be seen in the left panel of Figure \ref{fig:cmb_val_cov}, which shows the covariance matrix of simulations without foreground contaminants or deprojection.
-Then, introducing contaminant deprojection produces additional correlations that affect the lowest bandpowers and also involves a level of cross-talk between the three different cross-spectra.
+Second, introducing contaminant deprojection produces additional correlations that affect the lowest bandpowers and also involves a level of cross-talk between the three different cross-spectra.
 
@@
 
- \subsection{Performance and scalability}
- \label{ssec:scalability}
-     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
-In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
-We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
-We use one interactive Cori-Haswell node at the National Energy Research Scientific Computing Center (NERSC).
-This kind of nodes have two 16-core Intel Xeon (E5-2698 v3) processors at 2.3 GHz, and 128 GB DDR4 of RAM.
-We executed the code using 32 \texttt{OpenMP} threads.
-In Figure~\ref{fig:scaling_lmax} we can see the results for this test.
-We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 20).
-In all cases when we compare with the reference line $\propto \ell_{max}^{3}$ we can see that the code scales as expected.
-We fixed the proportionality constant for the reference line using the timing of the auto-power of the spin-0 field at $\ell_{max}=96$.
-We can also see that for very small $N_{side}$ data handling time starts to be comparable to the pure power-spectra computation time, increasing the expected execution time.
-
-     \begin{figure}
-     \centering
-     \includegraphics[width=0.9\columnwidth]{time_scaling}
-     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
-We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}
-     \label{fig:scaling_lmax}
-     \end{figure}
-
-    In addition, we also check the parallel performance of \nmt.
-We proceed again, as in the previous case, to compute the auto- and cross-power spectra of spin-0 and spin-2 fields with no contaminants for maps with $N_{side}=256$ and $N_{side}=1024$ and we vary the number of \texttt{OpenMP} threads from 1 to 32 in powers of 2.
-The results can be seen in Figure~\ref{fig:scaling_nthreads}.
-We can see that the operations involving spin-0 fields seem to be more affected by the number of threads used, however, the spin-2 auto-power spectra computing time seems to be very stable for any number of threads larger than 2.
-\todo{Discuss this further?
-Take the second part out?
-Ideas?}
-
-    \begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{time_scaling_threads}
-     \caption{Computing time for auto-power spectra of spin-0 (red lines) and spin-2 (cyan lines) contaminant-free maps, and their cross-power spectra (blue lines) as a function of the number of \texttt{OpenMP} threads used for the computation.
-We check maps with $N_{side}=256$ (solid lines) and $N_{side}=1024$ (broken lines).}
-     \label{fig:scaling_nthreads}
-     \end{figure}
   \section{Discussion}\label{sec:discussion}
@@
 
+    \nmt~does not currently produce a covariance matrix estimate. This
+    is done for several reasons. First, for many problems, such as
+    large-scale structure power spectrum estimation, the covariance
+    matrix contains import contributions from the connected four-point
+    function that depend on the precise nature of the problem and go
+    beyond the estimation step that \nmt~strives to do well. Second,
+    with all additional options included, the exact analytical
+    estimates are either computational infeasible or overly
+    approximate. Since majority of the forthcoming experiments will
+    have a large number of mock datasets available for testing
+    systematic effects, these maps also offer a natural path towards
+    estimating the data covariance.
+
     \nmt~is open-source software and publicly available at \url{add_url}.
@@
 \subsection*{Acknowledgements}
-We would like to thank Rupert Allison, Jo Dunkley, Franz Elsner, Boris Leistedt, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya Peiris, Naomi Robertson and Ben Thorne for useful comments and discussions.
-DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
+We would like to thank Rupert Allison, Jo Dunkley, Franz Elsner, Boris
+Leistedt, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya
+Peiris, Naomi Robertson and Ben Thorne for useful comments and
+discussions. DA acknowledges support from the Beecroft trust and from
+the Science and Technology Facilities Council (STFC) through an Ernest
+Rutherford Fellowship, grant reference ST/P004474/1. This research
+used resources of the National Energy Research Scientific Computing
+Center (NERSC), a U.S. Department of Energy Office of Science User
+Facility operated under Contract No. DE-AC02-05CH11231.
 
@@
         &\hat{\sf Y}^{s\dag}_{\ell m}=(-1)^{m+s}\hat{\sf Y}^{-s}_{\ell -m}\\
-        &\int d\nv \hat{\sf Y}^s_{\ell m}\hat{\sf Y}^{s\dag}_{\ell'm'}=\hat{\sf 1}\delta_{\ell\ell'}\delta_{mm'},\\
+        &\int d\nv \hat{\sf Y}^s_{\ell m}\hat{\sf Y}^{s\dag}_{\ell'm'}=\hat\mI\delta_{\ell\ell'}\delta_{mm'},\\
         &\int d\nv \left(\hat{\sf Y}^{s\dag}_{\bf l}(\nv)\hat{\sf Y}^{s}_{{\bf l}_1}(\nv)\right)\,\hat{\sf Y}^0_{{\bf l}_2}(\nv)\equiv\hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2},
@@
       \end{equation}
+
+\section{Linear contamination from non-linear
+  contamination} \label{app:lincont}
+\as{It is up to David to decide whether he buys this argument or
+  not. I'm fine with not and then we soften then claim in 2.2. Doesn't
+  really affect any of the paper otherwise.}
+Imagine a contaminant field $c$ (e.g. a point-spread function (PSF) size)
+that affects the observable field $o$ (e.g. the number count
+densities) through a non-linear, but \emph{local} on some scale local
+functional $F$:
+\begin{equation}
+  \delta_o(\nv) \rightarrow \delta_o + F [ \delta_c ]
+\end{equation}
+where we can, without loss of generality, take out the mean effects
+(i.e. there is some mean PSF size and fluctuations around this mean
+size produce additional fluctuations in the number density of
+fluctuations).
+
+In general, for sufficiently non-linear functionals $F$, the Taylor
+expansion around $\delta_c(\nv)$ is not valid. However, one can look
+at the effect of smoothing by a kernel of size $S(R)$, which in
+Fourier space is equivalent by multiplication by $e^{-k^2R^2}$. We get
+\begin{equation}
+  \delta^R_o(\nv) \rightarrow \delta^R_o  + S(R)[F [ \delta_c
+  ]] =  \delta^R_o + F'_R [ \delta^R_c ],
+\end{equation}
+where we used a shorthand $\delta_x^R=S(R)[\delta_x]$ and  $F'_R$ is now a \emph{different} functional that takes a smoothed
+$\delta_c$ and returns out a smoothed $\delta_o$.  The crucial point
+is that $F'_R$ must still be
+local, since one cannot turn a local functional into a non-local one
+by application of smoothing.
+
+This means, that for a well behaved $\delta_c$, such that variances of
+both $\delta^R_c$ and $\delta^R_o$ are small for a large enough $R$,
+the Taylor expansion is valid for $F'_R$ (even if it is not for $F$),
+giving
+\begin{equation}
+  \delta^R_o(\nv) \rightarrow \delta^R_o  + b \delta^R_c + c \nabla^2 \delta^R_c
+\end{equation}
+or in Fourier space
+\begin{equation}
+    \delta_o(\nv) \rightarrow \delta_o  + b \delta_c + c k^2 \delta_c
+\end{equation}
+for $k\ll R^{-1}$. In other words a ``local'' contaminant becomes
+``linearly biased'' in the $k\rightarrow 0 $ regime, very much like
+very non-linear physics of galaxy formations results in
+linearly biased galaxy field on large scales.
 
@@
              \begin{equation}
-               \sum_{\bf x}\,_sE^\dag_{\bf l}({\bf x})\,_sE_{\bf k}({\bf x})=N_xN_y\delta_{{\bf l},{\bf k}}\hat{\sf 1}.
+               \sum_{\bf x}\,_sE^\dag_{\bf l}({\bf x})\,_sE_{\bf k}({\bf x})=N_xN_y\delta_{{\bf l},{\bf k}}\hat\mI.
              \end{equation}
@@
       &{\sf M}^{00}_{\ell k}=\frac{k}{2\pi}\int_0^\pi \frac{d\theta}{\pi}\,\tilde{C}^{vw}_{q(\ell,k,\theta)}\\
-      &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\hat{\sf 1},
+      &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\hat\mI,
       \hspace{12pt}
@@
 \end{document}
+
+ \subsection{Performance and scalability}
+ \label{ssec:scalability}
+     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
+In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
+We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
+We use one interactive Cori-Haswell node at the National Energy Research Scientific Computing Center (NERSC).
+This kind of nodes have two 16-core Intel Xeon (E5-2698 v3) processors at 2.3 GHz, and 128 GB DDR4 of RAM.
+We executed the code using 32 \texttt{OpenMP} threads.
+In Figure~\ref{fig:scaling_lmax} we can see the results for this test.
+We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 20).
+In all cases when we compare with the reference line $\propto \ell_{max}^{3}$ we can see that the code scales as expected.
+We fixed the proportionality constant for the reference line using the timing of the auto-power of the spin-0 field at $\ell_{max}=96$.
+We can also see that for very small $N_{side}$ data handling time starts to be comparable to the pure power-spectra computation time, increasing the expected execution time.
+
+     \begin{figure}
+     \centering
+     \includegraphics[width=0.9\columnwidth]{time_scaling}
+     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
+We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}
+     \label{fig:scaling_lmax}
+     \end{figure}
+
+    In addition, we also check the parallel performance of \nmt.
+We proceed again, as in the previous case, to compute the auto- and cross-power spectra of spin-0 and spin-2 fields with no contaminants for maps with $N_{side}=256$ and $N_{side}=1024$ and we vary the number of \texttt{OpenMP} threads from 1 to 32 in powers of 2.
+The results can be seen in Figure~\ref{fig:scaling_nthreads}.
+We can see that the operations involving spin-0 fields seem to be more affected by the number of threads used, however, the spin-2 auto-power spectra computing time seems to be very stable for any number of threads larger than 2.
+\todo{Discuss this further?
+Take the second part out?
+Ideas?}
+
+    \begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{time_scaling_threads}
+     \caption{Computing time for auto-power spectra of spin-0 (red lines) and spin-2 (cyan lines) contaminant-free maps, and their cross-power spectra (blue lines) as a function of the number of \texttt{OpenMP} threads used for the computation.
+We check maps with $N_{side}=256$ (solid lines) and $N_{side}=1024$ (broken lines).}
+     \label{fig:scaling_nthreads}
+     \end{figure}

### 2018-08-06 doc/Paper/main.tex
debug toggle

@@
 \linenumbers
+
+\usepackage{etoolbox}
+
+\newtoggle{DEBUG}
+\toggletrue{DEBUG}
+
+\iftoggle{DEBUG}{
 \newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
@@
 \newcommand{\dam}[1]{{\textcolor{green}{[\textbf{DA}: #1]}}}
-\newcommand{\js}[1]
-{{\textcolor{orange}
-{[\textbf{JS}: #1]}}}
+\newcommand{\js}[1]{{\textcolor{orange}{[\textbf{JS}: #1]}}}
 \newcommand{\todo}[1]{{\bf TODO: #1}}
+}{
+\newcommand{\chg}[1]{}
+\newcommand{\as}[1]{}
+\newcommand{\dam}[1]{}
+\newcommand{\js}[1]{}
+\newcommand{\todo}[1]{}
+}
 
@@
   spectra in the most complex scenarios with no detectable
-  bias. \nmt~can be found at \url{add_url}, and is provided with
+  bias. \nmt~can be found at \url{https://github.com/LSSTDESC/NaMaster}, and is provided with
   comprehensive documentation and a number of code examples to
@@
   \section{Code structure}\label{sec:structure}
-    All of the methods described above have been implemented in a software package called \nmt\footnote{The code source is hosted in \url{add_url}, and its documentation can be found at \url{http://namaster.readthedocs.io/}.}.
+    All of the methods described above have been implemented in a software package called \nmt\footnote{The code source is hosted in \url{https://github.com/LSSTDESC/NaMaster}, and its documentation can be found at \url{http://namaster.readthedocs.io/}.}.
 We briefly describe the code structure here.
@@
 
-    \nmt~is open-source software and publicly available at \url{add_url}.
-To maximize the usefulness of this software for the community, it is accompanied by an extensive documentation and example code\footnote{See \url{add_url}.}.
-\nmt~is written in {\tt C} and {\tt OpenMP}-parallelized to maximize its performance, although we encourage its use through the companion {\tt python} wrapper, the fastest-growing language of choice in the analysis of astronomical data.
-The code is in constant development, and we will strive to implement further functionality to e.g.
-estimate covariance matrices, extend the range of supported sky pixelizations and enhance its computational performance.
+    \nmt~is open-source software and publicly available at \url{https://github.com/LSSTDESC/NaMaster}. To maximize the
+    usefulness of this software for the community, it is accompanied
+    by an extensive documentation and example code\footnote{See \url{
+        https://github.com/LSSTDESC/NaMaster}.}. \nmt~is written in
+    {\tt C} and {\tt OpenMP}-parallelized to maximize its performance,
+    although we encourage its use through the companion {\tt python}
+    wrapper, the fastest-growing language of choice in the analysis of
+    astronomical data. The code is in constant development, and we
+    will strive to implement further functionality to e.g. estimate
+    covariance matrices, extend the range of supported sky
+    pixelizations and enhance its computational performance.
 

### 2018-08-06 src/healpix_extra.c
all non-SHT healpy_extra unit tests

@@ -22,7 +22,7 @@ void he_write_healpix_map(flouble **tmap,int nfields,long nside,char *fname)
     tunit[ii]=my_malloc(256);
     sprintf(ttype[ii],"map %d",ii+1);
     sprintf(tform[ii],"1E");
-    sprintf(tunit[ii],"uK");
+    sprintf(tunit[ii],"  ");
   }
 
   fits_create_file(&fptr,fname,&status);
@@ -434,7 +434,7 @@ void he_in_ring(int nside,int iz,flouble phi0,flouble dphi,
   int nir_here;
   flouble phi_low,phi_hi,shift;
 
-  conservative=1;//Do we take intersected pixels which
+  conservative=1;//Do we take intersected pixels whose
                  //centers do not fall within range?
   take_all=0;//Take all pixels in ring?
   to_top=0;

### 2018-08-06 test/nmt_test_hpex.c

@@ -1,0 +1,183 @@
+#include "namaster.h"
+#include "ctest.h"
+#include "utils.h"
+#include "nmt_test_utils.h"
+#include <chealpix.h>
+
+CTEST(nmt,he_io) {
+
+  set_error_policy(THROW_ON_ERROR);
+
+  printf("\nError messages expected: \n");
+
+  int ii;
+  long nside=4;
+  long npix=he_nside2npix(nside);
+  ASSERT_EQUAL(npix,12*nside*nside);
+
+  double **mps0=my_malloc(1*sizeof(double *));
+  double **mps2=my_malloc(1*sizeof(double *));
+  for(ii=0;ii<1;ii++)
+    mps0[ii]=my_calloc(npix,sizeof(double));
+  for(ii=0;ii<2;ii++)
+    mps2[ii]=my_calloc(npix,sizeof(double));
+
+  //Writing
+  try {he_write_healpix_map(mps0,1,nside,"!mps.fits");}
+  catch(1) {}
+  ASSERT_EQUAL(0,exception_status);
+  try {he_write_healpix_map(mps2,1,nside,"!mps.fits");}
+  catch(1) {}
+  ASSERT_EQUAL(0,exception_status);
+  try {he_write_healpix_map(mps2,2,nside,"!mps.fits");}
+  catch(1) {}
+  ASSERT_EQUAL(0,exception_status);
+
+  for(ii=0;ii<1;ii++)
+    free(mps0[ii]);
+  for(ii=0;ii<2;ii++)
+    free(mps2[ii]);
+  
+  //Reading
+  long ns2;
+  int nfd,isnest;
+  //Correct file params
+  he_get_file_params("mps.fits",&ns2,&nfd,&isnest);
+  ASSERT_EQUAL(nside,ns2);
+  ASSERT_EQUAL(2,nfd);
+  ASSERT_EQUAL(0,isnest);
+  //Read map
+  mps0[0]=he_read_healpix_map("mps.fits",&ns2,0);
+  ASSERT_EQUAL(nside,ns2);
+  mps2[0]=he_read_healpix_map("mps.fits",&ns2,0);
+  ASSERT_EQUAL(nside,ns2);
+  //Go one column too far
+  try {mps2[1]=he_read_healpix_map("mps.fits",&ns2,2);}
+  catch(1) {}
+  ASSERT_EQUAL(1,exception_status);
[…]

### 2018-08-06 test/nmt_test_utils.c

@@ -34,7 +34,7 @@ void test_compare_arrays(int n,double *y,int narr,int iarr,char *fname)
   fclose(fi);
 }
 
-CTEST(nmt,drc3jj) {
+CTEST_SKIP(nmt,drc3jj) {
   int sizew=1000;
   double *w3=my_malloc(sizew*sizeof(double));
   int r,l1min,l1max;
@@ -62,16 +62,16 @@ CTEST(nmt,drc3jj) {
   printf("\nError messages expected: \n");
   try { r=drc3jj(100,200,2,-2,&l1min,&l1max,w3,100); }
   catch(1) {}
-  ASSERT_EQUAL(exception_status,1);
+  ASSERT_EQUAL(1,exception_status);
   try { r=drc3jj(2,3,3,0,&l1min,&l1max,w3,sizew); }
   catch(1) {}
-  ASSERT_EQUAL(exception_status,1);
+  ASSERT_EQUAL(1,exception_status);
   
   set_error_policy(EXIT_ON_ERROR);
   free(w3);
 }
 
-CTEST(nmt,mp_pinv) {
+CTEST_SKIP(nmt,mp_pinv) {
   gsl_matrix *M=gsl_matrix_alloc(3,3);
 
   //Non-degenerate matrix
@@ -123,7 +123,7 @@ CTEST(nmt,mp_pinv) {
 #define M2_POISSONL 1000. //variance top hat
 #define M2_POISSONS 0.5 //variance top hat
 #define M2_GAUSS 1. //variance Gaussian
-CTEST(nmt,rngs) {
+CTEST_SKIP(nmt,rngs) {
   gsl_rng *r=init_rng(1234);
   int ii;
   double m_01=0,s_01=0;
@@ -178,14 +178,14 @@ CTEST(nmt,rngs) {
   end_rng(r);
 }
 
-CTEST(nmt,my_linecount) {
+CTEST_SKIP(nmt,my_linecount) {
   FILE *f=my_fopen("test/cls.txt","r");
   int cnt=my_linecount(f);
   ASSERT_EQUAL(cnt,768);
   fclose(f);
 }
 
-CTEST(nmt,errors) {
+CTEST_SKIP(nmt,errors) {
   set_error_policy(THROW_ON_ERROR);
 
   printf("\nError messages expected: \n");
@@ -193,16 +193,16 @@ CTEST(nmt,errors) {
   //File doesn't exist
   try { FILE *f=my_fopen("test/cls.txtb","r"); }
[…]

### 2018-08-06 test/nmt_test_hpex.c
basic SHT tests

@@ -4,8 +4,103 @@
 #include "nmt_test_utils.h"
 #include <chealpix.h>
 
-CTEST(nmt,he_io) {
+CTEST(nmt,alm2cl)
+{
+  int ii;
+  long nside=256;
+  int lmax=3*nside-1;
+  long npix=he_nside2npix(nside);
+  double **maps=my_malloc(3*sizeof(double *));
+  fcomplex **alms=my_malloc(3*sizeof(fcomplex *));
+  double **cls=my_malloc(7*sizeof(double *));
+  
+  for(ii=0;ii<7;ii++)
+    cls[ii]=my_malloc((lmax+1)*sizeof(double));
 
+  //Read data
+  for(ii=0;ii<3;ii++) {
+    long ns;
+    maps[ii]=he_read_healpix_map("test/maps.fits",&ns,ii);
+    ASSERT_EQUAL(nside,ns);
+    alms[ii]=my_malloc(he_nalms(lmax)*sizeof(fcomplex));
+  }
+
+  //SHT
+  he_map2alm(nside,lmax,1,0,maps,alms,3);
+  he_map2alm(nside,lmax,1,2,&(maps[1]),&(alms[1]),3);
+
+  //Power spectra
+  he_alm2cl(&(alms[0]),&(alms[0]),0,0,&(cls[0]),lmax);
+  he_alm2cl(&(alms[0]),&(alms[1]),0,1,&(cls[1]),lmax);
+  he_alm2cl(&(alms[1]),&(alms[1]),1,1,&(cls[3]),lmax);
+  for(ii=0;ii<7;ii++)
+    test_compare_arrays(lmax+1,cls[ii],7,ii,"test/benchmarks/cls_afst.txt",1E-5);
+
+  //Anafast
+  he_anafast(&(maps[0]),&(maps[0]),0,0,&(cls[0]),nside,lmax,3);
+  he_anafast(&(maps[0]),&(maps[1]),0,1,&(cls[1]),nside,lmax,3);
+  he_anafast(&(maps[1]),&(maps[1]),1,1,&(cls[3]),nside,lmax,3);
+  for(ii=0;ii<7;ii++)
+    test_compare_arrays(lmax+1,cls[ii],7,ii,"test/benchmarks/cls_afst.txt",1E-5);
+  
+  for(ii=0;ii<3;ii++) {
+    free(maps[ii]);
+    free(alms[ii]);
+  }
+
+  for(ii=0;ii<7;ii++)
+    free(cls[ii]);
+
+  free(cls);
[…]

### 2018-08-06 test/nmt_test_master.c

@@ -21,7 +21,7 @@ CTEST_SKIP(nmt,cell_00_full) {
 
   nmt_compute_coupled_cell(f0,f0,cell,3);
   nmt_decouple_cl_l(w00,cell,cell_noise,cell_noise,cell_out);
-  test_compare_arrays(bin->n_bands,cell_out[0],1,0,"test/benchmarks/cl_00_benchmark.txt");
+  test_compare_arrays(bin->n_bands,cell_out[0],1,0,"test/benchmarks/cl_00_benchmark.txt",1E-3);
   
   free(cell[0]); free(cell);
   free(cell_noise[0]); free(cell_noise);
@@ -60,7 +60,7 @@ CTEST_SKIP(nmt,cell_02_full) {
   nmt_compute_coupled_cell(f0,f2,cell,3);
   nmt_decouple_cl_l(w02,cell,cell_noise,cell_noise,cell_out);
   for(ipix=0;ipix<2;ipix++)
-    test_compare_arrays(bin->n_bands,cell_out[ipix],2,ipix,"test/benchmarks/cl_02_benchmark.txt");
+    test_compare_arrays(bin->n_bands,cell_out[ipix],2,ipix,"test/benchmarks/cl_02_benchmark.txt",1E-3);
   
   for(ipix=0;ipix<2;ipix++) {
     free(cell[ipix]);
@@ -106,7 +106,7 @@ CTEST_SKIP(nmt,cell_22_full) {
   nmt_compute_coupled_cell(f2,f2,cell,3);
   nmt_decouple_cl_l(w22,cell,cell_noise,cell_noise,cell_out);
   for(ipix=0;ipix<4;ipix++)
-    test_compare_arrays(bin->n_bands,cell_out[ipix],4,ipix,"test/benchmarks/cl_22_benchmark.txt");
+    test_compare_arrays(bin->n_bands,cell_out[ipix],4,ipix,"test/benchmarks/cl_22_benchmark.txt",1E-3);
   
   for(ipix=0;ipix<4;ipix++) {
     free(cell[ipix]);

### 2018-08-06 test/nmt_test_utils.h

@@ -6,6 +6,6 @@
 
 int *test_get_sequence(int n0,int nf);
 
-void test_compare_arrays(int n,double *y,int narr,int iarr,char *fname);
+void test_compare_arrays(int n,double *y,int narr,int iarr,char *fname,double rtol);
 
 #endif //_NMT_TEST_UTILS_

### 2018-08-16 doc/Paper/main.tex
Addressed some of Boris' and Scott's comments

@@
 \usepackage{dsfont}
+\usepackage{courier}
 
@@
 language=python,                % choose the language of the code
-basicstyle=\footnotesize,       % the size of the fonts that are used for the code
+basicstyle=\footnotesize\ttfamily,       % the size of the fonts that are used for the code
 numbers=left,                   % where to put the line-numbers
@@
 %\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
+ %\hypersetup{draft}
 \newcommand{\nv}{\hat{\bf n}}
@@
   current and future cosmological observations. This paper presents
-  \nmt, a public, validated, robust, accurate and easy-to-use software
+  \nmt, a public, validated, accurate and easy-to-use software
   package that, for the first time, provides a unified framework to
@@
   bias. \nmt~can be found at \url{https://github.com/LSSTDESC/NaMaster}, and is provided with
-  comprehensive documentation and a number of code examples to
-  maximize its usefulness in the community.
+  comprehensive documentation and a number of code examples.
 \end{abstract}
@@
   estimation easier. This has motivated a large body of literature
-  aimed at designing optimal and efficient power spectrum estimators.
+  aimed at designing optimal and efficient power spectrum estimators and 
+  used extensively in Cosmic Microwave Background (CMB) two-point measurements.
 
     While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information, their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ and $\ell_{\rm max}^4$ respectively with the maximum multipole $\ell_{\rm max}$.
-For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves a the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
-This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
+For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
+This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
 Unfortunately, in spite of the advantages of this method, no public, validated and well-maintained codes exist supporting all of these capabilities.
@@
         The pseudo-$C_\ell$ estimates will be optimal when these steps are close to what an optimal quadratic estimator would do.
-First, this requires that $\mC$ be close to diagonal, which is true when either the noise is large, or when the underlying power spectrum is close to white.
+First, this requires that $\mC$ be close to diagonal, which is true when either the noise is large and uncorrelated, or when the underlying power spectrum is close to white.
 Multiplication by the mask mixes modes, and replacement of multiplication by the response matrix with band-power averaging, only makes sense, when the mask is ``well-behaved'', i.e.
@@
       \end{align}
-      \emph{per purified field}. \todo{check this}.
+      per purified field. \todo{check this}.
 
@@
           where $W^{ab}_\ell$ is the product of the harmonic transform
-          of the beams for maps $a$ and $b$. Note that due to hairy
-          ball theorem, it is impossible to have an anisotropic beam
+          of the beams for maps $a$ and $b$. Note that since
+          any continuous tangent vector field on the sphere 
+          must, at some location, be zero (hairy ball theorem)
+          it is impossible to have an anisotropic beam
           that is uniform over the entire sphere.  However, it is
@@
           beam smoothing (to both data maps and contaminant maps) and taking
-          the ration. \as{How does this interact with E/B purification?}
+          the ratio. \as{How does this interact with E/B purification?}
 
@@
       \end{enumerate}
-      For this reason, \nmt\, also supports the computation of power spectra from flat-sky maps.
+      For this reason, \nmt, also supports the computation of power spectra from flat-sky maps.
 Appendix \ref{app:flat} describes the dictionary between spin-$s$ fields and their Fourier/harmonic coefficients defined in flat and curved skies, and we present the flat-sky versions of the pseudo-$C_\ell$ methods described in the previous sections here.
@@
 
-f2=NmtField(mask,[map_q,map_u],
-            templates=[[dust_q,dust_u]],
-            purify_b=True,beam=beam_ell)
-f0=NmtField(mask,[map_t])
-
-b=NmtBin(nside,nlb=10)
-
-wsp=NmtWorkspace()
-wsp.compute_coupling_matrix(f0,f2,b)
-
-cl_bias=deprojection_bias(f0,f2,cl_theory)
-
-cl_coupled=compute_coupled_cell(f0,f2)
-
-cl_decoupled=wsp.decouple_cell(cl_coupled,
-                               cl_bias)
+f2 = NmtField(mask, [map_q,map_u],
+            templates = [[dust_q,dust_u]],
+            purify_b = True,beam=beam_ell)
+f0 = NmtField(mask, [map_t])
+
+b = NmtBin(nside, nlb = 10)
+
+wsp = NmtWorkspace()
+wsp.compute_coupling_matrix(f0, f2, b)
+
+cl_bias = deprojection_bias(f0, f2, cl_theory)
+
+cl_coupled = compute_coupled_cell(f0, f2)
+
+cl_decoupled = wsp.decouple_cell(cl_coupled, cl_bias)
     \end{lstlisting}

### 2018-08-17 doc/Paper/main.tex
addressed Reviewer 2's comments

@@
  %\hypersetup{draft}
-\newcommand{\nv}{\hat{\bf n}}
+\newcommand{\nv}{\hat{\boldsymbol{\theta}}}
 \newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
@@
 \newcommand{\vb}{{\mathbf{b}}}
-\newcommand{\mC}{{\sf C}}
 \newcommand{\mM}{{\sf M}}
@@
 \begin{abstract}
-  The pseudo-$C_\ell$ is an algorithm for estimation of the angular
-  power and cross-power spectra that is very fast and in realistic
-  cases also nearly optimal. The algorithm can be extended to deal
-  with contaminant deprojection and $E/B$ purification, and can
-  therefore be applied in a wide variety of scenarios of interest for
-  current and future cosmological observations. This paper presents
-  \nmt, a public, validated, accurate and easy-to-use software
-  package that, for the first time, provides a unified framework to
-  compute angular cross-power spectra of any pair of spin-0 or spin-2
-  fields, contaminated by an arbitrary number of linear systematics
-  and requiring $B$- or $E$-mode purification, both in the sphere or
-  in the flat-sky approximation. We describe the mathematical
-  background of the estimator, including all the features above, and
-  its software implementation in \nmt. We construct a validation suite
-  that aims to resemble the types of observations that next-generation
-  large-scale structure and ground-based CMB experiments will face,
-  and use it to show that the code is able to recover the input power
-  spectra in the most complex scenarios with no detectable
-  bias. \nmt~can be found at \url{https://github.com/LSSTDESC/NaMaster}, and is provided with
-  comprehensive documentation and a number of code examples.
+  The pseudo-$C_\ell$ is an algorithm for estimation of the angular power and cross-power spectra that is very fast and in realistic cases also nearly optimal.
+The algorithm can be extended to deal with contaminant deprojection and $E/B$ purification, and can therefore be applied in a wide variety of scenarios of interest for current and future cosmological observations.
+This paper presents \nmt, a public, validated, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both in the sphere or in the flat-sky approximation.
+We describe the mathematical background of the estimator, including all the features above, and its software implementation in \nmt.
+We construct a validation suite that aims to resemble the types of observations that next-generation large-scale structure and ground-based CMB experiments will face, and use it to show that the code is able to recover the input power spectra in the most complex scenarios with no detectable bias.
+\nmt~can be found at \url{https://github.com/LSSTDESC/NaMaster}, and is provided with comprehensive documentation and a number of code examples.
 \end{abstract}
@@
   \section{Introduction}\label{sec:intro}
-  Two-point functions have proven to be the most useful summary
-  statistic for cosmology in terms of data compression, due to the
-  near-Gaussian nature of the modes of the cosmic density field that
-  can be easily analyzed to extract cosmological
-  information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since
-  the only true observables in any astronomical dataset are the
-  intensity and polarization as a function of frequency $\nu$ and sky
-  position $\nv$, angular 2-point correlators play a particularly
-  central role in most cosmological analyses, both in the form of
-  configuration-space two-point correlation functions $\xi(\theta)$
-  and as Fourier-space\footnote{In this paper we use the term
-    ``Fourier'' to describe both the flat space Fourier expansion as
-    well as harmonic-space expansion on the sphere. We will also use
-    the term ``power spectrum'' to describe both the auto power
-    spectrum of a field and cross-power spectra between two fields. }
-  power spectra $C_\ell$
-  \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.  Although
-  correlation functions are often easier to estimate from complex
-  observations~\citep{1993ApJ...412...64L}, a Fourier space analysis
-  is advantageous for many other reasons. The most important is that
-  Fourier space is a natural space for working with statistics of
-  translation-invariant fields. Therefore, Fourier space analysis has
-  a clearer separation between linear and non-linear scales, as well
-  as Gaussian and non-Gaussian
-  modes~\citep{2008PhRvD..77j3013H}. Moreover, individual band-power
-  measurements are typically significantly less correlated compared to
-  configuration-space analysis which makes covariance matrix
-  estimation easier. This has motivated a large body of literature
-  aimed at designing optimal and efficient power spectrum estimators and 
-  used extensively in Cosmic Microwave Background (CMB) two-point measurements.
+    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}.
+Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier-space\footnote{In this paper we use the term ``Fourier'' to describe both the flat space Fourier expansion as well as harmonic-space expansion on the sphere.
+We will also use the term ``power spectrum'' to describe both the auto power spectrum of a field and cross-power spectra between two fields.
+} power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.
+Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a Fourier space analysis is advantageous for many other reasons.
+The most important is that Fourier space is a natural space for working with statistics of translation-invariant fields.
+Therefore, Fourier space analysis has a clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}.
+Moreover, individual band-power measurements are typically significantly less correlated compared to configuration-space analysis which makes covariance matrix estimation easier.
+This has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators and used extensively in Cosmic Microwave Background (CMB) two-point measurements.
 
@@
 This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
-Unfortunately, in spite of the advantages of this method, no public, validated and well-maintained codes exist supporting all of these capabilities.
+Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward, no public, validated and well-maintained code exists supporting all of these capabilities consistently.
 This paper presents a public software package, \nmt, that unifies and extends all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.
@@
 Section \ref{sec:validation} presents the stress tests used to validate the code.
-We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicability in typical user-case scenarios in both fields.
+We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data, so this section illustrates not just its validity but also applicability in typical user-case scenarios in both fields.
 We conclude in Section \ref{sec:discussion}.
@@
       \begin{align}\label{eq:sht1}
-        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,_{s_a}{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
+        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
         {\bf a}(\nv)&=\mathcal{S}^{-1}\left[{\bf a}_{\ell m}\right]^{s_a}_{\nv}\equiv\sum_{\ell m}\,{\sf Y}^{s_a}_{\ell m}(\nv)\,{\bf a}_{\ell m}.
       \end{align}
-      Note that, in general we will use a vector notation such that, for a complex spin-$s_a$ field $a$, we form the vector ${\bf a}\equiv({\rm Re}(a),{\rm Im}(a))$.
-The harmonic coefficients above are decomposed in a similar manner into $E$ and $B$ modes: ${\bf a}_{\ell m}\equiv(a^E_{\ell m},a^B_{\ell m})$.
+      Note that, in general we will use vector notation such that, for a complex spin-$s_a$ field $a$, we form the vector ${\bf a}\equiv({\rm Re}(a),{\rm Im}(a))$.
+The harmonic coefficients above are decomposed in a similar manner into $E$ and $B$ modes: ${\bf a}_{\ell m}\equiv(a^E_{\ell m},a^B_{\ell m})$ (e.g.
+see Appendix \ref{app:shts} and \cite{1997PhRvD..55.1830Z}).
 The matrices ${\sf Y}^s_{\ell m}$ are defined in terms of the spin-weighed spherical harmonics, and are described in detail in Appendix \ref{app:shts}.
@@
       Note that for a general non-zero spin, vectors $\va$ and $\vb$ have two-components, each of which is a complex number.
-Nevertheless the rotational invariance requires that elements of the $2\times2$ matrix $\mC$ be real.
+Nevertheless the rotational invariance requires that elements of the $2\times2$ matrix ${\sf C}_\ell$ be real.
 
@@
 In general, the input data for measuring the power spectrum come pixelized on a sphere.
-Instead of dealing with an infinitely sampled points, we have a finite number of measurements
+Instead of dealing with an infinitely sampled field, we have a finite number of measurements
         \begin{equation}
@@
         where $\va_i^{\rm true}$ is is the true underlying field whose power spectrum we would like to estimate and ${\bf n}_i$ is the noise vector, which in general can be non-white and non-homogeneous.
-We will denote the pixel covariance of the noise component as ${\sf N}$.
+We will denote the pixel covariance of the noise component as $\tilde{\sf N}$\footnote{Note that the matrices $\tilde{\sf C}$ and $\tilde{\sf N}$ used in this section are the data and noise pixel-pixel covariance matrices, and therefore are different from the power spectrum matrices, ${\sf C}_\ell$ and ${\sf N}_\ell$, used in the rest of the paper (e.g.
+Eq.
+\ref{eq:cldef}).}.
 Note that for the sake of power-spectrum estimation the there is no conceptual difference between inhomogeneous noise and a survey mask: masked areas can be simply though of as areas of infinite noise.
@@
 
-        The underlying power spectrum in Eq.~\ref{eq:cldef} means that there is a non-trivial true underlying covariance matrix $\mC^{\rm true} = \left<\va^{\rm true} \va^{{\rm true}\dagger}\right>$, which in addition to the noise component leads to the actual data covariance $\mC = \left<\va \va^\dagger\right>=\mC^{\rm true} + {\sf N}$.
+        The underlying power spectrum in Eq.~\ref{eq:cldef} means that there is a non-trivial true underlying covariance matrix $\tilde{\sf C}^{\rm true} = \left<\va^{\rm true} \va^{{\rm true}\dagger}\right>$, which in addition to the noise component leads to the actual data covariance $\tilde{\sf C} = \left<\va \va^\dagger\right>=\tilde{\sf C}^{\rm true} + \tilde{\sf N}$.
 
@@
         \begin{equation}
-          {\sf F_{ij}} B_j = \va^\dagger \mC^{-1}
-          \left(\frac{\partial \mC}{\partial B_i} \right)
-          \mC^{-1} \va - \rm{Tr}\left[\left(\frac{\partial \mC}{\partial B_i}
-          \right) {\sf N} \right],
+          {\sf F_{ij}} B_j = \va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
+          \rm{Tr}\left[{\sf P}_i\,{\sf N} \right],
         \end{equation}
         where ${\sf F}$ is the Fisher matrix,the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
-        \begin{equation}
-          \mC = \sum_i \left(\frac{\partial \mC}{\partial B_i}\right) + {\sf N},
-        \end{equation}
-        and the last term correspond to bias of convolved estimates by the presence of noise.
-Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices $\partial {\sf C}/\partial B_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
+        \begin{equation}\label{eq:bpws_qmv}
+          \tilde{\sf C} = \sum_i {\sf P}_i\,B_i + {\sf N},
+        \end{equation}
+        and the last term in Eq.
+\ref{eq:bpws_qmv} correspond to bias of convolved estimates due to the presence of noise.
+Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices ${\sf P}_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
 The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix.
@@
         The pseudo-$C_\ell$ estimates will be optimal when these steps are close to what an optimal quadratic estimator would do.
-First, this requires that $\mC$ be close to diagonal, which is true when either the noise is large and uncorrelated, or when the underlying power spectrum is close to white.
-Multiplication by the mask mixes modes, and replacement of multiplication by the response matrix with band-power averaging, only makes sense, when the mask is ``well-behaved'', i.e.
+First, this requires that $\tilde{\sf C}$ be close to diagonal, which is true when either the noise is large and uncorrelated, or when the underlying power spectrum is close to white.
+Multiplication by the mask mixes modes, and replacement of multiplication by the response matrix with band-power averaging, makes sense only when the mask is ``well-behaved'', i.e.
 compact with no high-frequency structure in it.
@@
         where we have defined the vectorization operator \citep{2008PhRvD..77j3013H}
-        \begin{equation}
-          {\rm vec}\left[\left(\begin{array}{cc}
+        \begin{align}
+          &{\rm vec}\left[\left(\begin{array}{cc}
                           C^{EE} & C^{EB}\\
@@
           \end{array}
-          \right).
-        \end{equation}
+          \right),\\
+          &{\rm vec}\left[\left(\begin{array}{c}
+                                 C^{TE}\\C^{TB}
+                                \end{array}\right)\right]\equiv
+            \left(\begin{array}{c}
+                    C^{TE}\\C^{TB}
+            \end{array}\right),
+            \hspace{6pt}
+            {\rm vec}\left[C^{TT}\right]={\rm C}^{TT},
+        \end{align}
+        and $T$ here stands for any spin-$0$ field.
 
@@
       \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
-        Given the loss of information implicit in masking the originally full-sky field, it is in general not possible to invert Eq.
+        Given the loss of information implicit in masking the originally curved-sky field, it is in general not possible to invert the mode-coupling matrix in Eq.
 \ref{eq:modecoup} directly.
@@
           {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
-          &=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right].
-        \end{align}
+          \left\langle{\rm vec}\left[\tilde{\sf B}^{ab}_q\right]\right\rangle&=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
+        \end{align}
+        where in the second line we have taken the expectation value of first one.
 One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function).
 Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
+        \begin{figure*}
+          \centering
+          \includegraphics[width=0.9\textwidth]{maps_contamination}
+          \caption{Depiction of the contaminant deprojection process.
+{\sl Top left:} underlying true signal map (galaxy overdensity used in Section \ref{sssec:validation.suite.lss}).
+{\sl Top right:} sky contaminant (dust contamination used in Section \ref{sssec:validation.suite.lss}).
+{\sl Bottom left:} contaminated map obtained by adding the previous two maps.
+{\sl Bottom right:} cleaned map obtained through Eq.
+\ref{eq:clean_f}.
+Since only one contaminant is considered in this case, the cleaning is almost perfect.
+Residual contamination at the power spectrum level must still be subtracted in order to avoid a biased estimate.}
+          \label{fig:lss_cont}
+        \end{figure*}
         \begin{equation}\label{eq:decoup}
@@
     \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
-    In a typical data-analysis case, we deal with contaminants that
-    pollute the measured signal. Usually, we know something about
-    these contaminants. As an example, in case of CMB analysis, we
-    might have have a map of a given foreground at a different
-    frequency which can be marginalized out. But even when we do not,
-    we often have a tracer map of a given effect, for example, the map
-    of dust might tell us something about systematic errors related to
-    imperfect correction to dust reddening. Even if the dust map is
-    non-linearly and perhaps stochastically related to the amount of
-    reddening, as long as relation is local in space, we get a linear
-    contamination on large scales, very much like complex astrophysics
-    of galaxy formation nevertheless reduces to linear biasing on very
-    large scales (see Appendix \ref{app:lincont}). Thus, we generically deal with a
-    linear model of the kind:
+      In a typical data-analysis case, we deal with contaminants that pollute the measured signal.
+Usually, we know something about these contaminants.
+As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which can be marginalized out.
+But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added PSF fluctuations could be used to remove the possible correlations between measured ellipticities.
+Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}).
+Thus, we generically deal with a linear model of the kind:
       \begin{equation}\label{eq:aobs}
@@
 The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}.
+The process is illustrated in Figure \ref{fig:lss_cont}.
 A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}.
@@
       \begin{align}\label{eq:deproj_coeff}
-        &\hat{\alpha}_i=\sum_iF_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
-        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv).
-      \end{align}
-      The clean map is then given by
+        &\hat{\alpha}_i=F_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
+        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv),
+      \end{align}
+      where there is an implicit summation over the repeated index $j$ in Eq. \ref{eq:deproj_coeff} (which we will omit from now on). The clean map is then given by
+      \begin{align}
+        {\bf a}^c(\nv)
+        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')
+      \end{align}
+      which, when substituting Eq. \ref{eq:aobs}, yields
       \begin{align}\label{eq:clean}
-        {\bf a}^c(\nv)
-        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')\\\nonumber
-        &={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv'),
-      \end{align}
-      where there is an implicit summation over $i$ and $j$ (which we will omit from now on).
+        {\bf a}^c(\nv)&={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv').
+      \end{align}      
       The harmonic coefficients of ${\bf a}^c$ are therefore given by
@@
 
-      Due to the residual statistical contamination in Eq.
-\ref{eq:clean}, a direct computation of the pseudo-$C_\ell$ of the cleaned map would yield a biased estimator beyond the mode coupling induced by the mask $v$.
+      Although in general we expect the second term in Eq.
+\ref{eq:clean} to be very small, since the signal and contaminant maps are uncorrelated, due to this residual statistical contamination, a direct computation of the pseudo-$C_\ell$ of the cleaned map would yield a biased estimator beyond the mode coupling induced by the mask $v$.
 This bias can however be estimated analytically and corrected for.
@@
 
-      Two further points must be addressed before moving on. First,
-      the computation of the deprojection bias requires an estimate of
-      the true power spectrum ${\sf C}^{ab}_\ell$ (see
-      Eqs. \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}). In practice, this
-      can be done through an iterative approach, starting from an
-      initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the
-      estimated power spectrum in the previous iteration. A
-      sufficiently good guess will likely be enough to remove the bias
-      in a single iteration.  Alternatively, the ensemble average of
-      Eq. \ref{eq:dpbias0} can be seen as a convolution of
-      ${\sf C}^{ab}_\ell$, which could be treated by computing the
-      convolution matrix as a modified version of the standard
-      mode-coupling matrix that accounts for mode
-      deprojection. Estimating this convolution matrix in this case
-      becomes significantly more computationally intensive, and
-      therefore an iterative approach is preferred.
+      Two further points must be addressed before moving on.
+First, the computation of the deprojection bias requires an estimate of the true power spectrum ${\sf C}^{ab}_\ell$ (see Eqs.
+\ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}).
+In practice, this can be done through an iterative approach, starting from an initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the estimated power spectrum in the previous iteration.
+A sufficiently good guess will likely be enough to remove the bias in a single iteration.
+Alternatively, the ensemble average of Eq.
+\ref{eq:dpbias0} can be seen as a convolution of ${\sf C}^{ab}_\ell$, which could be treated by computing the convolution matrix as a modified version of the standard mode-coupling matrix that accounts for mode deprojection.
+Estimating this convolution matrix in this case becomes significantly more computationally intensive, and therefore an iterative approach is preferred.
 
@@
     \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
-    As described in
-    \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B},
-    $B$-mode purification refers to the map-level removal of the
-    contamination from $E$-modes in a the $B$-mode component of a
-    given map caused by an incomplete sky coverage and vice-versa. The
-    procedure is particularly useful in situations in which the
-    $E$-mode component of the signal is significantly larger than the
-    $B$ modes, as is the case in CMB. In this case, removing the
-    leakage at the power spectrum level (i.e. the standard
-    pseudo-$C_\ell$ approach) produces a sub-optimal estimator in
-    which the variance in the $B$-mode power spectrum is dominated by
-    the variance of the leaked $E$-modes. This section presents the
-    purification method and the associated modifications to the
-    pseudo-$C_\ell$ algorithm as described in the references above. In
-    this section we describe the algorithm in terms of purifying
-    $B$-modes from $E$-mode contamination, but we note that the
-    reverse, i.e. purifying $E$-modes from $B$-mode contamination is
-    conceptually exactly the same. Our discussion is also
-    specific for the spin-2 fields.
+      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa\footnote{Note that this is a different effect from the leakage between different polarized and unpolarized channels in CMB observations due to instrumental effects (asymmetric beams, pair differencing, polarization angle calibration etc.)}.
+The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB.
+In this case, removing the leakage at the power spectrum level (i.e.
+the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes.
+This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above.
+In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e.
+purifying $E$-modes from $B$-mode contamination is conceptually exactly the same.
+Our discussion is also specific for the spin-2 fields.
 
@@
 
-      Let $w(\nv)$ be a sky window function defining the sky region to be analyzed (and the weight to be applied in each pixel). The standard pseudo $B$-mode of a field ${\bf P}$ is then given by
+      Let $w(\nv)$ be a sky window function defining the sky region to be analyzed (and the weight to be applied in each pixel).
+The standard pseudo $B$-mode of a spin-2 field ${\bf P}$ is then given by
       \begin{align}\nonumber
@@
       \begin{equation}\label{eq:pureb}
-        B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}}\left(\tilde{P}_1\right)^B_{\bf l}+\alpha_{\ell,2}\left(\tilde{P}_0\right)^B_{\bf l},
-      \end{equation}
-      where $(a)^B_{\bf l}$ stands for the $B$-mode component of field ${\bf a}$, and we have defined the fields $\tilde{P}_n=(\eth^{2-n}w)^*(Q+iU)$, where $Q$ and $U$ are the real and imaginary parts of the field $P$ (see Appendix \ref{app:shts} for the definitions of $\eth$ and $\alpha_{\ell,s}$).
+        B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2\frac{\beta_{\ell,2}}{\beta_{\ell,1}}\left(\tilde{P}_1\right)^B_{\bf l}+\beta_{\ell,2}\left(\tilde{P}_0\right)^B_{\bf l},
+      \end{equation}
+      where $(a)^B_{\bf l}$ stands for the $B$-mode component of field ${\bf a}$, and we have defined the fields $\tilde{P}_n=(\eth^{2-n}w)^*(Q+iU)$, where $Q$ and $U$ are the real and imaginary parts of the field $P$ (see Appendix \ref{app:shts} for the definitions of $\eth$ and $\beta_{\ell,s}$).
 
@@
       \begin{align}
-        \eth^nw={\cal S}^{-1}\left[\left(-\frac{w_{\bf l}}{\alpha_{\ell,n}},0\right)\right],
+        \eth^nw={\cal S}^{-1}\left[\left(-\frac{w_{\bf l}}{\beta_{\ell,n}},0\right)\right],
       \end{align}
@@
       \begin{align}\nonumber
-        &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}+2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}\alpha_{\ell'',1}}\wtj{\ell}{\ell'}{\ell''}{1}{-2}{1}\\
-        &\,\,\,+\frac{\alpha_{\ell,2}}{\alpha_{\ell'',2}}\wtj{\ell}{\ell'}{\ell''}{0}{-2}{2}.
+        &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}+2\frac{\beta_{\ell,2}}{\beta_{\ell,1}\beta_{\ell'',1}}\wtj{\ell}{\ell'}{\ell''}{1}{-2}{1}\\
+        &\,\,\,+\frac{\beta_{\ell,2}}{\beta_{\ell'',2}}\wtj{\ell}{\ell'}{\ell''}{0}{-2}{2}.
       \end{align}
@@
 a stochastic contaminant that we do not have a template for).
-Furthermore, the signal component will often be smoothed on the smallest scales by an instrumental beam.
+Furthermore, the signal component will often be smoothed on the smallest scales by an instrumental beam (although this latter effect is less relevant for galaxy surveys).
 We discuss how to deal with both of these complications here.
@@
           \end{equation}
-          where $W^{ab}_\ell$ is the product of the harmonic transform
-          of the beams for maps $a$ and $b$. Note that since
-          any continuous tangent vector field on the sphere 
-          must, at some location, be zero (hairy ball theorem)
-          it is impossible to have an anisotropic beam
-          that is uniform over the entire sphere.  However, it is
-          possible that the beam is both anisotropic and spatially
-          varying. In this case, the pseudo-$C_\ell$ power spetrum
-          estimation will produce biased estimates that will be
-          suppressed by an \emph{effective} beam. Correcting for the
-          beam in such cases can also be done analytically, but we
-          leave it for future work.  In practice, if required, the
-          suppression factor can be found by applying the PCL
-          estimator to a a set of mock input maps with the same
-          underlying signal but  with and without application of the
-          beam smoothing (to both data maps and contaminant maps) and taking
-          the ratio. \as{How does this interact with E/B purification?}
+          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$.
+Note that since any continuous tangent vector field on the sphere must, at some location, be zero (the so-called hairy ball theorem), it is impossible to have an anisotropic beam that is uniform over the entire sphere.
+However, it is possible that the beam is both anisotropic and spatially-varying.
+In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam.
+Correcting for the beam in such cases can also be done analytically, but we leave this for future work.
+In practice, if required, the
+          suppression factor can be found by applying the PCL estimator to a a set of mock input maps with the same underlying signal but  with and without application of the beam smoothing (to both data maps and contaminant maps) and taking the ratio.
 
@@
          \paragraph*{Uncorrelated noise.}
-           If the noise is homogeneous but uncorrelated (i.e.
+           If the noise is inhomogeneous but uncorrelated (i.e.
 $\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle=\mI\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
@@
                      {\rm PCL}_\ell({\bf n},{\bf n})&=\mI\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
-                                                    &=\mI\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
-                   \end{align}
-                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=\alpha\sigma_n^{-2}(\nv)$.
+                                                    &=\mI A^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
+                   \end{align}
+                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=A\,\sigma_n^{-2}(\nv)$.
              \item The additional bias from contaminant deprojection can also be computed analytically, starting from Eq. \ref{eq:dpbias0}, as:
@@
     \subsection{Flat-sky pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.flat}
-      Using the full-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
+      Using the curved-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
       \begin{enumerate}
-        \item Full-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch.
+        \item Curved-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch.
 This can lead to an inefficient use of memory and storage space, because the standard numerical implementations of spherical transforms operate on full-sky data.
@@
 ${\cal O}(N_{\rm pix}\log(N_{\rm pix}))$ for DFTs) and because by SHTs involve operations over the full sphere.
-        \item Fourier transforms over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
+        \item Fourier transforms over a flat patch offer exact quadrature, which makes them inherently more numerically stable\footnote{Note that it is also possible to achieve exact quadrature with certain spherical pixelization schemes (e.g.
+\cite{2011IJMPD..20.1053D}).}.
       \end{enumerate}
-      For this reason, \nmt, also supports the computation of power spectra from flat-sky maps.
+      For this reason, \nmt\, also supports the computation of power spectra from flat-sky maps.
 Appendix \ref{app:flat} describes the dictionary between spin-$s$ fields and their Fourier/harmonic coefficients defined in flat and curved skies, and we present the flat-sky versions of the pseudo-$C_\ell$ methods described in the previous sections here.
@@
         \end{equation}
+        Here, ${\cal D}$ stands for a standard 2D discrete Fourier transform, and ${\sf R}$ is a rotation matrix in 2 dimensions.
 Note that, in the context of flat-sky fields, the vector ${\bf l}$ is a 2D wave vector, and must not be confused with the abbreviation ${\bf l}\equiv(\ell,m)$ used in previous sections.
-        It is then straightforward to estimate the covariance of two masked field at the same wavenumber:
+The quantity ${\bf a}_{\bf l}$ is therefore the Fourier transform of the real-space flat-sky field ${\bf a}({\bf x})$, where ${\bf x}$ is the angular displacement with respect to a given reference point.
+Using Eq.
+\ref{eq:pcl_flat_masked} it is then straightforward to estimate the covariance of two masked fields at the same wavenumber:
         \begin{equation}\label{eq:flat_coup}
@@
 
-        At this stage it is natural to connect directly with the final bandpowers since, unlike in the full-sky case, there is no natural minimal binning ($\Delta\ell=1$) for the un-masked power spectrum.
+        At this stage it is natural to connect directly with the final bandpowers since, unlike in the curved-sky case, there is no natural minimal binning ($\Delta\ell=1$) for the un-masked power spectrum.
 In this case we will define a bandpower ${\sf B}_q$, indexed by an integer $q$ as the average of the covariance above over a set of ${\bf l}$ values $S_q$:
@@
 Taking the continuum approximation is unfortunately not accurate enough in practice, as described in \cite{2016arXiv161204664A}, and a fully discrete approach is usually necessary.
-This slower performance is normally compensated by the smaller pixel numbers and faster DFTs that can be used in the flat sky, so that the flat-sky approximation is still preferable over a full-sky treatment for small sky patches.
+This slower performance is normally compensated by the smaller pixel numbers and faster DFTs that can be used in the flat sky, so that the flat-sky approximation is still preferable over a curved-sky treatment for small sky patches.
 Appendix \ref{app:flat.pcl} discusses these issues in detail, providing the continuum-limit expressions mentioned above.
@@
     \begin{itemize}
-      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations.
+      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for curved-sky calculations.
 Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citep{1999ASPC..172..487P}.
-      \item The basic full-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citep{2013A&A...554A.112R}.
+      \item The basic curved-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citep{2013A&A...554A.112R}.
       \item In flat sky, SHTs are replaced by 2D DFTs, for which \nmt~uses the {\tt FFTW} library~\citep{FFTW05}.
@@
             templates = [[dust_q,dust_u]],
-            purify_b = True,beam=beam_ell)
+            purify_b = True, beam=beam_ell)
 f0 = NmtField(mask, [map_t])
@@
       \item All preliminary I/O operations (reading maps, masks etc.) take place implicitly in line 3.
-      \item In line 5 we generate a field with spin 2.
+      \item In line 5 we create a field with spin 2.
 The field is defined in terms of a mask and maps of the $(Q,U)$ Stokes parameters.
@@
       \item The pseudo-$C_\ell$ is computed in line 17.
-      \item The deconvolved bandpowers are finally estimated using the workspace in line 19.
+      \item The deconvolved bandpowers are finally estimated using the workspace in line 19. Note that the mode-coupling matrix must have been precomputed in line 13 for this to be successful.
     \end{enumerate}
@@
 This section presents these two examples and the results of this validation in terms of accuracy and computational performance.
-Results will be presented for both full-sky and flat-sky realizations.
+Results will be presented for both curved-sky and flat-sky realizations.
 
@@
 All other cross-correlations are zero.
-The dashed lines show the noise power spectra, associated to shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
+The dashed lines show the noise power spectra, associated with shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
         \end{figure}
@@
           \includegraphics[width=0.9\textwidth]{maps_lss_sph}
-          \caption{Inputs for the large-scale structure, full-sky validation set.
+          \caption{Inputs for the large-scale structure, curved-sky validation set.
 From top to bottom and left to right: sky mask (1), Gaussian realization of the galaxy overdensity $\delta$ (2) and cosmic shear $\gamma_i$ (3,4), dust contaminant for $\delta$ (5), star contaminant for $\delta$ (6), PSF contaminant for $\gamma$ (7, $Q$ component) and small-scale contaminant for $\gamma$ (8, $Q$ component).}
@@
         For this validation suite we therefore use a mock dataset mimicking a typical observation of an LSST-like survey in a particular redshift bin.
-Slightly different set-ups are used for the full-sky and flat-sky validation suites.
+Slightly different set-ups are used for the curved-sky and flat-sky validation suites.
 The characteristics of these data are as follows:
@@
 \ref{fig:lss_cls}.
-Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the full-sky and flat-sky cases respectively.
-          \item {\bf Mask:} for the full-sky data, we use a mask built
-            as a combination of the sky coverage output by the LSST
-            {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a
-            more conservative galactic cut based on the reddening map
-            of \citet{1998wfsc.conf..297S}. To add an extra layer of
-            complexity, we drill 100 additional 1-degree holes on the
-            resulting combined mask. We also explore the improvement
-            stemming from the down-weighting of higher-noise in an
-            inverse-variance way in the presence of inhomogeneous
-            noise (see description below), which is the reason why the
-            default version of this mask is not binary, but tracers
-            the inverse of the noise variance (see top left panel of
-            Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we
-            use a high-resolution mask constructed from the
-            bright-object mask used for the first public data release
-            from the HSC collaboration~\citep{2018PASJ...70S...4A}.
+Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the curved-sky and flat-sky cases respectively.
+          \item {\bf Mask:} for the curved-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}.
+To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask.
+We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig.
+\ref{fig:lss_mocks_sph}).
+For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citep{2018PASJ...70S...4A}.
 The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales.
@@
           \item {\bf Contaminants:} we contaminate the signal maps described above with several different types of residuals.
-Maps of these are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.
-\ref{fig:lss_cont_cls} for the full-sky case.
+In all cases we add contaminants linearly, to avoid deviating from the model described in Section \ref{ssec:maths.deproj}.
+Maps of these contaminants are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.
+\ref{fig:lss_cont_cls} for the curved-sky case.
           \begin{itemize}
-            \item {\sl Dust.} For the full-sky suite, we include a linear contaminant for $\delta_g$ proportional to the fluctuations around the mean of the dust reddening map of~\citet{1998wfsc.conf..297S}.
+            \item {\sl Dust.} For the curved-sky suite, we include a linear contaminant for $\delta_g$ proportional to the fluctuations around the mean of the dust reddening map of~\citet{1998wfsc.conf..297S}.
 This contamination affects mostly the largest scales, and simulates the effect of an imperfect correction of the dust absorption on the galaxy number density.
 The amplitude of the dust fluctuations is chosen so that their power spectrum within the unmasked region be $\sim20\%$ of the galaxy power spectrum at $\ell\sim120$.
-For the flat-sky suite, we generate a flat-sky Gaussian realization of a field with a power spectrum $C_\ell\propto\ell^{-2.4}$, corresponding to the large-$\ell$ behaviour of the full-sky dust map.
+For the flat-sky suite, we generate a flat-sky Gaussian realization of a field with a power spectrum $C_\ell\propto\ell^{-2.4}$, corresponding to the large-$\ell$ behaviour of the curved-sky dust map.
 The amplitude was chosen such that the contaminant power spectrum be $\sim10\%$ of the signal at $\ell\sim400$.
             \item {\sl Stars.} We add a second contaminant to $\delta_g$ associated with star contamination.
-For the full-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citep{2014SPIE.9150E..14C}.
+For the curved-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citep{2014SPIE.9150E..14C}.
 We modify this contaminant slightly by adding a white-noise component that dominates the small-scale spectrum beyond $\ell\simeq400$.
@@
             where $J_1(x)$ is the order-1 Bessel function, and $2J_1(\ell\theta)/\ell\theta$ is the Fourier transform of a circular aperture with radius $\theta$.
-For this we assume $\theta_{\rm FoV}=1.75^\circ$.
+For this we assume $\theta_{\rm FoV}=1.75^\circ$, and therefore we try to mimic the signature of PSF fluctuations that vary between different pointings with a 3.5$^\circ$ diameter.
 The amplitude of the power spectrum is chosen so that the contaminant amounts to $30\%$ of the signal power spectrum at $\ell\sim50$.
-We use the same prescription for both the full-sky and the flat-sky cases.
+We use the same prescription for both the curved-sky and the flat-sky cases.
             \item {\sl Small-scale contamination.} Finally we include an additional small-scale contaminant to weak lensing with a flat power spectrum, unit $E$-$B$ ratio, and an amplitude fixed so the contaminant's power spectrum is $20\%$ of the signal at $\ell=500$.
@@
           \end{itemize}
-          \item {\bf Noise:} for the full-sky case we generate inhomogeneous noise realizations.
+          \item {\bf Noise:} for the curved-sky case we generate inhomogeneous noise realizations.
 To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is then intrinsic ellipticity scatter per component.
@@
 
-          \item {\bf Pixelization:} the main suite of full-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$.
+          \item {\bf Pixelization:} the main suite of curved-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$.
 This resolution allows us to validate the code on a large number of realizations without facing strong computational challenges on the main scales of interest ($\ell\lesssim2000$).
@@
           \includegraphics[width=0.99\columnwidth]{cls_cmb}
-          \caption{$E$-mode and $B$-mode power spectra (blue and yellow respectively) for the cosmological CMB signal (solid lines), for the full-sky foreground contaminants (dashed lines) and for the flat-sky foreground contaminants (dot-dashed lines).
+          \caption{$E$-mode and $B$-mode power spectra (blue and yellow respectively) for the cosmological CMB signal (solid lines), for the curved-sky foreground contaminants (dashed lines) and for the flat-sky foreground contaminants (dot-dashed lines).
 The noise power spectra used for the full- and flat-sky simulations are shown as black dashed and dot-dashed lines respectively.
@@
           \includegraphics[width=0.9\textwidth]{maps_cmb}
-          \caption{{\sl Left panels:} mask (top), and $Q$ and $U$ components (middle, bottom) of the foreground contaminant used in the CMB full-sky validation set.
+          \caption{{\sl Left panels:} mask (top), and $Q$ and $U$ components (middle, bottom) of the foreground contaminant used in the CMB curved-sky validation set.
 {\sl Right panels:} same as left panels for the flat-sky validation set.}
@@
 This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citep{2016PhRvL.116c1302B}, Advanced ACTPol~\citep{2016SPIE.9910E..14D,2017JCAP...06..031L} or the Simons Array~\citep{2016JLTP..184..805S,2017ApJ...848..121P}, as well as next-generation facilities such as the Simons Observatory\todo{cite if out} or CMB Stage-4~\citep{2016arXiv161002743A}.
-Potential future space missions, such as LiteBIRD~\citep{2018JLTP..tmp..124S} will be able to make low-resolution, full-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
+Potential future space missions, such as LiteBIRD~\citep{2018JLTP..tmp..124S} will be able to make low-resolution, curved-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
 Ground-based experiments, on the other hand, will be systematics-limited at low $\ell$, and will be able to make higher-resolution maps, for which pseudo-$C_\ell$ methods are better suited.
@@
 \ref{fig:cmb_cls}.
-          \item {\bf Mask:} for the full-sky case we assume a survey targeting a 4000 deg$^2$ footprint.
+          \item {\bf Mask:} for the curved-sky case we assume a survey targeting a 4000 deg$^2$ footprint.
 The corresponding mask is analytically apodized using the ``$C^2$'' apodization method of \cite{2009PhRvD..79l3515G} with an apodization scale of 10 degrees.
+This apodization method is based on smoothing the transition at the mask edge with a sinusoidal function that interpolates between 0 and 1 in a differentiable manner.
 The observed area encompasses one of the lowest-foreground regions observable from the Southern Hemisphere.
@@
 The assumed masks are shown in the top panels of Figure~\ref{fig:cmb_mocks}.
-          \item {\bf Noise:} for the full-sky and flat-sky case we assume white noise levels of 1 and 0.5$\mu K\,{\rm arcmin}$ in intensity respectively.
+          \item {\bf Noise:} for the curved-sky and flat-sky case we assume white noise levels of 1 and 0.5$\mu K\,{\rm arcmin}$ in intensity respectively.
 We also include a $1/f$-like component associated with atmospheric noise with a characteristic scale $\ell_{\rm knee}$ and tilt $\alpha_{\rm knee}$ such that the noise power spectrum can be written as
@@
           where $\sigma_N^2$ is the white-noise variance in one steradian.
-We use $\alpha_{\rm knee}=2.4$ and $\ell_{\rm knee}=10$ and 300 for the full-sky and flat-sky cases respectively.
+We use $\alpha_{\rm knee}=2.4$ and $\ell_{\rm knee}=10$ and 300 for the curved-sky and flat-sky cases respectively.
 The beam-deconvolved noise power spectra in both cases are shown in Fig.
 \ref{fig:cmb_cls}.
-          \item {\bf Beam:} we assume a Gaussian beam with full-width at half-maximum of $\theta_{\rm FWHM}=20$ and 1.4 arcmin for the full-sky and flat-sky cases respectively.
+          \item {\bf Beam:} we assume a Gaussian beam with full-width at half-maximum of $\theta_{\rm FWHM}=20$ and 1.4 arcmin for the curved-sky and flat-sky cases respectively.
 This resolution corresponds to what could be achieved at 150 GHz with with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case.
@@
 \ref{fig:cmb_cls}, the combination of beam size and $\ell_{\rm knee}$ makes the two experiments sensitive to the $B$-mode power spectrum in different ranges of scales, corresponding to the primordial $B$-mode signal around the recombination bump (S.A.) and to the high-ell lensing signal (L.A.).
-          \item {\bf Contaminants:} we assume a dust-like residual for both the full-sky an flat-sky cases.
-For the full-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citep{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$.
+          \item {\bf Contaminants:} we assume a dust-like residual for both the curved-sky an flat-sky cases.
+For the curved-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citep{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$.
 For the flat-sky signal we generate a Gaussian realization of a dust-like component with power spectra
@@
 \ref{fig:cmb_cls}.
-          \item {\bf Pixelization:} the full-sky case, targetting the $B$-mode recombination bump at $\ell\sim100$, uses a HEALPix resolution parameter $N_{\rm side}=256$, corresponding to $14$-arcmin pixels and sufficient to measure power spectra to multipoles $\ell\sim500$.
+          \item {\bf Pixelization:} the curved-sky case, targetting the $B$-mode recombination bump at $\ell\sim100$, uses a HEALPix resolution parameter $N_{\rm side}=256$, corresponding to $14$-arcmin pixels and sufficient to measure power spectra to multipoles $\ell\sim500$.
 The flat-sky case, targetting the lensing $B$-modes, uses square 2-arcmin pixels.
@@
 
-    \subsection{Full-sky large-scale structure validation}\label{ssec:validation.lss_full}
+    \subsection{Curved-sky large-scale structure validation}\label{ssec:validation.lss_full}
       \begin{figure*}
@@
         \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
-        \caption{{\sl Left panel:} average over all simulations in our full-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
+        \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
 The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
@@
         \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
-        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our full-sky LSS validation suite (blue histograms).
+        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our curved-sky LSS validation suite (blue histograms).
 Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot.
@@
 \ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra (darker dashed lines).
-The middle panel of the same figure shows the average residual with respect to the analytical prediction normalize by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).
+The middle panel of the same figure shows the average residual with respect to the analytical prediction normalized by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).
 We see that the latter are accurately recovered by \nmt~in all cases, including all null correlations ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$), and that the normalized residuals fluctuate around zero with typical variations of $\Delta C_\ell/\sigma_\ell\lesssim3$, corresponding to expected $\lesssim3\sigma$ fluctuations.
-For comparison, bottom panel of the figure show the mean relative difference for power spectra of contaminated maps without deprojection, where we can appreciate that the residuals, $\Delta C'_{\ell}/\sigma'_{\ell}$, appear to not be centered around zero in some cases, i.e., the estimated power-spectra are biased.
+For comparison, the bottom panel of the figure shows the mean relative difference for power spectra of contaminated maps without deprojection, where we can appreciate that the residuals, $\Delta C'_{\ell}/\sigma'_{\ell}$, appear to not be centered around zero in some cases, i.e., the estimated power-spectra are biased.
 
@@
         \caption{{\sl Left:} average over all simulations in our
-          full-sky CMB validation suite of the three spin-2 power
+          curved-sky CMB validation suite of the three spin-2 power
           spectra ($EE$, $EB$ and $BB$, solid lines) using mode
@@
       degrees of freedom. Figure \ref{fig:lss_val_chi2} shows the
-      distribution of $\chi^2_i$ across simulations approximating
+      distribution of $\chi^2_i$ across simulations approximating the
       covariance matrix as diagonal (blue histograms),
@@
 
-    \subsection{Full-sky CMB \texorpdfstring{$B$}{B}-mode validation}\label{ssec:validation.cmb_full}
+    \subsection{Curved-sky CMB \texorpdfstring{$B$}{B}-mode validation}\label{ssec:validation.cmb_full}
       \begin{figure*}
@@
         \includegraphics[width=0.99\columnwidth]{val_covar_cmb_sph}
-        \caption{{\sl Left:} covariance matrix for power spectra computed in the absence of foreground contaminants computed from the 1000 simulations of our full-sky CMB validation suite.
+        \caption{{\sl Left:} covariance matrix for power spectra computed in the absence of foreground contaminants computed from the 1000 simulations of our curved-sky CMB validation suite.
 The incomplete sky coverage causes a noticeable anti-correlation of neighboring bandpowers at low $\ell$ for the three power spectra ($EE$, $EB$ and $BB$), but beyond that, the covariance matrix is close to diagonal.
@@
         \includegraphics[width=0.80\textwidth]{val_chi2_cmb_sph}
-        \caption{Distribution of $\chi^2$ values found for the 1000 simulations in our full-sky CMB validation suite (blue histograms).
+        \caption{Distribution of $\chi^2$ values found for the 1000 simulations in our curved-sky CMB validation suite (blue histograms).
 Results are shown for the three polarized spectra ($EE$, $EB$ and $BB$), labeled in the top right corner of each panel.
@@
 
-    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalizations (flat-sky, deprojection and purification).
+    This paper presents \nmt, the first public, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalizations (flat-sky, deprojection and purification).
 We have described the steps taken to validate the code, showing that it is able to provide unbiased results in a wide variety of scenarios.
@@
     \begin{equation}
-      _sY_{\ell m}\equiv \alpha_{\ell,s} \eth^s Y_{\ell m},\hspace{6pt}
-      _{-s}Y_{\ell m}\equiv \alpha_{\ell,s} (-1)^s\bar{\eth}^s Y_{\ell m},
-    \end{equation}
-    where $\alpha_{\ell,s}\equiv\sqrt{(\ell-s)!/(\ell+s)!}$ and $Y_{\ell m}$ are the standard spherical harmonics. These functions satisfy the property: $(_sY_{\ell m})^*=(-1)^{s+m}\,_{-s}Y_{\ell-m}$.
+      _sY_{\ell m}\equiv \beta_{\ell,s} \eth^s Y_{\ell m},\hspace{6pt}
+      _{-s}Y_{\ell m}\equiv \beta_{\ell,s} (-1)^s\bar{\eth}^s Y_{\ell m},
+    \end{equation}
+    where $\beta_{\ell,s}\equiv\sqrt{(\ell-s)!/(\ell+s)!}$ and $Y_{\ell m}$ are the standard spherical harmonics. These functions satisfy the property: $(_sY_{\ell m})^*=(-1)^{s+m}\,_{-s}Y_{\ell-m}$.
 
@@
       _s{\bf Y}^E_{\ell m}&\equiv {\bf D}^E_sY_{\ell m}
-      \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
+      \equiv-\frac{\beta_{\ell,s}}{2}\left(\begin{array}{c}
                               \eth^s+\bar{\eth}^s\\
@@
           _s{\bf Y}^B_{\ell m}&\equiv {\bf D}^B_sY_{\ell m}
-          \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
+          \equiv-\frac{\beta_{\ell,s}}{2}\left(\begin{array}{c}
                               i(\eth^s-\bar{\eth}^s)\\
@@
 
-      In analogy with our full-sky nomenclature, let us now write $a$ as a vector such that in real space ${\bf a}({\bf x})\equiv({\rm Re}(a),{\rm Im}(a))$, and in Fourier space ${\bf a}_{\bf l}\equiv(\,_sE_{\bf l},\,_sB_{\bf l})$.
+      In analogy with our curved-sky nomenclature, let us now write $a$ as a vector such that in real space ${\bf a}({\bf x})\equiv({\rm Re}(a),{\rm Im}(a))$, and in Fourier space ${\bf a}_{\bf l}\equiv(\,_sE_{\bf l},\,_sB_{\bf l})$.
 We can rewrite the Eq.
@@
 
-  \section{Full-sky validation for different pixel resolutions}\label{app:full_sky_nside}
+  \section{Curved-sky validation for different pixel resolutions}\label{app:full_sky_nside}
     We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of the HEALPix resolution parameter $N_{\rm side}$.
-In this appendix we show that we recover the correct power spectra using the inputs from our full-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).
+In this appendix we show that we recover the correct power spectra using the inputs from our curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).
 In particular we tested $N_{\rm side}$ ranging from 8 to 8192.
@@
       \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
-      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the full-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
+      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
       \label{fig:validation.fullsky.lss.4096}
@@
      One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
-In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
+In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the curved-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
 We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
@@
      \includegraphics[width=0.9\columnwidth]{time_scaling}
-     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
+     \caption{Computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
 We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}

### 2018-08-17 sandbox_validation/data/flatmaps.py

@@ -174,8 +174,8 @@ class FlatMapInfo(object) :
             ax=fig.add_subplot(111,projection=self.wcs)
         if title is not None :
             ax.set_title(title,fontsize=tfs)
-        image= ax.imshow(map_in.reshape([self.ny,self.nx]),
-			origin='lower', interpolation='nearest')
+        image= ax.imshow(map_in.reshape([self.ny,self.nx]),vmin=colorMin,vmax=colorMax,
+			 origin='lower', interpolation='nearest')
         if addColorbar :
             plt.colorbar(image)
         ax.set_xlabel(xlabel,fontsize=fs)

### 2018-08-20 doc/Paper/main.tex

--- 2018-08-16 doc/Paper/main.tex
@@
       \begin{align}\label{eq:sht1}
-        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,_{s_a}{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
+        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
         {\bf a}(\nv)&=\mathcal{S}^{-1}\left[{\bf a}_{\ell m}\right]^{s_a}_{\nv}\equiv\sum_{\ell m}\,{\sf Y}^{s_a}_{\ell m}(\nv)\,{\bf a}_{\ell m}.
@@
         \begin{equation}
-          {\sf F_{ij}} B_j = \va^\dagger \mC^{-1}
-          \left(\frac{\partial \mC}{\partial B_i} \right)
-          \mC^{-1} \va - \rm{Tr}\left[\left(\frac{\partial \mC}{\partial B_i}
-          \right) {\sf N} \right],
+          {\sf F_{ij}} B_j = \va^\dagger \mC^{-1}\,{\sf P}_i\,\mC^{-1} \va -
+          \rm{Tr}\left[{\sf P}_i\,{\sf N} \right],
         \end{equation}
@@
         \begin{equation}
-          \mC = \sum_i \left(\frac{\partial \mC}{\partial B_i}\right) + {\sf N},
+          \mC = \sum_i {\sf P}_i\,B_i + {\sf N},
         \end{equation}
         and the last term correspond to bias of convolved estimates by the presence of noise.
-Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices $\partial {\sf C}/\partial B_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
+Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices ${\sf P}_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
 The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix.
@@
       \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
-        Given the loss of information implicit in masking the originally full-sky field, it is in general not possible to invert Eq.
+        Given the loss of information implicit in masking the originally full-sky field, it is in general not possible to invert the mode-coupling matrix in Eq.
 \ref{eq:modecoup} directly.
@@
 Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the full-sky and flat-sky cases respectively.
-          \item {\bf Mask:} for the full-sky data, we use a mask built
-            as a combination of the sky coverage output by the LSST
-            {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a
-            more conservative galactic cut based on the reddening map
-            of \citet{1998wfsc.conf..297S}. To add an extra layer of
-            complexity, we drill 100 additional 1-degree holes on the
-            resulting combined mask. We also explore the improvement
-            stemming from the down-weighting of higher-noise in an
-            inverse-variance way in the presence of inhomogeneous
-            noise (see description below), which is the reason why the
-            default version of this mask is not binary, but tracers
-            the inverse of the noise variance (see top left panel of
-            Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we
-            use a high-resolution mask constructed from the
-            bright-object mask used for the first public data release
-            from the HSC collaboration~\citep{2018PASJ...70S...4A}.
+          \item {\bf Mask:} for the full-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}.
+To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask.
+We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig.
+\ref{fig:lss_mocks_sph}).
+For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citep{2018PASJ...70S...4A}.
 The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales.
@@
           \item {\bf Contaminants:} we contaminate the signal maps described above with several different types of residuals.
-Maps of these are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.
+In all cases we add contaminants linearly, to avoid deviating from the model described in Section \ref{ssec:maths.deproj}.
+Maps of these contaminants are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.
 \ref{fig:lss_cont_cls} for the full-sky case.
@@
             where $J_1(x)$ is the order-1 Bessel function, and $2J_1(\ell\theta)/\ell\theta$ is the Fourier transform of a circular aperture with radius $\theta$.
-For this we assume $\theta_{\rm FoV}=1.75^\circ$.
+For this we assume $\theta_{\rm FoV}=1.75^\circ$, and therefore we try to mimic the signature of PSF fluctuations that vary between different pointings with a 3.5$^\circ$ diameter.
 The amplitude of the power spectrum is chosen so that the contaminant amounts to $30\%$ of the signal power spectrum at $\ell\sim50$.
@@
 The corresponding mask is analytically apodized using the ``$C^2$'' apodization method of \cite{2009PhRvD..79l3515G} with an apodization scale of 10 degrees.
+This apodization method is based on smoothing the transition at the mask edge with a sinusoidal function that interpolates between 0 and 1 in a differentiable manner.
 The observed area encompasses one of the lowest-foreground regions observable from the Southern Hemisphere.

### 2018-08-20 src/nmt_mask.c
small bug in apotype error reporting

@@ -142,6 +142,6 @@ void nmt_apodize_mask(long nside,flouble *mask_in,flouble *mask_out,flouble apos
     else if(!strcmp(apotype,"Smooth")) 
       apodize_mask_smooth(nside,mask_in,mask_out,aposize);
     else
-      report_error(NMT_ERROR_APO,"Unknown apodization type %s. Allowed: \"Smooth\", \"C1\", \"C2\"\n");
+      report_error(NMT_ERROR_APO,"Unknown apodization type %s. Allowed: \"Smooth\", \"C1\", \"C2\"\n",apotype);
   }
 }

### 2018-08-20 src/nmt_mask_flat.c

@@ -164,7 +164,7 @@ void nmt_apodize_mask_flat(int nx,int ny,flouble lx,flouble ly,
       apodize_mask_smooth(fs,mask_in,mask_out,aposize);
     else {
       nmt_flatsky_info_free(fs);
-      report_error(NMT_ERROR_APO,"Unknown apodization type %s. Allowed: \"Smooth\", \"C1\", \"C2\"\n");
+      report_error(NMT_ERROR_APO,"Unknown apodization type %s. Allowed: \"Smooth\", \"C1\", \"C2\"\n",apotype);
     }
     nmt_flatsky_info_free(fs);
   }

### 2018-08-20 .travis/install.sh
trying to make travis skip...

@@ -36,6 +36,7 @@ cd $HOME
 git clone https://github.com/LSSTDESC/NaMaster.git
 cd NaMaster
 ./configure --prefix=$TRAVIS_BUILD_DIR
+make clean
 make
 sudo make install
 make check

### 2018-08-22 doc/Paper/main.tex

--- 2018-08-17 doc/Paper/main.tex
@@
       \end{align}
-      per purified field. \todo{check this}.
+      per purified field.
 
@@
         \end{figure}
-        The first validation suite exemplifies the use of
-        pseudo-$C_\ell$ algorithms in the analysis of large-scale
-        structure data. Current and next-generation imaging surveys
-        such as DES~\citep{2016MNRAS.460.1270D},
-        KiDS~\citep{2017A&A...604A.134D},
-        HSC~\citep{2018PASJ...70S...4A},
-        LSST~\citep{2008arXiv0805.2366I} and
-        Euclid~\citep{2011arXiv1110.3193L} will constrain the growth
-        and geometry of structure through cosmic time by using the
-        overdensity of galaxies as a proxy for the matter
-        inhomogeneities (a spin-0 field) and their correlated shape
-        distortions (a spin-2 field at first order), mostly caused by
-        weak lensing. These measurements will mostly follow a
-        tomographic approach, using the angular auto- and cross-power
-        spectra between these observables across different bins of
-        photometric redshift
-        \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
-
-        Large-scale structure data are often characterized by
-        inhomogeneous sky coverages and complex mask structures,
-        associated with the presence of bright stars and observational
-        artifacts. They are also affected by many sky contaminants,
-        associated to Galactic sources (dust absorption, star density)
-        and observing conditions (seeing, airmass
-        etc.) as discussed in \citet{2016ApJS..226...24L}. The typically flatter power
-        spectra for both galaxy clustering and lensing (e.g. when
-        compared with the damping tail of CMB temperature
-        fluctuations), makes the pseudo-$C_\ell$ approach ideal for
-        this kind of analysis \citep{2013MNRAS.435.1857L,2016ApJS..226...24L}. This science case therefore allows us to validate several aspects of the code simultaneously:
+        The first validation suite exemplifies the use of pseudo-$C_\ell$ algorithms in the analysis of large-scale structure data.
+Current and next-generation imaging surveys such as DES~\citep{2016MNRAS.460.1270D},
+        KiDS~\citep{2017A&A...604A.134D}, HSC~\citep{2018PASJ...70S...4A}, LSST~\citep{2008arXiv0805.2366I} and Euclid~\citep{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing.
+These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
+
+        Large-scale structure data are often characterized by inhomogeneous sky coverages and complex mask structures, associated with the presence of bright stars and observational artifacts.
+They are also affected by many sky contaminants, associated to Galactic sources (dust absorption, star density) and observing conditions (seeing, airmass etc.) as discussed in \citet{2016ApJS..226...24L}.
+The typically flatter power spectra for both galaxy clustering and lensing (e.g.
+when compared with the damping tail of CMB temperature fluctuations), makes the pseudo-$C_\ell$ approach ideal for this kind of analysis \citep{2013MNRAS.435.1857L,2016ApJS..226...24L}.
+This science case therefore allows us to validate several aspects of the code simultaneously:
         \begin{itemize}
@@
       \end{equation}
-      The full covariance matrix is very close to diagonal, as shown
-      in the right panel of Fig. \ref{fig:lss_val_cls}. Under the
-      assumption that the estimator is unbiased
-      ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$
-      should follow a $\chi^2$ distribution with a $N_{\rm dof}$
-      degrees of freedom. Figure \ref{fig:lss_val_chi2} shows the
-      distribution of $\chi^2_i$ across simulations approximating the
-      covariance matrix as diagonal (blue histograms),
-      compared with the predicted $\chi^2$ distribution for
-      $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the
-      six different power spectrum combinations. Both distributions
-      are found to be compatible. Moreover, the dot-dashed lines show
-      the values of $\chi^2$ found for the mean residual with respect
-      to the input power spectra (middle panel of
-      Fig. \ref{fig:lss_val_cls}). The probability to exceed (PTE)
-      associated with these $\chi^2$ values are higher than $10\%$ in
-      all cases, and therefore our validation suite does not show any
-      statistically significant bias in the estimator. This bias, if
-      any, must be at least $\sqrt{1000}\sim30$ times smaller than the
-      statistical uncertainties associated with our validation suite,
-      which are representative of the statistical power of
-      next-generation galaxy surveys.
+      The full covariance matrix is very close to diagonal, as shown in the right panel of Fig.
+\ref{fig:lss_val_cls}.
+Under the assumption that the estimator is unbiased ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$ should follow a $\chi^2$ distribution with a $N_{\rm dof}$ degrees of freedom.
+Figure \ref{fig:lss_val_chi2} shows the distribution of $\chi^2_i$ across simulations approximating the covariance matrix as diagonal (blue histograms), compared with the predicted $\chi^2$ distribution for $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the six different power spectrum combinations.
+Both distributions
+      are found to be compatible.
+Moreover, the dot-dashed lines show the values of $\chi^2$ found for the mean residual with respect to the input power spectra (middle panel of Fig.
+\ref{fig:lss_val_cls}).
+The probability to exceed (PTE) associated with these $\chi^2$ values are higher than $10\%$ in all cases, and therefore our validation suite does not show any statistically significant bias in the estimator.
+This bias, if any, must be at least $\sqrt{1000}\sim30$ times smaller than the statistical uncertainties associated with our validation suite, which are representative of the statistical power of next-generation galaxy surveys.
 
@@
       \end{figure*}
-      We have carried out a similar set of tests on an ensemble of
-      1000 simulations from the CMB validation suite described in
-      Section \ref{sssec:validation.suite.cmb}. In our fiducial run we
-      both deprojected the foreground contaminants and subsequently
-      used $B$-mode purification to decouple $E$ and $B$-modes at the
-      map level, since this is the main feature of the code we want to
-      validate. The left panel of Figure \ref{fig:cmb_val_cls} shows
-      the mean of the $EE$, $EB$ and $BB$ power spectra computed for
-      all simulations (solid lines) compared with the expectation from
-      the input spectra for $EE$ and $BB$ (dashed lines, mostly lying
-      under the solid ones). The average residuals normalized over the
-      error on the mean, shown in the lower panel of the same Figure,
-      fluctuate around zero with an amplitude of up to $3\sigma$, as
-      would be expected in the absence of a bias in the
-      estimator.
-
-      The above figure was created using E/B mode
-      purification. Disabling this option has no impact on the bias of
-      the estimates, but increases its variance as we show in the
-      Figure \ref{fig:cmb_val_sigma}.
+      We have carried out a similar set of tests on an ensemble of 1000 simulations from the CMB validation suite described in Section \ref{sssec:validation.suite.cmb}.
+In our fiducial run we both deprojected the foreground contaminants and subsequently used $B$-mode purification to decouple $E$ and $B$-modes at the map level, since this is the main feature of the code we want to validate.
+The left panel of Figure \ref{fig:cmb_val_cls} shows the mean of the $EE$, $EB$ and $BB$ power spectra computed for all simulations (solid lines) compared with the expectation from the input spectra for $EE$ and $BB$ (dashed lines, mostly lying under the solid ones).
+The average residuals normalized over the error on the mean, shown in the lower panel of the same figure, fluctuate around zero with an amplitude of up to $3\sigma$, as would be expected in the absence of a bias in the estimator.
+
+      The above figure was created using $E/B$ mode purification.
+Disabling this option has no impact on the bias of the estimates, but increases its variance as we show in the Figure \ref{fig:cmb_val_sigma}.
 We plot diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra.
@@
 As shown in Section \ref{sec:validation}, we are able to recover the input power spectrum in all cases (which include varying degrees of contamination, purification, different spins etc.) with no detectable bias down to $\lesssim1/30$-th of the statistical uncertainties.
+
+    Besides describing and validating the code, and in order to generalize the pseudo-$C_\ell$ estimator to all the cases supported by \nmt, this paper has also presented a number of results that are, to the best of our knowledge, new.
+These include:
+    \begin{itemize}
+      \item Extending the analytical treatment of contaminant deprojection presented in \cite{2016MNRAS.456.2095E} to fields of arbitrary spin.
+      \item Combining contaminant deprojection and $E/B$ purification.
+      \item Consistently deriving all of this functionality (standard pseudo-$C_\ell$, deprojection and purification for arbitrary spins) in the flat-sky approximation.
+    \end{itemize}
 
@@
 
-    \nmt~does not currently produce a covariance matrix estimate. This
-    is done for several reasons. First, for many problems, such as
-    large-scale structure power spectrum estimation, the covariance
-    matrix contains import contributions from the connected four-point
-    function that depend on the precise nature of the problem and go
-    beyond the estimation step that \nmt~strives to do well. Second,
-    with all additional options included, the exact analytical
-    estimates are either computational infeasible or overly
-    approximate. Since majority of the forthcoming experiments will
-    have a large number of mock datasets available for testing
-    systematic effects, these maps also offer a natural path towards
-    estimating the data covariance.
-
-    \nmt~is open-source software and publicly available at \url{https://github.com/LSSTDESC/NaMaster}. To maximize the
-    usefulness of this software for the community, it is accompanied
-    by an extensive documentation and example code\footnote{See \url{
-        https://github.com/LSSTDESC/NaMaster}.}. \nmt~is written in
-    {\tt C} and {\tt OpenMP}-parallelized to maximize its performance,
-    although we encourage its use through the companion {\tt python}
-    wrapper, the fastest-growing language of choice in the analysis of
-    astronomical data. The code is in constant development, and we
-    will strive to implement further functionality to e.g. estimate
-    covariance matrices, extend the range of supported sky
-    pixelizations and enhance its computational performance.
+    \nmt~does not currently produce a covariance matrix estimate.
+Although computationally efficent analytical estimates of the Gaussian covariance matrix can be derived under certain approximations \citep{2004MNRAS.349..603E,2005MNRAS.360.1262B}, this is not widely supported by the code for several reasons.
+First, for many problems, such as large-scale structure power spectrum estimation, the covariance matrix contains important contributions from the connected four-point function that depend on the precise nature of the problem and go beyond the estimation step that \nmt~strives to do well.
+Second, with all additional options included, especially if a large number of contaminant templates are deprojected, the exact analytical estimates are either computational unfeasible or overly approximate.
+Since a majority of the forthcoming experiments will rely on a large number of mock datasets available for testing systematic effects, these maps also offer a natural path towards estimating the data covariance.
+
+    \nmt~is open-source software and publicly available at \url{https://github.com/LSSTDESC/NaMaster}.
+To maximize the usefulness of this software for the community, it is accompanied by an extensive documentation and example code\footnote{See \url{https://github.com/LSSTDESC/NaMaster}.}.
+\nmt~is written in {\tt C} and {\tt OpenMP}-parallelized to maximize its performance, although we encourage its use through the companion {\tt python} wrapper, the fastest-growing language of choice in the analysis of astronomical data.
+The code is in constant development, and we will strive to implement further functionality to e.g.
+estimate covariance matrices, extend the range of supported sky pixelizations and enhance its computational performance.
 
 \subsection*{Acknowledgements}
-We would like to thank Rupert Allison, Jo Dunkley, Franz Elsner, Boris
-Leistedt, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya
-Peiris, Naomi Robertson and Ben Thorne for useful comments and
-discussions. DA acknowledges support from the Beecroft trust and from
-the Science and Technology Facilities Council (STFC) through an Ernest
-Rutherford Fellowship, grant reference ST/P004474/1. This research
-used resources of the National Energy Research Scientific Computing
-Center (NERSC), a U.S. Department of Energy Office of Science User
-Facility operated under Contract No. DE-AC02-05CH11231.
+We would like to thank Rupert Allison, Scott Dodelson, Jo Dunkley, Franz Elsner, Boris Leistedt, Francois Lanusse, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya Peiris, Naomi Robertson and Ben Thorne for useful comments and discussions.
+DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
+This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
+Department of Energy Office of Science User Facility operated under Contract No.
+DE-AC02-05CH11231.
 
@@
 
-\section{Linear contamination from non-linear
-  contamination} \label{app:lincont}
-\as{It is up to David to decide whether he buys this argument or
-  not. I'm fine with not and then we soften then claim in 2.2. Doesn't
-  really affect any of the paper otherwise.}
-Imagine a contaminant field $c$ (e.g. a point-spread function (PSF) size)
-that affects the observable field $o$ (e.g. the number count
-densities) through a non-linear, but \emph{local} on some scale local
-functional $F$:
+  \section{Linear contamination from non-linear contamination} \label{app:lincont}
+    Consider a contaminant field $c$ (e.g.
+a point-spread function (PSF) size) that affects the observable field $o$ (e.g.
+the number count densities) through a non-linear, but \emph{local} on some scale local functional $F$:
     \begin{equation}
@@
     \end{equation}
-where we can, without loss of generality, take out the mean effects
-(i.e. there is some mean PSF size and fluctuations around this mean
-size produce additional fluctuations in the number density of
-fluctuations).
-
-In general, for sufficiently non-linear functionals $F$, the Taylor
-expansion around $\delta_c(\nv)$ is not valid. However, one can look
-at the effect of smoothing by a kernel of size $S(R)$, which in
-Fourier space is equivalent by multiplication by $e^{-k^2R^2}$. We get
-\begin{equation}
-  \delta^R_o(\nv) \rightarrow \delta^R_o  + S(R)[F [ \delta_c
-  ]] =  \delta^R_o + F'_R [ \delta^R_c ],
-\end{equation}
-where we used a shorthand $\delta_x^R=S(R)[\delta_x]$ and  $F'_R$ is now a \emph{different} functional that takes a smoothed
-$\delta_c$ and returns out a smoothed $\delta_o$.  The crucial point
-is that $F'_R$ must still be
-local, since one cannot turn a local functional into a non-local one
-by application of smoothing.
-
-This means, that for a well behaved $\delta_c$, such that variances of
-both $\delta^R_c$ and $\delta^R_o$ are small for a large enough $R$,
-the Taylor expansion is valid for $F'_R$ (even if it is not for $F$),
-giving
+    where we can, without loss of generality, take out the mean effects (i.e.
+there is some mean PSF size and fluctuations around this mean size produce additional fluctuations in the number density of fluctuations).
+
+    In general, for sufficiently non-linear functionals $F$, the Taylor expansion around $\delta_c(\nv)$ is not valid.
+However, one can look at the effect of smoothing by a kernel of size $S(R)$, which in Fourier space is equivalent by multiplication by $e^{-k^2R^2}$.
+We get
+    \begin{equation}
+      \delta^R_o(\nv) \rightarrow \delta^R_o  + S(R)[F [ \delta_c  ]] =  \delta^R_o + F'_R [ \delta^R_c ],
+    \end{equation}
+    where we used a shorthand $\delta_x^R=S(R)[\delta_x]$ and  $F'_R$ is now a \emph{different} functional that takes a smoothed $\delta_c$ and returns out a smoothed $\delta_o$.
+The crucial point is that $F'_R$ must still be local, since one cannot turn a local functional into a non-local one by application of smoothing.
+
+    This means, that for a well behaved $\delta_c$, such that variances of both $\delta^R_c$ and $\delta^R_o$ are small for a large enough $R$, the Taylor expansion is valid for $F'_R$ (even if it is not for $F$), giving
     \begin{equation}
@@
     \end{equation}
-for $k\ll R^{-1}$. In other words a ``local'' contaminant becomes
-``linearly biased'' in the $k\rightarrow 0 $ regime, very much like
-very non-linear physics of galaxy formations results in
-linearly biased galaxy field on large scales.
+    for $k\ll R^{-1}$.
+In other words a ``local'' contaminant becomes ``linearly biased'' in the $k\rightarrow 0 $ regime, very much like very non-linear physics of galaxy formations results in linearly biased galaxy field on large scales.
 
@@
     \subsection{Spin-\texorpdfstring{$s$}{s} fields in flat sky}\label{app:flat.fields}
-      In the flat sky we will label a position in the plane by two
-      coordinates $(x,y)\equiv{\bf x}$. These coordinates can be
-      directly related to increments in the spherical coordinates
-      $(\theta,\varphi)$ by considering a sufficiently small map
-      centered around the equator (i.e. $\sin\theta\sim 1$). In
-      this case we identify the $x$ coordinate with a latitude shift
-      ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
-      shifts ($\delta\varphi=\delta y$) \todo{make a plot of
-        this?}. \as{no, but isn't it other way round? Latitute is $y$
-        in my brain? :)} \dam{it isn't, have a look here: \url{https://healpix.jpl.nasa.gov/html/intronode12.htm}.} Using this dictionary, the differential operator $\eth$ now takes the form:
+      In the flat sky we will label a position in the plane by two coordinates $(x,y)\equiv{\bf x}$.
+These coordinates can be directly related to increments in the spherical coordinates $(\theta,\varphi)$ by considering a sufficiently small map centered around the equator (i.e.
+$\sin\theta\sim 1$).
+In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
+      shifts ($\delta\varphi=\delta y$) the differential operator $\eth$ now takes the form: %\todo{make a plot of this?}.
+\as{no, but isn't it other way round?
+Latitute is $y$ in my brain?
+:)} \dam{it isn't, have a look here \url{https://healpix.jpl.nasa.gov/html/intronode12.htm}.} Using this dictionary,
       \begin{equation}
@@
     \subsection{Discrete description and DFTs}\label{app:flat.dft}
-      As we discuss in \todo{where}, it is more convenient to describe power spectrum estimation methods in a discretized flat sky.
+      As we discuss in \ref{sssec:maths.flat.pcl101}, it is more convenient to describe power spectrum estimation methods in a discretized flat sky.
 Let the patch of the sky under inspection be contained by a rectangle of sides $L_x$ and $L_y$ (in units of radians), and let us discretize this rectangle by dividing it into an $N_x\times N_y$ grid with pixels of area $\Delta{\bf x}^2\equiv\Delta x\Delta y=(L_x/N_x)(L_y/N_y)$.

### 2018-08-22 doc/Paper/main.tex
Addressed comment about timing

@@
 We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles.
+The typical running time for one realization is $\sim 600$s using 32 \texttt{OpenMP} threads.
+Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation.
+More details about timing and scalability of \nmt can be found in Appendix \ref{sec:scalability}.
 
@@
 
-\end{document}
-
- \subsection{Performance and scalability}
- \label{ssec:scalability}
+\section{Performance and scalability}
+\label{sec:scalability}
      One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
@@
 
+\end{document}
+
     In addition, we also check the parallel performance of \nmt.

### 2018-08-25 doc/Paper/main.tex
Fixes a few typos

@@
 \begin{abstract}
-  The pseudo-$C_\ell$ is an algorithm for estimation of the angular power and cross-power spectra that is very fast and in realistic cases also nearly optimal.
+  The pseudo-$C_\ell$ is an algorithm for estimating the angular power and cross-power spectra that is very fast and in realistic cases also nearly optimal.
 The algorithm can be extended to deal with contaminant deprojection and $E/B$ purification, and can therefore be applied in a wide variety of scenarios of interest for current and future cosmological observations.
-This paper presents \nmt, a public, validated, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both in the sphere or in the flat-sky approximation.
+This paper presents \nmt, a public, validated, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both on the sphere or in the flat-sky approximation.
 We describe the mathematical background of the estimator, including all the features above, and its software implementation in \nmt.
@@
         and the last term in Eq.
-\ref{eq:bpws_qmv} correspond to bias of convolved estimates due to the presence of noise.
+\ref{eq:bpws_qmv} corresponds to a bias of the convolved estimates due to the presence of noise.
 Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices ${\sf P}_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
@@
 But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added PSF fluctuations could be used to remove the possible correlations between measured ellipticities.
-Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}).
+Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as that relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}).
 Thus, we generically deal with a linear model of the kind:
@@
 In practice, if required, the
-          suppression factor can be found by applying the PCL estimator to a a set of mock input maps with the same underlying signal but  with and without application of the beam smoothing (to both data maps and contaminant maps) and taking the ratio.
+          suppression factor can be found by applying the PCL estimator to a set of mock input maps with the same underlying signal but  with and without application of the beam smoothing (to both data maps and contaminant maps) and taking the ratio.
 
@@
 To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask.
-We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig.
+We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but traces the inverse of the noise variance (see top left panel of Fig.
 \ref{fig:lss_mocks_sph}).
@@
           \item {\bf Beam:} we assume a Gaussian beam with full-width at half-maximum of $\theta_{\rm FWHM}=20$ and 1.4 arcmin for the curved-sky and flat-sky cases respectively.
-This resolution corresponds to what could be achieved at 150 GHz with with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case.
+This resolution corresponds to what could be achieved at 150 GHz with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case.
 The choice of $\ell_{\rm knee}$ described above for the noise curves is also consistent with the atmospheric noise levels achievable with those types of telescopes \todo{cite Simons Observatory paper if out}.

### 2018-08-30 doc/Paper/main.tex
another small typo
addressed most of Reviewer 3's comments

--- 2018-08-22 doc/Paper/main.tex
@@
 
-    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information, their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ and $\ell_{\rm max}^4$ respectively with the maximum multipole $\ell_{\rm max}$.
+    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g.
+\cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}).
 For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
@@
         \begin{equation}
-          {\sf F_{ij}} B_j = \va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
-          \rm{Tr}\left[{\sf P}_i\,{\sf N} \right],
-        \end{equation}
-        where ${\sf F}$ is the Fisher matrix,the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
+          {\sf F_{ij}} B_j = \frac{1}{2}\va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
+          \frac{1}{2}\rm{Tr}\left[\tilde{\sf C}^{-1}{\sf P}_i\tilde{\sf C}^{-1}\,{\sf N} \right],
+        \end{equation}
+        where ${\sf F}$ is the Fisher matrix, the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
         \begin{equation}\label{eq:bpws_qmv}
@@
         \begin{equation}
-          {\bf a}^v_{\bf l}=\sum_{{\bf l}_1{\bf l}_2}={\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}{\bf a}_{{\bf l}_1}v_{{\bf l}_2},
+          {\bf a}^v_{\bf l}=\sum_{{\bf l}_1{\bf l}_2}{\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}{\bf a}_{{\bf l}_1}v_{{\bf l}_2},
         \end{equation}
@@
 
-        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq. \ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by:
+        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq.
+\ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by \citep{2002ApJ...567....2H,2003ApJS..148..161K}:
         \begin{align}\label{eq:m00}
@@
 The process is illustrated in Figure \ref{fig:lss_cont}.
-A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}.
+A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{uncorrelated weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}.
 Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
@@
 In practice, this can be done through an iterative approach, starting from an initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the estimated power spectrum in the previous iteration.
-A sufficiently good guess will likely be enough to remove the bias in a single iteration.
 Alternatively, the ensemble average of Eq.
@@
 if a set of templates are linearly related).
-The problem can however be solved self-consistently by using the Moore-Penrose pseudo-inverse.
+The problem can however be solved self-consistently by using the Moore-Penrose pseudo-inverse \citep{2017MNRAS.465.1847E}.
 This is trivial for a symmetric and positive-definite matrix such as ${\sf F}$, and corresponds to setting the inverse of all of its zero eigenvalues (or all those below a given threshold) to zero.
@@
 purifying $E$-modes from $B$-mode contamination is conceptually exactly the same.
-Our discussion is also specific for the spin-2 fields.
+Our discussion is also specific for spin-2 fields.
 
@@
       \begin{equation}
-        ({\bf D}^E_s\varphi,{\bf D}^B_s\psi)=\int d\nv\,({\bf D}^E_s\varphi)^\dag{\bf D}^B_s\psi=0
-      \end{equation}
+        ({\bf D}^E_s\varphi,{\bf D}^B_s\psi)=\int d\nv\,({\bf D}^E_s\varphi)^\dag{\bf D}^B_s\psi=0,
+      \end{equation}
+      where $\varphi$ and $\psi$ are two arbitrary scalar fields.
 This can be done by integrating by parts and noting that the celestial sphere has no boundaries.
@@
       \begin{equation}
-        B^p_{\bf l}=\int d\nv\left({\bf D}^B_2(w\,Y_{\bf l})\right)^\dag\,{\bf P}(\nv).
-      \end{equation}
-      Since ${\bf D}^B_2(wY_{\bf l})$ is a $B$-mode quantity, $B^p_{\bf l}$ should receive contributions only from $B$-modes.
+        B^p_{\bf l}=\int d\nv\left({\bf D}^B_s(w\,Y_{\bf l})\right)^\dag\,{\bf P}(\nv).
+      \end{equation}
+      Since ${\bf D}^B_s(wY_{\bf l})$ is a $B$-mode quantity, $B^p_{\bf l}$ should receive contributions only from $B$-modes.
 
@@
 Results will be presented for both curved-sky and flat-sky realizations.
+Note that, in all cases, we use the true signal power spectra as input to compute the bias from contaminant deprojection (see Eqs.
+\ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}).
+This allows us to verify that the code works as expected, and to isolate any residual bias associated with software bugs, rather than an imperfect guess of the true power spectrum.
 
@@
 In all cases we add contaminants linearly, to avoid deviating from the model described in Section \ref{ssec:maths.deproj}.
+The contaminant amplitudes are chosen so that any residual contamination in the relevant range of multipoles will be detected statistically by the validation suite.
 Maps of these contaminants are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.

### 2018-08-30 doc/Paper/main.tex
modified figs 2 and 5 according to ref. report

@@
           \caption{Signal and noise power spectra used in the LSS validation set.
-Red, green and blue curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode.
+Dark blue, cyan and red curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode.
 All other cross-correlations are zero.
@@
           \includegraphics[width=0.99\columnwidth]{cls_cont_lss}
-          \caption{{\sl Top:} signal power spectrum (red) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (green) and stars (blue).
-{\sl Middle:} $E$-mode lensing power spectrum (red) and the spectra of the PSF and small-scale contaminants used in our validation set (green and blue respectively).
+          \caption{{\sl Top:} signal power spectrum (solid dark blue) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (dashed cyan) and stars (dot-dashed red).
+{\sl Middle:} $E$-mode lensing power spectrum (solid dark blue) and the spectra of the PSF and small-scale contaminants used in our validation set (dashed cyan and dot-dashed red respectively).
 {\sl Bottom:} same as middle panel for the $B$-mode power spectrum.
 Note that in this case the input signal power spectrum is $0$.
-In all cases the yellow line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
+In all cases the orange dotted line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
         \end{figure}

### 2018-08-30 doc/Paper/main.tex
modified fig 12 according to ref. report

@@
         \caption{{\sl Left:} covariance matrix for power spectra computed in the absence of foreground contaminants computed from the 1000 simulations of our curved-sky CMB validation suite.
-The incomplete sky coverage causes a noticeable anti-correlation of neighboring bandpowers at low $\ell$ for the three power spectra ($EE$, $EB$ and $BB$), but beyond that, the covariance matrix is close to diagonal.
+The incomplete sky coverage causes a noticeable anti-correlation of neighboring bandpowers at low $\ell$ for the three power spectra ($EE$, $EB$ and $BB$, see inset for the $EE$ case), but beyond that, the covariance matrix is close to diagonal.
 {\sl Right:} the same as the left panel but this time for simulations containing foreground contaminants.

### 2018-08-30 sandbox_validation/plot_sph_pure.py

@@ -4,6 +4,7 @@ import scipy.stats as st
 from matplotlib import rc
 import matplotlib
 import pymaster as nmt
+from mpl_toolkits.axes_grid1.inset_locator import mark_inset,zoomed_inset_axes
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 
@@ -29,7 +30,7 @@ def read_cls(fname) :
 l_th,clEE_th,clEB_th,clBE_th,clBB_th=read_cls(prefix_pure+"_cl_th.txt")
 ndof=len(l_th)
 
-print "Reading"
+print("Reading")
 clEE_pure=[]; clEB_pure=[]; clBB_pure=[];
 clEE_nopu=[]; clEB_nopu=[]; clBB_nopu=[];
 clEE_noco=[]; clEB_noco=[]; clBB_noco=[];
@@ -48,7 +49,7 @@ clEE_nopu=np.array(clEE_nopu); clEB_nopu=np.array(clEB_nopu); clBB_nopu=np.array
 clEE_noco=np.array(clEE_noco); clEB_noco=np.array(clEB_noco); clBB_noco=np.array(clBB_noco);
 clEE_nodb=np.array(clEE_nodb); clEB_nodb=np.array(clEB_nodb); clBB_nodb=np.array(clBB_nodb);
 
-print "Computing statistics"
+print("Computing statistics")
 def compute_stats(y,y_th) :
     mean=np.mean(y,axis=0)
     cov=np.mean(y[:,:,None]*y[:,None,:],axis=0)-mean[:,None]*mean[None,:]
@@ -109,6 +110,14 @@ ax.set_xticks(ndof*(np.arange(3)+0.5))
 ax.set_yticks(ndof*(np.arange(3)+0.5))
 ax.set_xticklabels(['$EE$','$EB$','$BB$'])
 ax.set_yticklabels(['$EE$','$EB$','$BB$'])
+axins=zoomed_inset_axes(ax,2.5,loc=6)
+axins.imshow(cov_pure/np.sqrt(np.diag(cov_pure)[None,:]*np.diag(cov_pure)[:,None]),
+             interpolation='nearest',cmap=plt.cm.Greys)
+axins.get_xaxis().set_visible(False)
+axins.get_yaxis().set_visible(False)
+axins.set_xlim(0.,0.2*ndof)
+axins.set_ylim(0.2*ndof,0.)
[…]

### 2018-08-30 doc/Paper/main.tex
implemented Francois' typos

@@
       \item $E/B$ purification can only be carried out in the context of pseudo-$C_\ell$ estimators when the mask is differentiable up to its second derivatives, which can usually be achieved with an appropriate apodization.
-Although \nmt~provides tools to carry our this type of operations, some of the properties of the input mask (e.g.
+Although \nmt~provides tools to carry out this type of operations, some of the properties of the input mask (e.g.
 sharp corners) will in general spoil differentiability even after apodization.

### 2018-08-30 doc/Paper/main.tex
band-limit paragraph

--- 2018-08-30 doc/Paper/main.tex
@@
  %\hypersetup{draft}
-\newcommand{\nv}{\hat{\boldsymbol{\theta}}}
+\newcommand{\nv}{\hat{\bf n}}
 \newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
@@
 \newcommand{\vb}{{\mathbf{b}}}
+\newcommand{\mC}{{\sf C}}
 \newcommand{\mM}{{\sf M}}
@@
 \begin{abstract}
-  The pseudo-$C_\ell$ is an algorithm for estimation of the angular power and cross-power spectra that is very fast and in realistic cases also nearly optimal.
-The algorithm can be extended to deal with contaminant deprojection and $E/B$ purification, and can therefore be applied in a wide variety of scenarios of interest for current and future cosmological observations.
-This paper presents \nmt, a public, validated, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both in the sphere or in the flat-sky approximation.
-We describe the mathematical background of the estimator, including all the features above, and its software implementation in \nmt.
-We construct a validation suite that aims to resemble the types of observations that next-generation large-scale structure and ground-based CMB experiments will face, and use it to show that the code is able to recover the input power spectra in the most complex scenarios with no detectable bias.
-\nmt~can be found at \url{https://github.com/LSSTDESC/NaMaster}, and is provided with comprehensive documentation and a number of code examples.
+  The pseudo-$C_\ell$ is an algorithm for estimation of the angular
+  power and cross-power spectra that is very fast and in realistic
+  cases also nearly optimal. The algorithm can be extended to deal
+  with contaminant deprojection and $E/B$ purification, and can
+  therefore be applied in a wide variety of scenarios of interest for
+  current and future cosmological observations. This paper presents
+  \nmt, a public, validated, accurate and easy-to-use software
+  package that, for the first time, provides a unified framework to
+  compute angular cross-power spectra of any pair of spin-0 or spin-2
+  fields, contaminated by an arbitrary number of linear systematics
+  and requiring $B$- or $E$-mode purification, both in the sphere or
+  in the flat-sky approximation. We describe the mathematical
+  background of the estimator, including all the features above, and
+  its software implementation in \nmt. We construct a validation suite
+  that aims to resemble the types of observations that next-generation
+  large-scale structure and ground-based CMB experiments will face,
+  and use it to show that the code is able to recover the input power
+  spectra in the most complex scenarios with no detectable
+  bias. \nmt~can be found at \url{https://github.com/LSSTDESC/NaMaster}, and is provided with
+  comprehensive documentation and a number of code examples.
 \end{abstract}
@@
   \section{Introduction}\label{sec:intro}
-    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}.
-Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier-space\footnote{In this paper we use the term ``Fourier'' to describe both the flat space Fourier expansion as well as harmonic-space expansion on the sphere.
-We will also use the term ``power spectrum'' to describe both the auto power spectrum of a field and cross-power spectra between two fields.
-} power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.
-Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a Fourier space analysis is advantageous for many other reasons.
-The most important is that Fourier space is a natural space for working with statistics of translation-invariant fields.
-Therefore, Fourier space analysis has a clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}.
-Moreover, individual band-power measurements are typically significantly less correlated compared to configuration-space analysis which makes covariance matrix estimation easier.
-This has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators and used extensively in Cosmic Microwave Background (CMB) two-point measurements.
-
-    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g.
-\cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}).
+  Two-point functions have proven to be the most useful summary
+  statistic for cosmology in terms of data compression, due to the
+  near-Gaussian nature of the modes of the cosmic density field that
+  can be easily analyzed to extract cosmological
+  information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since
+  the only true observables in any astronomical dataset are the
+  intensity and polarization as a function of frequency $\nu$ and sky
+  position $\nv$, angular 2-point correlators play a particularly
+  central role in most cosmological analyses, both in the form of
+  configuration-space two-point correlation functions $\xi(\theta)$
+  and as Fourier-space\footnote{In this paper we use the term
+    ``Fourier'' to describe both the flat space Fourier expansion as
+    well as harmonic-space expansion on the sphere. We will also use
+    the term ``power spectrum'' to describe both the auto power
+    spectrum of a field and cross-power spectra between two fields. }
+  power spectra $C_\ell$
+  \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.  Although
+  correlation functions are often easier to estimate from complex
+  observations~\citep{1993ApJ...412...64L}, a Fourier space analysis
+  is advantageous for many other reasons. The most important is that
+  Fourier space is a natural space for working with statistics of
+  translation-invariant fields. Therefore, Fourier space analysis has
+  a clearer separation between linear and non-linear scales, as well
+  as Gaussian and non-Gaussian
+  modes~\citep{2008PhRvD..77j3013H}. Moreover, individual band-power
+  measurements are typically significantly less correlated compared to
+  configuration-space analysis which makes covariance matrix
+  estimation easier. This has motivated a large body of literature
+  aimed at designing optimal and efficient power spectrum estimators and 
+  used extensively in Cosmic Microwave Background (CMB) two-point measurements.
+
+    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information, their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ and $\ell_{\rm max}^4$ respectively with the maximum multipole $\ell_{\rm max}$.
 For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
 This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
-Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward, no public, validated and well-maintained code exists supporting all of these capabilities consistently.
+Unfortunately, in spite of the advantages of this method, no public, validated and well-maintained codes exist supporting all of these capabilities.
 This paper presents a public software package, \nmt, that unifies and extends all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.
@@
 Section \ref{sec:validation} presents the stress tests used to validate the code.
-We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data, so this section illustrates not just its validity but also applicability in typical user-case scenarios in both fields.
+We envisage this code to be useful in analysis of both the Cosmic Microwave Background data and in the tomographic large-scale structure from photometric data and this section illustrates not just its validity but also applicability in typical user-case scenarios in both fields.
 We conclude in Section \ref{sec:discussion}.
@@
       \begin{align}\label{eq:sht1}
-        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
+        {\bf a}_{\ell m}&\equiv\mathcal{S}\left[{\bf a}(\nv)\right]^{s_a}_{\ell m}\equiv\int d\nv\,_{s_a}{\sf Y}^{s_a\dag}_{\ell m}(\nv)\,{\bf a}(\nv), \\\label{eq:sht2}
         {\bf a}(\nv)&=\mathcal{S}^{-1}\left[{\bf a}_{\ell m}\right]^{s_a}_{\nv}\equiv\sum_{\ell m}\,{\sf Y}^{s_a}_{\ell m}(\nv)\,{\bf a}_{\ell m}.
       \end{align}
-      Note that, in general we will use vector notation such that, for a complex spin-$s_a$ field $a$, we form the vector ${\bf a}\equiv({\rm Re}(a),{\rm Im}(a))$.
-The harmonic coefficients above are decomposed in a similar manner into $E$ and $B$ modes: ${\bf a}_{\ell m}\equiv(a^E_{\ell m},a^B_{\ell m})$ (e.g.
-see Appendix \ref{app:shts} and \cite{1997PhRvD..55.1830Z}).
+      Note that, in general we will use a vector notation such that, for a complex spin-$s_a$ field $a$, we form the vector ${\bf a}\equiv({\rm Re}(a),{\rm Im}(a))$.
+The harmonic coefficients above are decomposed in a similar manner into $E$ and $B$ modes: ${\bf a}_{\ell m}\equiv(a^E_{\ell m},a^B_{\ell m})$.
 The matrices ${\sf Y}^s_{\ell m}$ are defined in terms of the spin-weighed spherical harmonics, and are described in detail in Appendix \ref{app:shts}.
@@
       Note that for a general non-zero spin, vectors $\va$ and $\vb$ have two-components, each of which is a complex number.
-Nevertheless the rotational invariance requires that elements of the $2\times2$ matrix ${\sf C}_\ell$ be real.
+Nevertheless the rotational invariance requires that elements of the $2\times2$ matrix $\mC$ be real.
+
+      \subsubsection{Pixelization and signal band-limits}
+
+      SHT transforms as written in the Eq.~\ref{eq:sht1} is, strictly
+      speaking, defined on a continuous field on the sky and results
+      in an infinite number of spherical harmonic coefficients. In
+      practice, we have to always deal with a finite number of pixels
+      at which the field is sampled and consequently use a finite
+      number of spherical coefficients. If the sampling of the
+      underlying signal is too sparse, one will get aliasing of higher
+      frequency modes. In flat spaces, Nyquist theorem provides a
+      clear prescription for uniform sampling on a grid. Since there
+      is no such natural sampling on a sphere, the situation is more
+      complex and depends on the pixelization scheme
+      \cite{Driscoll:1994:CFT:184069.184073,
+        DBLP:journals/corr/abs-1110-6297}. Nevertheless, clear
+      heuristic prescriptions exist. Most importantly, the pixel separation needs
+      to be smaller than the wavelength of the highest frequency
+      present in the underlying signal. As a natural consequence, the
+      smoothing of the map must always precede any downgrade in
+      resolution (or equivalently downsampling) if it is necessary to
+      do so for numerical expediency.  Similarly, any mask should be
+      applied at the highest possible resolution and the resulting
+      map smoothed before being down-graded.
+
+      For the same reasons, even if the signal is correctly sampled in
+      a band-limited fashion, one must always estimate power over the
+      entire power band, even if one is not interested in certain
+      region of $\ell$ space. Otherwise, the power spectrum estimator
+      will try to ``explain'' the variance associated with $\ell$
+      modes not considered by artificially inflating the power in the
+      estimated modes. For {\tt HEALPix} \citep{2005ApJ...622..759G}
+      pixelization commonly used in cosmology and also in this paper,
+      \cite{2013MNRAS.435.1857L} suggest the following
+      prescription. If one wants to measure power up to $\ell_{\rm
+        max}$, one should employ pixelisation of
+      \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up
+      to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm
+        max}$ after correcting for pixel-response suppression and
+      accounting for their contribution to the covariance matrix of
+      measurements.
+
+      In this paper we do not consider these issues any further as \nmt
+      is a generic tool and it is up to the user to ensure band-limit
+      and pixelization constraints are properly accounted for.
 
@@
 In general, the input data for measuring the power spectrum come pixelized on a sphere.
-Instead of dealing with an infinitely sampled field, we have a finite number of measurements
+Instead of dealing with an infinitely sampled points, we have a finite number of measurements
         \begin{equation}
@@
         where $\va_i^{\rm true}$ is is the true underlying field whose power spectrum we would like to estimate and ${\bf n}_i$ is the noise vector, which in general can be non-white and non-homogeneous.
-We will denote the pixel covariance of the noise component as $\tilde{\sf N}$\footnote{Note that the matrices $\tilde{\sf C}$ and $\tilde{\sf N}$ used in this section are the data and noise pixel-pixel covariance matrices, and therefore are different from the power spectrum matrices, ${\sf C}_\ell$ and ${\sf N}_\ell$, used in the rest of the paper (e.g.
-Eq.
-\ref{eq:cldef}).}.
+We will denote the pixel covariance of the noise component as ${\sf N}$.
 Note that for the sake of power-spectrum estimation the there is no conceptual difference between inhomogeneous noise and a survey mask: masked areas can be simply though of as areas of infinite noise.
@@
 
-        The underlying power spectrum in Eq.~\ref{eq:cldef} means that there is a non-trivial true underlying covariance matrix $\tilde{\sf C}^{\rm true} = \left<\va^{\rm true} \va^{{\rm true}\dagger}\right>$, which in addition to the noise component leads to the actual data covariance $\tilde{\sf C} = \left<\va \va^\dagger\right>=\tilde{\sf C}^{\rm true} + \tilde{\sf N}$.
+        The underlying power spectrum in Eq.~\ref{eq:cldef} means that there is a non-trivial true underlying covariance matrix $\mC^{\rm true} = \left<\va^{\rm true} \va^{{\rm true}\dagger}\right>$, which in addition to the noise component leads to the actual data covariance $\mC = \left<\va \va^\dagger\right>=\mC^{\rm true} + {\sf N}$.
 
@@
         \begin{equation}
-          {\sf F_{ij}} B_j = \frac{1}{2}\va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
-          \frac{1}{2}\rm{Tr}\left[\tilde{\sf C}^{-1}{\sf P}_i\tilde{\sf C}^{-1}\,{\sf N} \right],
-        \end{equation}
-        where ${\sf F}$ is the Fisher matrix, the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
-        \begin{equation}\label{eq:bpws_qmv}
-          \tilde{\sf C} = \sum_i {\sf P}_i\,B_i + {\sf N},
-        \end{equation}
-        and the last term in Eq.
-\ref{eq:bpws_qmv} correspond to bias of convolved estimates due to the presence of noise.
-Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices ${\sf P}_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
+          {\sf F_{ij}} B_j = \va^\dagger \mC^{-1}
+          \left(\frac{\partial \mC}{\partial B_i} \right)
+          \mC^{-1} \va - \rm{Tr}\left[\left(\frac{\partial \mC}{\partial B_i}
+          \right) {\sf N} \right],
+        \end{equation}
+        where ${\sf F}$ is the Fisher matrix,the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
+        \begin{equation}
+          \mC = \sum_i \left(\frac{\partial \mC}{\partial B_i}\right) + {\sf N},
+        \end{equation}
+        and the last term correspond to bias of convolved estimates by the presence of noise.
+Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices $\partial {\sf C}/\partial B_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
 The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix.
@@
         The pseudo-$C_\ell$ estimates will be optimal when these steps are close to what an optimal quadratic estimator would do.
-First, this requires that $\tilde{\sf C}$ be close to diagonal, which is true when either the noise is large and uncorrelated, or when the underlying power spectrum is close to white.
-Multiplication by the mask mixes modes, and replacement of multiplication by the response matrix with band-power averaging, makes sense only when the mask is ``well-behaved'', i.e.
+First, this requires that $\mC$ be close to diagonal, which is true when either the noise is large and uncorrelated, or when the underlying power spectrum is close to white.
+Multiplication by the mask mixes modes, and replacement of multiplication by the response matrix with band-power averaging, only makes sense, when the mask is ``well-behaved'', i.e.
 compact with no high-frequency structure in it.
@@
         \begin{equation}
-          {\bf a}^v_{\bf l}=\sum_{{\bf l}_1{\bf l}_2}{\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}{\bf a}_{{\bf l}_1}v_{{\bf l}_2},
+          {\bf a}^v_{\bf l}=\sum_{{\bf l}_1{\bf l}_2}={\sf D}^{s_a}_{{\bf l}{\bf l}_1{\bf l}_2}{\bf a}_{{\bf l}_1}v_{{\bf l}_2},
         \end{equation}
@@
         where we have defined the vectorization operator \citep{2008PhRvD..77j3013H}
-        \begin{align}
-          &{\rm vec}\left[\left(\begin{array}{cc}
+        \begin{equation}
+          {\rm vec}\left[\left(\begin{array}{cc}
                           C^{EE} & C^{EB}\\
@@
           \end{array}
-          \right),\\
-          &{\rm vec}\left[\left(\begin{array}{c}
-                                 C^{TE}\\C^{TB}
-                                \end{array}\right)\right]\equiv
-            \left(\begin{array}{c}
-                    C^{TE}\\C^{TB}
-            \end{array}\right),
-            \hspace{6pt}
-            {\rm vec}\left[C^{TT}\right]={\rm C}^{TT},
-        \end{align}
-        and $T$ here stands for any spin-$0$ field.
-
-        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq.
-\ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by \citep{2002ApJ...567....2H,2003ApJS..148..161K}:
+          \right).
+        \end{equation}
+
+        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq. \ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by:
         \begin{align}\label{eq:m00}
@@
       \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
-        Given the loss of information implicit in masking the originally curved-sky field, it is in general not possible to invert the mode-coupling matrix in Eq.
+        Given the loss of information implicit in masking the originally full-sky field, it is in general not possible to invert Eq.
 \ref{eq:modecoup} directly.
@@
           {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
-          \left\langle{\rm vec}\left[\tilde{\sf B}^{ab}_q\right]\right\rangle&=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
-        \end{align}
-        where in the second line we have taken the expectation value of first one.
+          &=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right].
+        \end{align}
         One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function).
 Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
-        \begin{figure*}
-          \centering
-          \includegraphics[width=0.9\textwidth]{maps_contamination}
-          \caption{Depiction of the contaminant deprojection process.
-{\sl Top left:} underlying true signal map (galaxy overdensity used in Section \ref{sssec:validation.suite.lss}).
-{\sl Top right:} sky contaminant (dust contamination used in Section \ref{sssec:validation.suite.lss}).
-{\sl Bottom left:} contaminated map obtained by adding the previous two maps.
-{\sl Bottom right:} cleaned map obtained through Eq.
-\ref{eq:clean_f}.
-Since only one contaminant is considered in this case, the cleaning is almost perfect.
-Residual contamination at the power spectrum level must still be subtracted in order to avoid a biased estimate.}
-          \label{fig:lss_cont}
-        \end{figure*}
         \begin{equation}\label{eq:decoup}
@@
     \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
-      In a typical data-analysis case, we deal with contaminants that pollute the measured signal.
-Usually, we know something about these contaminants.
-As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which can be marginalized out.
-But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added PSF fluctuations could be used to remove the possible correlations between measured ellipticities.
-Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}).
-Thus, we generically deal with a linear model of the kind:
+    In a typical data-analysis case, we deal with contaminants that
+    pollute the measured signal. Usually, we know something about
+    these contaminants. As an example, in case of CMB analysis, we
+    might have have a map of a given foreground at a different
+    frequency which can be marginalized out. But even when we do not,
+    we often have a tracer map of a given effect, for example, the map
+    of dust might tell us something about systematic errors related to
+    imperfect correction to dust reddening. Even if the dust map is
+    non-linearly and perhaps stochastically related to the amount of
+    reddening, as long as relation is local in space, we get a linear
+    contamination on large scales, very much like complex astrophysics
+    of galaxy formation nevertheless reduces to linear biasing on very
+    large scales (see Appendix \ref{app:lincont}). Thus, we generically deal with a
+    linear model of the kind:
       \begin{equation}\label{eq:aobs}
@@
 The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}.
-The process is illustrated in Figure \ref{fig:lss_cont}.
-A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{uncorrelated weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}.
+A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}.
 Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
       \begin{align}\label{eq:deproj_coeff}
-        &\hat{\alpha}_i=F_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
-        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv),
-      \end{align}
-      where there is an implicit summation over the repeated index $j$ in Eq. \ref{eq:deproj_coeff} (which we will omit from now on). The clean map is then given by
-      \begin{align}
+        &\hat{\alpha}_i=\sum_iF_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
+        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv).
+      \end{align}
+      The clean map is then given by
+      \begin{align}\label{eq:clean}
         {\bf a}^c(\nv)
-        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')
-      \end{align}
-      which, when substituting Eq. \ref{eq:aobs}, yields
-      \begin{align}\label{eq:clean}
-        {\bf a}^c(\nv)&={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv').
-      \end{align}      
+        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')\\\nonumber
+        &={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv'),
+      \end{align}
+      where there is an implicit summation over $i$ and $j$ (which we will omit from now on).
       The harmonic coefficients of ${\bf a}^c$ are therefore given by
@@
 
-      Although in general we expect the second term in Eq.
-\ref{eq:clean} to be very small, since the signal and contaminant maps are uncorrelated, due to this residual statistical contamination, a direct computation of the pseudo-$C_\ell$ of the cleaned map would yield a biased estimator beyond the mode coupling induced by the mask $v$.
+      Due to the residual statistical contamination in Eq.
+\ref{eq:clean}, a direct computation of the pseudo-$C_\ell$ of the cleaned map would yield a biased estimator beyond the mode coupling induced by the mask $v$.
 This bias can however be estimated analytically and corrected for.
@@
 
-      Two further points must be addressed before moving on.
-First, the computation of the deprojection bias requires an estimate of the true power spectrum ${\sf C}^{ab}_\ell$ (see Eqs.
-\ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}).
-In practice, this can be done through an iterative approach, starting from an initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the estimated power spectrum in the previous iteration.
-Alternatively, the ensemble average of Eq.
-\ref{eq:dpbias0} can be seen as a convolution of ${\sf C}^{ab}_\ell$, which could be treated by computing the convolution matrix as a modified version of the standard mode-coupling matrix that accounts for mode deprojection.
-Estimating this convolution matrix in this case becomes significantly more computationally intensive, and therefore an iterative approach is preferred.
+      Two further points must be addressed before moving on. First,
+      the computation of the deprojection bias requires an estimate of
+      the true power spectrum ${\sf C}^{ab}_\ell$ (see
+      Eqs. \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}). In practice, this
+      can be done through an iterative approach, starting from an
+      initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the
+      estimated power spectrum in the previous iteration. A
+      sufficiently good guess will likely be enough to remove the bias
+      in a single iteration.  Alternatively, the ensemble average of
+      Eq. \ref{eq:dpbias0} can be seen as a convolution of
+      ${\sf C}^{ab}_\ell$, which could be treated by computing the
+      convolution matrix as a modified version of the standard
+      mode-coupling matrix that accounts for mode
+      deprojection. Estimating this convolution matrix in this case
+      becomes significantly more computationally intensive, and
+      therefore an iterative approach is preferred.
 
@@
 if a set of templates are linearly related).
-The problem can however be solved self-consistently by using the Moore-Penrose pseudo-inverse \citep{2017MNRAS.465.1847E}.
+The problem can however be solved self-consistently by using the Moore-Penrose pseudo-inverse.
 This is trivial for a symmetric and positive-definite matrix such as ${\sf F}$, and corresponds to setting the inverse of all of its zero eigenvalues (or all those below a given threshold) to zero.
@@
     \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
-      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa\footnote{Note that this is a different effect from the leakage between different polarized and unpolarized channels in CMB observations due to instrumental effects (asymmetric beams, pair differencing, polarization angle calibration etc.)}.
-The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB.
-In this case, removing the leakage at the power spectrum level (i.e.
-the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes.
-This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above.
-In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e.
-purifying $E$-modes from $B$-mode contamination is conceptually exactly the same.
-Our discussion is also specific for spin-2 fields.
+    As described in
+    \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B},
+    $B$-mode purification refers to the map-level removal of the
+    contamination from $E$-modes in a the $B$-mode component of a
+    given map caused by an incomplete sky coverage and vice-versa. The
+    procedure is particularly useful in situations in which the
+    $E$-mode component of the signal is significantly larger than the
+    $B$ modes, as is the case in CMB. In this case, removing the
+    leakage at the power spectrum level (i.e. the standard
+    pseudo-$C_\ell$ approach) produces a sub-optimal estimator in
+    which the variance in the $B$-mode power spectrum is dominated by
+    the variance of the leaked $E$-modes. This section presents the
+    purification method and the associated modifications to the
+    pseudo-$C_\ell$ algorithm as described in the references above. In
+    this section we describe the algorithm in terms of purifying
+    $B$-modes from $E$-mode contamination, but we note that the
+    reverse, i.e. purifying $E$-modes from $B$-mode contamination is
+    conceptually exactly the same. Our discussion is also
+    specific for the spin-2 fields.
 
@@
       \begin{equation}
-        ({\bf D}^E_s\varphi,{\bf D}^B_s\psi)=\int d\nv\,({\bf D}^E_s\varphi)^\dag{\bf D}^B_s\psi=0,
-      \end{equation}
-      where $\varphi$ and $\psi$ are two arbitrary scalar fields.
+        ({\bf D}^E_s\varphi,{\bf D}^B_s\psi)=\int d\nv\,({\bf D}^E_s\varphi)^\dag{\bf D}^B_s\psi=0
+      \end{equation}
       This can be done by integrating by parts and noting that the celestial sphere has no boundaries.
@@
 
-      Let $w(\nv)$ be a sky window function defining the sky region to be analyzed (and the weight to be applied in each pixel).
-The standard pseudo $B$-mode of a spin-2 field ${\bf P}$ is then given by
+      Let $w(\nv)$ be a sky window function defining the sky region to be analyzed (and the weight to be applied in each pixel). The standard pseudo $B$-mode of a field ${\bf P}$ is then given by
       \begin{align}\nonumber
@@
       \begin{equation}
-        B^p_{\bf l}=\int d\nv\left({\bf D}^B_s(w\,Y_{\bf l})\right)^\dag\,{\bf P}(\nv).
-      \end{equation}
-      Since ${\bf D}^B_s(wY_{\bf l})$ is a $B$-mode quantity, $B^p_{\bf l}$ should receive contributions only from $B$-modes.
+        B^p_{\bf l}=\int d\nv\left({\bf D}^B_2(w\,Y_{\bf l})\right)^\dag\,{\bf P}(\nv).
+      \end{equation}
+      Since ${\bf D}^B_2(wY_{\bf l})$ is a $B$-mode quantity, $B^p_{\bf l}$ should receive contributions only from $B$-modes.
 
@@
       \begin{equation}\label{eq:pureb}
-        B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2\frac{\beta_{\ell,2}}{\beta_{\ell,1}}\left(\tilde{P}_1\right)^B_{\bf l}+\beta_{\ell,2}\left(\tilde{P}_0\right)^B_{\bf l},
-      \end{equation}
-      where $(a)^B_{\bf l}$ stands for the $B$-mode component of field ${\bf a}$, and we have defined the fields $\tilde{P}_n=(\eth^{2-n}w)^*(Q+iU)$, where $Q$ and $U$ are the real and imaginary parts of the field $P$ (see Appendix \ref{app:shts} for the definitions of $\eth$ and $\beta_{\ell,s}$).
+        B^p_{\bf l}=\left(\tilde{P}_2\right)^B_{\bf l}+2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}}\left(\tilde{P}_1\right)^B_{\bf l}+\alpha_{\ell,2}\left(\tilde{P}_0\right)^B_{\bf l},
+      \end{equation}
+      where $(a)^B_{\bf l}$ stands for the $B$-mode component of field ${\bf a}$, and we have defined the fields $\tilde{P}_n=(\eth^{2-n}w)^*(Q+iU)$, where $Q$ and $U$ are the real and imaginary parts of the field $P$ (see Appendix \ref{app:shts} for the definitions of $\eth$ and $\alpha_{\ell,s}$).
 
@@
       \begin{align}
-        \eth^nw={\cal S}^{-1}\left[\left(-\frac{w_{\bf l}}{\beta_{\ell,n}},0\right)\right],
+        \eth^nw={\cal S}^{-1}\left[\left(-\frac{w_{\bf l}}{\alpha_{\ell,n}},0\right)\right],
       \end{align}
@@
       \begin{align}\nonumber
-        &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}+2\frac{\beta_{\ell,2}}{\beta_{\ell,1}\beta_{\ell'',1}}\wtj{\ell}{\ell'}{\ell''}{1}{-2}{1}\\
-        &\,\,\,+\frac{\beta_{\ell,2}}{\beta_{\ell'',2}}\wtj{\ell}{\ell'}{\ell''}{0}{-2}{2}.
-      \end{align}
-      per purified field.
+        &\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}+2\frac{\alpha_{\ell,2}}{\alpha_{\ell,1}\alpha_{\ell'',1}}\wtj{\ell}{\ell'}{\ell''}{1}{-2}{1}\\
+        &\,\,\,+\frac{\alpha_{\ell,2}}{\alpha_{\ell'',2}}\wtj{\ell}{\ell'}{\ell''}{0}{-2}{2}.
+      \end{align}
+      per purified field. \todo{check this}.
 
@@
 a stochastic contaminant that we do not have a template for).
-Furthermore, the signal component will often be smoothed on the smallest scales by an instrumental beam (although this latter effect is less relevant for galaxy surveys).
+Furthermore, the signal component will often be smoothed on the smallest scales by an instrumental beam.
 We discuss how to deal with both of these complications here.
@@
           \end{equation}
-          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$.
-Note that since any continuous tangent vector field on the sphere must, at some location, be zero (the so-called hairy ball theorem), it is impossible to have an anisotropic beam that is uniform over the entire sphere.
-However, it is possible that the beam is both anisotropic and spatially-varying.
-In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam.
-Correcting for the beam in such cases can also be done analytically, but we leave this for future work.
-In practice, if required, the
-          suppression factor can be found by applying the PCL estimator to a a set of mock input maps with the same underlying signal but  with and without application of the beam smoothing (to both data maps and contaminant maps) and taking the ratio.
+          where $W^{ab}_\ell$ is the product of the harmonic transform
+          of the beams for maps $a$ and $b$. Note that since
+          any continuous tangent vector field on the sphere 
+          must, at some location, be zero (hairy ball theorem)
+          it is impossible to have an anisotropic beam
+          that is uniform over the entire sphere.  However, it is
+          possible that the beam is both anisotropic and spatially
+          varying. In this case, the pseudo-$C_\ell$ power spetrum
+          estimation will produce biased estimates that will be
+          suppressed by an \emph{effective} beam. Correcting for the
+          beam in such cases can also be done analytically, but we
+          leave it for future work.  In practice, if required, the
+          suppression factor can be found by applying the PCL
+          estimator to a a set of mock input maps with the same
+          underlying signal but  with and without application of the
+          beam smoothing (to both data maps and contaminant maps) and taking
+          the ratio. \as{How does this interact with E/B purification?}
 
@@
          \paragraph*{Uncorrelated noise.}
-           If the noise is inhomogeneous but uncorrelated (i.e.
+           If the noise is homogeneous but uncorrelated (i.e.
 $\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle=\mI\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
@@
                      {\rm PCL}_\ell({\bf n},{\bf n})&=\mI\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
-                                                    &=\mI A^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
-                   \end{align}
-                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=A\,\sigma_n^{-2}(\nv)$.
+                                                    &=\mI\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
+                   \end{align}
+                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=\alpha\sigma_n^{-2}(\nv)$.
              \item The additional bias from contaminant deprojection can also be computed analytically, starting from Eq. \ref{eq:dpbias0}, as:
@@
     \subsection{Flat-sky pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.flat}
-      Using the curved-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
+      Using the full-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
       \begin{enumerate}
-        \item Curved-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch.
+        \item Full-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch.
 This can lead to an inefficient use of memory and storage space, because the standard numerical implementations of spherical transforms operate on full-sky data.
@@
 ${\cal O}(N_{\rm pix}\log(N_{\rm pix}))$ for DFTs) and because by SHTs involve operations over the full sphere.
-        \item Fourier transforms over a flat patch offer exact quadrature, which makes them inherently more numerically stable\footnote{Note that it is also possible to achieve exact quadrature with certain spherical pixelization schemes (e.g.
-\cite{2011IJMPD..20.1053D}).}.
+        \item Fourier transforms over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
       \end{enumerate}
-      For this reason, \nmt\, also supports the computation of power spectra from flat-sky maps.
+      For this reason, \nmt, also supports the computation of power spectra from flat-sky maps.
 Appendix \ref{app:flat} describes the dictionary between spin-$s$ fields and their Fourier/harmonic coefficients defined in flat and curved skies, and we present the flat-sky versions of the pseudo-$C_\ell$ methods described in the previous sections here.
@@
         \end{equation}
-        Here, ${\cal D}$ stands for a standard 2D discrete Fourier transform, and ${\sf R}$ is a rotation matrix in 2 dimensions.
         Note that, in the context of flat-sky fields, the vector ${\bf l}$ is a 2D wave vector, and must not be confused with the abbreviation ${\bf l}\equiv(\ell,m)$ used in previous sections.
-The quantity ${\bf a}_{\bf l}$ is therefore the Fourier transform of the real-space flat-sky field ${\bf a}({\bf x})$, where ${\bf x}$ is the angular displacement with respect to a given reference point.
-Using Eq.
-\ref{eq:pcl_flat_masked} it is then straightforward to estimate the covariance of two masked fields at the same wavenumber:
+        It is then straightforward to estimate the covariance of two masked field at the same wavenumber:
         \begin{equation}\label{eq:flat_coup}
@@
 
-        At this stage it is natural to connect directly with the final bandpowers since, unlike in the curved-sky case, there is no natural minimal binning ($\Delta\ell=1$) for the un-masked power spectrum.
+        At this stage it is natural to connect directly with the final bandpowers since, unlike in the full-sky case, there is no natural minimal binning ($\Delta\ell=1$) for the un-masked power spectrum.
 In this case we will define a bandpower ${\sf B}_q$, indexed by an integer $q$ as the average of the covariance above over a set of ${\bf l}$ values $S_q$:
@@
 Taking the continuum approximation is unfortunately not accurate enough in practice, as described in \cite{2016arXiv161204664A}, and a fully discrete approach is usually necessary.
-This slower performance is normally compensated by the smaller pixel numbers and faster DFTs that can be used in the flat sky, so that the flat-sky approximation is still preferable over a curved-sky treatment for small sky patches.
+This slower performance is normally compensated by the smaller pixel numbers and faster DFTs that can be used in the flat sky, so that the flat-sky approximation is still preferable over a full-sky treatment for small sky patches.
 Appendix \ref{app:flat.pcl} discusses these issues in detail, providing the continuum-limit expressions mentioned above.
@@
     \begin{itemize}
-      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for curved-sky calculations.
+      \item {\tt HEALPix} \citep{2005ApJ...622..759G} is currently the only pixelization scheme supported for full-sky calculations.
 Through {\tt HEALPix}, \nmt~also makes use of {\tt cfitsio}~\citep{1999ASPC..172..487P}.
-      \item The basic curved-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citep{2013A&A...554A.112R}.
+      \item The basic full-sky operation carried out by \nmt~is the spherical harmonic transform. For this, the code makes use of {\tt libsharp}~\citep{2013A&A...554A.112R}.
       \item In flat sky, SHTs are replaced by 2D DFTs, for which \nmt~uses the {\tt FFTW} library~\citep{FFTW05}.
@@
             templates = [[dust_q,dust_u]],
-            purify_b = True, beam=beam_ell)
+            purify_b = True,beam=beam_ell)
 f0 = NmtField(mask, [map_t])
@@
       \item All preliminary I/O operations (reading maps, masks etc.) take place implicitly in line 3.
-      \item In line 5 we create a field with spin 2.
+      \item In line 5 we generate a field with spin 2.
 The field is defined in terms of a mask and maps of the $(Q,U)$ Stokes parameters.
@@
       \item The pseudo-$C_\ell$ is computed in line 17.
-      \item The deconvolved bandpowers are finally estimated using the workspace in line 19. Note that the mode-coupling matrix must have been precomputed in line 13 for this to be successful.
+      \item The deconvolved bandpowers are finally estimated using the workspace in line 19.
     \end{enumerate}
@@
 This section presents these two examples and the results of this validation in terms of accuracy and computational performance.
-Results will be presented for both curved-sky and flat-sky realizations.
-Note that, in all cases, we use the true signal power spectra as input to compute the bias from contaminant deprojection (see Eqs.
-\ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}).
-This allows us to verify that the code works as expected, and to isolate any residual bias associated with software bugs, rather than an imperfect guess of the true power spectrum.
+Results will be presented for both full-sky and flat-sky realizations.
 
@@
           \caption{Signal and noise power spectra used in the LSS validation set.
-Dark blue, cyan and red curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode.
+Red, green and blue curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode.
 All other cross-correlations are zero.
-The dashed lines show the noise power spectra, associated with shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
+The dashed lines show the noise power spectra, associated to shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
         \end{figure}
@@
           \includegraphics[width=0.9\textwidth]{maps_lss_sph}
-          \caption{Inputs for the large-scale structure, curved-sky validation set.
+          \caption{Inputs for the large-scale structure, full-sky validation set.
 From top to bottom and left to right: sky mask (1), Gaussian realization of the galaxy overdensity $\delta$ (2) and cosmic shear $\gamma_i$ (3,4), dust contaminant for $\delta$ (5), star contaminant for $\delta$ (6), PSF contaminant for $\gamma$ (7, $Q$ component) and small-scale contaminant for $\gamma$ (8, $Q$ component).}
@@
           \includegraphics[width=0.99\columnwidth]{cls_cont_lss}
-          \caption{{\sl Top:} signal power spectrum (solid dark blue) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (dashed cyan) and stars (dot-dashed red).
-{\sl Middle:} $E$-mode lensing power spectrum (solid dark blue) and the spectra of the PSF and small-scale contaminants used in our validation set (dashed cyan and dot-dashed red respectively).
+          \caption{{\sl Top:} signal power spectrum (red) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (green) and stars (blue).
+{\sl Middle:} $E$-mode lensing power spectrum (red) and the spectra of the PSF and small-scale contaminants used in our validation set (green and blue respectively).
 {\sl Bottom:} same as middle panel for the $B$-mode power spectrum.
 Note that in this case the input signal power spectrum is $0$.
-In all cases the orange dotted line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
+In all cases the yellow line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
         \end{figure}
-        The first validation suite exemplifies the use of pseudo-$C_\ell$ algorithms in the analysis of large-scale structure data.
-Current and next-generation imaging surveys such as DES~\citep{2016MNRAS.460.1270D},
-        KiDS~\citep{2017A&A...604A.134D}, HSC~\citep{2018PASJ...70S...4A}, LSST~\citep{2008arXiv0805.2366I} and Euclid~\citep{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing.
-These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
-
-        Large-scale structure data are often characterized by inhomogeneous sky coverages and complex mask structures, associated with the presence of bright stars and observational artifacts.
-They are also affected by many sky contaminants, associated to Galactic sources (dust absorption, star density) and observing conditions (seeing, airmass etc.) as discussed in \citet{2016ApJS..226...24L}.
-The typically flatter power spectra for both galaxy clustering and lensing (e.g.
-when compared with the damping tail of CMB temperature fluctuations), makes the pseudo-$C_\ell$ approach ideal for this kind of analysis \citep{2013MNRAS.435.1857L,2016ApJS..226...24L}.
-This science case therefore allows us to validate several aspects of the code simultaneously:
+        The first validation suite exemplifies the use of
+        pseudo-$C_\ell$ algorithms in the analysis of large-scale
+        structure data. Current and next-generation imaging surveys
+        such as DES~\citep{2016MNRAS.460.1270D},
+        KiDS~\citep{2017A&A...604A.134D},
+        HSC~\citep{2018PASJ...70S...4A},
+        LSST~\citep{2008arXiv0805.2366I} and
+        Euclid~\citep{2011arXiv1110.3193L} will constrain the growth
+        and geometry of structure through cosmic time by using the
+        overdensity of galaxies as a proxy for the matter
+        inhomogeneities (a spin-0 field) and their correlated shape
+        distortions (a spin-2 field at first order), mostly caused by
+        weak lensing. These measurements will mostly follow a
+        tomographic approach, using the angular auto- and cross-power
+        spectra between these observables across different bins of
+        photometric redshift
+        \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.
+
+        Large-scale structure data are often characterized by
+        inhomogeneous sky coverages and complex mask structures,
+        associated with the presence of bright stars and observational
+        artifacts. They are also affected by many sky contaminants,
+        associated to Galactic sources (dust absorption, star density)
+        and observing conditions (seeing, airmass
+        etc.) as discussed in \citet{2016ApJS..226...24L}. The typically flatter power
+        spectra for both galaxy clustering and lensing (e.g. when
+        compared with the damping tail of CMB temperature
+        fluctuations), makes the pseudo-$C_\ell$ approach ideal for
+        this kind of analysis \citep{2013MNRAS.435.1857L,2016ApJS..226...24L}. This science case therefore allows us to validate several aspects of the code simultaneously:
         \begin{itemize}
@@
         For this validation suite we therefore use a mock dataset mimicking a typical observation of an LSST-like survey in a particular redshift bin.
-Slightly different set-ups are used for the curved-sky and flat-sky validation suites.
+Slightly different set-ups are used for the full-sky and flat-sky validation suites.
 The characteristics of these data are as follows:
@@
 \ref{fig:lss_cls}.
-Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the curved-sky and flat-sky cases respectively.
-          \item {\bf Mask:} for the curved-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}.
-To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask.
-We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig.
-\ref{fig:lss_mocks_sph}).
-For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citep{2018PASJ...70S...4A}.
+Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the full-sky and flat-sky cases respectively.
+          \item {\bf Mask:} for the full-sky data, we use a mask built
+            as a combination of the sky coverage output by the LSST
+            {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a
+            more conservative galactic cut based on the reddening map
+            of \citet{1998wfsc.conf..297S}. To add an extra layer of
+            complexity, we drill 100 additional 1-degree holes on the
+            resulting combined mask. We also explore the improvement
+            stemming from the down-weighting of higher-noise in an
+            inverse-variance way in the presence of inhomogeneous
+            noise (see description below), which is the reason why the
+            default version of this mask is not binary, but tracers
+            the inverse of the noise variance (see top left panel of
+            Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we
+            use a high-resolution mask constructed from the
+            bright-object mask used for the first public data release
+            from the HSC collaboration~\citep{2018PASJ...70S...4A}.
 The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales.
@@
           \item {\bf Contaminants:} we contaminate the signal maps described above with several different types of residuals.
-In all cases we add contaminants linearly, to avoid deviating from the model described in Section \ref{ssec:maths.deproj}.
-The contaminant amplitudes are chosen so that any residual contamination in the relevant range of multipoles will be detected statistically by the validation suite.
-Maps of these contaminants are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.
-\ref{fig:lss_cont_cls} for the curved-sky case.
+Maps of these are shown in the 4 bottom panels of Figure~\ref{fig:lss_mocks_sph} and Figure~\ref{fig:lss_mocks_flat}, and their impact on the angular power spectrum is presented in Fig.
+\ref{fig:lss_cont_cls} for the full-sky case.
           \begin{itemize}
-            \item {\sl Dust.} For the curved-sky suite, we include a linear contaminant for $\delta_g$ proportional to the fluctuations around the mean of the dust reddening map of~\citet{1998wfsc.conf..297S}.
+            \item {\sl Dust.} For the full-sky suite, we include a linear contaminant for $\delta_g$ proportional to the fluctuations around the mean of the dust reddening map of~\citet{1998wfsc.conf..297S}.
 This contamination affects mostly the largest scales, and simulates the effect of an imperfect correction of the dust absorption on the galaxy number density.
 The amplitude of the dust fluctuations is chosen so that their power spectrum within the unmasked region be $\sim20\%$ of the galaxy power spectrum at $\ell\sim120$.
-For the flat-sky suite, we generate a flat-sky Gaussian realization of a field with a power spectrum $C_\ell\propto\ell^{-2.4}$, corresponding to the large-$\ell$ behaviour of the curved-sky dust map.
+For the flat-sky suite, we generate a flat-sky Gaussian realization of a field with a power spectrum $C_\ell\propto\ell^{-2.4}$, corresponding to the large-$\ell$ behaviour of the full-sky dust map.
 The amplitude was chosen such that the contaminant power spectrum be $\sim10\%$ of the signal at $\ell\sim400$.
             \item {\sl Stars.} We add a second contaminant to $\delta_g$ associated with star contamination.
-For the curved-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citep{2014SPIE.9150E..14C}.
+For the full-sky realizations this contaminant is proportional to the fluctuation around the mean of a star density map generated using the LSST {\tt CatSim} catalog~\citep{2014SPIE.9150E..14C}.
 We modify this contaminant slightly by adding a white-noise component that dominates the small-scale spectrum beyond $\ell\simeq400$.
@@
             where $J_1(x)$ is the order-1 Bessel function, and $2J_1(\ell\theta)/\ell\theta$ is the Fourier transform of a circular aperture with radius $\theta$.
-For this we assume $\theta_{\rm FoV}=1.75^\circ$, and therefore we try to mimic the signature of PSF fluctuations that vary between different pointings with a 3.5$^\circ$ diameter.
+For this we assume $\theta_{\rm FoV}=1.75^\circ$.
 The amplitude of the power spectrum is chosen so that the contaminant amounts to $30\%$ of the signal power spectrum at $\ell\sim50$.
-We use the same prescription for both the curved-sky and the flat-sky cases.
+We use the same prescription for both the full-sky and the flat-sky cases.
             \item {\sl Small-scale contamination.} Finally we include an additional small-scale contaminant to weak lensing with a flat power spectrum, unit $E$-$B$ ratio, and an amplitude fixed so the contaminant's power spectrum is $20\%$ of the signal at $\ell=500$.
@@
           \end{itemize}
-          \item {\bf Noise:} for the curved-sky case we generate inhomogeneous noise realizations.
+          \item {\bf Noise:} for the full-sky case we generate inhomogeneous noise realizations.
 To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is then intrinsic ellipticity scatter per component.
@@
 
-          \item {\bf Pixelization:} the main suite of curved-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$.
+          \item {\bf Pixelization:} the main suite of full-sky simulations used to validate the code use \texttt{HEALPix} resolution parameter $N_{\rm side}=1024$, corresponding to a pixel size of $\sim3.4'$.
 This resolution allows us to validate the code on a large number of realizations without facing strong computational challenges on the main scales of interest ($\ell\lesssim2000$).
@@
           \includegraphics[width=0.99\columnwidth]{cls_cmb}
-          \caption{$E$-mode and $B$-mode power spectra (blue and yellow respectively) for the cosmological CMB signal (solid lines), for the curved-sky foreground contaminants (dashed lines) and for the flat-sky foreground contaminants (dot-dashed lines).
+          \caption{$E$-mode and $B$-mode power spectra (blue and yellow respectively) for the cosmological CMB signal (solid lines), for the full-sky foreground contaminants (dashed lines) and for the flat-sky foreground contaminants (dot-dashed lines).
 The noise power spectra used for the full- and flat-sky simulations are shown as black dashed and dot-dashed lines respectively.
@@
           \includegraphics[width=0.9\textwidth]{maps_cmb}
-          \caption{{\sl Left panels:} mask (top), and $Q$ and $U$ components (middle, bottom) of the foreground contaminant used in the CMB curved-sky validation set.
+          \caption{{\sl Left panels:} mask (top), and $Q$ and $U$ components (middle, bottom) of the foreground contaminant used in the CMB full-sky validation set.
 {\sl Right panels:} same as left panels for the flat-sky validation set.}
@@
 This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citep{2016PhRvL.116c1302B}, Advanced ACTPol~\citep{2016SPIE.9910E..14D,2017JCAP...06..031L} or the Simons Array~\citep{2016JLTP..184..805S,2017ApJ...848..121P}, as well as next-generation facilities such as the Simons Observatory\todo{cite if out} or CMB Stage-4~\citep{2016arXiv161002743A}.
-Potential future space missions, such as LiteBIRD~\citep{2018JLTP..tmp..124S} will be able to make low-resolution, curved-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
+Potential future space missions, such as LiteBIRD~\citep{2018JLTP..tmp..124S} will be able to make low-resolution, full-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
 Ground-based experiments, on the other hand, will be systematics-limited at low $\ell$, and will be able to make higher-resolution maps, for which pseudo-$C_\ell$ methods are better suited.
@@
 \ref{fig:cmb_cls}.
-          \item {\bf Mask:} for the curved-sky case we assume a survey targeting a 4000 deg$^2$ footprint.
+          \item {\bf Mask:} for the full-sky case we assume a survey targeting a 4000 deg$^2$ footprint.
 The corresponding mask is analytically apodized using the ``$C^2$'' apodization method of \cite{2009PhRvD..79l3515G} with an apodization scale of 10 degrees.
-This apodization method is based on smoothing the transition at the mask edge with a sinusoidal function that interpolates between 0 and 1 in a differentiable manner.
 The observed area encompasses one of the lowest-foreground regions observable from the Southern Hemisphere.
@@
 The assumed masks are shown in the top panels of Figure~\ref{fig:cmb_mocks}.
-          \item {\bf Noise:} for the curved-sky and flat-sky case we assume white noise levels of 1 and 0.5$\mu K\,{\rm arcmin}$ in intensity respectively.
+          \item {\bf Noise:} for the full-sky and flat-sky case we assume white noise levels of 1 and 0.5$\mu K\,{\rm arcmin}$ in intensity respectively.
 We also include a $1/f$-like component associated with atmospheric noise with a characteristic scale $\ell_{\rm knee}$ and tilt $\alpha_{\rm knee}$ such that the noise power spectrum can be written as
@@
           where $\sigma_N^2$ is the white-noise variance in one steradian.
-We use $\alpha_{\rm knee}=2.4$ and $\ell_{\rm knee}=10$ and 300 for the curved-sky and flat-sky cases respectively.
+We use $\alpha_{\rm knee}=2.4$ and $\ell_{\rm knee}=10$ and 300 for the full-sky and flat-sky cases respectively.
 The beam-deconvolved noise power spectra in both cases are shown in Fig.
 \ref{fig:cmb_cls}.
-          \item {\bf Beam:} we assume a Gaussian beam with full-width at half-maximum of $\theta_{\rm FWHM}=20$ and 1.4 arcmin for the curved-sky and flat-sky cases respectively.
+          \item {\bf Beam:} we assume a Gaussian beam with full-width at half-maximum of $\theta_{\rm FWHM}=20$ and 1.4 arcmin for the full-sky and flat-sky cases respectively.
 This resolution corresponds to what could be achieved at 150 GHz with with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case.
@@
 \ref{fig:cmb_cls}, the combination of beam size and $\ell_{\rm knee}$ makes the two experiments sensitive to the $B$-mode power spectrum in different ranges of scales, corresponding to the primordial $B$-mode signal around the recombination bump (S.A.) and to the high-ell lensing signal (L.A.).
-          \item {\bf Contaminants:} we assume a dust-like residual for both the curved-sky an flat-sky cases.
-For the curved-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citep{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$.
+          \item {\bf Contaminants:} we assume a dust-like residual for both the full-sky an flat-sky cases.
+For the full-sky set-up, we use maps of the dust emission at 150 GHz generated by {\tt PySM}~\citep{2017MNRAS.469.2821T} and scaled so that the amplitude of the residual power spectrum is about 10\% of the lensing $B$-mode signal at $\ell=50$.
 For the flat-sky signal we generate a Gaussian realization of a dust-like component with power spectra
@@
 \ref{fig:cmb_cls}.
-          \item {\bf Pixelization:} the curved-sky case, targetting the $B$-mode recombination bump at $\ell\sim100$, uses a HEALPix resolution parameter $N_{\rm side}=256$, corresponding to $14$-arcmin pixels and sufficient to measure power spectra to multipoles $\ell\sim500$.
+          \item {\bf Pixelization:} the full-sky case, targetting the $B$-mode recombination bump at $\ell\sim100$, uses a HEALPix resolution parameter $N_{\rm side}=256$, corresponding to $14$-arcmin pixels and sufficient to measure power spectra to multipoles $\ell\sim500$.
 The flat-sky case, targetting the lensing $B$-modes, uses square 2-arcmin pixels.
@@
 
-    \subsection{Curved-sky large-scale structure validation}\label{ssec:validation.lss_full}
+    \subsection{Full-sky large-scale structure validation}\label{ssec:validation.lss_full}
       \begin{figure*}
@@
         \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
-        \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
+        \caption{{\sl Left panel:} average over all simulations in our full-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
 The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
@@
         \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
-        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our curved-sky LSS validation suite (blue histograms).
+        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our full-sky LSS validation suite (blue histograms).
 Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot.
@@
 We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles.
-The typical running time for one realization is $\sim 600$s using 32 \texttt{OpenMP} threads.
-Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation.
-More details about timing and scalability of \nmt can be found in Appendix \ref{sec:scalability}.
 
@@
 \ref{fig:lss_val_cls} shows the six deconvolved bandpowers averaged over simulations (solid lines) compared with the analytical prediction for the input signal power spectra (darker dashed lines).
-The middle panel of the same figure shows the average residual with respect to the analytical prediction normalized by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).
+The middle panel of the same figure shows the average residual with respect to the analytical prediction normalize by the error on the mean (given by the standard deviation of all $N_{\rm sim}=1000$ simulations divided by $\sqrt{N_{\rm sim}}$).
 We see that the latter are accurately recovered by \nmt~in all cases, including all null correlations ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$), and that the normalized residuals fluctuate around zero with typical variations of $\Delta C_\ell/\sigma_\ell\lesssim3$, corresponding to expected $\lesssim3\sigma$ fluctuations.
-For comparison, the bottom panel of the figure shows the mean relative difference for power spectra of contaminated maps without deprojection, where we can appreciate that the residuals, $\Delta C'_{\ell}/\sigma'_{\ell}$, appear to not be centered around zero in some cases, i.e., the estimated power-spectra are biased.
+For comparison, bottom panel of the figure show the mean relative difference for power spectra of contaminated maps without deprojection, where we can appreciate that the residuals, $\Delta C'_{\ell}/\sigma'_{\ell}$, appear to not be centered around zero in some cases, i.e., the estimated power-spectra are biased.
 
@@
         \caption{{\sl Left:} average over all simulations in our
-          curved-sky CMB validation suite of the three spin-2 power
+          full-sky CMB validation suite of the three spin-2 power
           spectra ($EE$, $EB$ and $BB$, solid lines) using mode
@@
       \end{equation}
-      The full covariance matrix is very close to diagonal, as shown in the right panel of Fig.
-\ref{fig:lss_val_cls}.
-Under the assumption that the estimator is unbiased ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$ should follow a $\chi^2$ distribution with a $N_{\rm dof}$ degrees of freedom.
-Figure \ref{fig:lss_val_chi2} shows the distribution of $\chi^2_i$ across simulations approximating the covariance matrix as diagonal (blue histograms), compared with the predicted $\chi^2$ distribution for $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the six different power spectrum combinations.
-Both distributions
-      are found to be compatible.
-Moreover, the dot-dashed lines show the values of $\chi^2$ found for the mean residual with respect to the input power spectra (middle panel of Fig.
-\ref{fig:lss_val_cls}).
-The probability to exceed (PTE) associated with these $\chi^2$ values are higher than $10\%$ in all cases, and therefore our validation suite does not show any statistically significant bias in the estimator.
-This bias, if any, must be at least $\sqrt{1000}\sim30$ times smaller than the statistical uncertainties associated with our validation suite, which are representative of the statistical power of next-generation galaxy surveys.
+      The full covariance matrix is very close to diagonal, as shown
+      in the right panel of Fig. \ref{fig:lss_val_cls}. Under the
+      assumption that the estimator is unbiased
+      ($\langle{\bf s}\rangle={\bf t}$), the values of $\chi^2_i$
+      should follow a $\chi^2$ distribution with a $N_{\rm dof}$
+      degrees of freedom. Figure \ref{fig:lss_val_chi2} shows the
+      distribution of $\chi^2_i$ across simulations approximating
+      covariance matrix as diagonal (blue histograms),
+      compared with the predicted $\chi^2$ distribution for
+      $N_{\rm dof}=681$ (marked by the vertical dashed lines) for the
+      six different power spectrum combinations. Both distributions
+      are found to be compatible. Moreover, the dot-dashed lines show
+      the values of $\chi^2$ found for the mean residual with respect
+      to the input power spectra (middle panel of
+      Fig. \ref{fig:lss_val_cls}). The probability to exceed (PTE)
+      associated with these $\chi^2$ values are higher than $10\%$ in
+      all cases, and therefore our validation suite does not show any
+      statistically significant bias in the estimator. This bias, if
+      any, must be at least $\sqrt{1000}\sim30$ times smaller than the
+      statistical uncertainties associated with our validation suite,
+      which are representative of the statistical power of
+      next-generation galaxy surveys.
 
@@
 
-    \subsection{Curved-sky CMB \texorpdfstring{$B$}{B}-mode validation}\label{ssec:validation.cmb_full}
+    \subsection{Full-sky CMB \texorpdfstring{$B$}{B}-mode validation}\label{ssec:validation.cmb_full}
       \begin{figure*}
@@
         \includegraphics[width=0.99\columnwidth]{val_covar_cmb_sph}
-        \caption{{\sl Left:} covariance matrix for power spectra computed in the absence of foreground contaminants computed from the 1000 simulations of our curved-sky CMB validation suite.
+        \caption{{\sl Left:} covariance matrix for power spectra computed in the absence of foreground contaminants computed from the 1000 simulations of our full-sky CMB validation suite.
 The incomplete sky coverage causes a noticeable anti-correlation of neighboring bandpowers at low $\ell$ for the three power spectra ($EE$, $EB$ and $BB$), but beyond that, the covariance matrix is close to diagonal.
@@
         \includegraphics[width=0.80\textwidth]{val_chi2_cmb_sph}
-        \caption{Distribution of $\chi^2$ values found for the 1000 simulations in our curved-sky CMB validation suite (blue histograms).
+        \caption{Distribution of $\chi^2$ values found for the 1000 simulations in our full-sky CMB validation suite (blue histograms).
 Results are shown for the three polarized spectra ($EE$, $EB$ and $BB$), labeled in the top right corner of each panel.
@@
       \end{figure*}
-      We have carried out a similar set of tests on an ensemble of 1000 simulations from the CMB validation suite described in Section \ref{sssec:validation.suite.cmb}.
-In our fiducial run we both deprojected the foreground contaminants and subsequently used $B$-mode purification to decouple $E$ and $B$-modes at the map level, since this is the main feature of the code we want to validate.
-The left panel of Figure \ref{fig:cmb_val_cls} shows the mean of the $EE$, $EB$ and $BB$ power spectra computed for all simulations (solid lines) compared with the expectation from the input spectra for $EE$ and $BB$ (dashed lines, mostly lying under the solid ones).
-The average residuals normalized over the error on the mean, shown in the lower panel of the same figure, fluctuate around zero with an amplitude of up to $3\sigma$, as would be expected in the absence of a bias in the estimator.
-
-      The above figure was created using $E/B$ mode purification.
-Disabling this option has no impact on the bias of the estimates, but increases its variance as we show in the Figure \ref{fig:cmb_val_sigma}.
+      We have carried out a similar set of tests on an ensemble of
+      1000 simulations from the CMB validation suite described in
+      Section \ref{sssec:validation.suite.cmb}. In our fiducial run we
+      both deprojected the foreground contaminants and subsequently
+      used $B$-mode purification to decouple $E$ and $B$-modes at the
+      map level, since this is the main feature of the code we want to
+      validate. The left panel of Figure \ref{fig:cmb_val_cls} shows
+      the mean of the $EE$, $EB$ and $BB$ power spectra computed for
+      all simulations (solid lines) compared with the expectation from
+      the input spectra for $EE$ and $BB$ (dashed lines, mostly lying
+      under the solid ones). The average residuals normalized over the
+      error on the mean, shown in the lower panel of the same Figure,
+      fluctuate around zero with an amplitude of up to $3\sigma$, as
+      would be expected in the absence of a bias in the
+      estimator.
+
+      The above figure was created using E/B mode
+      purification. Disabling this option has no impact on the bias of
+      the estimates, but increases its variance as we show in the
+      Figure \ref{fig:cmb_val_sigma}.
 We plot diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra.
@@
 
-    This paper presents \nmt, the first public, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalizations (flat-sky, deprojection and purification).
+    This paper presents \nmt, the first robust, validated and easy-to-use power spectrum estimation software package that implements the pseudo-$C_\ell$ method to compute angular cross-power spectra for any pair of spin-0 or spin-2 fields, including all their generalizations (flat-sky, deprojection and purification).
 We have described the steps taken to validate the code, showing that it is able to provide unbiased results in a wide variety of scenarios.
@@
 As shown in Section \ref{sec:validation}, we are able to recover the input power spectrum in all cases (which include varying degrees of contamination, purification, different spins etc.) with no detectable bias down to $\lesssim1/30$-th of the statistical uncertainties.
-
-    Besides describing and validating the code, and in order to generalize the pseudo-$C_\ell$ estimator to all the cases supported by \nmt, this paper has also presented a number of results that are, to the best of our knowledge, new.
-These include:
-    \begin{itemize}
-      \item Extending the analytical treatment of contaminant deprojection presented in \cite{2016MNRAS.456.2095E} to fields of arbitrary spin.
-      \item Combining contaminant deprojection and $E/B$ purification.
-      \item Consistently deriving all of this functionality (standard pseudo-$C_\ell$, deprojection and purification for arbitrary spins) in the flat-sky approximation.
-    \end{itemize}
 
@@
 
-    \nmt~does not currently produce a covariance matrix estimate.
-Although computationally efficent analytical estimates of the Gaussian covariance matrix can be derived under certain approximations \citep{2004MNRAS.349..603E,2005MNRAS.360.1262B}, this is not widely supported by the code for several reasons.
-First, for many problems, such as large-scale structure power spectrum estimation, the covariance matrix contains important contributions from the connected four-point function that depend on the precise nature of the problem and go beyond the estimation step that \nmt~strives to do well.
-Second, with all additional options included, especially if a large number of contaminant templates are deprojected, the exact analytical estimates are either computational unfeasible or overly approximate.
-Since a majority of the forthcoming experiments will rely on a large number of mock datasets available for testing systematic effects, these maps also offer a natural path towards estimating the data covariance.
-
-    \nmt~is open-source software and publicly available at \url{https://github.com/LSSTDESC/NaMaster}.
-To maximize the usefulness of this software for the community, it is accompanied by an extensive documentation and example code\footnote{See \url{https://github.com/LSSTDESC/NaMaster}.}.
-\nmt~is written in {\tt C} and {\tt OpenMP}-parallelized to maximize its performance, although we encourage its use through the companion {\tt python} wrapper, the fastest-growing language of choice in the analysis of astronomical data.
-The code is in constant development, and we will strive to implement further functionality to e.g.
-estimate covariance matrices, extend the range of supported sky pixelizations and enhance its computational performance.
+    \nmt~does not currently produce a covariance matrix estimate. This
+    is done for several reasons. First, for many problems, such as
+    large-scale structure power spectrum estimation, the covariance
+    matrix contains import contributions from the connected four-point
+    function that depend on the precise nature of the problem and go
+    beyond the estimation step that \nmt~strives to do well. Second,
+    with all additional options included, the exact analytical
+    estimates are either computational infeasible or overly
+    approximate. Since majority of the forthcoming experiments will
+    have a large number of mock datasets available for testing
+    systematic effects, these maps also offer a natural path towards
+    estimating the data covariance.
+
+    \nmt~is open-source software and publicly available at \url{https://github.com/LSSTDESC/NaMaster}. To maximize the
+    usefulness of this software for the community, it is accompanied
+    by an extensive documentation and example code\footnote{See \url{
+        https://github.com/LSSTDESC/NaMaster}.}. \nmt~is written in
+    {\tt C} and {\tt OpenMP}-parallelized to maximize its performance,
+    although we encourage its use through the companion {\tt python}
+    wrapper, the fastest-growing language of choice in the analysis of
+    astronomical data. The code is in constant development, and we
+    will strive to implement further functionality to e.g. estimate
+    covariance matrices, extend the range of supported sky
+    pixelizations and enhance its computational performance.
 
 \subsection*{Acknowledgements}
-We would like to thank Rupert Allison, Scott Dodelson, Jo Dunkley, Franz Elsner, Boris Leistedt, Francois Lanusse, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya Peiris, Naomi Robertson and Ben Thorne for useful comments and discussions.
-DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
+We would like to thank Rupert Allison, Jo Dunkley, Franz Elsner, Boris
+Leistedt, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya
+Peiris, Naomi Robertson and Ben Thorne for useful comments and
+discussions. DA acknowledges support from the Beecroft trust and from
+the Science and Technology Facilities Council (STFC) through an Ernest
+Rutherford Fellowship, grant reference ST/P004474/1. This research
+used resources of the National Energy Research Scientific Computing
+Center (NERSC), a U.S. Department of Energy Office of Science User
+Facility operated under Contract No. DE-AC02-05CH11231.
 
@@
     \begin{equation}
-      _sY_{\ell m}\equiv \beta_{\ell,s} \eth^s Y_{\ell m},\hspace{6pt}
-      _{-s}Y_{\ell m}\equiv \beta_{\ell,s} (-1)^s\bar{\eth}^s Y_{\ell m},
-    \end{equation}
-    where $\beta_{\ell,s}\equiv\sqrt{(\ell-s)!/(\ell+s)!}$ and $Y_{\ell m}$ are the standard spherical harmonics. These functions satisfy the property: $(_sY_{\ell m})^*=(-1)^{s+m}\,_{-s}Y_{\ell-m}$.
+      _sY_{\ell m}\equiv \alpha_{\ell,s} \eth^s Y_{\ell m},\hspace{6pt}
+      _{-s}Y_{\ell m}\equiv \alpha_{\ell,s} (-1)^s\bar{\eth}^s Y_{\ell m},
+    \end{equation}
+    where $\alpha_{\ell,s}\equiv\sqrt{(\ell-s)!/(\ell+s)!}$ and $Y_{\ell m}$ are the standard spherical harmonics. These functions satisfy the property: $(_sY_{\ell m})^*=(-1)^{s+m}\,_{-s}Y_{\ell-m}$.
 
@@
       _s{\bf Y}^E_{\ell m}&\equiv {\bf D}^E_sY_{\ell m}
-      \equiv-\frac{\beta_{\ell,s}}{2}\left(\begin{array}{c}
+      \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
                               \eth^s+\bar{\eth}^s\\
@@
           _s{\bf Y}^B_{\ell m}&\equiv {\bf D}^B_sY_{\ell m}
-          \equiv-\frac{\beta_{\ell,s}}{2}\left(\begin{array}{c}
+          \equiv-\frac{\alpha_{\ell,s}}{2}\left(\begin{array}{c}
                               i(\eth^s-\bar{\eth}^s)\\
@@
 
-  \section{Linear contamination from non-linear contamination} \label{app:lincont}
-    Consider a contaminant field $c$ (e.g.
-a point-spread function (PSF) size) that affects the observable field $o$ (e.g.
-the number count densities) through a non-linear, but \emph{local} on some scale local functional $F$:
+\section{Linear contamination from non-linear
+  contamination} \label{app:lincont}
+\as{It is up to David to decide whether he buys this argument or
+  not. I'm fine with not and then we soften then claim in 2.2. Doesn't
+  really affect any of the paper otherwise.}
+Imagine a contaminant field $c$ (e.g. a point-spread function (PSF) size)
+that affects the observable field $o$ (e.g. the number count
+densities) through a non-linear, but \emph{local} on some scale local
+functional $F$:
 \begin{equation}
@@
 \end{equation}
-    where we can, without loss of generality, take out the mean effects (i.e.
-there is some mean PSF size and fluctuations around this mean size produce additional fluctuations in the number density of fluctuations).
-
-    In general, for sufficiently non-linear functionals $F$, the Taylor expansion around $\delta_c(\nv)$ is not valid.
-However, one can look at the effect of smoothing by a kernel of size $S(R)$, which in Fourier space is equivalent by multiplication by $e^{-k^2R^2}$.
-We get
-    \begin{equation}
-      \delta^R_o(\nv) \rightarrow \delta^R_o  + S(R)[F [ \delta_c  ]] =  \delta^R_o + F'_R [ \delta^R_c ],
-    \end{equation}
-    where we used a shorthand $\delta_x^R=S(R)[\delta_x]$ and  $F'_R$ is now a \emph{different} functional that takes a smoothed $\delta_c$ and returns out a smoothed $\delta_o$.
-The crucial point is that $F'_R$ must still be local, since one cannot turn a local functional into a non-local one by application of smoothing.
-
-    This means, that for a well behaved $\delta_c$, such that variances of both $\delta^R_c$ and $\delta^R_o$ are small for a large enough $R$, the Taylor expansion is valid for $F'_R$ (even if it is not for $F$), giving
+where we can, without loss of generality, take out the mean effects
+(i.e. there is some mean PSF size and fluctuations around this mean
+size produce additional fluctuations in the number density of
+fluctuations).
+
+In general, for sufficiently non-linear functionals $F$, the Taylor
+expansion around $\delta_c(\nv)$ is not valid. However, one can look
+at the effect of smoothing by a kernel of size $S(R)$, which in
+Fourier space is equivalent by multiplication by $e^{-k^2R^2}$. We get
+\begin{equation}
+  \delta^R_o(\nv) \rightarrow \delta^R_o  + S(R)[F [ \delta_c
+  ]] =  \delta^R_o + F'_R [ \delta^R_c ],
+\end{equation}
+where we used a shorthand $\delta_x^R=S(R)[\delta_x]$ and  $F'_R$ is now a \emph{different} functional that takes a smoothed
+$\delta_c$ and returns out a smoothed $\delta_o$.  The crucial point
+is that $F'_R$ must still be
+local, since one cannot turn a local functional into a non-local one
+by application of smoothing.
+
+This means, that for a well behaved $\delta_c$, such that variances of
+both $\delta^R_c$ and $\delta^R_o$ are small for a large enough $R$,
+the Taylor expansion is valid for $F'_R$ (even if it is not for $F$),
+giving
 \begin{equation}
@@
 \end{equation}
-    for $k\ll R^{-1}$.
-In other words a ``local'' contaminant becomes ``linearly biased'' in the $k\rightarrow 0 $ regime, very much like very non-linear physics of galaxy formations results in linearly biased galaxy field on large scales.
+for $k\ll R^{-1}$. In other words a ``local'' contaminant becomes
+``linearly biased'' in the $k\rightarrow 0 $ regime, very much like
+very non-linear physics of galaxy formations results in
+linearly biased galaxy field on large scales.
 
@@
     \subsection{Spin-\texorpdfstring{$s$}{s} fields in flat sky}\label{app:flat.fields}
-      In the flat sky we will label a position in the plane by two coordinates $(x,y)\equiv{\bf x}$.
-These coordinates can be directly related to increments in the spherical coordinates $(\theta,\varphi)$ by considering a sufficiently small map centered around the equator (i.e.
-$\sin\theta\sim 1$).
-In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
-      shifts ($\delta\varphi=\delta y$) the differential operator $\eth$ now takes the form: %\todo{make a plot of this?}.
-\as{no, but isn't it other way round?
-Latitute is $y$ in my brain?
-:)} \dam{it isn't, have a look here \url{https://healpix.jpl.nasa.gov/html/intronode12.htm}.} Using this dictionary,
+      In the flat sky we will label a position in the plane by two
+      coordinates $(x,y)\equiv{\bf x}$. These coordinates can be
+      directly related to increments in the spherical coordinates
+      $(\theta,\varphi)$ by considering a sufficiently small map
+      centered around the equator (i.e. $\sin\theta\sim 1$). In
+      this case we identify the $x$ coordinate with a latitude shift
+      ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
+      shifts ($\delta\varphi=\delta y$) \todo{make a plot of
+        this?}. \as{no, but isn't it other way round? Latitute is $y$
+        in my brain? :)} \dam{it isn't, have a look here: \url{https://healpix.jpl.nasa.gov/html/intronode12.htm}.} Using this dictionary, the differential operator $\eth$ now takes the form:
       \begin{equation}
@@
 
-      In analogy with our curved-sky nomenclature, let us now write $a$ as a vector such that in real space ${\bf a}({\bf x})\equiv({\rm Re}(a),{\rm Im}(a))$, and in Fourier space ${\bf a}_{\bf l}\equiv(\,_sE_{\bf l},\,_sB_{\bf l})$.
+      In analogy with our full-sky nomenclature, let us now write $a$ as a vector such that in real space ${\bf a}({\bf x})\equiv({\rm Re}(a),{\rm Im}(a))$, and in Fourier space ${\bf a}_{\bf l}\equiv(\,_sE_{\bf l},\,_sB_{\bf l})$.
 We can rewrite the Eq.
@@
     \subsection{Discrete description and DFTs}\label{app:flat.dft}
-      As we discuss in \ref{sssec:maths.flat.pcl101}, it is more convenient to describe power spectrum estimation methods in a discretized flat sky.
+      As we discuss in \todo{where}, it is more convenient to describe power spectrum estimation methods in a discretized flat sky.
 Let the patch of the sky under inspection be contained by a rectangle of sides $L_x$ and $L_y$ (in units of radians), and let us discretize this rectangle by dividing it into an $N_x\times N_y$ grid with pixels of area $\Delta{\bf x}^2\equiv\Delta x\Delta y=(L_x/N_x)(L_y/N_y)$.
@@
 
-  \section{Curved-sky validation for different pixel resolutions}\label{app:full_sky_nside}
+  \section{Full-sky validation for different pixel resolutions}\label{app:full_sky_nside}
     We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of the HEALPix resolution parameter $N_{\rm side}$.
-In this appendix we show that we recover the correct power spectra using the inputs from our curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).
+In this appendix we show that we recover the correct power spectra using the inputs from our full-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).
 In particular we tested $N_{\rm side}$ ranging from 8 to 8192.
@@
       \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
-      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
+      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the full-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
       \label{fig:validation.fullsky.lss.4096}
@@
 
-\section{Performance and scalability}
-\label{sec:scalability}
+\end{document}
+
+ \subsection{Performance and scalability}
+ \label{ssec:scalability}
      One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
-In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the curved-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
+In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the full-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
 We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
@@
      \includegraphics[width=0.9\columnwidth]{time_scaling}
-     \caption{Computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
+     \caption{Computing time for the auto-power spectra of full-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
 We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}
@@
      \end{figure}
-
-\end{document}
 

### 2018-08-31 doc/Paper/main.tex

--- 2018-08-30 doc/Paper/main.tex
@@
 Nevertheless the rotational invariance requires that elements of the $2\times2$ matrix ${\sf C}_\ell$ be real.
+
+      \subsubsection{Pixelization and signal band-limits}
+
+      SHT transforms as written in the Eq.~\ref{eq:sht1} is, strictly
+      speaking, defined on a continuous field on the sky and results
+      in an infinite number of spherical harmonic coefficients. In
+      practice, we have to always deal with a finite number of pixels
+      at which the field is sampled and consequently use a finite
+      number of spherical coefficients. If the sampling of the
+      underlying signal is too sparse, one will get aliasing of higher
+      frequency modes. In flat spaces, Nyquist theorem provides a
+      clear prescription for uniform sampling on a grid. Since there
+      is no such natural sampling on a sphere, the situation is more
+      complex and depends on the pixelization scheme
+      \cite{Driscoll:1994:CFT:184069.184073,
+        DBLP:journals/corr/abs-1110-6297}. Nevertheless, clear
+      heuristic prescriptions exist. Most importantly, the pixel separation needs
+      to be smaller than the wavelength of the highest frequency
+      present in the underlying signal. As a natural consequence, the
+      smoothing of the map must always precede any downgrade in
+      resolution (or equivalently downsampling) if it is necessary to
+      do so for numerical expediency.  Similarly, any mask should be
+      applied at the highest possible resolution and the resulting
+      map smoothed before being down-graded.
+
+      For the same reasons, even if the signal is correctly sampled in
+      a band-limited fashion, one must always estimate power over the
+      entire power band, even if one is not interested in certain
+      region of $\ell$ space. Otherwise, the power spectrum estimator
+      will try to ``explain'' the variance associated with $\ell$
+      modes not considered by artificially inflating the power in the
+      estimated modes. For {\tt HEALPix} \citep{2005ApJ...622..759G}
+      pixelization commonly used in cosmology and also in this paper,
+      \cite{2013MNRAS.435.1857L} suggest the following
+      prescription. If one wants to measure power up to $\ell_{\rm
+        max}$, one should employ pixelisation of
+      \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up
+      to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm
+        max}$ after correcting for pixel-response suppression and
+      accounting for their contribution to the covariance matrix of
+      measurements.
+
+      In this paper we do not consider these issues any further as \nmt
+      is a generic tool and it is up to the user to ensure band-limit
+      and pixelization constraints are properly accounted for.
 

### 2018-08-31 doc/Paper/main.tex
merged Anze's band-limit section, fixed some TODOs

--- 2018-08-30 doc/Paper/main.tex
@@
 
-    \subsection{Spherical harmonic transform and pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.pcl101}
+    \subsection{Spherical harmonic transforms and pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.pcl101}
       Let ${\bf a}(\nv)$ be a spin-$s_a$ quantity defined on the sphere, where $\nv$ is the unit vector pointing in a particular direction $(\theta,\varphi)$.
@@
 The matrices ${\sf Y}^s_{\ell m}$ are defined in terms of the spin-weighed spherical harmonics, and are described in detail in Appendix \ref{app:shts}.
+Some technical details regarding band-limits and the practicalities of SHTs in a discretized sphere are discussed in \ref{app:shts.blim}.
 To simplify the notation, we will also often abbreviate the pair $(\ell,m)$ as ${\bf l}$.
@@
         where $\mI$ is the unit matrix, and $P^{vw}_\ell\equiv(2\ell+1)\,{\rm PCL}_\ell(v,w)$.
-
-      \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
-        Given the loss of information implicit in masking the originally curved-sky field, it is in general not possible to invert the mode-coupling matrix in Eq.
-\ref{eq:modecoup} directly.
-One solution to this problem is to convolve the theory prediction with the mode-coupling matrix at the likelihood level.
-A more usual approach is instead to bin the coupled pseudo-$C_\ell$ into bandpowers.
-A bandpower $q$ is defined as a set of $N_q$ multipoles $\vec{\ell}_q\equiv(\ell_q^1,...,\ell_q^{N_q})$ and a set of weights defined at those multipoles $w^\ell_q$, and normalized such that $\sum_{\ell\in\vec{\ell}_q}w_q^\ell=1$.
-The $q$-th bandpower for the coupled pseudo-$C_\ell$ is then defined as
-        \begin{align}\label{eq:bpws}
-          {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
-          \left\langle{\rm vec}\left[\tilde{\sf B}^{ab}_q\right]\right\rangle&=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
-        \end{align}
-        where in the second line we have taken the expectation value of first one.
-One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function).
-Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
         \begin{figure*}
@@
         \end{figure*}
+
+      \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
+        Given the loss of information implicit in masking the originally curved-sky field, it is in general not possible to invert the mode-coupling matrix in Eq.
+\ref{eq:modecoup} directly.
+One solution to this problem is to convolve the theory prediction with the mode-coupling matrix at the likelihood level.
+A more usual approach is instead to bin the coupled pseudo-$C_\ell$ into bandpowers.
+A bandpower $q$ is defined as a set of $N_q$ multipoles $\vec{\ell}_q\equiv(\ell_q^1,...,\ell_q^{N_q})$ and a set of weights defined at those multipoles $w^\ell_q$, and normalized such that $\sum_{\ell\in\vec{\ell}_q}w_q^\ell=1$.
+The $q$-th bandpower for the coupled pseudo-$C_\ell$ is then defined as
+        \begin{align}\label{eq:bpws}
+          {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
+          \left\langle{\rm vec}\left[\tilde{\sf B}^{ab}_q\right]\right\rangle&=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
+        \end{align}
+        where in the second line we have taken the expectation value of first one.
+One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function).
+Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
         \begin{equation}\label{eq:decoup}
@@
         Our second example corresponds to the measurement of $B$-modes in the polarized CMB emission, closely associated with the presence of primordial gravitational waves.
-This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citep{2016PhRvL.116c1302B}, Advanced ACTPol~\citep{2016SPIE.9910E..14D,2017JCAP...06..031L} or the Simons Array~\citep{2016JLTP..184..805S,2017ApJ...848..121P}, as well as next-generation facilities such as the Simons Observatory\todo{cite if out} or CMB Stage-4~\citep{2016arXiv161002743A}.
+This is one of the main science cases targeted by current ground-based Stage-3 experiments such as BICEP/Keck~\citep{2016PhRvL.116c1302B}, Advanced ACTPol~\citep{2016SPIE.9910E..14D,2017JCAP...06..031L} or the Simons Array~\citep{2016JLTP..184..805S,2017ApJ...848..121P}, as well as next-generation facilities such as the Simons Observatory~\citep{2018arXiv180807445T} or CMB Stage-4~\citep{2016arXiv161002743A}.
 Potential future space missions, such as LiteBIRD~\citep{2018JLTP..tmp..124S} will be able to make low-resolution, curved-sky maps of the polarized CMB to put constraints on the primordial $B$-mode signal from its reionization bump, at $\ell<30$, and will likely make use of pixel-based likelihoods and optimal quadratic methods.
@@
 This resolution corresponds to what could be achieved at 150 GHz with a small aperture experiment (0.4m, S.A.) and a large aperture experiment (6m, L.A.) in either case.
-The choice of $\ell_{\rm knee}$ described above for the noise curves is also consistent with the atmospheric noise levels achievable with those types of telescopes \todo{cite Simons Observatory paper if out}.
+The choice of $\ell_{\rm knee}$ described above for the noise curves is also consistent with the atmospheric noise levels achievable with those types of telescopes \citep{2018arXiv180807445T}.
 As shown in Fig.
@@
 
+      \subsection{Pixelization and signal band-limits}\label{app:shts.blim}
+        SHT transforms as written in the Eq.~\ref{eq:sht1} are, strictly speaking, defined on a continuous field on the sky and results in an infinite number of spherical harmonic coefficients.
+In practice, we have to always deal with discretized sky maps with a finite number of pixels at which the field is sampled, and consequently use a finite number of spherical coefficients.
+If the sampling of the underlying signal is too sparse, we will get aliasing of higher frequency modes.
+In flat spaces, Nyquist's theorem provides a clear prescription for uniform sampling on a grid.
+Since there is no such natural sampling on a sphere, the situation is more complex and depends on the pixelization scheme used \cite{Driscoll:1994:CFT:184069.184073,DBLP:journals/corr/abs-1110-6297}.
+Nevertheless, clear heuristic prescriptions exist.
+Most importantly, the typical pixel separation needs to be smaller than the wavelength of the highest frequency mode present in the underlying signal.
+As a natural consequence, the smoothing of the map must always precede any downgrade in resolution (or equivalently downsampling) if it is necessary to do so for numerical expediency.
+Similarly, any mask should be applied at the highest possible resolution and the resulting map smoothed before being down-graded.
+
+        For the same reasons, even if the signal is correctly sampled in a band-limited sense, one must always estimate power over the entire power band, even if one is not interested in certain region of $\ell$ space.
+Otherwise, the power spectrum estimator will try to ``explain'' the variance associated with $\ell$ modes not considered by artificially inflating the power in the estimated modes.
+For the {\tt HEALPix} pixelization scheme \citep{2005ApJ...622..759G}, commonly used in cosmology and also in this paper, \cite{2013MNRAS.435.1857L} suggest the following prescription.
+If one wants to measure power up to $\ell_{\rm max}$, one should employ pixelisation of \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm max}$ after correcting for pixel-response suppression and  accounting for their contribution to the covariance matrix of measurements.
+
+        In this paper we do not consider these issues any further as \nmt~is a generic tool and it is up to the user to ensure band-limit and pixelization constraints are properly accounted for.
+
   \section{Linear contamination from non-linear contamination} \label{app:lincont}
@@
 In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
-      shifts ($\delta\varphi=\delta y$) the differential operator $\eth$ now takes the form: %\todo{make a plot of this?}.
-\as{no, but isn't it other way round?
-Latitute is $y$ in my brain?
-:)} \dam{it isn't, have a look here \url{https://healpix.jpl.nasa.gov/html/intronode12.htm}.} Using this dictionary,
+      shifts ($\delta\varphi=\delta y$) the differential operator $\eth$ now takes the form:
       \begin{equation}

### 2018-08-31 sandbox_validation/timing/run_all.sh
full timing

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+for nside in 64 256 1024 4096
+do
+    for task in field mcm pure deproj pure_deproj
+    do
+	for ncores in 1 2 4 8 16
+	do
+	    if [ ${nside} -gt 256 ]
+	    then
+		ncomp=10
+	    else
+		ncomp=100
+	    fi
+	    addqueue -q cmb -s -n 1x${ncores} -m 4 ./timings -nside ${nside} -ncomp ${ncomp} -do_${task} -out ${task}_ns${nside}_nc${ncores}.txt
+	done
+    done
+done

### 2018-08-31 sandbox_validation/timing/src/main.c

@@ -8,6 +8,8 @@
 #include "timing_field.h"
 #include "timing_mcm.h"
 #include "timing_pure.h"
+#include "timing_deproj.h"
+#include "timing_pure_deproj.h"
 
 void time_stuff(timing_st *tim,int ncomp,FILE *f)
 {
@@ -16,20 +18,41 @@ void time_stuff(timing_st *tim,int ncomp,FILE *f)
   timing_st_free(tim);
 }
 
+void help_and_exit()
+{
+  fprintf(stderr,"Usage: ./timings <options>\n");
+  fprintf(stderr,"Available options:\n");
+  fprintf(stderr,"  -nside INT\n");
+  fprintf(stderr,"  -ncomp INT\n");
+  fprintf(stderr,"  -out STRING\n");
+  fprintf(stderr,"  -do_field\n");
+  fprintf(stderr,"  -do_mcm\n");
+  fprintf(stderr,"  -do_pure\n");
+  fprintf(stderr,"  -do_deproj\n");
+  fprintf(stderr,"  -do_pure_deproj\n");
+  fprintf(stderr,"  -h\n\n");
+  exit(1);
+}
+
 int main(int argc,char **argv)
 {
   FILE *fout;
   char fname_out[256]="none";
   int ncomp=10,nside=256;
-  int do_field=1,do_mcm=1,do_pure=1;
+  int do_field=0,do_mcm=0,do_pure=0,do_deproj=0,do_pure_deproj=0;
   char **i;
   for(i=argv+1;*i;i++) {
     if(!strcmp(*i,"-ncomp")) ncomp=atoi(*++i);
     else if(!strcmp(*i,"-nside")) nside=atoi(*++i);
     else if(!strcmp(*i,"-out")) sprintf(fname_out,"%s",*++i);
-    else if(!strcmp(*i,"-dont_field")) do_field=0;
-    else if(!strcmp(*i,"-dont_mcm")) do_mcm=0;
-    else if(!strcmp(*i,"-dont_pure")) do_pure=0;
+    else if(!strcmp(*i,"-do_field")) do_field=1;
+    else if(!strcmp(*i,"-do_mcm")) do_mcm=1;
+    else if(!strcmp(*i,"-do_pure")) do_pure=1;
+    else if(!strcmp(*i,"-do_deproj")) do_deproj=1;
+    else if(!strcmp(*i,"-do_pure_deproj")) do_pure_deproj=1;
[…]

### 2018-08-31 sandbox_validation/timing/src/timing_pure.c

@@ -21,7 +21,7 @@ void *setup_pure(int nside,int spin1,int spin2)
 }
 
 //Evaluator for pure timing
-// - Generates and frees an nmt_pure object
+// - Generates and frees an nmt_field object
 void func_pure(void *data)
 {
   pure_data *d=(pure_data *)data;

### 2018-08-31 sandbox_validation/timing/src/timing_pure.h

@@ -13,7 +13,7 @@ typedef struct {
 void *setup_pure(int nside,int spin1,int spin2);
 
 //Evaluator for pure timing
-// - Generates and frees an nmt_pure object
+// - Generates and frees an nmt_field object
 void func_pure(void *data);
 
 //Destructor for pure timing

### 2018-08-31 sandbox_validation/timing/src/timing_utils.c

@@ -53,7 +53,7 @@ double **init_maps(int nside,int nmaps)
 }
 
 //timing_st constructor
-timing_st *timing_st_init(char *name,int nside,int is_1point,
+timing_st *timing_st_init(char *name,int nside,int nruns,
 			  void *(*setup_time)(int,int,int),
 			  void (*func_time)(void *),
 			  void (*free_time)(void *))
@@ -61,7 +61,7 @@ timing_st *timing_st_init(char *name,int nside,int is_1point,
   timing_st *tim=my_malloc(sizeof(timing_st));
   sprintf(tim->name,"%s",name);
   tim->nside=nside;
-  tim->is_1point=is_1point;
+  tim->nruns=nruns;
   tim->setup=setup_time;
   tim->func=func_time;
   tim->free=free_time;
@@ -77,36 +77,41 @@ void timing_st_free(timing_st *tim)
   free(tim);
 }
 
+//timing_st individual timer
+static double timing_st_time_single(timing_st *tim,int ncomp,int spin1,int spin2)
+{
+  int ii;
+  void *data=tim->setup(tim->nside,spin1,spin2);
+  timer(0);
+  for(ii=0;ii<ncomp;ii++) {
+    tim->func(data);
+  }
+  timer(1);
+  tim->free(data);
+  return 1000.*(relend-relbeg)/ncomp;
+}
+
 //timing_st timer
 void timing_st_time(timing_st *tim,int ncomp)
 {
-  int pol2max=1;
-  if(tim->is_1point)
-    pol2max=0;
-
-  int pol2;
-  int ind_write=0;
-  for(pol2=0;pol2<=pol2max;pol2++) {
-    int pol1;
-    for(pol1=pol2;pol1<=1;pol1++) {
-      int ii;
-      void *data=tim->setup(tim->nside,2*pol1,2*pol2);
-      timer(0);
-      for(ii=0;ii<ncomp;ii++) {
-	tim->func(data);
-      }
-      timer(1);
-      tim->free(data);
-      tim->times[ind_write]=1000.*(relend-relbeg)/ncomp;
-      ind_write++;
-    }
+  if(tim->nruns==1) { //One field, just spin-2
+    tim->times[0]=timing_st_time_single(tim,ncomp,2,0);
+  }
+  else if(tim->nruns==2) { //One field, spin-0 and spin-2
+    tim->times[0]=timing_st_time_single(tim,ncomp,0,0);
[…]

### 2018-08-31 sandbox_validation/timing/src/timing_utils.h

@@ -1,6 +1,8 @@
 #ifndef _TIMING_UTILS_
 #define _TIMING_UTILS_
 
+#define NTEMP 5
+
 extern double relbeg,relend;
 void timer(int i);
 
@@ -16,7 +18,7 @@ double timing(int ncomp,int nside,int spin1,int spin2,
 typedef struct {
   char name[16];
   int nside;
-  int is_1point;
+  int nruns;
   void *(*setup)(int,int,int);
   void (*func)(void *);
   void (*free)(void *);
@@ -24,7 +26,7 @@ typedef struct {
 } timing_st;
 
 //timing_st constructor
-timing_st *timing_st_init(char *name,int nside,int is_1point,
+timing_st *timing_st_init(char *name,int nside,int nruns,
 			  void *(*setup_time)(int,int,int),
 			  void (*func_time)(void *),
 			  void (*free_time)(void *));

### 2018-08-31 sandbox_validation/timing/plot_timing.py
adding missing files

@@ -1,0 +1,63 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+NSIDES=np.array([64,256,1024])
+NCORES=np.array([16]) #[1,2,4,8,16])
+
+def read_data(prefix) :
+    n_ns=len(NSIDES)
+    n_nc=len(NCORES)
+    
+    data=np.zeros([n_nc,n_ns,3])
+
+    for i_c,nc in enumerate(NCORES) :
+        for i_n,nn in enumerate(NSIDES) :
+            print prefix+"_ns%d_nc%d.txt"%(nn,nc)
+            d=np.genfromtxt(prefix+"_ns%d_nc%d.txt"%(nn,nc))[3:]
+            print d
+            data[i_c,i_n,:]=d
+    return data
+
+t_data={}
+for k in ['field','mcm','deproj','pure','pure_deproj'] :
+    t_data[k]=read_data(k)
+    plt.plot(NSIDES,t_data[k][0,:,0],label=k)
+    plt.plot(NSIDES,t_data[k][0,1,0]*((NSIDES+0.)/NSIDES[1])**3,'k--')
+plt.loglog()
+plt.legend(loc='upper left')
+plt.show()
+
+absc=np.array(['Field, $s=0$','Field, $s=2$',
+               'Deproj., $s=0$','Deproj., $s=2$',
+               'Purification','Deproj. $+$ purif.',
+               'MCM, 0-0','MCM, 0-2','MCM, 0-3'])
+data_1024_plot=np.array([t_data['field'][0,2,0], #Field s0
+                         t_data['field'][0,2,1], #Field s2
+                         t_data['deproj'][0,2,0], #Deproj s0
+                         t_data['deproj'][0,2,1], #Deproj s2
+                         t_data['pure'][0,2,0], #Pure s2
+                         t_data['pure_deproj'][0,2,0], #Pure-deproj s2
+                         t_data['mcm'][0,2,0], #MCM-00
+                         t_data['mcm'][0,2,1], #MCM-02
+                         t_data['mcm'][0,2,2] #MCM-22
+                         ])
+data_256_plot=np.array([t_data['field'][0,1,0], #Field s0
+                        t_data['field'][0,1,1], #Field s2
+                        t_data['deproj'][0,1,0], #Deproj s0
+                        t_data['deproj'][0,1,1], #Deproj s2
+                        t_data['pure'][0,1,0], #Pure s2
[…]

### 2018-08-31 sandbox_validation/timing/src/timing_deproj.c

@@ -1,0 +1,62 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <namaster.h>
+#include <omp.h>
+#include "utils.h"
+#include "timing_utils.h"
+#include "timing_deproj.h"
+
+//Setup function for deproj timing
+// - Initializes a set of maps and a mask
+void *setup_deproj(int nside,int spin1,int spin2)
+{
+  int ii;
+
+  deproj_data *data=my_malloc(sizeof(deproj_data));
+  data->nside=nside;
+  if(spin1) {
+    data->nmaps=2;
+    data->pol=1;
+  }
+  else {
+    data->nmaps=1;
+    data->pol=0;
+  }
+
+  data->maps=init_maps(nside,data->nmaps);
+  data->temp=my_malloc(NTEMP*sizeof(double **));
+  for(ii=0;ii<NTEMP;ii++)
+    data->temp[ii]=init_maps(nside,data->nmaps);
+  data->mask=init_mask(nside);
+
+  return data;
+}
+
+//Evaluator for deproj timing
+// - Generates and frees an nmt_deproj object
+void func_deproj(void *data)
+{
+  deproj_data *d=(deproj_data *)data;
+  nmt_field *f=nmt_field_alloc_sph(d->nside,d->mask,d->pol,d->maps,NTEMP,d->temp,NULL,0,0,3,1E-5);
+  nmt_field_free(f);
+}
+
+//Destructor for deproj timing
+void free_deproj(void *data)
+{
+  deproj_data *d=(deproj_data *)data;
+  int ii;
+  for(ii=0;ii<NTEMP;ii++) {
+    int jj;
+    for(jj=0;jj<d->nmaps;jj++)
+      free(d->temp[ii][jj]);
+    free(d->temp[ii]);
+  }
+  free(d->temp);
+  for(ii=0;ii<d->nmaps;ii++)
+    free(d->maps[ii]);
+  free(d->maps);
+  free(d->mask);
+  free(data);
+}

### 2018-08-31 sandbox_validation/timing/src/timing_deproj.h

@@ -1,0 +1,25 @@
+#ifndef _TIMING_DEPROJ_
+#define _TIMING_DEPROJ_
+
+//Data type for deproj timing
+typedef struct {
+  int nmaps;
+  int nside;
+  int pol;
+  double **maps;
+  double ***temp;
+  double *mask;
+} deproj_data;
+
+//Setup function for deproj timing
+// - Initializes a set of maps and a mask
+void *setup_deproj(int nside,int spin1,int spin2);
+
+//Evaluator for deproj timing
+// - Generates and frees an nmt_field object with templates
+void func_deproj(void *data);
+
+//Destructor for deproj timing
+void free_deproj(void *data);
+
+#endif //_TIMING_DEPROJ_

### 2018-08-31 sandbox_validation/timing/src/timing_pure_deproj.c

@@ -1,0 +1,53 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <namaster.h>
+#include <omp.h>
+#include "utils.h"
+#include "timing_utils.h"
+#include "timing_pure_deproj.h"
+
+//Setup function for pure_deproj timing
+// - Initializes a set of maps and a mask
+void *setup_pure_deproj(int nside,int spin1,int spin2)
+{
+  int ii;
+  pure_deproj_data *data=my_malloc(sizeof(pure_deproj_data));
+  data->nside=nside;
+
+  data->maps=init_maps(nside,2);
+  data->mask=init_mask(nside);
+  data->temp=my_malloc(NTEMP*sizeof(double **));
+  for(ii=0;ii<NTEMP;ii++)
+    data->temp[ii]=init_maps(nside,2);
+
+  return data;
+}
+
+//Evaluator for pure_deproj timing
+// - Generates and frees an nmt_pure_deproj object
+void func_pure_deproj(void *data)
+{
+  pure_deproj_data *d=(pure_deproj_data *)data;
+  nmt_field *f=nmt_field_alloc_sph(d->nside,d->mask,1,d->maps,NTEMP,d->temp,NULL,0,1,3,1E-5);
+  nmt_field_free(f);
+}
+
+//Destructor for pure_deproj timing
+void free_pure_deproj(void *data)
+{
+  pure_deproj_data *d=(pure_deproj_data *)data;
+  int ii;
+  for(ii=0;ii<NTEMP;ii++) {
+    int jj;
+    for(jj=0;jj<2;jj++)
+      free(d->temp[ii][jj]);
+    free(d->temp[ii]);
+  }
+  free(d->temp);
+  for(ii=0;ii<2;ii++)
+    free(d->maps[ii]);
+  free(d->maps);
+  free(d->mask);
+  free(data);
+}

### 2018-08-31 doc/Paper/main.tex
adding to timing

@@
 Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation.
-More details about timing and scalability of \nmt can be found in Appendix \ref{sec:scalability}.
+More details about timing and scalability of \nmt~can be found in Appendix \ref{app:scalability}.
 
@@
         \includegraphics[width=0.99\columnwidth]{val_cmb_bpw}
-        \caption{{\sl Left:} average over all simulations in our
-          curved-sky CMB validation suite of the three spin-2 power
-          spectra ($EE$, $EB$ and $BB$, solid lines) using mode
-          deprojection and E/B mode purification.
+        \caption{{\sl Left:} average over all simulations in our curved-sky CMB validation suite of the three spin-2 power spectra ($EE$, $EB$ and $BB$, solid lines) using mode deprojection and $B$-mode purification.
 The analytical prediction for the input signal power spectra are shown as dashed lines, and line mostly under the solid lines for the non-zero cases ($EE$ and $BB$).
@@
     \end{equation}
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
+      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
+      \label{fig:validation.fullsky.lss.4096}
+    \end{figure}
 
@@
 However, we generate 100 simulations with $N_{\rm side}=4096$ and show the results in Figure~\ref{fig:validation.fullsky.lss.4096}.
-We can see that the power spectra are recovered without any noticeable biases in the expected range of validity $\ell \lesssim 2 N_{\rm side}$.
-
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
-      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
-      \label{fig:validation.fullsky.lss.4096}
-    \end{figure}
-
-\bsp
-
-\section{Performance and scalability}
-\label{sec:scalability}
-     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated.
+We can see that the power spectra are recovered without any noticeable bias in the expected range of validity $\ell \lesssim 2 N_{\rm side}$.
+
+ \section{Performance and scalability}\label{app:scalability}
+     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{\rm max}^{3})$, where $\ell_{\rm max}$ is the maximum multipole at which the power-spectra are calculated.
 In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the curved-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants.
-We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$.
+We change the resolution of the \texttt{HEALPix} maps by varying the $N_{\rm side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{\rm max}=3N_{\rm side}$.
 We use one interactive Cori-Haswell node at the National Energy Research Scientific Computing Center (NERSC).
@@
 We executed the code using 32 \texttt{OpenMP} threads.
-In Figure~\ref{fig:scaling_lmax} we can see the results for this test.
-We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 20).
-In all cases when we compare with the reference line $\propto \ell_{max}^{3}$ we can see that the code scales as expected.
-We fixed the proportionality constant for the reference line using the timing of the auto-power of the spin-0 field at $\ell_{max}=96$.
-We can also see that for very small $N_{side}$ data handling time starts to be comparable to the pure power-spectra computation time, increasing the expected execution time.
+In the top panel of Figure~\ref{fig:scaling_lmax} we can see the results for this test.
+We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 02).
+In all cases we find that the code scales as expected (the solid red line shows the expected $\propto\ell_{\rm max}^3$ behaviour).
+We can also see that, for very small $N_{side}$, data handling time starts to be comparable to the power-spectrum computation time, increasing the execution time with respect to the expectation.
+
+     Not all parts of the power spectrum estimation take the same amount of time.
+E.g.
+even though the scaling with $\ell_{\rm max}$ is the same, the computation of the mode-coupling matrix typically takes longer than a performing single SHT.
+Computation times can be significantly increased if deprojection over a large number of maps is required or if $E/B$ purification is required.
+The bottom panel of Figure \ref{fig:scaling_lmax} shows the execution times associated to the most relevant types of operations:
 
@@
 
+\bsp
 \end{document}

### 2018-08-31 sandbox_validation/timing/plot_timing.py
generated timing plot

@@ -1,8 +1,12 @@
+from matplotlib import rc
+import matplotlib
 import numpy as np
 import matplotlib.pyplot as plt
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
 
 NSIDES=np.array([64,256,1024])
-NCORES=np.array([16]) #[1,2,4,8,16])
+NCORES=np.array([16])
 
 def read_data(prefix) :
     n_ns=len(NSIDES)
@@ -12,25 +16,18 @@ def read_data(prefix) :
 
     for i_c,nc in enumerate(NCORES) :
         for i_n,nn in enumerate(NSIDES) :
-            print prefix+"_ns%d_nc%d.txt"%(nn,nc)
-            d=np.genfromtxt(prefix+"_ns%d_nc%d.txt"%(nn,nc))[3:]
-            print d
+            d=np.genfromtxt('output/'+prefix+"_ns%d_nc%d.txt"%(nn,nc))[3:]
             data[i_c,i_n,:]=d
     return data
 
 t_data={}
 for k in ['field','mcm','deproj','pure','pure_deproj'] :
     t_data[k]=read_data(k)
-    plt.plot(NSIDES,t_data[k][0,:,0],label=k)
-    plt.plot(NSIDES,t_data[k][0,1,0]*((NSIDES+0.)/NSIDES[1])**3,'k--')
-plt.loglog()
-plt.legend(loc='upper left')
-plt.show()
 
 absc=np.array(['Field, $s=0$','Field, $s=2$',
                'Deproj., $s=0$','Deproj., $s=2$',
-               'Purification','Deproj. $+$ purif.',
-               'MCM, 0-0','MCM, 0-2','MCM, 0-3'])
+               'Purification','Deproj. + purif.',
+               'MCM, $0$-$0$','MCM, $0$-$2$','MCM, $2$-$2$'])
 data_1024_plot=np.array([t_data['field'][0,2,0], #Field s0
                          t_data['field'][0,2,1], #Field s2
                          t_data['deproj'][0,2,0], #Deproj s0
@@ -52,12 +49,23 @@ data_256_plot=np.array([t_data['field'][0,1,0], #Field s0
                         t_data['mcm'][0,1,2] #MCM-22
                         ])
 
-plt.figure()
+def tickfs(ax,x=True,y=True,fs=12) :
[…]

### 2018-08-31 sandbox_validation/timing/run_all.sh

@@ -1,5 +1,7 @@
 #!/bin/bash
 
+mkdir -p output
+
 for nside in 64 256 1024 4096
 do
     for task in field mcm pure deproj pure_deproj
@@ -12,7 +14,7 @@ do
 	    else
 		ncomp=100
 	    fi
-	    addqueue -q cmb -s -n 1x${ncores} -m 4 ./timings -nside ${nside} -ncomp ${ncomp} -do_${task} -out ${task}_ns${nside}_nc${ncores}.txt
+	    addqueue -q cmb -s -n 1x${ncores} -m 4 ./timings -nside ${nside} -ncomp ${ncomp} -do_${task} -out output/${task}_ns${nside}_nc${ncores}.txt
 	done
     done
 done

### 2018-08-31 doc/Paper/main.tex
added timing plot

@@
      \centering
-     \includegraphics[width=0.9\columnwidth]{time_scaling}
+     \includegraphics[width=0.99\columnwidth]{time_scaling}
+     \includegraphics[width=0.99\columnwidth]{timing}
      \caption{Computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
@@
 \end{document}
-
-    In addition, we also check the parallel performance of \nmt.
-We proceed again, as in the previous case, to compute the auto- and cross-power spectra of spin-0 and spin-2 fields with no contaminants for maps with $N_{side}=256$ and $N_{side}=1024$ and we vary the number of \texttt{OpenMP} threads from 1 to 32 in powers of 2.
-The results can be seen in Figure~\ref{fig:scaling_nthreads}.
-We can see that the operations involving spin-0 fields seem to be more affected by the number of threads used, however, the spin-2 auto-power spectra computing time seems to be very stable for any number of threads larger than 2.
-\todo{Discuss this further?
-Take the second part out?
-Ideas?}
-
-    \begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{time_scaling_threads}
-     \caption{Computing time for auto-power spectra of spin-0 (red lines) and spin-2 (cyan lines) contaminant-free maps, and their cross-power spectra (blue lines) as a function of the number of \texttt{OpenMP} threads used for the computation.
-We check maps with $N_{side}=256$ (solid lines) and $N_{side}=1024$ (broken lines).}
-     \label{fig:scaling_nthreads}
-     \end{figure}

### 2018-08-31 sandbox_validation/timing/plot_timing.py
small fix

@@ -65,7 +65,6 @@ ax.set_xlabel('Task',fontsize=15)
 ax.set_ylabel('Time (s)',fontsize=15)
 ax.set_xticks(np.arange(len(absc)))
 ax.set_xticklabels(absc,rotation=20)
-tickfs(ax,x=False)
-tickfs(ax,y=False,fs=12)
+tickfs(ax)
 plt.savefig("../plots_paper/timing.pdf",bbox_inches='tight')
 plt.show()

### 2018-08-31 sandbox_validation/timing/plot_timing.py
small fix

@@ -57,7 +57,7 @@ def tickfs(ax,x=True,y=True,fs=12) :
         for tick in ax.yaxis.get_major_ticks():
             tick.label.set_fontsize(fs)
 
-plt.figure(figsize=(10,5))
+plt.figure(figsize=(8,4))
 ax=plt.gca()
 ax.bar(np.arange(len(absc)),data_1024_plot/1000.,0.75)
 ax.text(0.03,0.85,'$N_{\\rm side}=1024$\n 5 contaminant templates.',transform=ax.transAxes,fontsize=15)

### 2018-08-31 doc/Paper/main.tex
added last timing figure

@@
 Computation times can be significantly increased if deprojection over a large number of maps is required or if $E/B$ purification is required.
-The bottom panel of Figure \ref{fig:scaling_lmax} shows the execution times associated to the most relevant types of operations:
+The bottom panel of Figure \ref{fig:scaling_lmax} shows the execution times associated to the most relevant types of operations.
+These are (using the labels provided in the figure):
+     \begin{itemize}
+       \item {\bf Field $s=0$ and $s=2$}: generating a {\tt NmtField} object. This entails taking the SHT of a given spin-0 or 2 map as well as that of its associated weights map.
+       \item {\bf Deproj.}: generating a {\tt NmtField} object with contaminant deprojection.
+This is the same as the above in addition to computing the SHT of a set of contaminant templates.
+In this case we include 5 templates, and the additional computation time will increase linearly with the number of templates.
+       \item {\bf Purification}: generating a {\tt NmtField} object with $B$-mode purification.
+This is the same as ``Field $s=2$'' in addition to computing the pure $B$-modes.
+This entails estimating a more accurate SHT of the apodized mask as well as a combination of harmonic-space and real-space operations described in Section \ref{ssec:maths.pureb}.
+       \item {\bf Deproj.
+$+$ purif.}: generating a {\tt NmtField} object with both $B$-mode purification and contaminant deprojection.
+This entails all of the operations implied by ``Deproj.'' and ``Purification'' in addition to the computation of the $B$-mode-purified templates (see Eqs.
+\ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}).
+       \item {\bf MCM $s_1$-$s_2$}: generating the pseudo-$C_\ell$ mode-coupling matrix for a spin-$s_1$ field and a spin-$s_2$ field.
+     \end{itemize}
+     The times reported in this figure correspond to the time required to perform a single instance of the corresponding operation for fields with resolution $N_{\rm side}=1024$ using 16 \texttt{OpenMP} threads.
+Note that, even though the computation of the mode-coupling matrix dominates the total execution time in the simplest scenarios, in the most complex situation, involving both purification and deprojection, setting up a single field can take significantly longer.
 
@@
        \includegraphics[width=0.99\columnwidth]{timing}
-     \caption{Computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{max}$.
-We include the expected scaling for the pseudo-$C_{\ell}$ estimator (solid line) and force it to go through the point for the spin-0 auto-power at $\ell=96$ showing remarkable agreement.}
-     \label{fig:scaling_lmax}
+       \caption{{\sl Top panel}: computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{\rm max}$.
+We include the expected scaling $\propto\ell_{\rm max}$ for the pseudo-$C_{\ell}$ estimator (solid red line) for comparison.
+{\sl Bottom panel:} execution time for different tasks associated with the pseudo-$C_\ell$ estimator.
+We show times for fields with resolution $N_{\rm side}=1024$, which can be extrapolated to other resolutions using a $\propto N_{\rm side}^3$ scaling, as described above.
+The description of the different tasks shown is given in the main text.}\label{fig:scaling_lmax}
      \end{figure}

### 2018-08-31 doc/Paper/main.tex
merged, added CCL ref

@@
 }
-%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
- %\hypersetup{draft}
 \newcommand{\nv}{\hat{\boldsymbol{\theta}}}
@@
 For $\gamma$, we assume no contribution from intrinsic alignments.
-The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library \todo{cite if paper is out} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}.
+The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library\footnote{\url{https://github.com/LSSTDESC/CCL/pulls}} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}.
 These power spectra are shown in Fig.

### 2018-08-31 doc/Paper/main.tex

@@
 For $\gamma$, we assume no contribution from intrinsic alignments.
-The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library\footnote{\url{https://github.com/LSSTDESC/CCL/pulls}} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}.
+The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library\footnote{\url{https://github.com/LSSTDESC/CCL}} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}.
 These power spectra are shown in Fig.
@@
 Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the curved-sky and flat-sky cases respectively.
-          \item {\bf Mask:} for the curved-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}.
+          \item {\bf Mask:} for the curved-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citep{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}.
 To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask.
@@
 We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles.
-The typical running time for one realization is $\sim 600$s using 32 \texttt{OpenMP} threads.
-Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation.
+The typical running time for one realization is $\sim 200$s using 32 \texttt{OpenMP} threads.
+Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in the pipeline.
 More details about timing and scalability of \nmt~can be found in Appendix \ref{app:scalability}.
@@
 
-      The above figure was created using $E/B$ mode purification.
+      The above figure was created using $B$-mode purification.
 Disabling this option has no impact on the bias of the estimates, but increases its variance as we show in the Figure \ref{fig:cmb_val_sigma}.
@@
       \subsection{Pixelization and signal band-limits}\label{app:shts.blim}
-        SHT transforms as written in the Eq.~\ref{eq:sht1} are, strictly speaking, defined on a continuous field on the sky and results in an infinite number of spherical harmonic coefficients.
+        SHT transforms as defined in Eq.~\ref{eq:sht1} are, strictly speaking, defined for continuous fields on the sky and results in an infinite number of spherical harmonic coefficients.
 In practice, we have to always deal with discretized sky maps with a finite number of pixels at which the field is sampled, and consequently use a finite number of spherical coefficients.
@@
 In flat spaces, Nyquist's theorem provides a clear prescription for uniform sampling on a grid.
-Since there is no such natural sampling on a sphere, the situation is more complex and depends on the pixelization scheme used \cite{Driscoll:1994:CFT:184069.184073,DBLP:journals/corr/abs-1110-6297}.
+Since there is no such natural sampling on a sphere, the situation is more complex and depends on the pixelization scheme used \citep{Driscoll:1994:CFT:184069.184073,DBLP:journals/corr/abs-1110-6297}.
 Nevertheless, clear heuristic prescriptions exist.
@@
 Otherwise, the power spectrum estimator will try to ``explain'' the variance associated with $\ell$ modes not considered by artificially inflating the power in the estimated modes.
-For the {\tt HEALPix} pixelization scheme \citep{2005ApJ...622..759G}, commonly used in cosmology and also in this paper, \cite{2013MNRAS.435.1857L} suggest the following prescription.
-If one wants to measure power up to $\ell_{\rm max}$, one should employ pixelisation of \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm max}$ after correcting for pixel-response suppression and  accounting for their contribution to the covariance matrix of measurements.
+For the {\tt HEALPix} pixelization scheme \citep{2005ApJ...622..759G}, commonly used in cosmology and also in this paper, \cite{2013MNRAS.435.1857L} suggest the following prescription: if one wants to measure power up to $\ell_{\rm max}$, one should employ pixelisation of \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm max}$ after correcting for pixel-response suppression and accounting for their contribution to the covariance matrix of measurements.
 
@@
 $\sin\theta\sim 1$).
-In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
-      shifts ($\delta\varphi=\delta y$) the differential operator $\eth$ now takes the form:
+In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude shifts ($\delta\varphi=\delta y$).
+The differential operator $\eth$ now takes the form:
       \begin{equation}

### 2018-09-04 doc/Paper/main.tex
started implementing second round of comments

@@
 Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward, no public, validated and well-maintained code exists supporting all of these capabilities consistently.
-This paper presents a public software package, \nmt, that unifies and extends all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.
+This paper presents a public software package, \nmt, that unifies all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.
+The paper also presents a few novel results that were derived as part of the development of \nmt.
+These include extending the contaminant deprojection formalism to spin-$2$ fields, combining it with $E/B$ purification and providing a consistent derivation of the same functionality in the flat-sky approximation.
 
@@
         &\hat{\sf Y}^{s\dag}_{\ell m}=(-1)^{m+s}\hat{\sf Y}^{-s}_{\ell -m}\\
-        &\int d\nv \hat{\sf Y}^s_{\ell m}\hat{\sf Y}^{s\dag}_{\ell'm'}=\hat\mI\delta_{\ell\ell'}\delta_{mm'},\\
+        &\int d\nv\,\hat{\sf Y}^{s\dag}_{\ell m}\hat{\sf Y}^s_{\ell'm'}=\hat\mI\delta_{\ell\ell'}\delta_{mm'},\\
         &\int d\nv \left(\hat{\sf Y}^{s\dag}_{\bf l}(\nv)\hat{\sf Y}^{s}_{{\bf l}_1}(\nv)\right)\,\hat{\sf Y}^0_{{\bf l}_2}(\nv)\equiv\hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2},

### 2018-09-08 doc/Paper/main.tex

@@
         \end{equation}
-        where ${\sf F}$ is the Fisher matrix, the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, ie.
+        where ${\sf F}$ is the Fisher matrix, the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, i.e.
         \begin{equation}\label{eq:bpws_qmv}
@@
         \end{equation}
-        and the last term in Eq.
+        (which defines the response matrices ${\sf P}_i$) and the last term in Eq.
 \ref{eq:bpws_qmv} corresponds to a bias of the convolved estimates due to the presence of noise.
-Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices ${\sf P}_i$ and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
+Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
 The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix.
@@
 Usually, we know something about these contaminants.
-As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which can be marginalized out.
+As an example, in case of CMB analysis, we might have a map of a given foreground at a different frequency which can be marginalized out.
 But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added PSF fluctuations could be used to remove the possible correlations between measured ellipticities.
@@
 The problem can however be solved self-consistently by using the Moore-Penrose pseudo-inverse \citep{2017MNRAS.465.1847E}.
-This is trivial for a symmetric and positive-definite matrix such as ${\sf F}$, and corresponds to setting the inverse of all of its zero eigenvalues (or all those below a given threshold) to zero.
+This is trivial for a symmetric and positive-definite matrix such as ${\sf F}$, and corresponds to setting the inverse of all of its zero eigenvalues (or all those below a given threshold) to zero\footnote{The choice of eigenvalue threshold in \nmt~is accessible to the user.}.
 

### 2018-09-08 sandbox_validation/check_flat_pure.py
updated all scripts with new synfast

@@ -80,8 +80,9 @@ print(" - Res(x): %.3lf arcmin. Res(y): %.3lf arcmin."%(fmi.lx*60/fmi.nx,fmi.ly*
 print(" - lmax = %d, lmin = %d"%(int(ell_max),int(ell_min)))
 def get_fields() :
     st,sq,su=nmt.synfast_flat(int(fmi.nx),int(fmi.ny),fmi.lx_rad,fmi.ly_rad,
-                              [cltt*beam**2+nltt,clee*beam**2+nlee,
-                               clbb*beam**2+nlbb,clte*beam**2+nlte],pol=True)
+                              np.array([cltt*beam**2+nltt,clte*beam**2+nlte,0*cltt,
+                                        clee*beam**2+nlee,0*clee,
+                                        clbb*beam**2+nlbb]),[0,2])
     st=st.flatten(); sq=sq.flatten(); su=su.flatten()
     if w_cont :
         sq+=np.sum(fgp,axis=0)[0,:]; su+=np.sum(fgp,axis=0)[1,:];
@@ -98,7 +99,7 @@ def get_fields() :
 np.random.seed(1000)
 print("Fielding")
 f2=get_fields()
-    
+
 #Use initial fields to generate coupling matrix
 w22=nmt.NmtWorkspaceFlat();
 if not os.path.isfile(prefix+"_w22.dat") :

### 2018-09-08 sandbox_validation/plot_flat.py

@@ -33,7 +33,7 @@ nltt[0]=0; nlee[0]=0; nlbb[0]=0; nlte[0]=0;
 fmi,mask_hsc=fm.read_flat_map("data/mask_lss_flat.fits")
 dum,fgt=fm.read_flat_map("data/cont_lss_dust_flat.fits") #Dust
 st,sq,su=nmt.synfast_flat(int(fmi.nx),int(fmi.ny),fmi.lx_rad,fmi.ly_rad,
-                          [cltt+nltt,clee+nlee,clbb+nlbb,clte+nlte],pol=True)
+                          [cltt+nltt,clte+nlte,0*cltt,clee+nlee,0*clee,clbb+nlbb],[0,2])
 mask_hsc[:]=1
 ff0=nmt.NmtFieldFlat(fmi.lx_rad,fmi.ly_rad,mask_hsc.reshape([fmi.ny,fmi.nx]),
                      [st+fgt.reshape([fmi.ny,fmi.nx])],
@@ -54,8 +54,6 @@ fmi.view_map(ff0.get_maps().flatten(),ax=ax,addColorbar=False,title='Cleaned map
              colorMin=np.amin(st),colorMax=np.amax(st))
 plt.savefig("plots_paper/maps_contamination.pdf",bbox_inches='tight')
 plt.show()
-
-'''
 
 print("Reading")
 clTT_clean=[]; clTE_clean=[]; clTB_clean=[]; clEE_clean=[]; clEB_clean=[]; clBB_clean=[];
@@ -255,4 +253,3 @@ for a in ax :
 ax[0].legend(loc='upper left',fontsize=12,frameon=False)
 plt.savefig("plots_paper/val_chi2_lss_flat.pdf",bbox_inches='tight')
 plt.show()
-'''

### 2018-09-08 test/benchmarks/generate_benchmarks.py

@@ -57,7 +57,7 @@ wcs,msk=read_flat_map("msk_flat.fits")
 lx=np.fabs(nx*wcs.wcs.cdelt[0])*np.pi/180
 ly=np.fabs(ny*wcs.wcs.cdelt[1])*np.pi/180
 dt,dq,du=nmt.synfast_flat(int(nx),int(ny),lx,ly,
-                          [cltt+nltt,clee+nlee,clbb+nlbb,clte+nlte],pol=True)
+                          [cltt+nltt,clte+nlte,0*cltt,clee+nlee,0*clee,clbb+nlbb],[0,2])
 write_flat_map("mps_flat.fits",np.array([dt,dq,du]),wcs,["T","Q","U"])
 d_ell=20; lmax=500.; ledges=np.arange(int(lmax/d_ell)+1)*d_ell+2
 _,st=read_flat_map("tmp_flat.fits",0)

### 2018-09-08 test/sample_covariance.py

@@ -36,16 +36,16 @@ def compute_master(f_a,f_b,wsp) :
     return cl_decoupled
 
 #Let's generate one particular sample and its power spectrum.
-print "Field"
+print("Field")
 f0=get_sample_field()
 b=nmt.NmtBin(nside,nlb=20) #We will use 20 multipoles per bandpower.
-print "Workspace"
+print("Workspace")
 w=nmt.NmtWorkspace()
 w.compute_coupling_matrix(f0,f0,b)
 cl_0=compute_master(f0,f0,w)[0]
 
 #Let's now compute the Gaussian estimate of the covariance!
-print "Covariance"
+print("Covariance")
 #First we generate a NmtCovarianceWorkspace object to precompute
 #and store the necessary coupling coefficients
 cw=nmt.NmtCovarianceWorkspace()
@@ -53,12 +53,12 @@ cw.compute_coupling_coefficients(w,w) #<- This is the time-consuming operation
 covar=nmt.gaussian_covariance(cw,clarr,clarr,clarr,clarr)
 
 #Let's now compute the sample covariance
-print "Sample covariance"
+print("Sample covariance")
 nsamp=100
 covar_sample=np.zeros([len(cl_0),len(cl_0)])
 mean_sample=np.zeros(len(cl_0))
 for i in np.arange(nsamp) :
-    print i
+    print(i)
     f=get_sample_field()
     cl=compute_master(f,f,w)[0]
     covar_sample+=cl[None,:]*cl[:,None]

### 2018-09-08 test/sample_pureB.py

@@ -53,7 +53,7 @@ def compute_master(f_a,f_b,wsp) :
 #We now iterate over several simulations, computing the power spectrum for each of them
 data_np=[]; data_yp=[];
 for i in np.arange(nsim) :
-    print i,nsim
+    print(i,nsim)
     fnp,fyp=get_fields()
     data_np.append(compute_master(fnp,fnp,w_np))
     data_yp.append(compute_master(fyp,fyp,w_yp))

### 2018-09-08 doc/Paper/main.tex
modified acknoweldgements to conform with PB guidelines

@@
 \subsection*{Acknowledgements}
-  We would like to thank Rupert Allison, Scott Dodelson, Jo Dunkley, Franz Elsner, Boris Leistedt, Francois Lanusse, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya Peiris, Naomi Robertson and Ben Thorne for useful comments and discussions.
+  This paper has undergone internal review in the LSST Dark Energy Science Collaboration.
+The internal reviewers were Scott Dodelson, Francois Lanusse and Boris Leistedt.
+Besides them, we would like to thank Rupert Allison, Jo Dunkley, Franz Elsner, Daniel Lenz, Thibaut Louis, Mathew Madhavacheril, Sigurd N\ae ss, Hiranya Peiris, Naomi Robertson and Ben Thorne for useful comments and discussions.
+
+  \input{contributions}
+
   DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
-
-  \input{contributions}
-
   \input{desc-tex/ack/standard}

### 2018-09-14 doc/Paper/main.tex
added references to pcl code repos

@@
 
-    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g.
+    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T,2018arXiv180702484V} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g.
 \cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}).
-For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
+For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C,2005MNRAS.358..833T} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation).
 This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
-Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward, no public, validated and well-maintained code exists supporting all of these capabilities consistently.
+Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward\footnote{See, for instance: \href{http://www2.iap.fr/users/hivon/software/PolSpice/}{PolSpice}, \href{https://gitlab.in2p3.fr/tristram/Xpol}{Xpol}, \href{https://gitlab.in2p3.fr/tristram/Xpure}{Xpure} and \href{https://github.com/thibautlouis/hyperQube}{hyperQube}.}, no public, validated and well-maintained code exists supporting all of these capabilities consistently.
 This paper presents a public software package, \nmt, that unifies all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.
@@
       \subsubsection{Standard pseudo-$C_\ell$}\label{sssec:maths.flat.pcl101}
-        Using the definitions presented in Appendix \ref{app:flat}, it is easy to show that the Fourier coefficients of a masked field are given by
+        Using the definitions presented in Appendix \ref{app:flat}, it is easy to show that the Fourier coefficients of a masked field are given by \citep{2013MNRAS.435.2040L}
         \begin{equation}\label{eq:pcl_flat_masked}
@@
       \item In flat sky, SHTs are replaced by 2D DFTs, for which \nmt~uses the {\tt FFTW} library~\citep{FFTW05}.
+      \item The GNU Scientific Library \cite{gslbook} is also used for some numerical calculations.
     \end{itemize}
@@
  \section{Performance and scalability}\label{app:scalability}
+   \begin{figure}
+     \centering
+     \includegraphics[width=0.99\columnwidth]{time_scaling}
+     \includegraphics[width=0.99\columnwidth]{timing}
+     \caption{{\sl Top panel}: computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{\rm max}$.
+We include the expected scaling $\propto\ell_{\rm max}$ for the pseudo-$C_{\ell}$ estimator (solid red line) for comparison.
+{\sl Bottom panel:} execution time for different tasks associated with the pseudo-$C_\ell$ estimator.
+We show times for fields with resolution $N_{\rm side}=1024$, which can be extrapolated to other resolutions using a $\propto N_{\rm side}^3$ scaling, as described above.
+The description of the different tasks shown is given in the main text.}\label{fig:scaling_lmax}
+   \end{figure}
    One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{\rm max}^{3})$, where $\ell_{\rm max}$ is the maximum multipole at which the power-spectra are calculated.
@@
 
-     \begin{figure}
-       \centering
-       \includegraphics[width=0.99\columnwidth]{time_scaling}
-       \includegraphics[width=0.99\columnwidth]{timing}
-       \caption{{\sl Top panel}: computing time for the auto-power spectra of curved-sky spin-0 (data 00, hollow circles), and spin-2 (data 22, solid triangles) contamination-free maps, as wells as their cross-power (data 02, solid circles) as a function of $\ell_{\rm max}$.
-We include the expected scaling $\propto\ell_{\rm max}$ for the pseudo-$C_{\ell}$ estimator (solid red line) for comparison.
-{\sl Bottom panel:} execution time for different tasks associated with the pseudo-$C_\ell$ estimator.
-We show times for fields with resolution $N_{\rm side}=1024$, which can be extrapolated to other resolutions using a $\propto N_{\rm side}^3$ scaling, as described above.
-The description of the different tasks shown is given in the main text.}\label{fig:scaling_lmax}
-     \end{figure}
-
 \bsp

### 2018-09-14 .travis/install_deps.sh
modified travis files to fix small build error

@@ -15,7 +15,6 @@ cd libsharp
 autoconf -i
 ./configure --enable-pic
 make
-make install
 mv auto/bin $TRAVIS_BUILD_DIR
 mv auto/lib $TRAVIS_BUILD_DIR
 mv auto/include $TRAVIS_BUILD_DIR

### 2018-09-18 doc/Paper/main.tex
updated contentious sentence

@@
 This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}.
-Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward\footnote{See, for instance: \href{http://www2.iap.fr/users/hivon/software/PolSpice/}{PolSpice}, \href{https://gitlab.in2p3.fr/tristram/Xpol}{Xpol}, \href{https://gitlab.in2p3.fr/tristram/Xpure}{Xpure} and \href{https://github.com/thibautlouis/hyperQube}{hyperQube}.}, no public, validated and well-maintained code exists supporting all of these capabilities consistently.
+Due to the advantages of this method, a number of useful, publicly available codes currently exist to carry out some of these calculations \footnote{See, for instance: \href{http://www2.iap.fr/users/hivon/software/PolSpice/}{PolSpice}, \href{https://gitlab.in2p3.fr/tristram/Xpol}{Xpol}, \href{https://gitlab.in2p3.fr/tristram/Xpure}{Xpure} and \href{https://github.com/thibautlouis/hyperQube}{hyperQube}.}.
+Unfortunately, no public, validated and maintained code exists simultaneously supporting all of the capabilities described above in a consistent manner.
 This paper presents a public software package, \nmt, that unifies all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.

### 2018-09-18 .travis/install_deps.sh
changed url to GSL dependency to fix travis

@@ -22,5 +22,5 @@ cd ..
 
 #### Install GSL2.0+ ####
 
-wget http://mirror.rise.ph/gnu/gsl/gsl-2.4.tar.gz && tar xzf gsl-2.4.tar.gz && cd gsl-2.4 &&  ./configure --enable-shared && make && sudo make install && cd ..
+wget ftp://ftp.gnu.org/gnu/gsl/gsl-2.4.tar.gz && tar xzf gsl-2.4.tar.gz && cd gsl-2.4 &&  ./configure --enable-shared && make && sudo make install && cd ..
 

### 2018-09-18 doc/Paper/main.tex

@@
           &M^{0+}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0},\\\label{eq:m22}
-          &{\sf M}^{22}=\left(
+          &{\sf M}^{22}_{\ell\ell'}=\left(
                         \begin{array}{cccc}

### 2018-09-18 .travis/install_deps.sh
Updated to gsl-2.5 and added JS acks

@@ -22,5 +22,5 @@ cd ..
 
 #### Install GSL2.0+ ####
 
-wget ftp://ftp.gnu.org/gnu/gsl/gsl-2.4.tar.gz && tar xzf gsl-2.4.tar.gz && cd gsl-2.4 &&  ./configure --enable-shared && make && sudo make install && cd ..
+wget ftp://ftp.gnu.org/gnu/gsl/gsl-2.5.tar.gz && tar xzf gsl-2.5.tar.gz && cd gsl-2.5 &&  ./configure --enable-shared && make && sudo make install && cd ..
 

### 2018-09-18 .travis/install_deps.sh
Added passive ftp option

@@ -22,5 +22,5 @@ cd ..
 
 #### Install GSL2.0+ ####
 
-wget ftp://ftp.gnu.org/gnu/gsl/gsl-2.5.tar.gz && tar xzf gsl-2.5.tar.gz && cd gsl-2.5 &&  ./configure --enable-shared && make && sudo make install && cd ..
+wget ftp://ftp.gnu.org/gnu/gsl/gsl-2.5.tar.gz --passive-ftp && tar xzf gsl-2.5.tar.gz && cd gsl-2.5 &&  ./configure --enable-shared && make && sudo make install && cd ..
 

### 2019-02-13 doc/Paper/main.tex
Updated to gsl-2.5 and added JS acks

@@
-\RequirePackage{docswitch}
-\setjournal{\flag}
-
 \documentclass[usenatbib]{mnrasb}
+\bibliographystyle{mnrasb}
+\pdfoutput=1
+\pdfminorversion=5
+
+%\documentclass[usenatbib]{mnrasb}
 
 % Custom commands from LSST DESC, see texmf/styles/lsstdesc_macros.sty
-\usepackage{lsstdesc_macros}
+%\usepackage{lsstdesc_macros}
 \usepackage{graphicx}
@@
 
-\usepackage{lipsum}
-\usepackage{lineno}
-\linenumbers
-
-\usepackage{etoolbox}
-
-\newtoggle{DEBUG}
-\toggletrue{DEBUG}
-
-\iftoggle{DEBUG}{
-\newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
-\newcommand{\as}[1]{{\textcolor{red}{[\textbf{AS}: #1]}}}
-\newcommand{\dam}[1]{{\textcolor{green}{[\textbf{DA}: #1]}}}
-\newcommand{\js}[1]{{\textcolor{orange}{[\textbf{JS}: #1]}}}
-\newcommand{\todo}[1]{{\bf TODO: #1}}
-}{
-\newcommand{\chg}[1]{}
-\newcommand{\as}[1]{}
-\newcommand{\dam}[1]{}
-\newcommand{\js}[1]{}
-\newcommand{\todo}[1]{}
-}
+\title[A unified pseudo-$C_\ell$ framework]{A unified pseudo-$C_\ell$ framework}
+  \author[D. Alonso et al.]{David Alonso$^{1,2}$\thanks{david.alonso@physics.ox.ac.uk}, Javier Sanchez$^{3}$, An\v{z}e Slosar$^{4}$ \newauthor{(The LSST Dark Energy Science Collaboration)}\\
+  $^{1}$ School of Physics and Astronomy, Cardiff University, The Parade, Cardiff, CF24 3AA, United Kingdom\\
+  $^{2}$ Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom\\
+  $^{3}$ Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA\\
+  $^{4}$ Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}
 
 \begin{document}
-
-\title{A unified pseudo-$C_\ell$ framework}
-
-\maketitlepre
+  \date{\today}
+  \pagerange{1--27} \pubyear{2018}
+  \maketitle
 
@@
 
-\dockeys{methods: data analysis -- methods: numerical}
-
-\maketitlepost
+  \begin{keywords}
+    methods: data analysis -- methods: numerical
+  \end{keywords}
 
   \section{Introduction}\label{sec:intro}
-    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}.
+    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1994ApJ...430L..85G,1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}.
 Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier-space\footnote{In this paper we use the term ``Fourier'' to describe both the flat space Fourier expansion as well as harmonic-space expansion on the sphere.
@@
         In the standard incarnation of the optimal quadratic estimator~\citep{1998PhRvD..57.2117B}, the measurement is produced on inverse covariance weighted data, namely
-        \begin{equation}
+        \begin{equation}\label{eq:fisher_bpw}
           {\sf F_{ij}} B_j = \frac{1}{2}\va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
@@
         (which defines the response matrices ${\sf P}_i$) and the last term in Eq.
-\ref{eq:bpws_qmv} corresponds to a bias of the convolved estimates due to the presence of noise.
+\ref{eq:fisher_bpw} corresponds to a bias of the convolved estimates due to the presence of noise.
 Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix.
@@
         \end{align}
-        where in the second line we have taken the expectation value of first one.
+        where in the second line we have taken the expectation value of the first one.
 One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function).
@@
         \begin{equation}
-          {\cal M}^{s_as_b}_{qq'}=\sum_{\ell\in\vec{\ell}_q}\sum_{\ell'\in\vec{\ell}_{q'}}w_q^\ell\,{\sf M}^{s_as_b}_{\ell\ell'}.
+          {\cal M}^{s_as_b}_{qq'} \equiv \sum_{\ell\in\vec{\ell}_q}\sum_{\ell'\in\vec{\ell}_{q'}}w_q^\ell\,{\sf M}^{s_as_b}_{\ell\ell'}.
         \end{equation}
@@
 Usually, we know something about these contaminants.
-As an example, in case of CMB analysis, we might have a map of a given foreground at a different frequency which can be marginalized out.
-But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added PSF fluctuations could be used to remove the possible correlations between measured ellipticities.
+As an example, in case of CMB analysis, we might have a map of a given foreground at a different frequency which can be marginalized out\footnote{It is worth noting that the formalism described here assumes that a given contaminant template has no significant contribution from the signal itself.
+This might not be the case, for example, if one blindly took a high-frequency sky map to serve as a template of dust, since that may still contain a significant CMB contribution.
+Care must therefore be exercised when defining the contaminant templates to marginalize over.}.
+But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added point spread function (PSF) fluctuations could be used to remove the possible correlations between measured ellipticities.
 Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as that relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}).
@@
     \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
-      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa\footnote{Note that this is a different effect from the leakage between different polarized and unpolarized channels in CMB observations due to instrumental effects (asymmetric beams, pair differencing, polarization angle calibration etc.)}.
-The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB.
+      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa\footnote{Note that this is a different effect from the leakage between different polarized and unpolarized channels in CMB observations due to instrumental effects (asymmetric beams, pair differencing, polarization angle calibration etc.)}.
+The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in for the CMB.
 In this case, removing the leakage at the power spectrum level (i.e.
@@
       per purified field.
+
+      It is worth noting that, as described in \cite{2009PhRvD..79l3515G}, the performance of $E/B$ purification can be optimized by an appropriate choice of window function.
+The corresponding optimal spin-0, spin-1 and spin-2 window functions can be estimated using a preconditioned conjugate-gradient method.
+The resulting spin window functions are no longer related by simple covariant derivatives.
+The performance of this method, compared with simple apodization methods, will depend on the noise properties of the experiment and the footprint geometry.
+This optimal method is currently not available in \nmt, and will be implemented in future releases of the code.
 
@@
         \subsubsection{Beam deconvolution}\label{sssec:maths.beam_noise.beam}
-          Including the effect of an spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
+          Including the effect of a spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
           \begin{equation}
@@
           \centering
-          \includegraphics[width=0.9\textwidth]{maps_lss_sph}
+          \includegraphics[width=0.99\textwidth]{maps_lss_sph}
           \caption{Inputs for the large-scale structure, curved-sky validation set.
@@
 
-        For this validation suite we therefore use a mock dataset mimicking a typical observation of an LSST-like survey in a particular redshift bin.
+        For this validation suite we therefore use a mock dataset mimicking a typical observation of a LSST-like survey in a particular redshift bin.
 Slightly different set-ups are used for the curved-sky and flat-sky validation suites.
@@
           \item {\bf Noise:} for the curved-sky case we generate inhomogeneous noise realizations.
-To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is then intrinsic ellipticity scatter per component.
+To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is the intrinsic ellipticity scatter per component.
 These noise realizations are then locally rescaled by a factor $f(\nv)=f_{\rm star}(\nv)\,f_{\rm depth}(\nv)$ made up of two components:
@@
           \centering
-          \includegraphics[width=0.9\textwidth]{maps_cmb}
+          \includegraphics[width=0.99\textwidth]{maps_cmb}
           \caption{{\sl Left panels:} mask (top), and $Q$ and $U$ components (middle, bottom) of the foreground contaminant used in the CMB curved-sky validation set.
-{\sl Right panels:} same as left panels for the flat-sky validation set.}
+{\sl Right panels:} same as left panels for the flat-sky validation set.
+All maps have units of $\mu K_{\rm CMB}$.}
           \label{fig:cmb_mocks}
@@
         \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines).
-The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
+The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly superposed on the solid line (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$).
 The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero.
@@
 The residuals are scattered around zero with random fluctuations of up to $\sim3\sigma$.
-{\sl Right:} exact bandpower window functions for the fully accounting for the effects of mode-coupling (solid black) compared with the input top-hat bandpower weights (red).
+{\sl Right:} exact bandpower window functions (solid black), which fully account for the effects of mode-coupling, compared with the input top-hat bandpower weights (red).
 Using the correct window functions is in general important in order to avoid parameter biases.
@@
 In our fiducial run we both deprojected the foreground contaminants and subsequently used $B$-mode purification to decouple $E$ and $B$-modes at the map level, since this is the main feature of the code we want to validate.
-The left panel of Figure \ref{fig:cmb_val_cls} shows the mean of the $EE$, $EB$ and $BB$ power spectra computed for all simulations (solid lines) compared with the expectation from the input spectra for $EE$ and $BB$ (dashed lines, mostly lying under the solid ones).
+The left panel of Figure \ref{fig:cmb_val_cls} shows the mean of the $EE$, $EB$ and $BB$ power spectra computed for all simulations (solid lines) compared with the expectation from the input spectra for $EE$ and $BB$ (dashed lines, mostly superposed on the solid ones).
 The average residuals normalized over the error on the mean, shown in the lower panel of the same figure, fluctuate around zero with an amplitude of up to $3\sigma$, as would be expected in the absence of a bias in the estimator.
@@
         \includegraphics[width=0.99\columnwidth]{val_cl_cmb_flat}
-        \caption{{\sl Left and Right:} the analogues of the left panels of Figures \ref{fig:lss_val_cls} and \ref{fig:cmb_cls} for the flat-sky LSS and CMB validation sets respectively.} \label{fig:val_cls_flat}
+        \caption{{\sl Left and Right:} the analogues of the left panels of Figures \ref{fig:lss_val_cls} and \ref{fig:cmb_val_cls} for the flat-sky LSS and CMB validation sets respectively.} \label{fig:val_cls_flat}
       \end{figure*}
@@
 \ref{eq:bpresponse} differs in general from a simple binning operator, even if the bandpower weights in Eq.
-\ref{eq:bpws} are chosen to be top-hat functions with a fixed width $\Delta\ell$.
+\ref{eq:bpws} are chosen to be top-hat functions with a fixed width $\Delta\ell=12$.
 This is illustrated in the right panel of Figure \ref{fig:cmb_val_cls}, which shows, in black, the exact window function for different bandpowers (solid black), in comparison with the input top-hat bandpower weights (solid red).
@@
 
-  \input{contributions}
+  Author contributions are listed below. \\
+  David Alonso: Co-led project; wrote the code; contributed to validation. \\
+  Javier Sanchez: Co-led project; wrote code; contributed to validation. \\
+  An\v{z}e Slosar: Co-led project; contributed to validation. \\
 
   DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
-  \input{desc-tex/ack/standard}
-
-\bibliographystyle{mnrasb}
+JS acknowledges support from the United States Department of Energy.
+The DESC acknowledges ongoing support from the Institut National de Physique Nucl\'eaire et de Physique des Particules in France; the Science \& Technology Facilities Council in the United Kingdom; and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
+DESC uses resources of the IN2P3 Computing Center (CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la Recherche Scientifique; the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.\ Department of Energy under Contract No.\ DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants; and the UK particle physics grid, supported by the GridPP Collaboration.
+This work was performed in part under DOE Contract DE-AC02-76SF00515.
+
+\setlength{\bibhang}{2.0em}
+\setlength\labelwidth{0.0em}
 \bibliography{main}
